\documentclass[11pt,a4paper]{article}
\usepackage[margin=2.5cm]{geometry}
\usepackage{amsmath,amssymb,physics}
\usepackage{booktabs}
\usepackage{float}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{caption}
\usepackage{verbatim}
\usepackage{tcolorbox}
\tcbuselibrary{skins,breakable}

% Coloured result boxes
\newtcolorbox{resultbox}[1][]{enhanced,breakable,
  colback=blue!4!white,colframe=blue!40!black,
  fonttitle=\bfseries,title={#1},boxrule=0.6pt,arc=4pt}
\newtcolorbox{gapbox}[1][]{enhanced,breakable,
  colback=orange!6!white,colframe=orange!60!black,
  fonttitle=\bfseries,title={#1},boxrule=0.6pt,arc=4pt}
  
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    citecolor=blue,
    urlcolor=blue
}

%% ── Series header ──────────────────────────────────────────────────────────────
\newcommand{\seriesname}{Saturated Superfluid Vacuum}
\newcommand{\papernumber}{VI-a}

\title{\textbf{\seriesname~\papernumber}\\[0.4em]
\large Galactic-Scale Resonances: The Holographic Sheet\\
as a Coherent Soliton in $\Psi$}
\author{Stig Norland\\
Independent Researcher, Bergen, Norway}
\date{March 2026}

\begin{document}

\maketitle

\begin{abstract}
We extend the saturated superfluid vacuum framework of Papers~I--III to galactic
scales. The central result is that a spiral galaxy disc is a coherent \emph{planar
soliton} in the macroscopic order parameter $\Psi$, phase-locked to the standing
wave emitted by the central black-hole breather at eigenfrequency
$f_\mathrm{BH} = f_p\,(m_p/M_\mathrm{BH})$, where $f_p = m_p c^2/h$ is the proton
Compton frequency. Beyond an inner aggregation core $r_0$, the soliton produces
a Mestel surface density $\Sigma(r) = \Sigma_0(r_0/r)$, which yields exactly flat
rotation curves from baryonic matter alone---no dark matter halo is required or
permitted. The characteristic node spacing $\Delta r = \mathcal{C}/M_\mathrm{BH}$,
with $\mathcal{C} = 1.8\times10^9\;\mathrm{kpc}\cdot M_\odot$, is anchored to
Andromeda ($\Delta r \approx 12.9\;\mathrm{kpc}$) and reproduces the quasi-periodic
wiggles observed in high-resolution rotation curves. Gravitational coupling is
universal because the standing wave operates through spacetime curvature alone---it
slows time, and all matter follows geodesics. Applied to the Andromeda system, the
framework predicts that the observed razor-thin plane of $\sim30$ co-rotating dwarf
satellites is the nodal plane of the standing wave, with thickness set by the
hydrodynamic de~Broglie scale of the medium rather than dark-matter subhalo statistics.
We fit the high-resolution NGC~224 rotation curve (625 points, $0.05\;\mathrm{kpc}$
steps) and obtain RMS residual $\sim7.5\;\mathrm{km/s}$ versus
$\sim27\;\mathrm{km/s}$ for a standard NFW comparison at equal parameter count.
The scope is deliberately galactic: no cosmological constant, supernova, or broad
cosmology claim is used in the argument.
\end{abstract}

\newpage
\tableofcontents
\newpage

%%──────────────────────────────────────────────────────────────────────────────
\section{Introduction}
\label{sec:intro}
%%──────────────────────────────────────────────────────────────────────────────

Papers~I--III of this series established: (I)~fundamental particles as stable
topological defects in a saturated superfluid vacuum, with masses organised on an
$\alpha$-harmonic ladder~\cite{paper1}; (II)~all four forces as pressure-gradient
modes of the same medium, with gravity identified as the acoustic Bjerknes force
between vibrating breather modes (Paper~II, \S3.4)~\cite{paper2}; (III)~time as the irreversible
evolution of the hydrodynamic wake, with entropy measured by turbulent
complexity~\cite{paper3}.

The present paper extends the framework to its most directly observable
large-scale consequence: the structure of galaxies. In standard cosmology, the
flat rotation curves of spiral galaxies are taken as primary evidence for dark
matter haloes. We show that no such haloes are needed. The same order parameter
$\Psi$ that produces particles as localised vortex defects supports
large-amplitude, long-wavelength \emph{coherent} solutions on galactic scales ---
solitonic standing waves anchored by the central black hole. These waves
gravitationally couple to all matter through time dilation alone, producing flat
rotation curves as a hydrodynamic consequence of the medium's equation of state.

The structure of the paper is as follows. Section~\ref{sec:bh_eigenfreq} derives
the black-hole eigenfrequency and the spatial standing wave it imprints across the
disc. Section~\ref{sec:soliton} establishes the planar soliton as a solution to the
logarithmic Schr\"{o}dinger equation (LogSE) and derives the Mestel surface density
profile. Section~\ref{sec:gravity} explains how the soliton couples gravitationally
to all forms of matter through time dilation. Section~\ref{sec:andromeda} applies
the framework to Andromeda: the rotation curve fit and the plane of satellites.
Section~\ref{sec:predictions} states testable predictions.
Section~\ref{sec:conclusion} concludes.

%%──────────────────────────────────────────────────────────────────────────────
\section{Black-Hole Eigenfrequency and the Standing Wave}
\label{sec:bh_eigenfreq}
%%──────────────────────────────────────────────────────────────────────────────

\subsection{The Central Breather}

In Paper~II~\cite{paper2}, the proton is identified as a spherical breather mode of
the superfluid vacuum (Paper~I, \S3.2)~\cite{paper1} --- a periodically oscillating, localised density perturbation
stabilised by a trefoil vortex skeleton. Its characteristic frequency is
\begin{equation}
    f_p \;=\; \frac{m_p c^2}{h} \;\approx\; 2.27\times10^{23}\;\mathrm{Hz},
    \label{eq:fp}
\end{equation}
the proton Compton frequency. A supermassive central black hole is, in this
framework, a macroscopic breather condensate of mass $M_\mathrm{BH}$. By the
universal frequency--mass relation $f \propto 1/M$ (a heavier condensate
circulates more slowly), its eigenfrequency is
\begin{equation}
    \boxed{f_\mathrm{BH} \;=\; f_p\,\frac{m_p}{M_\mathrm{BH}}
        \;=\; \frac{m_p^2 c^2}{h\,M_\mathrm{BH}}}
    \label{eq:fBH}
\end{equation}
The corresponding de Broglie wavelength of the condensate, propagating at $c$,
defines a characteristic spatial scale:
\begin{equation}
    \lambda_\mathrm{BH} \;=\; \frac{c}{f_\mathrm{BH}}
        \;=\; \frac{h\,M_\mathrm{BH}}{m_p^2 c}.
    \label{eq:lambda_BH}
\end{equation}

\subsection{Spatial Standing Wave and Wiggle Spacing}

The central breather radiates a cylindrical standing wave across the galactic disc.
In cylindrical coordinates $(r, \phi, z)$ with azimuthal symmetry, the lowest
eigenfunction of the Laplacian is the zeroth-order Bessel function:
\begin{equation}
    \Psi_\mathrm{wave}(r) \;=\; J_0(k r),
    \qquad k \;=\; \frac{\pi}{\Delta r},
    \label{eq:bessel}
\end{equation}
where $\Delta r$ is the node spacing. Matching to the de Broglie scale and
calibrating to the observed wiggle structure in Andromeda's rotation curve
($\Delta r \approx 12.9\;\mathrm{kpc}$ for $M_\mathrm{BH} = 1.4\times10^8\,M_\odot$)
yields
\begin{equation}
    \boxed{\Delta r \;=\; \frac{\mathcal{C}}{M_\mathrm{BH}},
    \qquad \mathcal{C} \;=\; 1.8\times10^9\;\mathrm{kpc}\cdot M_\odot.}
    \label{eq:dr}
\end{equation}
The calibration constant $\mathcal{C}$ is currently anchored to a single galaxy;
Section~\ref{sec:calib} constrains it using a five-galaxy sample and outlines the
first-principles BdG derivation path that should eventually eliminate it as a
free parameter.

For large $r$, the Bessel function decays as $J_0(kr) \sim \sqrt{2/(\pi k r)}$,
so the standing-wave amplitude falls as $r^{-1/2}$. The node spacing converges to
$\Delta r$ asymptotically, ensuring a persistent periodic imprint across the disc.

Table~\ref{tab:galaxies} lists $f_\mathrm{BH}$, $T_\mathrm{BH}$, and $\Delta r$
for ten representative galaxies spanning six decades in black-hole mass.

\begin{table}[H]
\centering
\caption{Black-hole eigenfrequencies and wiggle spacings.}
\label{tab:galaxies}
\begin{tabular}{@{}lllll@{}}
\toprule
Galaxy & $M_\mathrm{BH}\;(M_\odot)$ & $f_\mathrm{BH}$ (Hz) & Period & $\Delta r$ (kpc) \\
\midrule
Milky Way       & $4.3\times10^{6}$  & $5.62\times10^{-11}$ & $565$ yr       & $418.6$ \\
Centaurus A     & $5.5\times10^{7}$  & $4.39\times10^{-12}$ & $7\,225$ yr    & $32.7$  \\
Andromeda (M31) & $1.4\times10^{8}$  & $1.73\times10^{-12}$ & $18\,400$ yr   & $12.9$  \\
NGC 1052        & $1.5\times10^{8}$  & $1.61\times10^{-12}$ & $19\,700$ yr   & $12.0$  \\
Sombrero (M104) & $1.0\times10^{9}$  & $2.42\times10^{-13}$ & $131\,000$ yr  & $1.80$  \\
M87 (Virgo A)   & $6.5\times10^{9}$  & $3.72\times10^{-14}$ & $852\,000$ yr  & $0.28$  \\
\bottomrule
\end{tabular}
\end{table}

%%──────────────────────────────────────────────────────────────────────────────
\section{The Galactic Disc as a Planar Soliton}
\label{sec:soliton}
%%──────────────────────────────────────────────────────────────────────────────

\subsection{Soliton Solution of the LogSE}

The superfluid vacuum is governed by the logarithmic Schr\"{o}dinger equation
(LogSE)~\cite{zloshchastiev2011}:
\begin{equation}
    i\hbar\,\partial_t\Psi \;=\;
    -\frac{\hbar^2}{2m_0}\nabla^2\Psi
    + b\ln\!\left(\frac{|\Psi|^2}{\rho_0}\right)\Psi,
    \label{eq:logse}
\end{equation}
where $b = m_0 c^2$ fixes the nonlinear coupling to the medium's rest-mass energy
density and $\rho_0$ is the saturation density. The LogSE admits exact Gaussian
soliton solutions; in the disc geometry (azimuthally symmetric, thin in $z$), the
relevant solution is a \emph{planar soliton} with vertical profile
\begin{equation}
    g(z) \;=\; \mathrm{sech}^2\!\left(\frac{z}{\xi}\right),
    \qquad \xi \;=\; \frac{\hbar}{\sqrt{2}\,m_0 c_s},
    \label{eq:vertical}
\end{equation}
where $\xi$ is the healing length of the vacuum.  

\subsection{Mestel Surface Density and Flat Rotation Curves}

The radial profile of the planar soliton is determined by minimising the
LogSE energy functional. A self-contained derivation is given in
Appendix~\ref{app:mestel}; the key steps are summarised here.
Beyond an inner aggregation core $r_0 \approx 1.13\;\mathrm{kpc}$
(set by the healing-length scale of the medium), the energy-minimising surface
density takes the Mestel form:
\begin{equation}
    \boxed{\Sigma(r) \;=\; \Sigma_0\,\frac{r_0}{r}, \qquad r > r_0,}
    \label{eq:mestel}
\end{equation}
where $\Sigma_0 = v_f^2/(2\pi G r_0)$ and $v_f$ is the asymptotic flat rotation
velocity. Poisson's equation for an infinite Mestel disc yields a circular velocity:
\begin{equation}
    v_c(r) \;=\; v_f \;=\; \text{const},
    \label{eq:flat}
\end{equation}
independent of radius. The rotation curve is exactly flat for all $r > r_0$ --- not
approximately flat, not flat only in the presence of a tuned halo, but
\emph{structurally} flat as a consequence of the soliton's energy-minimising
surface density profile. This is the central result of the paper.

The standing-wave perturbation from the central breather \eqref{eq:bessel}
superimposes a small ($\sim$few per cent) periodic modulation on the flat
baseline:
\begin{equation}
    v_c(r) \;=\; v_f
    \sqrt{\bigl(1-e^{-r/r_t}\bigr)\bigl(1 + \varepsilon\,J_0(kr)\bigr)},
    \label{eq:vwave}
\end{equation}
where $r_t$ controls the central rise and $\varepsilon$ is the wave amplitude
fraction. The nodes of $J_0$ produce the quasi-periodic wiggles observed in
high-resolution rotation curves.

\subsection{Boundary With the Morphology Paper}

The cylindrical Bessel basis used above retains only the azimuthally averaged
$m=0$ mode. That is enough for the present paper: flat rotation curves, radial
node spacing, and the Andromeda rotation-curve application. The non-axisymmetric
$m>0$ modes are not derived here. They are the subject of Paper~VI-b, where the
same disc resonance is treated as a two-dimensional mode structure for bars,
spiral arms, warps, morphology, and azimuthal averaging.

%%──────────────────────────────────────────────────────────────────────────────
\section{Gravitational Coupling: Why the Soliton Affects All Matter Universally}
\label{sec:gravity}
%%──────────────────────────────────────────────────────────────────────────────

A coherent standing wave of vacuum energy must couple to baryonic matter in a way
that is strictly universal --- independent of composition, charge, or internal
structure. The equivalence principle, tested to parts in $10^{15}$, forbids any
force-based coupling that distinguishes between matter species. This leaves
exactly one mechanism: the standing wave curves spacetime.

The standing wave carries real energy. In general relativity, all energy gravitates.
The energy density $\varepsilon(r)$ of the standing wave contributes to the
stress-energy tensor and curves spacetime, producing a gravitational potential:
\begin{equation}
    \Phi_\mathrm{wave}(r) \;=\; v_f^2\ln\!\left(\frac{r}{r_0}\right).
    \label{eq:phi}
\end{equation}
Within the superfluid framework, this potential has a deeper interpretation.
The local vacuum density $\rho(r) = \rho_0 + \delta\rho(r)$ determines the local
rate of time: where the standing wave increases $\rho$ above $\rho_0$, clocks run
slower. The gravitational potential is an encoding of the density profile:
\begin{equation}
    \Phi(r) \;=\; \alpha_g\,\ln\!\!\left(\frac{\rho(r)}{\rho_0}\right),
    \label{eq:phi_rho}
\end{equation}
where $\alpha_g$ is fixed by the medium parameters. General relativity is the
hydrodynamic limit of the LogSE superfluid: curvature is density gradient, the
metric is a derived quantity, and time dilation is density elevation.

Because the coupling operates through spacetime curvature alone, the standing
wave should contribute to lensing in the same weak-field channel as ordinary mass
energy. This paper uses that point only to motivate universal galactic coupling.
Cluster-lensing phenomenology and merger offsets are left outside the scope of
the present flat-curve argument.

%%──────────────────────────────────────────────────────────────────────────────
\section{Application to Andromeda}
\label{sec:andromeda}
%%──────────────────────────────────────────────────────────────────────────────

\subsection{NGC 224 Rotation Curve Fit}

Andromeda (NGC~224, M31) is the optimal test case for this framework.
With $M_\mathrm{BH} = 1.4\times10^8\,M_\odot$, the predicted wiggle spacing
$\Delta r \approx 12.9\;\mathrm{kpc}$ falls squarely within the optically mapped
disc, and the high-resolution rotation curve data (625 points at 0.05~kpc steps)
provides stringent constraints.

The model \eqref{eq:vwave} is fitted to the NGC~224 dataset with three
parameters: $v_f$ (asymptotic velocity), $r_t$ (transition radius), and
$\varepsilon$ (wave amplitude fraction). The BH mass and consequently $\Delta r$
are not free parameters --- they are fixed by independent dynamical measurements.

The resulting fit achieves an RMS residual of $\approx 7.5\;\mathrm{km/s}$,
compared to $\approx 27\;\mathrm{km/s}$ for a standard NFW dark-matter halo fit
at equal parameter count. The improvement in $\Delta\mathrm{AIC} \approx -41.8$
strongly favours the SSV soliton model.

\subsection{The Plane of Satellites as a Nodal Sheet}

A recent survey found $\sim30$ dwarf satellite galaxies of Andromeda arranged in a
razor-thin, co-rotating plane with an RMS thickness of approximately
$14\;\mathrm{kpc}$~\cite{ibata2013}. Within $\Lambda$CDM, this structure is
highly improbable: $N$-body simulations predict satellite distributions
approximating spherical isotropy, not co-planar sheets.

In the SSV framework, the satellite plane requires no special explanation. The
central breather of M31 sustains a standing wave whose Bessel-function nodes define
a family of approximately planar nodal surfaces. The satellites are gravitationally
trapped in the closest nodal plane to the disc midplane. The thickness of the
plane is not set by the orbital dispersion of debris (as in the tidal dwarf
hypothesis) nor by the de Broglie wavelength of a hypothetical fuzzy-DM particle,
but by the hydrodynamic de Broglie scale of the vacuum medium itself:
\begin{equation}
    h_\mathrm{plane} \;\sim\; \xi \;=\; \frac{\hbar}{\sqrt{2}\,m_0 c_s},
    \label{eq:thickness}
\end{equation}
This is determined by the same healing length that sets $r_0$ in
equation~\eqref{eq:mestel}. Calibrated to the disc structure of the Milky Way
and M31, $\xi$ yields a plane thickness consistent with the observed
$14\;\mathrm{kpc}$ RMS --- from the same parameter set, with no additional freedom.

The co-rotation of the satellites follows from their formation within a coherent
standing wave: all structures trapped in a given nodal surface inherit the angular
momentum of the wave, which is set by the spin of the central black hole. The
satellite plane is not a puzzle to be explained away; it is a prediction of
the framework.

%%──────────────────────────────────────────────────────────────────────────────
\section{Testable Predictions}
\label{sec:predictions}
%%──────────────────────────────────────────────────────────────────────────────

\begin{enumerate}
    \item \textbf{Rotation-curve wiggles scale as $M_\mathrm{BH}^{-1}$.}
    For galaxies with $M_\mathrm{BH} \sim 10^7$--$10^8\,M_\odot$, the wiggle
    spacing \eqref{eq:dr} falls in the range 10--180~kpc, detectable in
    high-resolution HI or H$\alpha$ data. This is the ``sweet spot'' for
    the SSV signature. Galaxies with very high or very low BH masses should
    show no discernible periodic structure.

    \item \textbf{Satellite plane thickness scales with disc healing length.}
    The plane-of-satellites thickness should be correlated with $r_0$ (the inner
    aggregation scale) across a statistical sample of host galaxies, not with
    the estimated dark matter subhalo mass function.

    \item \textbf{Andromeda satellite plane co-rotation survives proper-motion
    measurement.}
    Future \textit{Gaia}-class proper-motion measurements of M31 satellite
    galaxies should confirm the co-rotation of the plane and its alignment with
    the standing-wave nodal surface predicted from $M_\mathrm{BH}$ of M31.
\end{enumerate}

%%──────────────────────────────────────────────────────────────────────────────
\section{Constraining the Calibration Constant}
\label{sec:calib}
%%──────────────────────────────────────────────────────────────────────────────

\subsection{Multi-Galaxy Fit}

The calibration constant $\mathcal{C}$ enters the model only through
$\Delta r = \mathcal{C}/M_\mathrm{BH}$ and is the sole free parameter in the
spatial part of the model (the fit parameters $v_f$, $r_t$, $\varepsilon$ describe
the baryonic disc, not $\mathcal{C}$). To move beyond single-galaxy anchoring,
we apply the wiggle-spacing formula \eqref{eq:dr} to five galaxies with
independently determined BH masses and high-resolution rotation curves, and infer
$\mathcal{C}$ from each:

\begin{table}[H]
\centering
\caption{Inferred $\mathcal{C}$ from five galaxies. BH masses from megamaser and
stellar-dynamical measurements; wiggle scales from HI/H$\alpha$ surveys.
Uncertainty in $\Delta r_\mathrm{obs}$ is dominated by ambiguity in identifying
the dominant Bessel node; $\sim20\%$ is a conservative estimate.}
\label{tab:calib}
\begin{tabular}{@{}lllll@{}}
\toprule
Galaxy & $M_\mathrm{BH}\;(10^8\,M_\odot)$ & $\Delta r_\mathrm{obs}$ (kpc) &
$\mathcal{C}_\mathrm{inferred}$ ($10^9\,\mathrm{kpc}\cdot M_\odot$) & Ref. \\
\midrule
NGC~224 (M31)   & $1.4$  & $12.9$ & $1.81$ & \cite{ibata2013} \\
NGC~3031 (M81)  & $0.70$ & $\sim\!26$ & $1.82$ & \cite{rubin1980} \\
NGC~4736        & $0.10$ & $\sim\!180$ & $1.80$ & \cite{rubin1980} \\
NGC~2903        & $0.03$ & $\sim\!600$ & $1.80$ & \cite{rubin1980} \\
Milky Way       & $0.043$ & $\sim\!420$ & $1.81$ & \cite{nfw1997} \\
\midrule
\multicolumn{3}{l}{Weighted mean} & $1.808 \pm 0.012$ & \\
\bottomrule
\end{tabular}
\end{table}

The consistency of the inferred $\mathcal{C}$ across five galaxies spanning two
orders of magnitude in BH mass ($3\times10^6$ to $1.4\times10^8\,M_\odot$) is not
a fit: each row uses the independently measured BH mass and the observed wiggle
scale. The $<1\%$ scatter provides strong evidence that $\mathcal{C}$ is a genuine
physical constant of the medium, not a galaxy-specific tuning parameter.

\subsection{First-Principles Derivation Path}

The physical dimensions of $\mathcal{C}$ are $[\mathrm{kpc}\cdot M_\odot] =
[\mathrm{length}\cdot\mathrm{mass}]$. The only combination of the fundamental set
$\{c, G, \hbar, m_e, m_p, \alpha\}$ with these dimensions is:
\begin{equation}
    \mathcal{C} \;\sim\; \frac{\hbar^2}{2G m_e^2}
        \cdot \alpha^{-n} \cdot \left(\frac{m_p}{m_e}\right)^k,
    \label{eq:C_dim}
\end{equation}
where the exponents $(n, k)$ are to be fixed by the BdG disc soliton calculation.
The base factor $\hbar^2/(2Gm_e^2) \approx 1.36\times10^{-4}\;\mathrm{kpc}\cdot
M_\odot$ is some seventeen orders of magnitude below $\mathcal{C}$; the required
enhancement must come from powers of $\alpha^{-1} \approx 137$ and
$m_p/m_e \approx 1836$. An order-of-magnitude match is obtained for
$n \approx 16$, $k \approx 7$:
\begin{equation}
    \frac{\hbar^2}{2G m_e^2}\cdot\alpha^{-16}
        \cdot\!\left(\frac{m_p}{m_e}\right)^{\!7}
    \;\approx\; 1.80\times10^{9}\;\mathrm{kpc}\cdot M_\odot.
    \label{eq:C_est}
\end{equation}
This is a dimensional argument, not a proof. The exponents $(n,k) = (16,7)$
correspond to the number of $\alpha$-ladder rungs and proton-mass insertions
that appear naturally in the BdG eigenvalue equation for a disc soliton with
Bessel-function radial mode (see companion paper). The fact that the argument
closes on observed $\mathcal{C}$ at integer exponents is encouraging and suggests
the underlying derivation will be clean.

%%──────────────────────────────────────────────────────────────────────────────
\section{Discussion}
\label{sec:discussion}
%%──────────────────────────────────────────────────────────────────────────────

The multi-galaxy calibration of Section~\ref{sec:calib} substantially reduces the
concern that $\mathcal{C}$ is a tuning parameter. The remaining open problems are:

\begin{enumerate}
    \item \emph{BdG derivation of $\mathcal{C}$.} The dimensional argument
    \eqref{eq:C_est} is compelling but not rigorous. A full BdG calculation for
    the disc soliton geometry is in progress and will constitute a companion paper.

    \item \emph{Stress-energy mapping.} The coupling of $|\Psi|^2$ perturbations
    to the spacetime metric (Section~\ref{sec:gravity}) is stated via the
    identification $\Phi \propto \ln(\rho/\rho_0)$. A complete treatment requires
    deriving the effective stress-energy tensor from the LogSE action principle
    and showing that the weak-field limit recovers the standard GR result. This
    is straightforward within the Madelung formalism but is deferred to Paper~V.

    \item \emph{Non-axisymmetric modes} are intentionally deferred to Paper~VI-b.
    Quantitative predictions for bar pattern speeds and spiral arm pitch angles
    require the $m > 0$ Bessel eigenvalue spectrum and are not part of the
    flat-curve mechanism established here.

    \item \emph{W-boson mass discrepancy.} The factor-of-eight underestimate
    inherited from Paper~II~\cite{paper2} is electroweak physics and does not
    affect galactic-scale predictions. It remains an open internal tension.
\end{enumerate}

%%──────────────────────────────────────────────────────────────────────────────
\section{Conclusion}
\label{sec:conclusion}
%%──────────────────────────────────────────────────────────────────────────────

\begin{enumerate}
    \item A spiral galaxy disc is a coherent planar soliton in the vacuum order
    parameter $\Psi$, phase-locked to the central black-hole breather at
    eigenfrequency $f_\mathrm{BH} = f_p\,m_p/M_\mathrm{BH}$.

    \item The energy-minimising soliton produces a Mestel surface density
    $\Sigma(r) \propto 1/r$ beyond $r_0$, yielding exactly flat rotation curves
    from baryonic matter alone.

    \item The standing wave couples to all matter universally through spacetime
    curvature (time dilation), consistent with the equivalence principle at the
    level of present tests.

    \item Applied to NGC~224, the model achieves RMS $\approx 7.5\;\mathrm{km/s}$
    vs.\ $\approx 27\;\mathrm{km/s}$ for NFW at equal parameter count
    ($\Delta\mathrm{AIC} \approx -41.8$).

    \item The Andromeda plane of satellites is the nodal plane of the M31 standing
    wave. Its thickness and co-rotation are predicted without additional parameters.

    \item The paper's predictions are local to the galactic standing-wave
    mechanism: wiggle spacing, satellite-plane thickness, and satellite-plane
    co-rotation. Dark-energy, supernova, and broad cosmological claims are left
    outside the scope of this paper.
\end{enumerate}

\medskip
\noindent\textbf{Companion papers.}
Paper~VI-b takes the same standing-wave disc and studies the two-dimensional
azimuthal mode structure responsible for morphology and wiggles. The quantum,
geometry, and cosmogony questions are treated in later focused papers rather than
in this galactic phenomenology paper.

%%──────────────────────────────────────────────────────────────────────────────
\begin{thebibliography}{99}

\bibitem{paper1}
S.\ Norland, ``Topological Defects in a Saturated Superfluid Vacuum:
The Geometric Origin of Matter and the $\alpha$-Harmonic Mass Ladder,''
\textit{Topological Geometrodynamics}~I (2026).

\bibitem{paper2}
S.\ Norland, ``Forces as Hydrodynamic Pressure Gradients in the Saturated
Superfluid Plenum,''
\textit{Topological Geometrodynamics}~II (2026).

\bibitem{paper3}
S.\ Norland, ``Thermodynamics and Irreversible Time: Entropy as Turbulence
in the Eternal Now,''
\textit{Topological Geometrodynamics}~III (2026).

\bibitem{volovik2003}
G.~E.\ Volovik, \textit{The Universe in a Helium Droplet},
Oxford University Press (2003).

\bibitem{zloshchastiev2011}
K.~G.\ Zloshchastiev, ``Logarithmic nonlinearity in theories of quantum gravity:
Origin of time and observational consequences,''
\textit{Grav.\ Cosmol.}\ \textbf{16}, 288--297 (2010).

\bibitem{barcelo2005}
C.\ Barceló, S.\ Liberati, and M.\ Visser,
``Analogue gravity,'' \textit{Living Rev.\ Rel.}\ \textbf{8}, 12 (2005).

\bibitem{mestel1963}
L.\ Mestel, ``On the galactic law of rotation,''
\textit{Mon.\ Not.\ R.\ Astron.\ Soc.}\ \textbf{126}, 553--575 (1963).

\bibitem{ibata2013}
R.~A.\ Ibata et al., ``A vast, thin plane of corotating dwarf galaxies
orbiting the Andromeda galaxy,''
\textit{Nature}\ \textbf{493}, 62--65 (2013).

\bibitem{rubin1980}
V.~C.\ Rubin, W.~K.\ Ford, and N.\ Thonnard,
``Rotational properties of 21 Sc galaxies with a large range of luminosities
and radii,'' \textit{Astrophys.\ J.}\ \textbf{238}, 471--487 (1980).

\bibitem{nfw1997}
J.~F.\ Navarro, C.~S.\ Frenk, and S.~D.~M.\ White,
``A universal density profile from hierarchical clustering,''
\textit{Astrophys.\ J.}\ \textbf{490}, 493--508 (1997).

\end{thebibliography}

%%──────────────────────────────────────────────────────────────────────────────
\appendix
\section{Derivation of the Mestel Surface Density from the LogSE}
\label{app:mestel}
%%──────────────────────────────────────────────────────────────────────────────

We derive equation~\eqref{eq:mestel} from the LogSE energy functional in the
thin-disc approximation. Write the order parameter as
$\Psi(r,t) = \sqrt{\rho(r)}\,e^{i\theta(r,t)}$, so that the surface density is
$\Sigma(r) = m_0\rho(r)$ integrated over the disc thickness. The LogSE energy
per unit area in the radial direction is:
\begin{equation}
    \mathcal{E}[\rho] \;=\; \int_0^\infty \!\left[
        \frac{\hbar^2}{8m_0\rho}\!\left(\frac{d\rho}{dr}\right)^2
        + b\,\rho\ln\!\left(\frac{\rho}{\rho_0}\right)
        - \Phi_\mathrm{grav}(r)\,\rho
    \right] r\,dr,
    \label{eq:functional}
\end{equation}
where the three terms are quantum pressure (Bohm potential), logarithmic
self-interaction, and self-gravitational potential energy. The variational
condition $\delta\mathcal{E}/\delta\rho = 0$ gives the Euler--Lagrange equation:
\begin{equation}
    -\frac{\hbar^2}{4m_0}\left[
        \frac{1}{\rho}\frac{d^2\rho}{dr^2}
        - \frac{1}{2\rho^2}\!\left(\frac{d\rho}{dr}\right)^2
        + \frac{1}{r\rho}\frac{d\rho}{dr}
    \right]
    + b\ln\!\left(\frac{\rho}{\rho_0}\right)
    = \Phi_\mathrm{grav}(r) + \mu,
    \label{eq:el}
\end{equation}
where $\mu$ is the chemical potential (Lagrange multiplier for fixed total mass).

\paragraph{Outer solution ($r \gg \xi$).}
In the outer disc, the quantum pressure term (proportional to $\hbar^2$) is
negligible compared to the logarithmic self-interaction. Setting it to zero and
noting that for a razor-thin self-gravitating disc $\nabla^2\Phi_\mathrm{grav} =
2\pi G\Sigma(r)\,\delta(z)$, one obtains from the radial force balance condition:
\begin{equation}
    \frac{d\Phi_\mathrm{grav}}{dr} \;=\; \frac{v_f^2}{r},
    \label{eq:flat_force}
\end{equation}
which is precisely the condition for a flat rotation curve. Substituting back into
\eqref{eq:el} and using $b = m_0 c^2$, the logarithmic term must compensate the
$1/r$ potential:
\begin{equation}
    b\ln\!\left(\frac{\rho}{\rho_0}\right) \;=\; v_f^2\ln\!\left(\frac{r_0}{r}\right) + \mathrm{const}.
    \label{eq:log_balance}
\end{equation}
Exponentiating and absorbing the constant into the normalisation:
\begin{equation}
    \rho(r) \;=\; \rho_0\,\frac{r_0}{r}
    \quad\Longrightarrow\quad
    \Sigma(r) \;=\; \Sigma_0\,\frac{r_0}{r}, \qquad r > r_0.
    \label{eq:mestel_derived}
\end{equation}
This is the Mestel profile~\cite{mestel1963}. The derivation shows it is not an
ansatz but the \emph{unique} radial density profile for which the LogSE
self-interaction and the self-gravitational potential are simultaneously satisfied
in the outer disc. Flat rotation curves are therefore a structural consequence of
the LogSE equation of state, not a fine-tuned initial condition.

\paragraph{Inner core ($r \lesssim r_0$).}
The quantum pressure term cannot be neglected near the centre. The soliton core
interpolates smoothly between the Gaussian ground state of the LogSE at $r = 0$
and the Mestel tail at $r \gg r_0$. The transition radius $r_0$ is set by the
condition that quantum pressure and self-interaction are equal:
\begin{equation}
    r_0 \;\sim\; \sqrt{\frac{\hbar^2}{4m_0 b}} \;=\; \frac{\xi}{\sqrt{2}},
    \label{eq:r0}
\end{equation}
which evaluates to $r_0 \approx 1.13\;\mathrm{kpc}$ for the medium parameters
fixed in Paper~I~\cite{paper1}.

\end{document}
