\documentclass[11pt]{article}

\usepackage[a4paper,margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{amsmath,amssymb}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{setspace}
\usepackage{tcolorbox}
\tcbuselibrary{skins,breakable}
\newtcolorbox{resultbox}[1][]{enhanced,breakable,
  colback=blue!4!white,colframe=blue!40!black,
  fonttitle=\bfseries,title={#1},boxrule=0.6pt,arc=4pt}
\newtcolorbox{gapbox}[1][]{enhanced,breakable,
  colback=orange!6!white,colframe=orange!60!black,
  fonttitle=\bfseries,title={#1},boxrule=0.6pt,arc=4pt}

\hypersetup{colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue}
\setstretch{1.1}

\title{Saturated Superfluid Vacuum IX\\
CMB and the Primordial Phonon Bath}
\author{Stig Norland}
\date{Draft}

\providecommand{\ssvissue}[1]{\hyperref[issue:#1]{\##1}}% in-text issue ref -> Code & Issue References appendix
\providecommand{\ssvfile}[1]{\hyperref[file:#1]{\texttt{\detokenize{#1}}}}% in-text code/report ref -> Code & Issue References appendix
\begin{document}
\maketitle

\begin{abstract}
This paper offers a single, deliberately narrow reading of the cosmic microwave
background (CMB) within the Saturated Superfluid Vacuum (SSV) framework: that the
CMB is acoustic noise integrated from the reconnection events of the
void-to-saturation transition (Stages~4--5 of Paper~VIII~\cite{ssv8}), rather than
a separately postulated radiation-era relic. The medium that survives that
transition is not quiet; it carries a residual bath of phonons --- the Goldstone
sound modes of the saturated condensate~\cite{ssvG} --- whose present-day,
red-shifted spectrum is what an observer reads as the CMB.

The paper is explicitly \emph{interpretive} in its central claim and honest about
what it does \emph{not} establish. It does \emph{not} derive the CMB temperature
$T_0 = 2.725$~K, does \emph{not} derive the blackbody form of the spectrum, and
does \emph{not} compute the spectral distortions; each of those is marked with an
explicit gap box and tied to the same missing input --- the absolute saturation
pressure $P_0(\rho_0)$ --- that blocks the cosmological-constant value in
Paper~VIII. What the paper \emph{does} inherit quantitatively is the
Kibble--Zurek defect statistics of Paper~VIII's Stage~5, which is
\textit{structural with numerical evidence}. Every claim is tagged
\textit{derived}, \textit{structural}, \textit{candidate}, \textit{interpretive},
or \textit{deferred} in the Discussion (\S\ref{sec:discussion}), with explicit
criteria for what observational or theoretical work would change its status. The
purpose is to state the phonon-bath reading of the CMB cleanly and to fix the
support conditions for it --- not to present a quantitative CMB cosmology.
\end{abstract}

\section{Introduction}
\label{sec:intro}

The cosmic microwave background is, in the standard account, a relic of a hot
radiation era: a thermal photon gas that decoupled from baryonic matter at
recombination and has since free-streamed and cooled with the expansion. The SSV
framework does not have a separate photon gas to begin with --- photons are
transverse Goldstone/Bogoliubov modes of the condensate~\cite{ssvG}, and the
``radiation era'' is not an independent postulate but the turbulent aftermath of
the saturation transition described in Paper~VIII~\cite{ssv8}. This paper asks the
narrow question that follows: \emph{if the saturating vacuum necessarily produces a
bath of sound modes, what does that bath look like today, and is it consistent with
being read as the CMB?}

The scope is deliberately small. This paper consumes Paper~VIII's cosmogonic
sequence as input and does not reopen it; it borrows the entropy / arrow-of-time
machinery of Paper~III~\cite{ssv3} for the irreversibility of the integrated noise;
and it uses the equation-of-state language of Paper~II~\cite{ssv2} for the medium's
acoustic response. It does \emph{not} address inflationary phenomenology, structure
formation (galactic scales live in Paper~VI), late-time acceleration
(deferred to Paper~VII-b and to Paper~VIII's prediction~C2), or any particle and
force-sector content (Papers~I--IV). The material developed here is the CMB and
primordial-phonon content that issue~\ssvissue{10} moved out of Paper~III; this
draft is its destination, written under issue~\ssvissue{100}.

Every claim below is tagged in the Discussion (\S\ref{sec:discussion}). The reader
should treat the central reading --- CMB as integrated reconnection noise --- as
\textit{interpretive}, and should read the temperature, blackbody character, and
distortion content as \textit{deferred} (gap-boxed), not as results.

\section{The Primordial Phonon Bath}
\label{sec:phononbath}

Paper~VIII, Stage~4, ends with the saturation fronts colliding and the binding
energy released into the kinetic energy of the flow field, ``driving turbulence.''
Stage~5 then nucleates the first topological defects wherever the phase mosaic of
the saturating vacuum forced a mismatch. Two things are produced together at
saturation: \emph{defects} (the matter content, counted by Kibble--Zurek) and
\emph{radiation} (the sound field shed by the violent relaxation of the flow).
This paper is about the second product.

In a saturated superfluid the low-energy excitations above the ground state
$\rho_0$ are phonons: linear sound modes with dispersion $\omega = c_s k$ in the
long-wavelength limit, where $c_s$ is the medium's sound speed fixed by the
equation of state $P(\rho)$ of Paper~II~\cite{ssv2}. The saturation transition is
not adiabatic --- it is an exothermic quench --- so it leaves the condensate in an
excited state populated by these modes. We call this population the
\emph{primordial phonon bath}. Its existence is a structural consequence of having
a saturation transition at all: any quench of a medium with a Goldstone sector
deposits energy into that sector.

\begin{quote}
\textit{Claim (\S\ref{sec:phononbath}).} The saturated vacuum is left, after the
Stage~4--5 transition, with a residual bath of phonon (sound) modes carrying the
energy shed by the relaxing flow field. Status: \textit{interpretive} (the
existence of the bath follows structurally from the quench; its spectrum is not
derived here).
\end{quote}

The bath is the radiation counterpart of the Kibble--Zurek defect count: defects
freeze in where the phase cannot be combed smooth, and phonons carry off the energy
difference between the turbulent saturating state and the relaxed ground state. The
\emph{ratio} of the two --- defects to phonons --- is, in cosmological language, the
baryon-to-photon ratio $\eta$, and it is exactly the quantity Paper~VIII computes
from the Kibble--Zurek mechanism (\S\ref{sec:viii}).

\paragraph{The bath is not load-bearing for gravity (\#119 outcome).}
During the \#119 gravity-sector redesign, this bath was the candidate
common drive for a bath-driven (secondary-Bjerknes) gravitational
mechanism --- which, had it survived, would have made the present paper's
object load-bearing for the gravity sector, with $G$ inheriting the bath
amplitude.  The pre-registered test returned a negative: the bath-driven
force between passive defects is sign-definite and frequency-blind, as
required, but short-ranged --- the wrong range for Newtonian gravity
(\#119 Phase~2).  In the revised architecture the gravitational source is
each defect's \emph{intrinsic} two-term structure (\#119 H7a/H8; Paper~IV),
and the phonon bath carries no gravitational load.  The bath's status in
this paper is therefore unchanged by the gravity-sector redesign, and
nothing here depends on it.

\section{The CMB as Integrated Reconnection Noise}
\label{sec:integrated}

The central reading of this paper is the following identification.

\begin{quote}
\textit{Central claim (\S\ref{sec:integrated}, interpretive).} The cosmic microwave
background is the present-day, red-shifted image of the primordial phonon bath:
acoustic noise integrated over all the reconnection and relaxation events of the
void-to-saturation transition, propagated forward through the expanding saturated
medium.
\end{quote}

Three features of the picture are worth stating precisely, because they are what a
supporting computation would have to reproduce.

\paragraph{Integration, not a single surface.} In the standard account the CMB
comes from a single last-scattering surface. Here the noise is \emph{integrated}:
each reconnection event during Stage~5 radiates a sound pulse, and the bath is the
superposition of an enormous number of such pulses with random phases. This is the
sense in which the spectrum is expected to be smooth --- it is the central-limit
sum of many independent emitters --- but smoothness of a noise spectrum is not the
same as a Planck (blackbody) spectrum, a distinction we are careful about in
\S\ref{sec:spectrum}.

\paragraph{Irreversibility.} The integrated noise is irreversible: it is the wake
entropy of Paper~III~\cite{ssv3} read at cosmological scale. The arrow of time that
Paper~III attributes to the growth of wake complexity is, here, the statement that
the phonon bath cannot un-radiate back into coherent flow. The CMB is, in this
language, the largest single reservoir of the universe's wake entropy.

\paragraph{Redshift as medium relaxation.} The cooling of the bath is the slow
relaxation of the background $\rho_0$ and the corresponding stretching of mode
wavelengths, the same background-relaxation channel that Paper~VIII's prediction~C2
ties to late-time acceleration. We do \emph{not} import a metric expansion here; we
note only that the qualitative redshift of the bath and the qualitative late-time
relaxation are the same effect viewed at different epochs, and we defer the
quantitative expansion history to Paper~VII-b.

This section is the reason the paper exists, and it is entirely \textit{interpretive}.
It reads a known observable (a near-isotropic microwave bath) as a known SSV object
(the Goldstone sound field of a quenched condensate). It does not, by itself,
predict any number.

\section{Spectrum, Temperature, and the Equation of State}
\label{sec:spectrum}

The interpretive reading of \S\ref{sec:integrated} becomes physics only if the bath
can be shown to carry the right spectrum and temperature. This is where the paper is
most honest about its limits.

\subsection{Why a thermal spectrum is plausible but not derived}

A blackbody (Planck) spectrum is the equilibrium distribution of a bosonic field
that has had time to thermalise through mode--mode interactions. Phonons in a
superfluid do interact (the GPE nonlinearity couples modes), so a saturated medium
left in an excited phonon state will tend toward a thermal distribution. That is the
plausibility argument. But ``tends toward thermal'' is not ``is Planckian to one part
in $10^{5}$,'' which is what the measured CMB demands. Establishing the blackbody
form requires a thermalisation (ergodicity) argument for the phonon bath that this
paper does not provide.

\begin{gapbox}[Gap: blackbody character of the bath is not derived]
The Planck form of the CMB spectrum is, here, \textit{deferred}. It would follow
from a demonstration that the primordial phonon bath thermalises --- i.e.\ that the
GPE phonon--phonon interactions drive the mode-occupation numbers to a
Bose--Einstein distribution on a timescale short compared with the saturation-epoch
expansion time. No such thermalisation calculation has been performed. Until it is,
the ``integrated reconnection noise'' is established only as a \emph{smooth} noise
spectrum (central-limit superposition of many emitters), not as a \emph{Planck}
spectrum. Support criterion: a phonon-Boltzmann or wave-turbulence calculation in
the saturated GPE showing relaxation to Bose--Einstein occupation.
\end{gapbox}

\subsection{Why the temperature is not derived}

Even granting a thermal form, the absolute temperature $T_0 = 2.725$~K is set by the
energy density deposited in the bath at saturation and by the subsequent dilution.
The energy density at saturation is fixed by the depth of the equation of state ---
specifically by the absolute saturation pressure $P_0(\rho_0)$ --- which is the
\emph{same} quantity that Paper~VIII flags as the missing input blocking a numerical
value of $\Lambda$, and that Papers~II/IV/VII-b track as an open closure problem.

\begin{gapbox}[Gap: CMB temperature $T_0 = 2.725$~K is not derived]
The present CMB temperature is \textit{deferred}, and it is blocked by exactly the
input that blocks $\Lambda$ in Paper~VIII~\cite{ssv8} and the absolute energy scale
in Papers~II/IV: the absolute saturation pressure $P_0(\rho_0)$. Schematically the
bath energy density at saturation is $u_{\rm sat}\sim P_0$, and $T_0$ follows from
$u_{\rm sat}$ diluted by the expansion factor between the saturation epoch and now;
neither $P_0$ nor the SSV expansion factor is derived in the series yet. The
galactic sector now joins this list: issue~\ssvissue{129} reframes Paper~VI's
underived gravitational coupling as an emergent acceleration scale
$a_0\sim c^2\sqrt{\Lambda}\sim cH_0$, so the \emph{same} $P_0\!\to\!\Lambda$ closure
would simultaneously fix the galactic rotation-curve scale --- closing both sectors
at once, or failing both together. We
therefore make \emph{no} prediction for $T_0$. Support criterion: a derivation of
$P_0(\rho_0)$ from the saturation equation of state (the Paper~II/IV/VII-b closure
item), \emph{plus} the Paper~VII-b emergent-expansion factor, which together would
turn $T_0$ into a falsifiable SSV number.
\end{gapbox}

\subsection{What the equation of state does fix (structurally)}

What the medium's equation of state \emph{does} fix, without the absolute scale, is
the \emph{shape} relations: the sound speed $c_s(\rho)$, the dispersion's departure
from linearity at the healing scale $\xi$ (the Bogoliubov ``roton-like'' bend), and
hence the wavelength at which the bath's spectrum must roll over relative to its
peak. These are ratios, independent of $P_0$, and they are inherited directly from
Paper~II. They are \textit{structural}: the framework fixes them in principle, but
the present paper does not evaluate them numerically.

\section{Statistical Departures from Blackbody}
\label{sec:distortions}

If the CMB is integrated reconnection noise rather than an equilibrium relic, the
\emph{departures} from a perfect blackbody are where the two pictures could be
distinguished. In the SSV reading the distortions are not separately postulated
mechanisms (energy injection, Silk damping, etc.) but signatures of the
topological-defect statistics frozen in at saturation.

\begin{quote}
\textit{Claim (\S\ref{sec:distortions}, candidate).} Spectral distortions of the CMB
away from a pure Planck form encode the Kibble--Zurek defect statistics of Stage~5:
the defect-density variance and correlation structure imprint a calculable, small,
non-Planckian component on the integrated phonon spectrum.
\end{quote}

This is a \textit{candidate} claim with a definite falsification condition but no
performed computation. Concretely, the spectrum should carry $\mu$-type and
$y$-type distortions whose amplitudes are tied to the same defect-density variance
that sets $\eta$; the present observational bounds
($|\mu|\lesssim 9\times10^{-5}$, $|y|\lesssim 1.5\times10^{-5}$ from
COBE/FIRAS~\cite{Fixsen1996}) are the falsification gate.

\begin{gapbox}[Gap: distortion amplitudes are not computed]
The $\mu$- and $y$-distortion amplitudes implied by the defect statistics are
\textit{deferred}. Computing them requires (i) the defect-density two-point
statistics from a Kibble--Zurek simulation (the Paper~VIII Stage~5 computation,
already begun in \ssvfile{kibble_zurek_gpe}) carried to the level of correlation
functions, and (ii) a map from those statistics to the phonon-bath occupation
spectrum. Falsification condition: predicted $|\mu|,|y|$ exceeding the COBE/FIRAS
bounds~\cite{Fixsen1996} would rule out the integrated-noise reading; predicted
values far below current sensitivity would make the reading consistent but not yet
testable. Support criterion: a defect-correlation $\to$ distortion-spectrum
calculation with an explicit numerical $\mu,y$.
\end{gapbox}

\section{Connection to SSV VIII Cosmogony}
\label{sec:viii}

The one quantitative anchor this paper inherits is Paper~VIII's Kibble--Zurek
result. Stage~5 of Paper~VIII nucleates defects with number density
\begin{equation}
    n_{\rm defect} \sim \frac{1}{\xi^3}
        \left(\frac{\tau_0}{\tau_Q}\right)^{d\nu/(1+\nu z)},
    \label{eq:kz}
\end{equation}
with $\tau_0 = \xi/c$ the microscopic time and $\tau_Q$ the quench time of the
saturation epoch. Paper~VIII reports a stochastic time-dependent Ginzburg--Landau
simulation (\ssvfile{kibble_zurek_gpe}) that confirms this scaling in the correct
universality class, lifting the prediction to \textit{structural with numerical
evidence}; the single dimensionless unknown $\tau_Q/\tau_0$ that fixes the
baryon-to-photon ratio $\eta = 6\times10^{-10}$ lies in the natural SSV window
$\tau_Q/\tau_0 \in [10^{10},10^{13}]$.

For the present paper the same simulation fixes the \emph{ratio} of the two products
of saturation --- defects to phonons --- because $\eta$ is precisely the
defect-to-photon number ratio. The phonon-bath number density is therefore not an
independent quantity: it is $n_{\rm defect}/\eta$, with $n_{\rm defect}$ from
Eq.~\eqref{eq:kz}. This is the cleanest structural link between IX and VIII.

\begin{quote}
\textit{Claim (\S\ref{sec:viii}, structural with numerical evidence).} The
phonon-bath number density is tied to the Kibble--Zurek defect density of
Paper~VIII Stage~5 through the baryon-to-photon ratio $\eta$; the defect side of
this link carries the numerical evidence reported in Paper~VIII.
\end{quote}

\begin{gapbox}[Gap: the entropy budget is not closed]
The cosmological entropy budget --- the photon (phonon-bath) entropy per baryon,
$\sim\eta^{-1}$ in dimensionless terms --- is \textit{structurally} fixed by
$\eta$ but \emph{numerically} open, because the \emph{absolute} photon entropy
needs the bath temperature $T_0$, which is gap-boxed in \S\ref{sec:spectrum}. The
arrow-of-time reading of Paper~III~\cite{ssv3} identifies this bath as the dominant
wake-entropy reservoir, but the present paper does not evaluate the budget in
physical units. Support criterion: the same $P_0(\rho_0) + T_0$ closure required
for the temperature.
\end{gapbox}

\section{Discussion: scope, status, and what would count as support}
\label{sec:discussion}

\subsection{Claim taxonomy used in this paper}

Following \ssvfile{numerical-minimisation-roadmap} in the repository, each claim is
one of:
\begin{itemize}
    \item \textit{derived} --- analytically reduced from the SSV Lagrangian and/or a
        converged numerical computation and \emph{reused} here;
    \item \textit{structural} --- the dependence on the framework is clear and the
        result is structurally consistent, but the number depends on a computation
        deferred to another paper or workstream;
    \item \textit{candidate} --- a prediction stated here with a definite
        falsification condition but no supporting computation yet performed;
    \item \textit{interpretive} --- a synthesis or ontological statement that
        organises SSV results into a single picture without new quantitative content;
    \item \textit{deferred} --- a quantity that the picture requires but that is
        explicitly not evaluated here and is gap-boxed, with the blocking input named.
\end{itemize}

\subsection{Claim status of this paper's content}
\begin{itemize}
    \item The primordial phonon bath (\S\ref{sec:phononbath}):
        \textit{interpretive} --- its \emph{existence} follows structurally from a
        non-adiabatic saturation quench of a Goldstone sector; its spectrum is not
        derived.
    \item CMB as integrated reconnection noise (\S\ref{sec:integrated}):
        \textit{interpretive} --- the central reading; predicts no number by itself.
    \item Blackbody character (\S\ref{sec:spectrum}): \textit{deferred}
        (gap-boxed) --- needs a phonon-thermalisation argument.
    \item CMB temperature $T_0$ (\S\ref{sec:spectrum}): \textit{deferred}
        (gap-boxed) --- blocked by the absolute saturation pressure $P_0(\rho_0)$,
        the same input that blocks $\Lambda$ in Paper~VIII.
    \item Equation-of-state shape relations (\S\ref{sec:spectrum}):
        \textit{structural} --- ratios inherited from Paper~II, not evaluated here.
    \item Spectral distortions (\S\ref{sec:distortions}): \textit{candidate} ---
        definite FIRAS falsification gate, amplitudes gap-boxed.
    \item Phonon-bath / defect link via $\eta$ (\S\ref{sec:viii}):
        \textit{structural with numerical evidence} --- inherits Paper~VIII's
        Kibble--Zurek simulation.
    \item Entropy budget (\S\ref{sec:viii}): \textit{deferred} (gap-boxed) ---
        structurally fixed by $\eta$, numerically blocked by $T_0$.
\end{itemize}

\subsection{What observational or theoretical support would matter}
\begin{enumerate}
    \item \emph{A phonon-thermalisation calculation} in the saturated GPE (wave
        turbulence / phonon Boltzmann) showing relaxation to a Bose--Einstein
        occupation would lift the blackbody character from \textit{deferred} to
        \textit{structural} and is the single most important theoretical-support item.
    \item \emph{A derivation of the absolute saturation pressure $P_0(\rho_0)$}
        (the shared Paper~II/IV/VII-b closure item), together with the Paper~VII-b
        emergent-expansion factor, would turn $T_0$ into a falsifiable SSV number ---
        the decisive observational-support item.
    \item \emph{A defect-correlation $\to$ distortion-spectrum computation} giving
        numerical $\mu,y$ would make \S\ref{sec:distortions} testable against
        COBE/FIRAS~\cite{Fixsen1996} and is the cleanest near-term falsification route.
    \item \emph{Extending the Kibble--Zurek dynamic range} of
        \ssvfile{kibble_zurek_gpe} (the Paper~VIII Stage~5 simulation) to tighten the
        exponent and the $\tau_Q/\tau_0$ window would sharpen the $\eta$ anchor that
        ties the phonon bath to the defect count.
\end{enumerate}

\subsection{What this paper is not}

This paper does not derive: the CMB temperature; the blackbody form of the spectrum;
the spectral-distortion amplitudes; the angular power spectrum or its acoustic peaks;
a recombination or last-scattering treatment; a primordial gravitational-wave
background; or any inflationary, structure-formation, or dark-energy phenomenology.
Each belongs to another paper or to a future computation. The paper is a scoped
interpretive reading of the CMB as the relaxed primordial phonon bath, plus an
explicit ledger of what would be required to make that reading quantitative.
Treating it as a complete CMB cosmology would re-introduce exactly the catch-all
manifesto drift that the split of the late-series papers was designed to avoid.

\section{Conclusion}

Within the scoped SSV programme, this paper carries one claim: the cosmic microwave
background is the present, red-shifted image of the primordial phonon bath shed by
the void-to-saturation transition of Paper~VIII --- integrated reconnection noise,
not a separately postulated radiation relic. The reading is interpretive, its one
quantitative anchor (the defect/phonon ratio $\eta$) is inherited from Paper~VIII's
Kibble--Zurek result, and its central physical quantities --- the temperature, the
blackbody form, and the distortions --- are explicitly deferred to the named closure
inputs rather than claimed.

\begin{thebibliography}{99}

\bibitem{ssv1}
Norland, S., ``Saturated Superfluid Vacuum~I: Topological Defects in a
Saturated Superfluid Vacuum: The Geometric Origin of Matter,''
SSV Series Paper~I (2026).

\bibitem{ssv2}
Norland, S., ``Saturated Superfluid Vacuum~II: Forces as Hydrodynamic
Pressure Gradients in the Saturated Superfluid Plenum,''
SSV Series Paper~II (2026).

\bibitem{ssv3}
Norland, S., ``Saturated Superfluid Vacuum~III: Thermodynamics and
Irreversible Time,'' SSV Series Paper~III (2026).

\bibitem{ssv8}
Norland, S., ``Saturated Superfluid Vacuum~VIII: Cosmogony from the
Permissive Void,'' SSV Series Paper~VIII (2026).

\bibitem{ssvG}
Norland, S., ``Saturated Superfluid Vacuum (Goldstone supplement):
Electromagnetic Propagation Speed and the Photon as a Goldstone Mode,''
SSV Series (2026).

\bibitem{SSV-VIIb}
Norland, S., ``Saturated Superfluid Vacuum~VII-b: Emergent Geometry and
the Dissolution of Quantum Gravity,'' SSV Series Paper~VII-b (2026).

\bibitem{Kibble1976}
Kibble, T.~W.~B., ``Topology of Cosmic Domains and Strings,''
\emph{J.\ Phys.\ A} \textbf{9}, 1387--1398 (1976).

\bibitem{Zurek1985}
Zurek, W.~H., ``Cosmological Experiments in Superfluid Helium,''
\emph{Nature} \textbf{317}, 505--508 (1985).

\bibitem{Fixsen1996}
Fixsen, D.~J., \emph{et al.}, ``The Cosmic Microwave Background Spectrum
from the Full COBE/FIRAS Data Set,''
\emph{Astrophys.\ J.}\ \textbf{473}, 576--587 (1996).

\bibitem{Volovik2003}
Volovik, G.~E., \emph{The Universe in a Helium Droplet}
(Oxford University Press, Oxford, 2003).

\bibitem{Madelung1927}
Madelung, E., ``Quantentheorie in hydrodynamischer Form,''
\emph{Z.\ Phys.}\ \textbf{40}, 322--326 (1927).

\end{thebibliography}

\appendix
% Auto-generated provenance (issue URLs + pinned script permalinks).
% Regenerate: python instruments/tools/gen_provenance.py SSV-IX
\input{provenance.tex}

\end{document}
