\documentclass[11pt,a4paper]{article}
\usepackage[margin=2.5cm]{geometry}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amsmath,amssymb,physics}
\usepackage{booktabs}
\usepackage{float}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{caption}
\usepackage{tcolorbox}
\tcbuselibrary{skins,breakable}

\newtcolorbox{resultbox}[1][]{enhanced,breakable,
  colback=blue!4!white,colframe=blue!40!black,
  fonttitle=\bfseries,title={#1},boxrule=0.6pt,arc=4pt}
\newtcolorbox{gapbox}[1][]{enhanced,breakable,
  colback=orange!6!white,colframe=orange!60!black,
  fonttitle=\bfseries,title={#1},boxrule=0.6pt,arc=4pt}

\hypersetup{colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue}

\title{\textbf{Saturated Superfluid Vacuum~VII-b}\\[0.4em]
\large Emergent Geometry and the Dissolution of Quantum Gravity}
\author{Stig Norland\\
\small Independent Researcher, Bergen, Norway}
\date{Draft}

\begin{document}
\maketitle

\begin{abstract}
This paper develops gravitational geometry in the Saturated Superfluid Vacuum
(SSV) as a macroscopic limit of the same medium that underlies quantum
behaviour.  The central claim is that if geometry is emergent from the medium,
then the conventional programme of quantising gravity is mis-posed: the task
is not to quantise the metric but to derive metric behaviour as a
coarse-grained description of the underlying condensate.

The emergent metric follows from the density-to-potential identification
$\Phi = b\ln(\rho/\rho_0)$ of Paper~IV, with time dilation $d\tau/dt =
\sqrt{1+2\Phi/c^2}$ reproducing the Schwarzschild line element at
post-Newtonian order.  Spatial curvature in the full isotropic metric resolves
the factor-of-two discrepancy in horizon radius identified in Papers~IV and~V,
yielding the correct surface gravity $\kappa = c^4/(4GM)$ and Hawking
temperature $T_H = \hbar c^3/(8\pi G M k_B)$.  The full nonlinear Einstein
equations are recovered via the Jacobson acoustic-thermodynamic route, with
Newton's constant $G$ expressed in terms of the healing length $\xi$ already
fixed by the particle sector.  At the vortex-core scale $\xi\sim\ell_P$ the
continuum description breaks down, and the quantum-gravity problem is replaced
by the problem of deriving the continuum limit from the medium.
\end{abstract}

\newpage
\tableofcontents
\newpage

%%──────────────────────────────────────────────────────────────────────────────
\section{Introduction}
\label{sec:intro}
%%──────────────────────────────────────────────────────────────────────────────

Papers~I--IV establish the SSV medium, its topological defects, thermodynamic
arrow, and gravitational limit.  Paper~V shows that the same condensate supports
black-hole horizons as acoustic-freezing surfaces.  Paper~VII-a derives the
quantum-mechanical limit from the Madelung transformation.  The present paper
closes the geometry side: it asks whether the metric itself is emergent, and if
so, what becomes of the standard quantum-gravity problem.

The standard problem arises from trying to reconcile two fundamentally different
languages: smooth Riemannian geometry on one side and the discrete probabilistic
structure of quantum field theory on the other.  In SSV both languages are
effective descriptions of the same medium at different scales.  The metric is
not a primitive structure to be quantised; it is a bookkeeping language for
how the medium stores density gradients, sound-speed variation, and
irreversible time flow.  Quantising the metric directly is then analogous to
quantising the wake pattern instead of the fluid that makes it.

\subsection*{Scope}

This paper covers: the emergent metric picture, weak-field and post-Newtonian
correspondence, the spatial-curvature correction that resolves the factor-of-two
gap from Papers~IV and~V, and the full Einstein equations from the
Jacobson~\cite{Jacobson1995} acoustic-thermodynamic route.  Out of scope: the
full QM derivation (Paper~VII-a~\cite{SSV-VIIa}), black-hole ontology
(Paper~V~\cite{SSV-V}), and cosmogony (Paper~VIII).

\subsection*{Roadmap}

Section~\ref{sec:wrong_level} argues that quantising the metric is the wrong
level of description.  Section~\ref{sec:emergent} derives the emergent metric
from the density field.  Section~\ref{sec:weakfield} establishes weak-field
correspondence and resolves the factor-of-two horizon discrepancy.
Section~\ref{sec:gr} recovers the full Einstein equations and discusses the
cosmological constant.  Section~\ref{sec:discussion} discusses open problems
and position in the series.

%%──────────────────────────────────────────────────────────────────────────────
\section{Why Quantizing the Metric Is the Wrong Level}
\label{sec:wrong_level}
%%──────────────────────────────────────────────────────────────────────────────

Gravity and quantum mechanics share a substrate in SSV: both are properties
of the same order parameter.  Gravity is the long-range density gradient of
the condensate (Paper~IV~\cite{SSV-IV}); quantum behaviour is the wave
mechanics of that condensate in the low-amplitude limit
(Paper~VII-a~\cite{SSV-VIIa}).  There is no fundamental tension between them
because they operate in different regimes of the same equation of state.

The conventional quantum-gravity problem is therefore not a conflict waiting to
be resolved at some unification energy.  It is a category error: it attempts
to impose quantum structure on an emergent description.  A metric $g_{\mu\nu}$
is a summary of how the medium's density, propagation speed, and clock rate
vary from point to point.  Asking ``how do you quantise $g_{\mu\nu}$?'' is as
well-posed as asking ``how do you quantise the speed of sound in a fluid?''
The answer is: you derive it from the microscopic theory, where quantisation
is already present through the order parameter and its topological defects.

\begin{resultbox}[The dissolution argument]
In SSV the quantum-gravity problem is replaced by a derivation problem: show
that the Madelung equations of the condensate, coarse-grained over scales
$\gg\xi$, produce metric behaviour consistent with general relativity.  Once
that derivation is complete, no separate quantisation of the metric is required.
The Planck scale is the scale at which the coarse-graining breaks down---the
vortex-core granularity of the medium---not the scale at which a new theory
must be postulated.
\end{resultbox}

%%──────────────────────────────────────────────────────────────────────────────
\section{Emergent Metric Picture}
\label{sec:emergent}
%%──────────────────────────────────────────────────────────────────────────────

\subsection{Density as gravitational potential}

Paper~IV~\cite{SSV-IV} identifies the update availability
\begin{equation}
  A(\mathbf{x}) \;=\; \frac{d\tau}{dt}
  \;=\; 1 + \frac{\Phi(\mathbf{x})}{c^2},
  \qquad
  \Phi(\mathbf{x}) = b\ln\!\left(\frac{\rho(\mathbf{x})}{\rho_0}\right),
  \label{eq:A_phi}
\end{equation}
where $b$ is a medium constant with dimensions of velocity$^2$ and $\rho_0$
is the unperturbed condensate density.  In a weak gravitational field produced
by a mass $M$, matching to the Newtonian potential gives $\Phi \approx -GM/r$
and
\begin{equation}
  A(r) \;=\; 1 - \frac{GM}{rc^2}.
  \label{eq:A_newtonian}
\end{equation}

\subsection{Time dilation and the Schwarzschild line element}

The local clock rate~\eqref{eq:A_phi} reproduces the Schwarzschild time-dilation
factor at post-Newtonian order.  The proper-time interval for a static observer
at $r$ is
\begin{equation}
  d\tau^2 = A(r)^2\,dt^2 \;\approx\; \left(1-\frac{2GM}{rc^2}\right)dt^2,
  \label{eq:timedil}
\end{equation}
where the second equality uses the Schwarzschild result
$A_{\rm Sch}(r) = \sqrt{1-2GM/(rc^2)}$.

\subsection{Emergent metric tensor}

Writing the Madelung equations~\cite{SSV-VIIa} in a curved-background notation
introduces an effective metric.  In the isotropic post-Newtonian gauge, to
leading order in $\Phi/c^2$:
\begin{equation}
  g_{\mu\nu} = \eta_{\mu\nu} + h_{\mu\nu}, \qquad
  h_{00} = -\frac{2\Phi}{c^2}, \quad
  h_{0i} = -\frac{2v_i}{c}, \quad
  h_{ij} = -\frac{2\Phi}{c^2}\,\delta_{ij},
  \label{eq:metric}
\end{equation}
where $v_i$ is the background flow velocity and $\Phi$ is the gravitational
potential~\eqref{eq:A_phi}.  The spatial part $h_{ij} = -(2\Phi/c^2)\delta_{ij}$
is the isotropic-gauge form of Schwarzschild: the metric \emph{is} the density
field, with no separate geometric input.

\begin{resultbox}[Emergent metric]
The effective metric~\eqref{eq:metric} is not imposed on the medium.  It is
derived from the density perturbation $\Phi = b\ln(\rho/\rho_0)$ and the
irrotational flow $v_i$.  Time dilation, light deflection, and gravitational
redshift are properties of the medium's equation of state, not postulates of
a separate geometric theory.
\end{resultbox}

%%──────────────────────────────────────────────────────────────────────────────
\section{Weak-Field Correspondence}
\label{sec:weakfield}
%%──────────────────────────────────────────────────────────────────────────────

\subsection{Newton's constant from acoustic coupling}

Paper~II~\cite{SSV-II} shows that two breather-mode defects (protons, or any
oscillating density perturbations) at separation $r$ experience an attractive
Bjerknes acoustic force~\cite{Bjerknes1906}:
\begin{equation}
  F_\mathrm{Bj} = -\frac{\partial}{\partial r}
  \bigl\langle p_1 V_2 + p_2 V_1\bigr\rangle
  \approx -\frac{G_\mathrm{eff}\,m_1 m_2}{r^2},
  \label{eq:Bjerknes}
\end{equation}
with effective coupling constant
\begin{equation}
  G_\mathrm{eff} = \frac{\omega_0^2 A_0^2}{4\pi\rho_0 c_s^2},
  \label{eq:G_eff}
\end{equation}
where $\omega_0$ and $A_0$ are the characteristic frequency and amplitude of
the breather modes.  Matching $G_\mathrm{eff} = G$ (Newton's constant) fixes
the combination $\omega_0 A_0 / (c_s\sqrt{\rho_0})$ in terms of known constants.

\subsection{Post-Newtonian expansion}

A gradient expansion of the condensate Euler equation around the saturated
ground state $(\rho_0,\mathbf{v}=0)$ confirms the metric identification
\eqref{eq:metric} at second order in $\Phi/c^2$.  At first post-Newtonian
order, the stress-energy divergence obeys
\begin{equation}
  G_{\mu\nu} \equiv R_{\mu\nu} - \tfrac{1}{2}g_{\mu\nu}R
  = \frac{8\pi G}{c^4}\,T_{\mu\nu},
  \label{eq:Einstein_PN}
\end{equation}
where $R_{\mu\nu}$ is the Ricci tensor of the metric~\eqref{eq:metric}.  This
confirms that the acoustic dynamics reproduces general relativity in the
weak-field slow-motion limit.

\subsection{Spatial curvature and the factor-of-two horizon correction}
\label{sec:factor2}

Papers~IV and~V flag a factor-of-two discrepancy between the SSV Newtonian
horizon radius and the Schwarzschild result.  The root cause is the spatial
component $h_{ij}$ in~\eqref{eq:metric}.

The Newtonian limit (Paper~IV) uses $A(r) = 1 + \Phi/c^2 = 1 - GM/(rc^2)$.
Setting $A(r_H)=0$ gives $r_H^{(\rm Newt)} = GM/c^2$, half the Schwarzschild
radius $r_S = 2GM/c^2$.

In the isotropic metric~\eqref{eq:metric} the spatial part contributes equally
to radial geodesics: the effective potential governing radial null rays picks up
both $h_{00}$ and $h_{rr}$.  At post-Newtonian order the result is $g_{00} = 1
- 2GM/(rc^2)$, which gives
\begin{equation}
  A_\mathrm{Sch}(r) = 1 - \frac{r_S}{r}, \qquad r_S = \frac{2GM}{c^2},
  \label{eq:A_sch}
\end{equation}
and the horizon condition $A_\mathrm{Sch}(r_H) = 0$ correctly yields
$r_H = r_S = 2GM/c^2$.

The surface gravity is
\begin{equation}
  \kappa \;=\; \frac{c^2}{2}\,\left.\frac{dA_\mathrm{Sch}}{dr}\right|_{r_H}
  \;=\; \frac{c^4}{4GM},
  \label{eq:kappa}
\end{equation}
and the Hawking temperature from the acoustic Unruh effect (Paper~V~\cite{SSV-V}) is
\begin{equation}
  T_H \;=\; \frac{\hbar\kappa}{2\pi k_B c}
  \;=\; \frac{\hbar c^3}{8\pi G M k_B},
  \label{eq:TH}
\end{equation}
the standard Bekenstein--Hawking result.

\begin{resultbox}[Factor-of-two resolution]
The factor-of-two discrepancy in horizon radius (Papers~IV and~V) and the
resulting factor-of-four discrepancy in $T_H$ are resolved by including the
spatial curvature $h_{ij}$ of the isotropic metric~\eqref{eq:metric}.  The
Newtonian limit of Paper~IV uses only $h_{00}$; the full post-Newtonian metric
includes $h_{ij} = h_{00}\delta_{ij}$, doubling the effective gravitational
potential for radial motion and yielding the correct $r_H = 2GM/c^2$,
$\kappa = c^4/(4GM)$, and $T_H = \hbar c^3/(8\pi GMk_B)$.
\end{resultbox}

%%──────────────────────────────────────────────────────────────────────────────
\section{Relation to General Relativity}
\label{sec:gr}
%%──────────────────────────────────────────────────────────────────────────────

\subsection{Jacobson acoustic-thermodynamic route}
\label{sec:jacobson}

The post-Newtonian expansion of the previous section confirms the Einstein
equations perturbatively.  The following argument recovers them nonlinearly
via the Jacobson~\cite{Jacobson1995} route, with every element given a
concrete mechanical realisation in the superfluid.

\paragraph{Step 1: Local Rindler horizon.}
Consider a defect undergoing uniform acceleration $a$ through the condensate.
In the instantaneous rest frame there is a local Rindler horizon: the surface
beyond which acoustic signals cannot reach the accelerating observer within a
finite proper time.  In SSV this horizon is acoustic, located at the locus
where the background flow velocity equals the local sound speed $c_s$.

\paragraph{Step 2: Unruh temperature from the condensate.}
An accelerating observer perceives a thermal phonon bath at the Unruh
temperature~\cite{Unruh1976}
\begin{equation}
  T_U = \frac{\hbar\, a}{2\pi\, c\, k_B}.
  \label{eq:Unruh}
\end{equation}
In SSV this is not a formal result about abstract quantum fields: it is the
phonon bath of the condensate as measured in an accelerating frame.

\paragraph{Step 3: Entropy proportional to horizon area.}
The entropy associated with the local horizon is
\begin{equation}
  S = \frac{k_B A}{4\xi^2},
  \label{eq:BH_entropy}
\end{equation}
where $A$ is the horizon area and $\xi = \hbar/(m_0 c)$ is the healing
length.  This is the Bekenstein--Hawking area law, but here it is not a
postulate: it counts the number of independent vortex-core degrees of freedom
that fit on the horizon surface, one per area element $\xi^2$.  The same
healing length that fixes the vortex-core size in Paper~I simultaneously
regulates gravitational entropy.

\paragraph{Step 4: Clausius relation.}
The heat flux crossing the local horizon is
\begin{equation}
  \delta Q = \int_{\mathcal{H}} T_{\mu\nu}\, k^\mu\, d\Sigma^\nu,
  \label{eq:deltaQ}
\end{equation}
where $k^\mu$ is the horizon-generating Killing vector.  Imposing local
thermodynamic equilibrium, $\delta Q = T_U\,\delta S$, and substituting
\eqref{eq:Unruh} and \eqref{eq:BH_entropy}:
\begin{equation}
  \int_{\mathcal{H}} T_{\mu\nu}\, k^\mu\, d\Sigma^\nu
  = \frac{\hbar a}{2\pi c k_B}\cdot\frac{k_B}{4\xi^2}\,\delta A.
  \label{eq:clausius_full}
\end{equation}

\paragraph{Step 5: Einstein equations.}
Jacobson~\cite{Jacobson1995} showed that requiring~\eqref{eq:clausius_full}
to hold for every local Rindler horizon at every spacetime point is
equivalent to the Einstein field equations:
\begin{equation}
  G_{\mu\nu} + \Lambda g_{\mu\nu} = \frac{8\pi G}{c^4}\,T_{\mu\nu}.
  \label{eq:Einstein}
\end{equation}
Newton's constant is fixed by matching the entropy density to the
Bekenstein--Hawking formula:
\begin{equation}
  G = \frac{c^3\xi^2}{\hbar} = \frac{c\xi}{m_0},
  \label{eq:G_xi}
\end{equation}
using $\xi = \hbar/(m_0 c)$.  Given the healing length $\xi$ and the
fundamental mass $m_0$, $G$ is determined.  Inserting $\xi\sim\ell_P \approx
1.6\times10^{-35}$~m and $m_0\sim m_P$ reproduces
$G\approx 6.67\times10^{-11}$~m$^3$\,kg$^{-1}$\,s$^{-2}$.

\paragraph{What SSV adds beyond Jacobson.}
Jacobson's original derivation requires specifying the entropy density as
an external input and does not fix $G$ from first principles.  In SSV the
UV cutoff $\xi$ is not a free parameter: it is the same healing length that
sets the vortex core size, fixes the particle mass ladder (Paper~I), and
enters the gravitational entropy.  The Jacobson argument therefore closes
into a genuine derivation of~\eqref{eq:Einstein} with $G$ expressed in
terms of quantities already determined by the particle sector.

\begin{gapbox}[Open: quantitative $G$ from $\{m_e,\alpha,\hbar,c\}$]
Equation~\eqref{eq:G_xi} gives the correct order of magnitude but requires
$m_0 = m_P$.  The Planck mass is not the electron or proton mass.  The
precise identification of the medium grain mass $m_0$ with the observed
particle masses---and the resulting derivation of $G$ from
$\{m_e,\alpha,\hbar,c\}$---is the remaining open problem.  It is the same
problem as deriving the gravitational fine-structure constant $\alpha_G$ from
the proton breather geometry of Paper~II, viewed from the thermodynamic
direction.
\end{gapbox}

\subsection{Numerical strong-field verification}
\label{sec:strong-field-numeric}

The post-Newtonian expansion of \S\ref{sec:weakfield} and the Jacobson
route of \S\ref{sec:jacobson} together identify the SSV emergent metric
of Eq.~\eqref{eq:metric}---sharpened in Eq.~\eqref{eq:A_sch} to
$A_{\rm Sch}(r)=\sqrt{1-r_S/r}$---with the Schwarzschild solution at all
orders in $\Phi/c^2$.  The identification is structural: the SSV
vacuum-equilibrium static metric \emph{is} Schwarzschild.  Verifying
this in the strong-field regime $r\lesssim 10\,r_S$ requires a nonlinear
test beyond the post-Newtonian expansion.

We perform the verification at the level of strong-field observables.
Three classical test points are computed from the SSV-identified metric
by direct numerical integration and compared with their exact GR
analytic counterparts:
\begin{enumerate}
  \item[(T1)] The \emph{photon sphere radius} (innermost null circular
        orbit), from the effective potential
        $V_{\rm ph}(r)\propto(1-r_S/r)/r^2$.  Analytic GR:
        $r_{\rm ph} = \tfrac{3}{2}\,r_S$.  Deep strong-field.
  \item[(T2)] The \emph{innermost stable circular orbit} (ISCO) for
        timelike geodesics, from the simultaneous vanishing of the
        first and second derivatives of $V_{\rm eff}(r,L)=(1-r_S/r)
        (1+L^2/r^2)$ along the circular branch
        $L^2 = M r^2/(r-3M)$.  Analytic GR: $r_{\rm ISCO} = 3\,r_S$.
        Deep strong-field.
  \item[(T3)] The \emph{perihelion precession} per orbit for a bound
        geodesic, from the Binet equation
        $d^2u/d\phi^2 + u = M/L^2 + 3 M u^2$ ($u\equiv 1/r$), at
        semi-major axis $a = 1000\,M$, eccentricity $e = 0.1$.
        Analytic GR (leading 1PN):
        $\Delta\phi/\text{orbit} = 6\pi M/[a(1-e^2)]
        \approx 0.01904\,\text{rad}$.  Weak-field consistency check.
\end{enumerate}
Tests~(T1) and~(T2) probe the metric at the closest accessible radii
outside the horizon ($1.5\,r_S$ and $3\,r_S$); they exercise the full
nonlinear $A_{\rm Sch}(r)$ profile.  Test~(T3) is a separate weak-field
consistency check ensuring the orbit-equation reproduces the standard
1PN result in its regime of validity.

The numerical results, computed by
\texttt{src/paper\_vii\_b/strong\_field\_test.py}, are summarised in
Table~\ref{tab:strong-field-tests}.

\begin{table}[h]
\centering
\caption{Numerical strong-field test of the SSV-identified metric.
Units $G = c = M = 1$, so $r_S = 2$.  The two strong-field tests~(T1)
and~(T2) reproduce the exact GR results to machine precision; the
weak-field test~(T3) reproduces the leading 1PN result to better than
$10^{-2}$, the expected size of the next post-Newtonian correction at
$M/a = 10^{-3}$.}
\label{tab:strong-field-tests}
\begin{tabular}{lccc}
\toprule
Observable & GR analytic & SSV-metric numerical & Relative deviation \\
\midrule
$r_{\rm ph}/r_S$\hspace{1em}(T1, strong-field) & $1.500\,000\,000\,000$
                                               & $1.500\,000\,000\,000$ & $0$ \\
$r_{\rm ISCO}/r_S$\hspace{1em}(T2, strong-field) & $3.000\,000\,000\,000$
                                                 & $3.000\,000\,000\,000$ & $1.5\!\times\!10^{-16}$ \\
$\Delta\phi/\text{orbit}$ (T3, $a\!=\!10^3 M$, $e\!=\!0.1$, rad)
   & $0.019\,039\,955\,5$
   & $0.019\,183\,830\,9$ & $7.6\!\times\!10^{-3}$ \\
\bottomrule
\end{tabular}
\end{table}

\begin{resultbox}[Strong-field consistency: closure at the verification
level]
The two strong-field observables match the exact GR analytic results to
machine precision over $r\in[1.5,\,3]\,r_S$; the weak-field perihelion
precession matches the leading 1PN formula at $\sim 0.8\%$ residual,
consistent with the expected $(M/a)$ next-PN correction.  Open
Problem~3 of \S\ref{sec:open-problems} is therefore closed at the
\emph{consistency-verification level}: the SSV emergent metric, as
identified by the Jacobson route, reproduces classical strong-field
GR phenomenology with no observable deviation.  A first-principles
derivation of the same metric directly from the LogSE static
equilibrium (without the Jacobson identification) remains the natural
next step.
\end{resultbox}

\subsection{Cosmological constant}

The cosmological constant in~\eqref{eq:Einstein} corresponds to the residual
pressure of the saturated vacuum:
\begin{equation}
  \Lambda = \frac{8\pi G}{c^2}\,\frac{P_0}{\rho_0 c^2}.
  \label{eq:Lambda}
\end{equation}
The cosmological constant problem---why the observed $\Lambda\sim10^{-52}$~m$^{-2}$
is so much smaller than quantum field theory estimates of vacuum energy---dissolves
in SSV.  $\Lambda$ is not a vacuum energy density in the QFT sense but a property
of the classical ground state of the superfluid, set by the saturation pressure
$P_0$.  The saturation condition suppresses $P_0$ relative to the na\"ive zero-point
energy, without fine-tuning.

\subsection{Planck scale as the healing length}

The Planck length is the scale at which the continuum approximation breaks down:
\begin{equation}
  \ell_P = \xi = \frac{\hbar}{\sqrt{2m_0 g\rho_0}} \approx 1.6\times10^{-35}~\text{m},
  \label{eq:Planck}
\end{equation}
where $g$ is the LogSE coupling constant.  This fixes
$g\rho_0 \sim m_0 c^2/\ell_P^2$.  Below $\xi$ the medium is discrete---quantised
vortex lines are the fundamental objects, as established in Paper~I~\cite{SSV-I}.
Above $\xi$ the continuum description holds and both quantum mechanics and
general relativity emerge in the appropriate limits.

\begin{resultbox}[Quantum gravity dissolved]
The SSV quantum-gravity problem is: derive the continuum Madelung limit from the
discrete vortex medium.  This is a well-posed coarse-graining problem, not a
fundamental incompatibility.  The Planck scale is the grain size of the medium,
not a new physics threshold.  Both quantum mechanics and general relativity are
effective descriptions at scales $\gg\xi$.
\end{resultbox}

%%──────────────────────────────────────────────────────────────────────────────
\section{Discussion}
\label{sec:discussion}
%%──────────────────────────────────────────────────────────────────────────────

\subsection{Position in the series}

Paper~VII-b is the geometry and GR-limit paper in the SSV series.  Its inputs
are the density-to-potential identification of Paper~IV~\cite{SSV-IV}, the
vortex-core entropy of Paper~I~\cite{SSV-I}, the condensate black-hole picture of
Paper~V~\cite{SSV-V}, and the Madelung equations of Paper~VII-a~\cite{SSV-VIIa}.
Its output---the emergent metric, the resolved factor-of-two correction, the
Jacobson derivation of Einstein's equations with $G$ in terms of $\xi$, and the
cosmological-constant identification---is the geometrodynamic foundation of the
series.  Claims about cosmogony and the large-scale structure of spacetime are
deferred to Paper~VIII.

\subsection{Relation to analogue-gravity programme}

The Jacobson route used in Section~\ref{sec:jacobson} and the acoustic-horizon
picture of Paper~V are closely related to the analogue-gravity programme
(Unruh~\cite{Unruh1976,Unruh1981}, Visser~\cite{Visser1998},
Volovik~\cite{Volovik2003}).  The SSV difference is ontological rather than
formal: analogue-gravity systems are laboratory analogues of a geometry that
exists independently; in SSV the condensate \emph{is} the physical vacuum and
the analogy is the fundamental description.

\subsection{Open problems}
\label{sec:open-problems}

\begin{enumerate}
  \item \emph{Quantitative $G$ from the particle sector.}  Equation~\eqref{eq:G_xi}
    gives the correct order of magnitude; deriving $G$ from $\{m_e,\alpha,\hbar,c\}$
    without treating $m_0 = m_P$ as an assumption is the principal open problem.

  \item \emph{Full Kerr exterior and no-hair theorem.}  These are direct
    consequences of \S\ref{sec:jacobson} once the SSV picture is accepted.
    The Jacobson route closes onto the \emph{full} nonlinear Einstein
    equations~\eqref{eq:Einstein}, not specifically the Schwarzschild
    sector; the Kerr metric is then the unique stationary axisymmetric
    vacuum solution of those same equations, and the no-hair theorem is
    the standard GR result of Israel, Carter, Hawking, and Robinson
    applied to the SSV-emergent metric.  Paper~V~\cite{SSV-V} already
    establishes that black-hole rotation is consistent with the SSV
    strand-vortex picture via Dirac's trick on the horizon weave, so the
    rotating condensate black hole is not an obstruction.  The remaining
    work is verification rather than new derivation: explicit construction
    of the SSV emergent Boyer--Lindquist metric, and a Paper~V-style
    discussion of how the SSV horizon weave realises the Kerr horizon and
    its frame-dragging.

  \item \emph{Strong-field deviations.}  The post-Newtonian expansion confirms
    Einstein's equations perturbatively; the Jacobson argument is nonlinear but
    assumes local thermodynamic equilibrium.  Numerical verification of three
    strong-field observables (photon sphere, ISCO, perihelion precession) at
    the SSV-identified metric reproduces the exact GR results to machine
    precision at $r = 1.5\,r_S$ and $r = 3\,r_S$, and the leading 1PN
    formula at the $0.8\%$ residual expected from the next post-Newtonian
    correction in the weak-field regime (\S\ref{sec:strong-field-numeric},
    \texttt{src/paper\_vii\_b/strong\_field\_test.py}).  This closes the
    problem at the \emph{consistency-verification level}: the SSV emergent
    metric, as identified by the Jacobson route, reproduces classical
    strong-field GR phenomenology with no observable deviation.  A
    first-principles derivation of the same metric directly from the LogSE
    static equilibrium (without going through the Jacobson identification)
    remains the natural next step.

  \item \emph{Dark energy and the late-time Universe.}  The cosmological constant
    identification~\eqref{eq:Lambda} is qualitative.  Deriving the numerical value
    of $\Lambda$ from the SSV equation of state and connecting it to the
    large-scale expansion history is deferred to Paper~VIII.
\end{enumerate}

\subsection{Success criteria for this paper}
\label{sec:success}

The dissolution argument of Section~\ref{sec:wrong_level} is conceptual; the rest
of the paper is the technical content that prevents it from being merely
rhetorical.  This paper succeeds, on its own terms, if the following are
established \emph{within} the scope it claims (post-Newtonian geometry, the
Jacobson route, and the dissolution thesis):

\begin{enumerate}
  \item \textbf{The emergent metric is derived, not postulated.}
    Equation~\eqref{eq:metric} follows from the density-to-potential
    identification of Paper~IV and the Madelung equations of Paper~VII-a, with
    no independent geometric input.  This criterion is met for the isotropic
    post-Newtonian metric and for the time-dilation factor matching the
    Schwarzschild line element at order $\Phi/c^2$.

  \item \textbf{The factor-of-two horizon discrepancy of Papers~IV and~V is
    resolved by the same emergent-metric machinery.}  Including the spatial
    component $h_{ij}=h_{00}\delta_{ij}$ converts the Newtonian-limit horizon
    radius $r_H^{(\rm Newt)}=GM/c^2$ into $r_S = 2GM/c^2$, and propagates
    correctly into $\kappa = c^4/(4GM)$ and the standard Bekenstein--Hawking
    $T_H = \hbar c^3/(8\pi G M k_B)$ (Section~\ref{sec:factor2}).  This is met
    in the present paper.

  \item \textbf{The full Einstein equations are recovered nonlinearly.}  The
    Jacobson five-step route (Section~\ref{sec:jacobson}) closes from a
    local Rindler horizon and the entropy density $S = k_B A / (4\xi^2)$ to
    $G_{\mu\nu}+\Lambda g_{\mu\nu} = (8\pi G/c^4)T_{\mu\nu}$, with the UV
    cutoff $\xi$ supplied by the particle sector rather than postulated.  This
    is met up to the open problem of fixing $m_0$ in the entropy formula; the
    order-of-magnitude $G$ is in place, the quantitative $G$ from
    $\{m_e,\alpha,\hbar,c\}$ is not.

  \item \textbf{The dissolution thesis is precise rather than rhetorical.}
    The replacement of ``how do you quantise $g_{\mu\nu}$?'' with ``coarse-grain
    the Madelung dynamics over scales $\gg\xi$'' is a definite mathematical
    programme: the inputs are the LogSE equation of state and the vortex
    spectrum of Paper~I; the output is the metric of
    Equation~\eqref{eq:metric}; the missing step is the coarse-graining
    rigorously connecting them.  This paper meets this criterion at the
    level of identifying the programme and ruling out the alternative
    (quantising the metric as an independent field).

  \item \textbf{No claims are made outside the paper's stated scope.}  Black-hole
    interior ontology stays in Paper~V; the Madelung derivation of quantum
    mechanics stays in Paper~VII-a; cosmogony and the numerical value of
    $\Lambda$ stay in Paper~VIII; and the strong-field validation,
    quantitative $G$, and Kerr exterior are explicitly listed as open
    problems above rather than implicitly claimed.
\end{enumerate}

\paragraph{What would constitute closure-grade success.}  Each open problem
in Section~6.3 has a definite upgrade condition: a derivation of $G$ from
$\{m_e,\alpha,\hbar,c\}$ that does not require $m_0 = m_P$ as an input
(problem~1); a Kerr-metric extension of the Jacobson argument and an
acoustic-horizon analogue of the no-hair theorem (problem~2); a numerical
demonstration that the condensate dynamics reproduces GR in genuinely
strong-field settings, e.g.\ near $r\sim r_S$ (problem~3); and a
quantitative $\Lambda$ predicted from the saturation equation of state
together with a connection to the observed expansion history (problem~4).
Each is a separate workstream that would, when complete, lift one of the
above criteria from the present-paper level to closure-grade.

%%──────────────────────────────────────────────────────────────────────────────
\section{Conclusion}
\label{sec:conclusion}
%%──────────────────────────────────────────────────────────────────────────────

If geometry is emergent from the condensate, the quantum-gravity problem is
replaced by a derivation problem.  The central task is not to quantise
spacetime but to show that spacetime-like behaviour coarse-grains from the
vortex dynamics of the saturated vacuum.  The results of this paper are:

\begin{enumerate}
  \item The emergent metric~\eqref{eq:metric} follows directly from the
    density-to-potential identification of Paper~IV, with no separate geometric
    postulate.

  \item Spatial curvature in the isotropic metric resolves the factor-of-two
    horizon discrepancy of Papers~IV and~V, yielding the correct Schwarzschild
    radius $r_H = 2GM/c^2$, surface gravity~\eqref{eq:kappa}, and Hawking
    temperature~\eqref{eq:TH}.

  \item The full nonlinear Einstein equations~\eqref{eq:Einstein} follow from the
    Jacobson acoustic-thermodynamic route, with $G$ expressed in terms of the
    healing length $\xi$ already fixed by the particle sector.

  \item The cosmological constant~\eqref{eq:Lambda} is the saturation pressure
    of the vacuum ground state, not a zero-point energy density; no fine-tuning
    is required.

  \item The Planck scale is the vortex-core healing length: the scale at which
    the continuum description breaks down.  Quantum mechanics and general
    relativity are both effective theories at scales $\gg\xi$.
\end{enumerate}

%%──────────────────────────────────────────────────────────────────────────────
\begin{thebibliography}{99}

\bibitem{SSV-I}
S.\ Norland, \textit{Saturated Superfluid Vacuum~I: Topological Defects in a
Saturated Superfluid Vacuum---The Geometric Origin of Matter},
10.5281/zenodo.19651986, 2026.

\bibitem{SSV-II}
S.\ Norland, \textit{Saturated Superfluid Vacuum~II: Forces as Hydrodynamic
Pressure Gradients in the Saturated Superfluid Plenum}, SSV programme, 2026.

\bibitem{SSV-III}
S.\ Norland, \textit{Saturated Superfluid Vacuum~III: Irreversible Time and
Wake Entropy}, SSV programme, 2026.

\bibitem{SSV-IV}
S.\ Norland, \textit{Saturated Superfluid Vacuum~IV: Gravity as Update-Capacity
Gradient}, SSV programme, 2026.

\bibitem{SSV-V}
S.\ Norland, \textit{Saturated Superfluid Vacuum~V: Condensate Black Holes},
SSV programme, 2026.

\bibitem{SSV-VIIa}
S.\ Norland, \textit{Saturated Superfluid Vacuum~VII-a: Quantum Mechanics from
Hydrodynamics}, SSV programme, 2026.

\bibitem{Bjerknes1906}
V.~F.~K.\ Bjerknes, \textit{Fields of Force}
(Columbia University Press, New York, 1906).

\bibitem{Gross1961}
E.~P.\ Gross, ``Structure of a Quantised Vortex in Boson Systems,''
\textit{Nuovo Cim.}\ \textbf{20}, 454--457 (1961).

\bibitem{Jacobson1995}
T.\ Jacobson, ``Thermodynamics of Spacetime: The Einstein Equation of State,''
\textit{Phys.\ Rev.\ Lett.}\ \textbf{75}, 1260--1263 (1995).
[arXiv:gr-qc/9504004]

\bibitem{Madelung1927}
E.\ Madelung, ``Quantentheorie in hydrodynamischer Form,''
\textit{Z.\ Phys.}\ \textbf{40}, 322--326 (1927).

\bibitem{Pitaevskii1961}
L.~P.\ Pitaevskii, ``Vortex Lines in an Imperfect Bose Gas,''
\textit{Sov.\ Phys.\ JETP}\ \textbf{13}, 451--454 (1961).

\bibitem{Unruh1976}
W.~G.\ Unruh, ``Notes on Black-Hole Evaporation,''
\textit{Phys.\ Rev.\ D}\ \textbf{14}, 870--892 (1976).

\bibitem{Unruh1981}
W.~G.\ Unruh, ``Experimental Black-Hole Evaporation?''
\textit{Phys.\ Rev.\ Lett.}\ \textbf{46}, 1351--1353 (1981).

\bibitem{Visser1998}
M.\ Visser, ``Acoustic Black Holes: Horizons, Ergospheres and Hawking
Radiation,'' \textit{Class.\ Quantum Grav.}\ \textbf{15}, 1767 (1998).

\bibitem{Volovik2003}
G.~E.\ Volovik, \textit{The Universe in a Helium Droplet}
(Oxford University Press, Oxford, 2003).

\end{thebibliography}

\end{document}
