\documentclass[11pt]{article}

\usepackage[a4paper,margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{amsmath,amssymb}
\usepackage{physics}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{setspace}
\usepackage{mdframed}

\hypersetup{colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue}
\setstretch{1.1}

\newmdenv[linewidth=0.8pt,innertopmargin=5pt,innerbottommargin=5pt,%
  innerleftmargin=8pt,innerrightmargin=8pt]{gapbox}
\newmdenv[linewidth=1.1pt,innertopmargin=5pt,innerbottommargin=5pt,%
  innerleftmargin=8pt,innerrightmargin=8pt]{resultbox}

\title{Saturated Superfluid Vacuum V\\
Condensate Black Holes}
\author{Stig Norland}
\date{Draft}

\begin{document}
\maketitle

\begin{abstract}
This paper develops the ontology of black holes in the Saturated Superfluid Vacuum (SSV)
framework.  In SSV, the physical vacuum is a real medium described by a macroscopic order
parameter $\Psi$.  Matter consists of topological defects in $\Psi$; gravity arises from the
interference cross-term among those defects, depleting local update capacity and bending every
trajectory that crosses the resulting gradient (Paper IV).  The present paper asks what
happens at the extreme end of gravitational collapse.

The central argument proceeds in four steps.  First, because matter is made of vortex defects
with finite core size $\xi$, collapse terminates at the close-packing density of those cores
rather than at an infinite-density singularity.  Second, the surface at which outward acoustic
progress relative to the exterior vanishes --- the acoustic-freezing surface --- is identified
as the black-hole horizon.  It is a real physical surface of the medium, not a coordinate
artifact: at $r = r_H$, where the update availability $A(r) = 0$, every outward-directed
excitation is frozen in coordinate time.  Third, spin is a topological property of the
condensate rather than ordinary material rotation.  The circulation integral
$\oint\nabla\theta\cdot d\mathbf{l} = 2\pi n$ is a homotopy invariant; acoustic freezing of the
interior does not erase it, and angular momentum is conserved as topological winding.  Fourth,
because there is no singularity, information is not destroyed.  Infalling structure modifies
the condensate state and its boundary; the information is redistributed into present-state
wake complexity in the sense of Paper III, not annihilated.

A fifth result connects to Paper III's update-capacity language: the condensate's acoustic
horizon has a definite surface gravity $\kappa$, and a stationary observer near it must
accelerate to remain in place.  By the Unruh effect, such an observer registers thermal
radiation at temperature $T \propto \hbar c^3/(GM k_B)$, recovering the form of the Hawking
temperature.  The numerical coefficient carries a known factor-of-four gap, of the same origin
as the light-deflection gap in Paper IV: the leading-order Newtonian potential misses the
spatial-curvature contribution, which is deferred to Paper VII-b.

The paper claims the ontology and the mechanism.  It does not claim a full Kerr derivation,
an exact Bekenstein--Hawking coefficient, a black-hole population model, or galaxy-scale or
cosmological applications.
\end{abstract}

\newpage
\tableofcontents
\newpage

\section{Introduction}

This paper treats black holes in SSV as coherent condensate objects with finite interiors,
acoustic-freezing horizons, topological spin, and information-storing boundary structure.

The purpose is deliberately narrow.  The question is ontological: what is a black hole if the
vacuum is a saturated superfluid medium rather than an empty spacetime background?

The SSV answer, developed in the following sections, is that gravitational collapse does not
end at an infinite-density singularity.  It ends at a finite-density coherent state of the
vacuum medium, bounded by a surface at which outward acoustic progress tends to zero.  This
surface---the acoustic-freezing horizon---is not a coordinate artifact but a real physical
feature of the medium.  Spin is preserved as topological winding of the condensate phase.
Information is not annihilated; it is written into the present-state complexity of the
condensate and its boundary.

\subsection*{Dependencies}

This paper draws on Papers I--IV in the SSV series.  From Paper~I~\cite{SSV-I}: the LogSE
field equation, the vortex-defect ontology, and the healing length $\xi$.  From
Paper~II~\cite{SSV-II}: the gravitational potential $\Phi$ as an interference cross-term and
the fixed value of $G$.  From Paper~III~\cite{SSV-III}: proper time as the local change-rate
of $\Psi$, and entropy as wake-complexity.  From Paper~IV~\cite{SSV-IV}: the update
availability $A = d\tau/dt = 1 + \Phi/c^2$ and the derivation of gravitational time dilation,
deflection, and redshift from the one potential $\Phi$.

\subsection*{What this paper does not claim}

\begin{itemize}
  \item No full Kerr metric derivation.  Angular momentum in SSV is
  identified as topological winding; the exterior metric at all orders is deferred to
  Paper~VII-b~\cite{SSV-VIIb}.
  \item No exact Bekenstein--Hawking coefficient.  The area entropy is derived in form; the
  numerical prefactor requires the spatial-curvature correction and the identification
  $\xi = l_P$, both deferred.
  \item No black-hole population model, no claims about astrophysical near-extremal spin
  clustering.
  \item No galaxy-scale standing-wave applications and no broad cosmology.
\end{itemize}

\subsection*{Roadmap}

Section~2 fixes the minimal SSV background needed for strong-gravity analysis.
Section~3 argues that collapse terminates at finite density.  Section~4 derives the horizon
as an acoustic-freezing surface.  Section~5 establishes spin as topological winding.
Section~6 connects horizon structure to the wake-entropy framework of Paper~III.
Section~7 derives the Hawking temperature as an acoustic Unruh effect.
Section~8 collects open problems.

\section{SSV Setting for Strong Gravity}\label{sec:setting}

Only the parts of the SSV framework required for strong-gravity analysis are recalled here.
The full derivations are in Papers I--IV.

The vacuum order parameter is
\begin{equation}
  \Psi(\mathbf{x},t) = \sqrt{\rho(\mathbf{x},t)}\,e^{i\theta(\mathbf{x},t)},
\end{equation}
with density $\rho$, phase $\theta$, and superfluid velocity $\mathbf{v}_s = (\hbar/m)\nabla\theta$
away from vortex cores.  The field equation is the logarithmic Schr\"odinger equation
(LogSE)~\cite{SSV-I}
\begin{equation}\label{eq:logse}
  i\hbar\,\partial_t\Psi =
  \left[-\frac{\hbar^2}{2m}\nabla^2 + b\hbar\ln\!\left(\frac{|\Psi|^2}{\rho_0}\right)\right]\Psi,
\end{equation}
with coupling constant $b$ and background density $\rho_0$.  The Madelung decomposition of
\eqref{eq:logse} gives a continuity equation
\begin{equation}
  \partial_t\rho + \nabla\cdot(\rho\,\mathbf{v}_s) = 0,
\end{equation}
and a Bernoulli equation whose nonlinear potential $b\hbar\ln(\rho/\rho_0)$ sets the bulk
sound speed
\begin{equation}\label{eq:cs}
  c_s = \sqrt{\frac{b\hbar}{m}}
\end{equation}
and the healing length
\begin{equation}\label{eq:xi}
  \xi = \frac{\hbar}{mc_s} = \sqrt{\frac{\hbar}{mb}}.
\end{equation}

\subsection{Update capacity and the gravitational potential}

Paper~IV introduces the update availability $A(\mathbf{x},t)$, defined as the local rate of
proper-time advance per unit coordinate time,
\begin{equation}\label{eq:A}
  A = \frac{d\tau}{dt} = 1 + \frac{\Phi}{c^2},
\end{equation}
where $\Phi$ is the gravitational potential derived from the interference cross-term among
embedded defects.  For a spherically symmetric mass $M$ at $r \gg \xi$,
\begin{equation}\label{eq:phi}
  \Phi(r) = -\frac{GM}{r}
\end{equation}
to leading order~\cite{SSV-II,SSV-IV}.  Substituting into \eqref{eq:A},
\begin{equation}\label{eq:A_r}
  A(r) = 1 - \frac{GM}{rc^2}.
\end{equation}
This expression has the Schwarzschild structure: $A$ decreases from unity at $r\to\infty$ to
zero at a finite radius, and turns negative for smaller $r$.

\subsection{Outward acoustic propagation}

An acoustic mode emitted radially outward from position $r$ has, to leading order in
$\Phi/c^2$, a coordinate velocity
\begin{equation}\label{eq:vout}
  v_{\text{out}}(r) = c_s\cdot A(r).
\end{equation}
For long-wavelength excitations of the condensate, $c_s \to c$ (the Bogoliubov modes
propagate at $c$ in the long-wavelength limit~\cite{SSV-I}), so \eqref{eq:vout} becomes
\begin{equation}\label{eq:voutc}
  v_{\text{out}}(r) = c\cdot A(r).
\end{equation}
The sign and magnitude of $v_{\text{out}}$ track $A$: positive for $r > r_H$, zero at $r = r_H$,
and negative inside.  The acoustic horizon is therefore the surface $A(r_H) = 0$.

\section{Gravitational Collapse Without Singularity}\label{sec:collapse}

The Penrose singularity theorem~\cite{Penrose1965} establishes, within general relativity,
that gravitational collapse of ordinary matter must terminate in a geodesic-incompleteness
singularity once a trapped surface has formed.  SSV departs from this conclusion not by
evading the theorem but by operating outside its domain.  The medium described by the LogSE
is not a classical or quantum-field-theory matter source embedded in a fixed spacetime; it is
the spacetime-generating medium itself.  The singularity theorems do not apply.

\subsection{Close-packing of vortex cores}

In SSV, all matter consists of topological vortex defects in $\Psi$.  Each defect has a core
of radius $\sim\xi$ at which $|\Psi|\to 0$ with nonzero phase winding.  Two defects of the
same winding sign repel when their cores approach within $\sim\xi$: the energy cost of
overlapping density minima diverges as the cores overlap, because the phase-gradient kinetic
energy grows as the inter-core separation shrinks below $\xi$.

The maximum defect density is therefore set by close packing of cores,
\begin{equation}\label{eq:rhomax}
  \rho_{\max} \sim \frac{E_{\text{core}}}{\xi^3},
\end{equation}
where $E_{\text{core}} \sim \hbar c_s/\xi$ is the core energy per unit length times $\xi$.
Once the medium reaches $\rho_{\max}$, further compression would require overlapping cores,
which is topologically forbidden.  Collapse therefore terminates at finite density.

\begin{gapbox}
\textbf{Gap (Paper I program):} The exact value of $\rho_{\max}$, the threshold density
at which repulsion arrests collapse, requires a full 3D static minimisation of the LogSE with
defect close-packing boundary conditions.  This is the open computation of the Paper~I
programme~\cite{SSV-I}.  The qualitative claim---finite termination density---does not
depend on that number.
\end{gapbox}

\subsection{The condensate endpoint}

The endpoint of collapse is a finite-density coherent condensate state of the medium.  Its
defining properties are:
\begin{enumerate}
  \item Interior density $\rho_{\text{int}}$ finite and $\lesssim \rho_{\max}$.
  \item Phase $\theta(\mathbf{x})$ and density $\rho(\mathbf{x})$ well-defined throughout the
    interior.
  \item A causal boundary---the acoustic-freezing horizon---separating interior from exterior.
\end{enumerate}

The condensate picture resolves the apparent paradox of supermassive black holes.  A black
hole of mass $M = 10^9\,M_\odot$ has Schwarzschild radius
$r_S = 2GM/c^2 \approx 3\times 10^{12}$~m and average density
\begin{equation}
  \bar\rho = \frac{3M}{4\pi r_S^3} \approx 10^{-5}\;\text{kg}\,\text{m}^{-3},
\end{equation}
far below nuclear density.  There is no requirement that the condensate be locally
Planck-dense throughout its volume; the defining feature is not local density but the
formation of a horizon-scale acoustic-freezing surface.

\section{Horizon as Acoustic-Freezing Surface}\label{sec:horizon}

\subsection{The freezing condition}

Setting $A(r_H) = 0$ in \eqref{eq:A_r}:
\begin{equation}\label{eq:rH}
  r_H = \frac{GM}{c^2}.
\end{equation}
At $r = r_H$, every outward acoustic mode has $v_{\text{out}} = 0$: the acoustic horizon.
For $r < r_H$, $A < 0$, so $v_{\text{out}} < 0$: even nominally outward-directed modes are
dragged inward in coordinate time.  The region $r < r_H$ is causally disconnected from the
exterior.

\begin{gapbox}
\textbf{Gap (Paper VII-b):} The Newtonian potential \eqref{eq:phi} misses the spatial
curvature contribution to $\Phi$, which is the same factor deferred in Paper~IV~\S6.  The
full GR result is $r_S = 2GM/c^2$, a factor of two larger than \eqref{eq:rH}.  The mechanism
and qualitative structure are unaffected; the exact coefficient is established once
Paper~VII-b~\cite{SSV-VIIb} derives the spatial half of the metric.
\end{gapbox}

\subsection{Freezing versus pinning}

The acoustic-freezing condition has a specific meaning.  It does not assert that the
condensate interior is spatially immobile in any absolute sense.  It asserts that
\emph{outward signal propagation relative to the exterior observer} tends to zero.

The distinction is essential:
\begin{itemize}
  \item \emph{Freezing}: the failure of outward acoustic propagation, measured in coordinate
    time.  This is what the horizon imposes.
  \item \emph{Pinning}: absolute spatial immobility.  This is not claimed.
\end{itemize}

A frozen condensate can still translate as a whole (its centre of mass can move), can carry
angular momentum (Section~\ref{sec:spin}), and can store information at its boundary
(Section~\ref{sec:entropy}).  These are global and topological properties that are not
ordinary outward sound pulses escaping from the interior.

\subsection{Coordinate time and proper time at the horizon}

From Paper~III, proper time is the local change-rate of $\Psi$.  At $r = r_H$, $A = 0$: the
interior clock advances at zero rate relative to the exterior coordinate.  In the language of
Paper~III, the interior update-rate $\mathcal{R}[\Psi]$, as observed from outside, is zero.

Interior evolution does continue in the condensate's own local time.  The condensate medium
has definite $\Psi$ at every interior point; it is not static in any intrinsic sense.  The
freezing is a causal isolation, not a cessation of interior change.

\subsection{Summary statement}

The claim of this section is therefore:
\begin{quote}
  The black-hole horizon is the surface at which the medium's update availability $A(r)$
  vanishes.  Outward acoustic progress freezes at this surface, causally isolating the
  interior condensate from the exterior.  The surface is real and physical: it is a
  topographic feature of the update-capacity field $A(\mathbf{x})$, not a choice of
  coordinates.
\end{quote}

\section{Spin Without Acoustic Time}\label{sec:spin}

The most immediate objection to the condensate picture is: if time freezes at the horizon,
why does a black hole spin?

The answer is that spin in SSV is not modeled as ordinary internal material rotation
requiring freely propagating acoustic time through the interior.  It is a topological
property of the condensate phase field.

\subsection{Circulation and winding number}

The superfluid velocity is $\mathbf{v}_s = (\hbar/m)\nabla\theta$.  The circulation around
a closed loop $\mathcal{C}$ is
\begin{equation}\label{eq:circulation}
  \Gamma_\mathcal{C} = \oint_\mathcal{C}\mathbf{v}_s\cdot d\mathbf{l}
  = \frac{\hbar}{m}\oint_\mathcal{C}\nabla\theta\cdot d\mathbf{l}
  = \frac{2\pi\hbar}{m}\,n_\mathcal{C},
\end{equation}
where $n_\mathcal{C}\in\mathbb{Z}$ is the winding number of the phase around $\mathcal{C}$.
This is quantised and topologically protected: $n_\mathcal{C}$ is a homotopy invariant that
changes only when a vortex line crosses $\mathcal{C}$.

\subsection{Angular momentum as topological winding}

For a condensate with axial symmetry and winding number $n$ around the symmetry axis, the
angular momentum is
\begin{equation}\label{eq:angmom}
  J = \int_V \rho\,(\mathbf{r}\times\mathbf{v}_s)\,d^3x
  = \frac{\hbar\,n}{m}\,M_{\text{cond}},
\end{equation}
where $M_{\text{cond}} = \int\rho\,d^3x$ is the condensate mass.  The dimensionless spin
parameter
\begin{equation}
  a \equiv \frac{Jc}{GM^2} = \frac{\hbar nc^2}{Gm M_{\text{cond}}^2}\cdot M_{\text{cond}}
  = \frac{\hbar nc^2}{Gm M_{\text{cond}}}
\end{equation}
is a topological label of the condensate state.

\subsection{Topological protection through the horizon}

The winding number $n_\mathcal{C}$ for any loop $\mathcal{C}$ threading the condensate can
change only if a vortex line crosses $\mathcal{C}$.  For a loop encircling the condensate at
$r > r_H$, the relevant mechanism would be expulsion of a vortex filament through the
acoustic-freezing surface.

A vortex filament crossing the acoustic horizon must carry energy from the interior to the
exterior.  But the horizon is precisely the surface from which no acoustic energy propagates
outward.  A filament terminating at the horizon or expelled through it requires the medium
to move outward across $r_H$---which the acoustic-freezing condition forbids for ordinary
dynamical modes.

\paragraph{Penrose process and superradiance: yes, by emergent Kerr structure.}
\label{para:penrose-ssv}
External-perturbation extraction is the more interesting case, and the answer
is clean: the analogue Penrose process and the analogue superradiance both
occur in SSV, by the same mechanism as in the standard GR Kerr exterior.
The reason is structural.  The SSV-emergent metric of a rotating condensate
black hole is the Kerr metric (Paper~VII-b~\cite{SSV-VIIb}, \S6.3,
problem~2; the Jacobson route gives the full nonlinear Einstein equations,
and Kerr is the unique stationary axisymmetric vacuum solution).  Standard GR
processes that depend on the Kerr ergoregion, namely the Penrose
process~\cite{Penrose1971} and bosonic superradiance, therefore lift to the
SSV-emergent geometry without modification at the leading order treated
here.

What is distinct in SSV is the \emph{identification} of the extracted
quantities.  In Penrose's process, the particle that returns to infinity
carries more energy than the incoming probe; in SSV this energy is acoustic
energy of the surrounding medium, and the angular momentum carried by the
returning particle is condensate winding transferred from the horizon
weave to the external defect.  The horizon weave's topological winding
decreases by the amount transferred, exactly as the Kerr horizon's
angular-momentum parameter $a$ decreases in the GR calculation.  Total
winding is still conserved globally because the external defect carries off
the difference.

The topological-protection claim of the previous paragraph is compatible:
it concerns \emph{spontaneous} expulsion, where no external driving carries
the winding away.  Penrose extraction and superradiance require explicit
external coupling to the ergoregion, and that coupling is what transfers the
winding outward.  The two statements describe different regimes and do not
conflict.

\subsection{Separation of two claims}

Two claims that can be conflated should be kept separate:
\begin{itemize}
  \item \emph{Spin persistence:} a condensate black hole carries angular momentum as
    topological winding.  This is the claim of the present paper.
  \item \emph{Population claim:} astrophysical black holes cluster near particular spin
    values.  This requires separate observational and dynamical support and is explicitly
    outside the scope of this paper.
\end{itemize}

\section{Horizon Entropy and Information Storage}\label{sec:entropy}

\subsection{From wake complexity to horizon entropy}

Paper~III establishes that entropy in SSV is the irreducible wake-complexity of the present
state of $\Psi$:
\begin{equation}
  S_{\text{wake}} = k_B\,W[\Psi(t)],
\end{equation}
where $W[\Psi]$ measures the amount of present-state structure required to encode the
surviving wake of prior events.  For an isolated black-hole condensate, the relevant present
structure is the condensate interior plus its horizon boundary.

Infalling matter modifies the condensate state.  From the exterior, this modification is
written into the boundary structure of the condensate at $r \approx r_H$, because that is the
last surface at which incoming information can leave a physical impression before being
causally isolated.  The information is not destroyed; it is redistributed into the wake
complexity of the horizon-scale boundary.

\subsection{Area entropy from condensate surface counting}

The surface gravity argument of Section~\ref{sec:hawking} will establish that the relevant
physical scale at the horizon is the healing length $\xi$.  The horizon surface of area
$A_H$ can be partitioned into cells of size $\xi^2$ (the smallest independent region of the
condensate surface).  The number of independent configurational states of the condensate
boundary is, schematically,
\begin{equation}\label{eq:omega}
  \Omega_H \sim \exp\!\left(\frac{A_H}{\xi^2}\right).
\end{equation}
From the Boltzmann entropy,
\begin{equation}\label{eq:SH}
  S_H = k_B\ln\Omega_H \sim k_B\,\frac{A_H}{\xi^2}.
\end{equation}
The Bekenstein--Hawking area law~\cite{Bekenstein1973,Hawking1975} is
$S_{\text{BH}} = k_B c^3 A_H/(4G\hbar) = k_B A_H/(4l_P^2)$, with
$l_P = \sqrt{\hbar G/c^3}$ the Planck length.  Equation~\eqref{eq:SH} reproduces the
area dependence.  The coefficient requires $\xi = l_P$, which holds when the constituent
boson mass equals the Planck mass:
\begin{equation}\label{eq:mP}
  \xi = l_P \;\iff\; \frac{\hbar}{mc_s} = \sqrt{\frac{\hbar G}{c^3}}
  \;\iff\; m = \sqrt{\frac{\hbar c}{G}} = m_P.
\end{equation}

\begin{gapbox}
\textbf{Gap (Paper I program):}  The identification $m = m_P$---and hence the exact
coefficient in~\eqref{eq:SH}---requires deriving the constituent boson mass of the
fundamental SSV condensate from the LogSE.  This is part of the open computation of the
Paper~I programme.  The area dependence of $S_H$ is established without it.
\end{gapbox}

\begin{gapbox}
\textbf{Gap (Paper VII-b):}  The horizon area $A_H$ also carries the factor-of-four
correction from the spatial-curvature contribution to the horizon radius:
$A_H^{\text{GR}} = 4\pi r_S^2 = 4\pi(2GM/c^2)^2$ versus
$A_H^{\text{SSV}} = 4\pi r_H^2 = 4\pi(GM/c^2)^2$.  The exact Bekenstein--Hawking
coefficient absorbs both this factor and~\eqref{eq:mP}; both are deferred to Papers I and
VII-b.
\end{gapbox}

\subsection{No information destruction}

The condensate picture dissolves the information-loss problem rather than solving it.  The
problem assumes that collapse terminates at a singularity where physical structure is
destroyed.  In SSV, there is no such endpoint.  The condensate retains a definite $\Psi$
everywhere inside $r_H$.  Infalling structure changes the condensate state; the exterior
loses \emph{access} to the fine interior organisation, but access loss is not physical
annihilation.

This is consistent with the wake-entropy framework of Paper~III.  The distinction between
``information inaccessible to the exterior'' and ``information destroyed'' is the distinction
between high coarse-grained entropy (the exterior observer cannot resolve the fine wake
structure) and literal erasure of the structure from $\Psi$.  The medium keeps its records;
the observer simply cannot read them from outside.

\section{Thermal Emission from the Acoustic Horizon}\label{sec:hawking}

\subsection{The acoustic Unruh effect}

Unruh~\cite{Unruh1976,Unruh1981} showed that an observer with proper acceleration $a$ in
flat spacetime registers thermal radiation at temperature
\begin{equation}\label{eq:Tunruh}
  T_U = \frac{\hbar a}{2\pi k_B c}.
\end{equation}
The same effect arises at any acoustic horizon in a medium where the flow velocity equals the
local sound speed~\cite{Unruh1981,BLV2005}: a stationary observer just outside the sonic
horizon must accelerate to maintain position against the flow, and registers thermal phonons
at the temperature set by its acceleration.

In SSV, the acoustic-freezing surface at $r = r_H$ plays the exact role of the sonic horizon.
An observer who remains stationary at $r$ slightly above $r_H$ must sustain a proper
acceleration to resist the infall tendency of the medium.  This acceleration is set by the
gradient of the update availability.

\subsection{Surface gravity and Hawking temperature}

A stationary observer at radius $r$ must accelerate with proper acceleration
\begin{equation}\label{eq:a_r}
  a(r) = c^2\left|\frac{dA}{dr}\right| = c^2\,\frac{GM}{r^2 c^2} = \frac{GM}{r^2}
\end{equation}
to remain at fixed $r$ in the potential $\Phi = -GM/r$.  This is the Newtonian
gravitational acceleration, confirmed from the same $A(r)$ that yields the time dilation of
Paper~IV.

At the horizon $r = r_H = GM/c^2$:
\begin{equation}\label{eq:surface_gravity}
  \kappa \equiv c^2\left|\frac{dA}{dr}\right|_{r_H}
  = c^2\cdot\frac{GM}{r_H^2 c^2} = \frac{c^4}{GM}.
\end{equation}
The surface gravity $\kappa$ is the coordinate acceleration needed to hover at the horizon.
Substituting into~\eqref{eq:Tunruh},
\begin{equation}\label{eq:TH_SSV}
  T_H^{\text{SSV}} = \frac{\hbar\kappa}{2\pi k_B c} = \frac{\hbar c^3}{2\pi k_B GM}.
\end{equation}
The standard Hawking temperature~\cite{Hawking1974,Hawking1975} is
\begin{equation}\label{eq:TH_GR}
  T_H^{\text{GR}} = \frac{\hbar c^3}{8\pi k_B GM}.
\end{equation}
The two expressions agree in their dependence on $\hbar$, $c$, $G$, $M$, and $k_B$.  The
numerical factor differs by four.

\begin{gapbox}
\textbf{Gap (Paper VII-b):}  The factor-of-four discrepancy has the same origin as the
light-deflection and horizon-location gaps: the leading-order potential $\Phi = -GM/r$ places
the horizon at $r_H = GM/c^2$ instead of the GR value $r_S = 2GM/c^2$, and the GR surface
gravity includes the redshift factor at the horizon.  Both corrections are recovered once the
spatial-curvature contribution to the metric is derived in Paper~VII-b.
\end{gapbox}

\subsection{Qualitative consequences}

Even without the numerical coefficient, the SSV derivation establishes three points.

\textit{Hawking radiation is not a quantum-gravity phenomenon in SSV.}  It is the acoustic
Unruh effect of a condensate with a real physical horizon---the same effect that has been
observed in analogue acoustic black holes in Bose--Einstein condensates~\cite{Steinhauer2016}.
No new physics beyond the LogSE and the update-capacity framework is needed.

\textit{The temperature is set by the horizon gradient.}  From~\eqref{eq:TH_SSV},
$T_H \propto c^4/(GM)$.  Smaller (lighter) black holes have higher Hawking temperatures.
As the condensate radiates, $M$ decreases, $r_H$ decreases, and $T_H$ increases---the
well-known Hawking evaporation runaway.

\textit{The endpoint of evaporation is not a naked singularity.}  As $r_H\to\xi$, the
condensate reaches the Planck scale.  At this point the close-packing regime of
Section~\ref{sec:collapse} is reached and the condensate cannot compress further.  The
evaporation terminates at a Planck-scale remnant rather than a singularity; the
next subsection establishes that the remnant is structurally stable rather than
dissolving back into the medium.

\subsection{The Planck-scale evaporation endpoint}
\label{sec:evaporation-endpoint}

The qualitative termination above leaves a structural question open: does the
Planck-scale remnant dissolve completely into the saturated medium, or does it
stabilise as a topologically protected defect?  The close-packing equation of
state of \S\ref{sec:collapse} answers this definitively.  The LogSE
equation of state has three properties that together fix the endpoint as a
\emph{stable} Planck-scale remnant.

\paragraph{Argument 1: the LogSE chemical-potential floor.}
The logarithmic potential $V(\rho) = -b\rho\ln(\rho/\rho_0)$ has chemical
potential
\begin{equation}
  \mu(\rho) \;=\; \frac{\delta E}{\delta\rho}
                \;=\; -b\bigl[\ln(\rho/\rho_0)+1\bigr],
  \label{eq:logSE-chempot}
\end{equation}
which diverges to $+\infty$ as $\rho\to 0$.  Any region whose density tries to
fall below $\rho_0$ pays a logarithmic chemical-potential cost; the surrounding
saturated medium at $\mu(\rho_0) = -b$ therefore exerts an inward pressure on
any sub-saturated region.  A remnant attempting to ``dissolve'' would need to
expand below saturation in a finite region, against this confining pressure.
The LogSE equation of state forbids smooth dissolution.

\paragraph{Argument 2: topological-charge conservation.}
The original collapsing condensate carries integer winding number $n$ around
its axis (Eq.~\eqref{eq:angmom}, \S\ref{sec:spin}).  Hawking evaporation
radiates phonons in the longitudinal pressure channel; these are pure phase
disturbances of the saturated medium and carry zero winding.  Throughout
evaporation, $n$ is therefore conserved.  At the endpoint a fully dissolved
($n = 0$) state cannot be reached from a non-trivial parent ($n \ne 0$)
without an unwinding event that passes through $\rho \to 0$ over an extended
region---precisely what Argument 1 forbids.  Any black hole with $J \ne 0$
must therefore leave a winding-bearing remnant.

\paragraph{Argument 3: horizon vanishing at the Planck scale.}
The acoustic-freezing mechanism of \S\ref{sec:horizon} requires a smooth
$A(r)$ profile with $A\to 0$ over a region of scale $\gtrsim\xi$ for the
continuum acoustic approximation to apply.  When $r_H\lesssim\xi$ the smooth
horizon can no longer be supported by the medium: the very mechanism that
produces Hawking radiation switches off.  The remnant ceases to be a
``black hole'' in the acoustic-horizon sense and becomes a static
Planck-scale condensate defect, no longer radiating.  This terminates
evaporation at a definite mass scale.

\paragraph{Endpoint mass and structure.}
Combining the three arguments, the evaporation endpoint is a static
Planck-scale defect with
\begin{equation}
  M_{\rm rem} \;\sim\; \frac{c^2\,\xi}{G} \;\sim\; m_P,
  \qquad
  r_{\rm rem} \;\sim\; \xi \;\sim\; \ell_P,
  \label{eq:M_remnant}
\end{equation}
carrying the parent black hole's topological winding $n$ and confined by the
chemical-potential pressure of the surrounding saturated medium.  The defect
cannot annihilate spontaneously: doing so would require either dissolution
(forbidden by Argument 1) or topological-charge violation (forbidden by
Argument 2).  Annihilation requires meeting an oppositely-wound partner,
exactly as for any topologically protected defect in a superfluid.

\begin{resultbox}
\textbf{Structural answer: stable Planck-scale remnant.}\\[2pt]
The close-packing equation of state of \S\ref{sec:collapse} forces the
evaporation endpoint to be a \emph{stable} Planck-scale defect, not a
dissolved state.  Three independent arguments converge:
\begin{enumerate}\setlength\itemsep{2pt}
  \item the LogSE chemical potential diverges as $\rho\to 0$, so the
        saturated medium confines any sub-saturated region;
  \item the topological winding $n$ is conserved under Hawking evaporation
        (phonons carry no winding); a winding-bearing parent cannot
        evaporate to the trivial state without violating Argument 1;
  \item the acoustic-horizon mechanism switches off at $r_H\sim\xi$, so
        the remnant ceases to radiate at the Planck mass.
\end{enumerate}
Status: \emph{structural}.  The mass scale
$M_{\rm rem}\sim m_P$ inherits the
$\xi\sim\ell_P$ identification (Paper~VII-b~\cite{SSV-VIIb}, candidate-grade);
the no-dissolution conclusion is independent of the quantitative
close-packing density $\rho_{\max}$ tracked under the Paper~I
programme~\cite{SSV-I}.  The qualitative answer to the issue is therefore
definite without the deferred numerics.
\end{resultbox}

\paragraph{Observational consequence: Planck-mass relics.}
Every black hole that completes Hawking evaporation leaves behind a single
Planck-mass topological defect.  Primordial black holes (PBHs) with initial
mass $M_*\sim 5\times10^{14}\,\mathrm{g}$ would have finished evaporating
in the present cosmological epoch, depositing $\sim M_*/m_P \sim 10^{19}$
times less remnant mass than parent mass --- a remnant density today of
$\Omega_{\rm rem}\sim 10^{-19}\,\Omega_{\rm PBH,initial}$, well below
current dark-matter constraints unless the initial PBH abundance is
near the present dark-matter density.  Heavier PBHs ($M\gg M_*$) have not
yet reached the endpoint and contribute negligibly to the relic density.
The framework therefore predicts a definite but typically small remnant
contribution to dark matter, with the size set by the PBH formation
spectrum rather than by the endpoint mechanism itself.

\section{Discussion and Open Problems}\label{sec:discussion}

\subsection{Correspondence with general relativity}

The condensate picture reproduces the following features of GR black holes:
\begin{itemize}
  \item A causal horizon that prevents outward signal propagation.
  \item Gravitational time dilation growing without bound as $r\to r_H$ (since $A\to 0$).
  \item Hawking radiation with the correct parametric dependence on $\hbar$, $c$, $G$, $M$.
  \item Angular momentum as a conserved parameter characterising the exterior field.
  \item Horizon entropy proportional to surface area.
\end{itemize}

The correspondence not claimed here:
\begin{itemize}
  \item No Kerr metric derivation.  The exterior field of a spinning condensate is
    characterised by $M$ and $J$, but the exact spacetime geometry (including
    frame-dragging) requires the full emergent-metric treatment of Paper~VII-b.
  \item No exact Bekenstein--Hawking coefficient.  This requires $m = m_P$ and the
    spatial-curvature correction to $r_H$, both deferred.
  \item No PPN extension beyond leading order.
\end{itemize}

\subsection{The no-hair structure}

In GR, the no-hair theorem states that the exterior of a stationary black hole is completely
characterised by mass $M$, angular momentum $J$, and electric charge $Q$.  In SSV, the
exterior acoustic field generated by the interference cross-term is determined by the same
three quantities:
\begin{itemize}
  \item $M$ from the total condensate mass (integrated interference cross-term, Paper~II).
  \item $J$ from the topological winding number $n$, via \eqref{eq:angmom}.
  \item $Q$ from a global $U(1)$ phase of $\Psi$, if the condensate carries charge.
\end{itemize}
Any other structure of the condensate interior would need to appear in the exterior field as
a multipole moment above the monopole and dipole---but the LogSE cross-term in the
far-field limit gives only a monopole (the Newtonian $1/r$ potential) and a dipole
(the frame-dragging-like term from the axial circulation).  Higher multipoles are
exponentially suppressed at $r\gg r_H$.

\begin{gapbox}
\textbf{Gap (Paper VII-b):}  A formal proof that the exterior field has only $M$, $J$, $Q$
to all orders, and the derivation of the full Kerr-like exterior metric, are deferred to
Paper~VII-b.
\end{gapbox}

\subsection{Relation to analogue-gravity programme}

The identification of the SSV horizon as an acoustic-freezing surface is not merely an
analogy.  In SSV, the acoustic horizon \emph{is} the physical black-hole horizon.  The
condensate medium that produces gravity is the same medium whose excitations propagate at $c$
and whose phonon spectrum terminates at the acoustic-freezing surface.  The analogue-gravity
programme~\cite{BLV2005} provides the mathematical tools for the acoustic Hawking effect;
SSV provides the physical ontology that makes the correspondence exact rather than
approximate.

\subsection{Open problems}

The following technical tasks remain for later work:
\begin{enumerate}
  \item \textbf{Close-packing density:} Numerical determination of $\rho_{\max}$ from the
    3D static minimisation of the LogSE with vortex close-packing boundary conditions
    (Paper~I programme).
  \item \textbf{Boson mass:} Derivation of $m = m_P$ (or the corresponding $\xi = l_P$)
    from the fundamental condensate specification (Paper~I programme).
  \item \textbf{Spatial curvature:} The factor-of-two correction to $r_H$ and the
    factor-of-four correction to $T_H$ and $S_H$, from the full emergent metric
    (Paper~VII-b).
  \item \textbf{Kerr exterior:} The exterior metric of a spinning condensate, including
    frame-dragging and the exact no-hair result (Paper~VII-b).
  \item \textbf{Evaporation endpoint:} The qualitative final state of Hawking
    evaporation as $r_H\to\xi$ is established as a stable Planck-scale
    topological remnant in \S\ref{sec:evaporation-endpoint}, on three
    structural grounds (LogSE chemical-potential floor, topological
    winding conservation, acoustic-horizon vanishing at $r_H\sim\xi$).
    The quantitative close-packing density $\rho_{\max}$ and the precise
    relation $\xi = \ell_P$ remain in the Paper~I programme and
    Paper~VII-b respectively; neither affects the no-dissolution
    conclusion.
  \item \textbf{Superradiance:} Whether angular momentum can be extracted from the condensate
    by external perturbations, and whether a Kerr-like superradiance instability exists.
\end{enumerate}

\section{Conclusion}

In SSV, black holes are finite coherent condensates of the vacuum medium.  Collapse
terminates not at a singularity but at the close-packing density of topological vortex cores.
The horizon is the acoustic-freezing surface $A(r_H) = 0$, a real topographic feature of the
medium's update-capacity field.  Spin is topological winding of the condensate phase,
protected by homotopy against acoustic freezing.  Information is written into the
present-state wake complexity of the condensate and its boundary rather than destroyed.  The
horizon emits thermal phonons at the acoustic Unruh temperature $T_H \propto \hbar c^3/(GMk_B)$,
recovering Hawking radiation as an ordinary consequence of the medium's acoustic structure.

Three gapboxes carry forward the same spatial-curvature correction deferred in Paper~IV:
the exact horizon radius, the Bekenstein--Hawking coefficient, and the Hawking temperature
prefactor each require the emergent-metric treatment of Paper~VII-b.  Two further gapboxes
are inside the Paper~I numerical programme: the close-packing density and the identification
$m = m_P$.  Nothing in the present paper depends on those numbers for its qualitative claims.

\begin{thebibliography}{99}

\bibitem{SSV-I}
S.~Norland, \textit{Saturated Superfluid Vacuum I: Topological Defects in a Saturated
Superfluid Vacuum---The Geometric Origin of Matter and the $\alpha$-Harmonic Mass Ladder},
10.5281/zenodo.19651986, 2026.

\bibitem{SSV-II}
S.~Norland, \textit{Saturated Superfluid Vacuum II}, SSV programme, 2026.

\bibitem{SSV-III}
S.~Norland, \textit{Saturated Superfluid Vacuum III: Irreversible Time and Wake Entropy},
SSV programme, 2026.

\bibitem{SSV-IV}
S.~Norland, \textit{Saturated Superfluid Vacuum IV: Gravity as Update-Capacity Gradient},
SSV programme, 2026.

\bibitem{SSV-VIIb}
S.~Norland, \textit{Saturated Superfluid Vacuum VII-b: Emergent Geometry and the Dissolution
of Quantum Gravity}, SSV programme, forthcoming.

\bibitem{Penrose1965}
R.~Penrose, ``Gravitational collapse and space-time singularities,''
\textit{Phys.~Rev.~Lett.}~\textbf{14}, 57--59 (1965).

\bibitem{Bekenstein1973}
J.~D.~Bekenstein, ``Black holes and entropy,''
\textit{Phys.~Rev.~D}~\textbf{7}, 2333--2346 (1973).

\bibitem{Hawking1974}
S.~W.~Hawking, ``Black hole explosions?,''
\textit{Nature}~\textbf{248}, 30--31 (1974).

\bibitem{Hawking1975}
S.~W.~Hawking, ``Particle creation by black holes,''
\textit{Commun.~Math.~Phys.}~\textbf{43}, 199--220 (1975).

\bibitem{Unruh1976}
W.~G.~Unruh, ``Notes on black-hole evaporation,''
\textit{Phys.~Rev.~D}~\textbf{14}, 870--892 (1976).

\bibitem{Unruh1981}
W.~G.~Unruh, ``Experimental black-hole evaporation?,''
\textit{Phys.~Rev.~Lett.}~\textbf{46}, 1351--1353 (1981).

\bibitem{BLV2005}
C.~Barcel\'o, S.~Liberati, and M.~Visser, ``Analogue gravity,''
\textit{Living Rev.~Rel.}~\textbf{8}, 12 (2005).

\bibitem{Penrose1971}
R.~Penrose, ``Extraction of rotational energy from a black hole,''
\textit{Nat.~Phys.~Sci.}~\textbf{229}, 177--179 (1971).

\bibitem{Steinhauer2016}
J.~Steinhauer, ``Observation of quantum Hawking radiation and its entanglement in an analogue
black hole,'' \textit{Nat.~Phys.}~\textbf{12}, 959--965 (2016).

\end{thebibliography}

\end{document}
