\documentclass[11pt]{article}

\usepackage[a4paper,margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{amsmath,amssymb}
\usepackage{physics}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{setspace}
\usepackage{tcolorbox}
\tcbuselibrary{skins,breakable}

\newtcolorbox{gapbox}[1][]{enhanced,breakable,
  colback=orange!6!white,colframe=orange!60!black,
  fonttitle=\bfseries,title={#1},boxrule=0.6pt,arc=4pt}

\newtcolorbox{falsbox}[1][]{enhanced,breakable,
  colback=red!5!white,colframe=red!55!black,
  fonttitle=\bfseries,title={#1},boxrule=0.6pt,arc=4pt}

\hypersetup{colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue}
\setstretch{1.1}

\title{Saturated Superfluid Vacuum IV\\
Gravity as Update-Capacity Gradient}
\author{Stig Norland}
\date{Draft}

\providecommand{\ssvfile}[1]{\hyperref[file:#1]{\texttt{\detokenize{#1}}}}% in-text code/report ref -> Code & Issue References appendix
\begin{document}
\maketitle

\begin{abstract}
This paper develops gravity in the Saturated Superfluid Vacuum (SSV)
framework as the bending of motion through gradients of local time delay.
It builds on Paper~III, where physical time is identified with the
local rate of change of the order parameter $\Psi$ --- the medium's local
\emph{update capacity}.  The chain is short: loading the medium consumes
update capacity and so slows local time; that slowing is
uneven and therefore has a gradient; and any motion through the gradient is
bent toward the region of greatest delay.  That bending is gravity.  From a
single potential $\Phi$ follow free fall, gravitational time
dilation, light deflection, and redshift, each at leading order in
$\Phi/c^2$.  The treatment is weak-field throughout: it delivers the time
sector of gravity in full --- including the time-delay (Einstein 1911) half
of light deflection --- with the spatial sector and the recovery of full
general relativity left to Paper~VII-b.

\textbf{Status change (issue \#119, 2026-06).}  The original account of
\emph{what builds} $\Phi$ --- the time-averaged interference of the
pressure waves that breathers mutually radiate (Paper~II's Bjerknes
mechanism, \S\ref{sec:interference} here) --- is \emph{falsified as
written}: beyond one wavelength the time-averaged cross-term oscillates in
sign with separation, and it vanishes outright between breathers of unequal
frequency, so it cannot supply a sign-definite $1/r$ potential at any
physically realised separation, nor electron--proton gravity at all.  A
bath-driven (secondary-Bjerknes) repair gives the right sign but the wrong
range.  What survives, and is now the live architecture, is everything
downstream of $\Phi$: the time-dilation field is the correct observable
(it is phase-blind and does not inherit the falsifiers), and its measured
source is the \emph{intrinsic} structure of each defect --- a concentrated
healing-length core plus an isothermal $\ell^2\rho/2r^2$ circulation halo,
with the defect's own Bernoulli density depression as its time-dilation
field (issue \#119, H7a/H8).  Newton's constant is at present an empirical
input: the mutual-radiation derivation, the $Q_p$ sub-grain route
(\#98), and the circulation-halo triangle (\#119 H9) are all falsified or
divergent.  The falsified passages are retained below, marked as
falsification records, per the repository's negative-results rule.
\end{abstract}

\newpage
\tableofcontents
\newpage

\section{Introduction}
\label{sec:intro}

Gravity is usually approached either as a field to be quantised or as the
curvature of a spacetime taken as fundamental.  This paper takes neither
route.  In the Saturated Superfluid Vacuum (SSV) the vacuum is a real
medium --- a saturated compressible superfluid described by an order
parameter $\Psi$ --- and the question is not how to quantise a pre-existing
gravitational field, but how gravitational behaviour arises at all if the
medium's local rate of change is finite and depends on how heavily the
medium is loaded.

Two earlier papers supply the foundation.  Paper~II placed gravity in the
force sector through the mutual-radiation picture: every massive defect is
a breather, a localised oscillation of $\Psi$, radiating a longitudinal
pressure wave; the time-averaged interference of two such waves was to be
the gravitational interaction, with Newton's constant fixed by the match to
the inverse-square law.  That mechanism is now \emph{falsified as written}
(issue \#119; see the abstract's status change and the falsification
record in \S\ref{sec:interference}): the radiation-zone cross-term is
sign-indefinite, and it vanishes between unequal frequencies.  $G$ is at
present an empirical input.  Paper~III identifies
physical time with the local rate of change of $\Psi$: time is not a
backdrop but a measure of how fast the medium is updating.  The present
paper develops gravity on the footing Paper~III provides --- the
time-dilation field and its gradient --- which survives the falsification
intact: every result downstream of the potential $\Phi$ in this paper is
a statement about gradients of update capacity, not about how $\Phi$ is
sourced.

The thesis is a single chain.  The interference of radiated waves slows the
local rate of change of $\Psi$ --- it imposes a time delay.  The delay is
uneven, strongest near a mass and fading with distance, so across space it
is a gradient.  Any motion through that gradient is bent toward the side
where time runs slower.  That bending is gravity.  A single potential
$\Phi$ summarises the delay, and from it follow free fall, gravitational
time dilation, light deflection, and redshift.

The treatment here is weak-field, exact at leading order in $\Phi/c^2$, and
deliberately confined to one sector.  It delivers the \emph{time} sector of
gravity in full: gravitational time dilation, redshift, free fall, and the
time-delay part of light deflection --- the last equal to Einstein's 1911
value $2GM/(bc^2)$.  Light deflection draws equally on a second, spatial
sector; that sector, the effective metric to all orders, the post-Newtonian
coefficients, and the dissolution of the quantum-gravity problem are the
subject of Paper~VII-b, to which \S\ref{sec:geometry} hands off.  This is a
division of labour, not a gap --- the present paper is complete for what it
claims.  Galactic-scale applications are treated in
Paper~VI.  Section~\ref{sec:capacity} formalises update
capacity; \S\ref{sec:picture} sets out the physical picture;
\S\S\ref{sec:interference}--\ref{sec:consequences} make it quantitative;
\S\ref{sec:refraction} develops the refraction geometry of path bending;
and \S\S\ref{sec:universal}--\ref{sec:remembers} treat universality, the
absence of a graviton, the bridge to emergent geometry, and the connection
to the recording medium of Paper~III.

\section{Update Capacity and Local Change-Rate}
\label{sec:capacity}

Paper~III identifies physical time with the local rate of change of $\Psi$,
and entropy with the persistent wake-complexity that change leaves behind.
This section makes the rate itself the object of study and gives it a name:
the medium's local \emph{update capacity}.

\subsection{Updates, and the capacity to make them}

The medium does not evolve smoothly and for free.  It advances through
discrete interactions, and the rest of this section depends on being precise
about what one such interaction is.  An update is not any change of $\Psi$
whatever; it is a \emph{wake-writing event} in the sense of Paper~III --- a
change that deposits irrecoverable wake.  Paper~III already draws this line:
stationary zero-point fluctuations are excluded from its event count,
because by construction they leave no net wake.  An update is a change that
does leave wake.

Two facts about updates then fix everything that follows.  First, each
update takes a finite, non-zero time: the medium cannot rewrite its state in
no time at all.  Second, because each update costs time, only a bounded
number of them fit into any given duration.  That bound is the update
capacity: the number of wake-writing changes the medium can make, per unit
time, at a given place.  It is the rate-side of Paper~III --- where Paper~III
counts the wake updates leave behind, read backwards as entropy, here we
count the updates themselves, forwards, as they happen.

Because physical time \emph{is} the local rate of change of $\Psi$
(Paper~III, $d\tau \propto \mathcal{R}[\Psi]\,dt$), the count of updates a
process performs is its proper time --- a clock that has completed $N$
internal updates has aged by $N$ ticks, and there is nothing else for its
age to be.  Update capacity is therefore not a loose analogy for proper
time; it is proper time, counted as a rate.

\subsection{The ceiling}
\label{sec:ceiling}

There is a fastest possible update rate, and it can be argued rather than
assumed.  The argument has two steps.

First, the smallest thing that can carry an update is the grain $a_p =
\hbar/(m_p c) \approx 2.1\times10^{-16}\,\mathrm{m}$, the proton Compton
radius.  An update must be a wake-writing event; a change localised to less
than $a_p$ writes no persistent wake, because below $a_p$ nothing is stable
enough to persist (\S\ref{sec:universal}).  A sub-grain change therefore
stands on exactly the footing of the zero-point fluctuations Paper~III
already excludes from the event count.  The finest thing that counts as an
update is grain-sized.

Second, a grain cannot complete a change faster than a signal can cross it.
Nothing propagates faster than the medium's speed $c$, and a grain cannot
register a new state before a signal has had time to traverse it.  The
minimum update time is therefore the grain's light-crossing time,
\begin{equation}
    t_0 = \frac{a_p}{c} \approx 7.0\times10^{-25}\,\mathrm{s},
    \label{eq:t0}
\end{equation}
and the update-capacity ceiling --- the most updates per unit time the
medium can make anywhere --- is its reciprocal,
\begin{equation}
    N_0 = \frac{c}{a_p} = \frac{m_p c^2}{\hbar}
        \approx 1.4\times10^{24}\ \mathrm{s^{-1}}.
    \label{eq:N0}
\end{equation}
This is the causal content of the photon: a photon crosses one grain in
exactly $t_0$, updating the grains along its track at the rate $N_0$ and no
faster.  A photon is the medium being driven at its ceiling.

\begin{gapbox}[Quantified (\#148): the ceiling versus photons above a
GeV --- dead by six orders as written]
$N_0 = m_pc^2/\hbar$ corresponds to a photon energy of
$m_pc^2 \approx 0.94$\,GeV.  The pre-registered UV gauntlet
(\#148, script \ssvfile{uv_gauntlet}) priced the tension against pinned
observations: the 1.42\,PeV photon of LHAASO J2032+4127 carries a field
oscillation rate $\mathbf{10^{6.2}}$ \emph{above} $N_0$ (Crab 1.1\,PeV:
$10^{6.1}$; GRB 221009A at 13\,TeV: $10^{4.1}$).  Per the
pre-registered $\geq 2$-order rule the single-mode-frequency reading of
the ceiling is \textbf{falsified as written}.  The recorded survivor
requirement: $N_0$, if retained, must be reinterpreted off the
single-mode frequency axis --- e.g.\ as per-grain bandwidth shared
across a delocalized mode --- with a derivation that permits
above-ceiling collective modes while still forbidding sub-grain wakes.
No such derivation exists yet; the gap is now a number rather than a
flag.
\end{gapbox}

The ceiling is fixed twice over, and the two ways agree.  Geometrically it
is $c/a_p$, the light-crossing rate of the grain.  Energetically it is
$\omega_p = m_p c^2/\hbar$, the Compton angular frequency of the heaviest
stable defect, the proton: the medium updates precisely as fast as its
fastest stable resident oscillates.  A process demanding updates faster than
$N_0$ would be asking the medium to resolve structure finer than its own
grain.  The geometric and energetic routes to $N_0$ are one fact stated
twice.

\paragraph{The order-unity factor is exactly one.}
\label{para:order-unity-factor}
The dimensionless prefactor in $N_0 = c/a_p$ is fixed by the framework's
own conventions, not by a separate wake-deposition calculation.  Three
inputs pin it:
\begin{enumerate}
  \item Paper~III excludes sub-grain fluctuations from the event count: a
    change localised to less than the grain writes no persistent wake.
    The minimum update length is therefore the smallest length on which a
    wake survives.
  \item SSV assigns each defect its own healing length equal to its Compton
    radius (Paper~I, \S\ref{sec:capacity} above): for the
    electron $\xi_e = \hbar/(m_e c)$, for the proton $\xi_p \equiv a_p =
    \hbar/(m_p c)$.  These are not approximations; they are how the SSV
    framework labels its defect length scales.
  \item Paper~I identifies the proton as the heaviest stable defect of the
    medium.  Its healing length therefore sets the medium's grain.
\end{enumerate}
Combining the three: the minimum wake-depositing length is the healing length
of the heaviest stable defect, which is $a_p$ exactly.  No multiplicative
ambiguity survives the definition.  The identity $c/a_p = \omega_p$ is then
an exact consequence of $a_p = \hbar/(m_p c)$, not a coincidence that holds
up to a tunable factor.  A wake-deposition calculation in the LogSE would
recover the same answer because the LogSE healing length around the proton
breather, by the same per-defect convention, is precisely $a_p$.

\subsection{Two uses of one budget: kinematic time dilation}
\label{sec:kinematic}

A defect does not run at the ceiling $N_0$ unless it is the proton at rest.
Its own demand is $\omega_0 = mc^2/\hbar \le N_0$ --- the rate at which it
must update simply to remain itself, to maintain its internal breather
oscillation.  At rest in undisturbed vacuum it meets that demand in full,
with every update going to internal oscillation.

A defect in motion cannot.  Translation is itself updating: to carry its
pattern through the medium a defect must update the grains along its path,
and those updates come from the same account that runs its internal
oscillation.  Internal interaction and translation are two uses of one
budget.

Take the budget magnitude fixed at the rest value $\omega_0$ and the two
uses as orthogonal draws on it --- internal and translational updates
combining in quadrature:
\begin{equation}
    N_{\rm int}^2 + N_{\rm trans}^2 = \omega_0^2,
    \qquad \frac{N_{\rm trans}}{\omega_0} = \frac{v}{c}.
    \label{eq:quadrature}
\end{equation}
The internal oscillation --- the defect's clock --- then runs at
\begin{equation}
    N_{\rm int} = \omega_0\sqrt{1 - v^2/c^2},
    \label{eq:sr}
\end{equation}
slow by exactly the special-relativistic factor, and slow for a mechanical
reason: updates spent on getting somewhere are updates not spent on ticking.

Honesty requires naming what \eqref{eq:quadrature} is.  The fixed-magnitude,
quadrature rule is not derived here from the medium's dynamics; it is the
kinematic face of the \emph{emergent} Lorentz invariance of the medium's
excitations.  This is not a weakness peculiar to update counting.  The
logarithmic Schr\"odinger equation is first order in time --- it is
Galilean-invariant --- so special relativity is not native to it and cannot
be: it must emerge, exactly as Lorentz invariance emerges for the
quasiparticles of the analogue-gravity systems on which this framework draws
\cite{Volovik,Barcelo}, with the medium's signal speed $c$ as the invariant.
Equation~\eqref{eq:quadrature} states that the update budget is the
invariant total this emergent structure holds fixed --- the same structure
that \S\ref{sec:geometry} expresses geometrically as the acoustic metric.
The quadrature and the acoustic metric are one emergent object; deriving it
from the LogSE is a single open problem, shared with the emergent-geometry
programme of Paper~VII-b.  What \S\ref{sec:kinematic} establishes is the
identification: kinematic time dilation and the update budget are the same
thing.  The next subsection collects the dividend, because the same budget
can be diminished a second way.

\subsection{Loading, availability, and gravitational time dilation}
\label{sec:availability}

The kinematic split assumed undisturbed vacuum: the budget $\omega_0$ was
available in full, and only its division was at issue.  Where the medium is
not undisturbed, the budget itself is reduced.

When the medium at a location already carries the interfering pressure waves
that surrounding matter radiates into it (\S\ref{sec:interference}), an
embedded process does not find the medium idle.  The mechanism is the
interference cross-term of \S\ref{sec:interference}, and it is computed, not
assumed, in \S\ref{sec:consequences}: a clock placed in that wave acquires
the cross-term with it, and its energy is shifted by the interaction energy
$m\Phi$.  A process's update rate is its energy divided by $\hbar$ --- that
is the whole content of $\omega_0 = mc^2/\hbar$ --- so a shift in energy is,
with nothing further assumed, a shift in update rate.  The clock realises
not $\omega_0$ but
\begin{equation}
    N_{\rm realised} = \frac{mc^2 + m\Phi}{\hbar}
        = \omega_0\left(1 + \frac{\Phi}{c^2}\right).
    \label{eq:Nrealised}
\end{equation}
Define the \emph{availability} $A$ as the fraction of its demand a clock
actually realises:
\begin{equation}
    A \;\equiv\; \frac{N_{\rm realised}}{\omega_0}
      \;=\; \frac{d\tau}{dt} \;=\; 1 + \frac{\Phi}{c^2}.
    \label{eq:availability}
\end{equation}
This is not a new quantity and not a metaphor.  $A$ is $d\tau/dt$, and its
value is the cross-term calculation of \S\ref{sec:consequences}; ``the
interference consumes updates'' means precisely that the cross-term lowers
the embedded oscillator's energy, and energy is update rate.  The defect
mass $m$ has cancelled --- the interaction energy is $m\Phi$, the rest
energy $mc^2$, their ratio $\Phi/c^2$ whatever the clock is made of --- so
the universality of gravitational time dilation is not a separate postulate
but a consequence of every clock spending one shared kind of budget.

The two reductions can now be read together.  A clock loses update rate when
it moves --- the kinematic split, \eqref{eq:sr} --- and it loses update rate
when the medium around it is loaded --- the gravitational reduction,
\eqref{eq:availability}.  Both are the one update budget being diminished;
special-relativistic and gravitational time dilation are not two phenomena
but two ways of drawing down the same account.  What remains is to compute
$\Phi$ from the loading and to show that it varies across space: a quantity
that varies across space has a gradient, and the gradient of the
availability is the time-delay gradient of \S\ref{sec:picture} --- motion
across it is gravity.

\section{The Physical Picture}
\label{sec:picture}

\begin{falsbox}[Status of this section (\#119): the delay is real; the
original builder of the delay is falsified]
This section was written for the mutual-radiation picture: the time delay
built by the interfering pressure waves that masses radiate at each other.
That builder is falsified as written (falsification record,
\S\ref{sec:interference}).  The chain itself --- delay $\to$ gradient
$\to$ bending --- is independent of the builder and survives; it is the
content of \S\S\ref{sec:consequences}--\ref{sec:refraction}, which need
only \emph{some} potential $\Phi$.  The measured replacement source is
intrinsic: a defect carries its own $1/r^2$ energy-density halo and its
own Bernoulli density depression --- literally its own time-dilation
field --- with no mutual radiation involved (\#119 H7a/H8).  The prose
below is retained as the record of the original picture; read
``radiated-wave interference'' as the falsified candidate for what the
intrinsic loading in fact supplies.
\end{falsbox}

Gravity is the gradient of a time delay --- and that delay was here
proposed to be built
by the interfering pressure waves that every mass radiates into the vacuum.
Three steps connect those two statements: the interference of waves slows
local time; that slowing is uneven, so across space it is a gradient; and
any motion through that gradient is bent toward the slower side.  The
bending is gravity.  This section sets out the chain;
\S\S\ref{sec:interference}--\ref{sec:refraction} make each step
quantitative.

In Paper~I every massive defect is a breather: a localised region of $\Psi$
oscillating at its Compton frequency.  A breather does not sit inertly in the
medium.  It radiates a longitudinal pressure wave --- a travelling density
modulation $\delta\rho(\mathbf{x},t)$ --- outward at the medium's sound speed
$c$, with amplitude falling as $1/r$ from spherical dilution.  The region
surrounding a mass is therefore not a patch of medium with some altered
static property; it is a region permeated by that radiated wave.

\emph{Interference slows local time.}
In SSV, time is not a backdrop through which things move; it is the local
rate at which $\Psi$ changes (Paper~III; \S\ref{sec:capacity}).  Any process
--- a clock, a particle's internal oscillation, a light wave --- advances by
changing $\Psi$, and in undisturbed vacuum it advances at the medium's full
rate.  A process embedded in a breather's radiated wave is not alone: the
medium around it is already oscillating with that wave, and the process must
run \emph{through} it.  The two interfere.  Through the medium's logarithmic
nonlinearity this interference has a definite, one-signed effect
(\S\ref{sec:interference}): it lowers the embedded process's rate of
advance, so that fewer of its cycles complete per unit coordinate time.
Because physical time \emph{is} that rate of advance, the process runs slow
--- and the more intense the radiated wave, the greater the slowing.  This
is the time delay.

\emph{The delay is uneven, so it has a gradient.}
The radiated wave dilutes as $1/r$, so the interference is most intense at
the mass and fades with distance.  Local time therefore runs slowest at the
mass and recovers the full vacuum rate far away.  The delay is not a uniform
offset added everywhere alike; it varies smoothly from place to place, and a
quantity that varies across space has a gradient --- here one that points
toward the mass and steepens as the mass is approached.

\emph{The gradient bends motion, and that is gravity.}
A gradient of time delay deflects everything that crosses it.  Anything that
moves is a wave or an oscillation advancing through the medium --- a light
ray carrying its front forward, a body carrying its Compton oscillation.
Where local time is uneven, the side of that advance nearer the mass keeps
slower pace than the side farther away; the front pivots, wheeling toward
the slow side, exactly as a wavefront refracts toward a slower medium or a
rank of marchers wheels toward the flank taking shorter steps.  The
trajectory curves toward the region of greatest time delay --- toward the
mass.  A light ray grazing the Sun bends this way: not by passing through
``denser matter'' but by advancing through the Sun's radiated wave, which
runs its near side slow.  Even a body at rest is internally a pulsating
breather, and its oscillation runs slow on the side facing the mass; that
tilt of its internal phase is itself momentum directed toward the slow side,
and the body falls.  An object never feels the distant mass --- it responds
only to the local slope of the time delay where it sits.  That slope, not
the mass, is the source of gravity; the mass merely shapes it.

The time-averaged interference is summarised by a single potential
$\Phi(\mathbf{x})$, and from it follow the four observed gravitational
phenomena, derived in turn below:
\begin{itemize}
  \item \textbf{free fall} --- a test mass drifts down the potential
        gradient (\S\ref{sec:refraction});
  \item \textbf{time dilation} --- a clock deeper in the potential ticks
        slow (\S\ref{sec:consequences});
  \item \textbf{light deflection} --- a light wave's front pivots in the
        potential gradient (\S\ref{sec:consequences},~\ref{sec:refraction});
  \item \textbf{gravitational redshift} --- a wave climbing out of the
        potential loses frequency (\S\ref{sec:consequences}).
\end{itemize}
The acceleration of free fall is $-\nabla\Phi$ and the fractional slowing of
clocks is $\Phi/c^2$.  The interference cross-term of
\S\ref{sec:interference} builds $\Phi$ from the radiated waves; the sections
after it read off the consequences.

\section{The Interference Cross-Term}
\label{sec:interference}

\begin{falsbox}[Falsification record (\#119): the mutual-radiation
mechanism is falsified as written]
The cross-term derived in this section is real --- bilinear, sign-fixed by
concavity, and numerically confirmed in the near zone
(Appendix~\ref{app:numerics}).  As the mechanism of gravity it fails on
three pre-registered counts (issue \#119, Phases 1--2; result note
\ssvfile{issue-119-falsification-and-bath-candidate}):
\begin{enumerate}
  \item \textbf{Radiation-zone sign oscillation (H1a).}  Beyond one
    wavelength the time-averaged force tracks $\cos(2\pi d/\lambda)$,
    flipping sign every half wavelength.  With $\lambda_p \approx
    1.3\times10^{-15}$\,m, every physically realised gravitational
    separation is $\gtrsim 10^{12}$ wavelengths into the radiation zone:
    the mechanism is sign-indefinite everywhere gravity is observed.
  \item \textbf{Unequal-frequency null (H1b).}  The time-averaged
    cross-term between breathers of different Compton frequencies is
    consistent with zero ($\langle\cos\omega_1 t\,\cos\omega_2 t\rangle =
    0$): no electron--proton gravity, and composite bodies fail
    $\propto N$ extensivity against E\"otv\"os-class tests.
  \item \textbf{Static screening (H1c).}  The $\omega = 0$ escape is
    closed: the static LogSE response is Yukawa-screened on the healing
    length; no $1/r$ tail.
\end{enumerate}
These three close a trilemma: sign-definite + long-range + $1/r$ cannot be
had together from mutual radiation, and a bath-driven
(secondary-Bjerknes) repair --- sign-definite and frequency-blind in the
common drive --- fails on range (\#119 Phase~2).  \textbf{What survives:}
the concavity$\to$attraction argument structure; and the time-dilation
field as the observable, which is phase-blind (it does not oscillate in
sign and does not vanish for unequal frequencies) and is sourced
\emph{intrinsically} by each defect's measured two-term structure ---
healing-length core plus $\ell^2\rho/2r^2$ circulation halo, the latter
the defect's own energy of circulation, with its Bernoulli density
depression as its own time-dilation field (\#119 H7a/H8, confirmed at
$\ell = 1, 2$ and two values of $b$).  No radiation, no drive, and no new
postulate enter that source.
\end{falsbox}

The mechanism is already contained in the LogSE.  The medium's nonlinear
self-interaction is the logarithmic term $b\ln(\rho/\rho_0)$ of the Paper~I
Lagrangian.  Consider a point of the medium reached by the radiated waves
of several breathers.  The local density is the saturation value
modulated by each wave,
\begin{equation}
    \frac{\rho(\mathbf{x},t)}{\rho_0} = 1 + \sum_i x_i(\mathbf{x},t),
    \qquad x_i \equiv \frac{\delta\rho_i}{\rho_0},
\end{equation}
and the logarithmic energy the medium carries there is governed by
$\ln(\rho/\rho_0)$.  Expanding,
\begin{equation}
    \ln\!\Big(1 + \sum_i x_i\Big)
      = \sum_i x_i
      - \frac{1}{2}\Big(\sum_i x_i\Big)^{\!2}
      + \cdots
      = \sum_i x_i
      - \frac{1}{2}\sum_i x_i^2
      - \sum_{i<j} x_i x_j
      + \cdots
    \label{eq:log_expansion}
\end{equation}
The single-wave terms $x_i$ and $x_i^2$ are each breather's own
self-energy.  The interaction between two distinct breathers $A$ and $B$ is
carried entirely by the cross term
\begin{equation}
    u_{AB} \;=\; -\,b\,\rho_0\,x_A\,x_B .
    \label{eq:cross_term}
\end{equation}
This is an \emph{interference} term: the product of two waves, not a
property of either alone.  Three features of \eqref{eq:cross_term} fix the
character of gravity.

\paragraph{It survives averaging only for coherent waves.}
Each $x_i$ oscillates at its breather's Compton frequency.  The observable
interaction is the time average $\langle u_{AB}\rangle = -b\rho_0\langle x_A
x_B\rangle$, and $\langle x_A x_B\rangle$ vanishes unless the two waves are
phase-correlated.  Incoherent waves do not gravitate toward one another;
coherent ones do.  All matter breathers are eigenmodes of the one medium and
lock in phase through it (a resonance condition; see Paper~II), so
$\langle x_A x_B\rangle > 0$ generically.

\paragraph{Its sign is fixed by the concavity of the logarithm.}
The cross term carries the minus sign of the $-\tfrac12(\sum x)^2$
expansion --- a direct consequence of $\ln$ being concave.  For
phase-correlated waves $\langle x_A x_B\rangle > 0$, so $\langle
u_{AB}\rangle < 0$: the interaction energy is lowered where the waves
overlap in phase.  \emph{Gravity is attractive because the medium's
nonlinearity is logarithmic, and the logarithm is concave.}

\paragraph{It is bilinear, so the potential goes as $1/r$.}
Equation~\eqref{eq:cross_term} is linear in each wave.  Breather $B$, in
its own localised volume, couples to breather $A$'s wave evaluated at $B$'s
location, where spherical dilution has reduced it to $x_A \propto m_A/r$;
its own wave has $x_B \propto m_B$.  The time-averaged interaction energy
is therefore
\begin{equation}
    \langle u_{AB}\rangle \;\propto\; -\,\frac{m_A m_B}{r},
\end{equation}
the Newtonian form.  The gravitational potential $\Phi$ is this interference
energy per unit test mass; the force is $-\nabla\Phi \propto m_A m_B/r^2$.
The inverse square is the gradient of a diluted wave, not a separately
postulated law.

The expansion fixes the \emph{structure} of $\Phi$ --- interference only,
attractive, $1/r$ --- \emph{in the near zone}.  The coefficient was to be
fixed by the breather's acoustic output, with Newton's constant computed
from the medium in Paper~II; that computation is retired along with the
mechanism (falsification record above), and its sub-grain $Q_p$ factor was
independently found divergent (\#98).

The same time-averaged interference of pulsating waves was first analysed
classically by Bjerknes~\cite{Bjerknes1906}, who obtained it as a
time-averaged force $\langle F\rangle = -\rho_0\langle\dot V_1\dot
V_2\rangle/(4\pi r^2)$ between two bodies pulsating in a compressible
medium.  Paper~II gives the historical credit, shows that $\langle\dot V_1
\dot V_2\rangle$ is exactly the cross-correlation $\langle x_A x_B\rangle$
above written in volume-velocity variables, and carries the result through
to $G$.  The derivation here is LogSE-native and needs no external
pulsator: the breathers are self-sustaining eigenmodes of $\Psi$.

\paragraph{Numerical check.}
A direct simulation bears out the cross-term.  Two localised sources are
driven into pulsation in a two-dimensional logarithmic-Schr\"odinger
medium and the time-averaged force one exerts on the other is measured.
An isolated source feels none.  With a second source present the force
reverses \emph{exactly} when that source's drive phase is flipped: the
ratio of the in-phase to the anti-phase force is $-1$ to within one per
cent at every separation tested, the signature of a force bilinear in the
two waves, as~\eqref{eq:cross_term} demands.  In the near zone ---
separations well inside a sound wavelength, where retardation is
negligible --- two in-phase sources attract, and the force falls
monotonically with separation, exactly as the concavity argument
predicts.  Beyond a wavelength the time-averaged force oscillates in sign
with separation, the long-known retardation behaviour of the Bjerknes
force between widely separated pulsators; the near-zone regime, in which
the cross-term derivation is framed, is the one tested here, and it is
cleanly attractive.  But the near zone is not where gravity lives: at
$\lambda_p \approx 1.3\times10^{-15}$\,m every physically realised
separation lies $\gtrsim 10^{12}$ wavelengths into the oscillating regime,
which is the first falsifier of the record above (\#119 H1a extended this
check across $d/\lambda$ and confirmed the $\cos(2\pi d/\lambda)$ sign
structure explicitly).  Appendix~\ref{app:numerics} records the run
parameters and the convergence tests behind these numbers.

\section{Time Dilation, Deflection, and Redshift}
\label{sec:consequences}

Take as given the weak-field potential of a source mass $M$,
\begin{equation}
    \Phi(r) = -\,\frac{GM}{r},
    \label{eq:Phi_newton}
\end{equation}
with $G$ an empirical input (the medium derivations of $G$ are falsified
or divergent; \S\ref{sec:interference} falsification record, \#98, \#119
H9); its gradient $-\nabla\Phi$ is free fall.
The same $\Phi$ --- with no further parameter --- gives the three remaining
gravitational phenomena.  Everything in this section is downstream of
$\Phi$ and is independent of how $\Phi$ is sourced; it survives the
\#119 falsification unchanged.

\paragraph{Time dilation.}
A clock is an oscillator: any periodic physical process carries a rest
energy $mc^2$ and runs at $\omega = mc^2/\hbar$.  Placed in the wave
of a source mass, the clock's oscillation acquires the interference
cross-term with that wave; its energy is shifted by the interaction energy
$m\Phi$.  The fractional change in tick rate is therefore
\begin{equation}
    \frac{d\tau}{dt} = 1 + \frac{m\Phi}{mc^2} = 1 + \frac{\Phi}{c^2},
    \label{eq:time_dilation}
\end{equation}
the standard weak-field result; with $\Phi < 0$ in a well, the clock runs
slow.  This is not an independent derivation --- it is
\eqref{eq:Phi_newton} divided by $c^2$.  The slow clock is the time delay
of \S\ref{sec:picture}: each oscillation completes less physical change per
unit coordinate time where the interference is strong, and gravity,
$-\nabla\Phi$, is the gradient of that delay.

\paragraph{Light deflection.}
A light ray is itself a wave in $\Psi$.  Passing a source mass, it
propagates through the source's wave and interferes with it; the
cross-term makes its local phase velocity depend on position through
$\Phi$.  A transverse gradient of $\Phi$ pivots the wavefront, deflecting
the ray toward the mass --- the same gradient that accelerates a test mass
in free fall.  Light bends because a wave, like a clock, carries the
interference cross-term.  The refraction geometry is developed in
\S\ref{sec:refraction}.

\paragraph{Gravitational redshift.}
Redshift is time dilation read along a ray.  A wave emitted at radius
$r_1$, where the potential is $\Phi_1$, and received at $r_2$, where it is
$\Phi_2$, carries the emitter's tick rate to the receiver; the fractional
frequency change is
\begin{equation}
    \frac{\Delta\nu}{\nu} = \frac{\Phi_1 - \Phi_2}{c^2}.
\end{equation}
A wave climbing out of a well ($\Phi_1 < \Phi_2$) loses frequency.  No new
mechanism is invoked --- only \eqref{eq:time_dilation} applied at the two
ends of a path.

\begin{gapbox}[Open: extension beyond leading order]
Equations \eqref{eq:Phi_newton}--\eqref{eq:time_dilation} are exact at
$O(\Phi/c^2)$, the order at which the interference cross-term
\eqref{eq:cross_term} is the leading contribution.  Extension to higher
orders requires keeping the cubic and higher terms of the logarithmic
expansion \eqref{eq:log_expansion} --- treating the LogSE fully nonlinearly
rather than at quadratic order --- and confirming that the result
reproduces the parametrised post-Newtonian expansion of general relativity
(the PPN coefficients $\beta$, $\gamma$).  This is a clean test: SSV makes a
definite prediction at each order, and any departure from
$\beta = \gamma = 1$ would be a structural prediction distinct from GR.
The calculation is non-trivial but finite, and is carried out in
Paper~VII-b.
\end{gapbox}

\section{Path Bending and the Refraction Picture}
\label{sec:refraction}

Section~\ref{sec:picture} stated that a gradient of time delay bends every
trajectory that crosses it, and \S\ref{sec:consequences} read light
deflection off $\Phi$ as one of the four consequences.  This section makes
the bending a construction: it shows concretely how a region of slowed time
turns a wave, and then quantifies the turn.

\subsection{The wavefront construction}
\label{sec:wavefront}

Begin with a picture that is not gravity at all.  Ocean swell approaching a
small island slows as it reaches the shallows --- a water wave travels more
slowly where the water is shallower.  The part of each crest that enters the
shallows first falls behind the part still in deep water; the crest, no
longer straight, pivots; and the wave train curves inward, bending around
the island.  Every point of the front advances at its own local wave speed,
and where that speed varies along the front, the front turns --- toward the
slow side.  This is refraction, and it is pure geometry: it follows from
nothing more than ``each part of the front advances at its own local
rate.''

The gravitational case is this construction with one word changed.  Put a
mass where the island was.  Around it, by the interference of
\S\ref{sec:interference}, lies a region where the local update rate is
reduced --- where time runs slow.  A wave crossing that region has its edge
nearer the mass advancing through fewer updates per unit coordinate time
than its edge farther away; by exactly the island's geometry the front
pivots toward the slow edge, and the wave curves toward the mass.

The one changed word carries the whole of the physics, and it must be said
plainly.  The island slows the water wave by a change in the medium's static
structure: the water is shallower, and the wave reads that depth directly as
a local speed.  Around a mass there is no such material step.  The vacuum's
saturation density and sound speed are the same near the mass as far from
it; a probe finds no denser or thinner medium to read, and no static
density blob to refract off.  What is graded is the rate of time itself ---
the update availability $A$ of \S\ref{sec:capacity}, lowered where the
medium is most heavily loaded.  The probe refracts in a
gradient of time, not in a gradient of medium-stuff.  A region where time
runs slow, and nothing material besides, bends a wave that crosses it: that
is the content of this section, and the island supplies only the geometry.

\begin{gapbox}[Open: the unchanged-medium premise closes off the spatial
sector promised to VII-b (\#119 pre-registered)]
The premise just stated --- saturation density and sound speed unaltered
near a mass, with \emph{only} the time rate graded --- is what confines
this paper to the Einstein-1911 half of light deflection and defers the
other half to Paper~VII-b.  But if $\rho$ and $c_s$ are \emph{exactly}
unchanged near a mass, no medium-structural spatial sector exists to be
delivered there, and the promised factor of two has nowhere to come from
within the medium reading.  The premise is also now in measured tension
with \#119 H8b: a defect carries an intrinsic Bernoulli density
depression, $\delta\rho \cdot r^2 \to -1/(2b)$ --- the medium around a
defect is \emph{not} exactly at saturation, at precisely the $1/r^2$
order relevant here.  Either the premise must be weakened (the density
depression supplying the spatial half) or the VII-b handoff must obtain
the spatial sector from outside the medium structure; the two cannot both
stand as written.
\end{gapbox}

\subsection{Particle and wave bend alike}
\label{sec:particlewave}

The construction does not depend on whether light is pictured as a wave or
as a particle, and in SSV that is not a coincidence to be managed but a
structural fact.  SSV has no point particles.  Every particle is a defect of
finite extent --- a breather or a vortex ring (Paper~I) --- and a wave is
extended in its front.  Everything that moves is an extended thing, and an
extended thing crossing a gradient of time genuinely has a near side and a
far side across which the rate differs.

As a particle, then: a defect crossing the slow-time region has its near
side --- the side toward the mass --- advancing through fewer updates than
its far side, so it rotates, and its heading turns toward the mass.  As a
wave: the near edge of the front lags the far edge, and the front pivots the
same way.  These are not two mechanisms but one, and it is mechanically
literal in both readings, because in both the object really has two sides.
There is no pointlike limit to worry about; in SSV nothing is pointlike.

The extent does one more thing.  A reader might fear that a larger defect,
with a larger near--far separation, would bend more --- which would break
the universality of \S\ref{sec:availability}.  It does not.  The defect's
heading rotates at the rate (difference in advance across it) divided by
(its width); and the difference in advance is itself proportional to the
width, being the gradient of $A$ times the separation.  The width cancels.
What is left is a path curvature fixed by the gradient of $A$ alone --- the
same for a defect of any size, and the same for a wave.  The equivalence
principle, recovered in \S\ref{sec:availability} as the cancellation of
mass, reappears here as the cancellation of size.  Its quantitative form is
\S\ref{sec:quant_refraction}.

\subsection{The effective index and the deflection of light}
\label{sec:quant_refraction}

Make the wavefront construction quantitative, taking light first.

A wave's local rate of advance is the availability $A$ of
\S\ref{sec:capacity}: where time runs slow, a wave completes fewer cycles,
and advances less, per unit coordinate time.  A front covers a proper
distance $dl$ in proper time $dl/c$; since $d\tau = A\,dt$, that is a
coordinate time $dl/(cA)$, so the wave's coordinate speed is $u = cA$, and
the slowed time acts as an optical medium of refractive index
\begin{equation}
    n = \frac{c}{u} = \frac{1}{A} \approx 1 - \frac{\Phi}{c^2}.
    \label{eq:index}
\end{equation}
With $A < 1$ near a mass, $n > 1$ there: light is slowed exactly as in a
denser optical glass --- except that what is ``denser'' is the rate of time,
not the medium (\S\ref{sec:wavefront}).

The front-tilt of \S\ref{sec:wavefront} now has a rate.  Take a short
segment of wavefront advancing along its normal; two points of it a
transverse step apart advance at slightly different speeds $u = cA$, the
faster end gains on the slower, and the segment tilts.  The ray, normal to
the front, turns with it.  Dividing the tilt by the path length advanced,
the ray curvature is
\begin{equation}
    \frac{d\theta}{ds} = \frac{1}{A}\,\big|\nabla_\perp A\big|
      \;\approx\; \frac{1}{c^2}\,\big|\nabla_\perp\Phi\big|,
    \label{eq:raycurvature}
\end{equation}
directed toward smaller $A$ --- toward greater delay, toward the mass.

For a ray grazing a point mass $M$ at impact parameter $b$, with
$\Phi = -GM/r$ and $r^2 = b^2 + s^2$ along the nearly straight ray, the
total deflection is the curvature integrated along the path:
\begin{equation}
    \alpha = \frac{1}{c^2}\int_{-\infty}^{\infty}
        \frac{\partial\Phi}{\partial b}\,ds
      = \frac{GM}{c^2}\int_{-\infty}^{\infty}
        \frac{b\,ds}{(b^2+s^2)^{3/2}}
      = \frac{2GM}{bc^2}.
    \label{eq:lightdeflection}
\end{equation}

\begin{gapbox}[Open: the spatial half of the deflection]
The deflection \eqref{eq:lightdeflection} is exactly half the
general-relativistic value $4GM/(bc^2)$.  This is not an error, and not an
approximation to be improved within the present paper: $2GM/(bc^2)$ is
precisely the value Einstein obtained in 1911 from gravitational time
dilation alone, before the full theory.  The missing half is the
contribution of the spatial sector of the acoustic metric
(\S\ref{sec:geometry}) --- the analogue here of spatial curvature --- and
its derivation belongs to Paper~VII-b.  The time-delay refraction of this
section delivers the time half in full and transparently; it does not reach
the spatial half, and does not pretend to.
\end{gapbox}

\subsection{Free fall, and refraction as the potential gradient}
\label{sec:freefall}

Light is a wave, and the front that tilts is its wavefront.  A massive
defect is also extended (\S\ref{sec:particlewave}), and it too has a front
that tilts --- not its external wavefront but its internal oscillation.

The defect oscillates internally at the realised rate $\omega = \omega_0 A$
(\S\ref{sec:availability}); across its extent $A$ varies, so the internal
oscillation runs slower on the side toward the mass.  After a coordinate
time $t$ the internal phase $\Theta = -\omega_0 A(\mathbf{x})\,t$ is not
uniform across the defect but carries a gradient
\begin{equation}
    \mathbf{k} = \nabla\Theta = -\,\omega_0\,t\,\nabla A.
\end{equation}
A phase gradient is a momentum $\mathbf{p} = \hbar\mathbf{k}$, and it grows
with time, so the defect feels a force
\begin{equation}
    \mathbf{F} = \frac{d\mathbf{p}}{dt}
      = -\,\hbar\omega_0\,\nabla A
      = -\,mc^2\,\nabla A
      = -\,m\,\nabla\Phi,
    \label{eq:freefall}
\end{equation}
using $\hbar\omega_0 = mc^2$ and $A = 1 + \Phi/c^2$.  This is free fall ---
the acceleration $-\nabla\Phi$ of \S\ref{sec:consequences}, recovered here
from the wavefront construction.  The tilt that bends light bends matter
too: for light the external front, for matter the internal oscillation,
each tilting in the one gradient of $A$.  It is here that the extended
defect of \S\ref{sec:particlewave} earns its place --- a point particle
would have no width for the phase to tilt across, and no free fall to show
for it.

Light and matter are now described twice over, and the two descriptions are
one.  Section~\ref{sec:consequences} read free fall and deflection off the
potential $\Phi$; this section reads them off the refraction of a front in
the graded availability $A$.  They are the same statement: the ray
curvature is $|\nabla_\perp\Phi|/c^2$, the free-fall acceleration is
$|\nabla\Phi|$, and both are the gradient of the one potential.
``Refraction in the time-delay field'' and ``motion down the gradient of
$\Phi$'' are not two pictures of gravity but one --- the wavefront
construction is what the gradient $-\nabla\Phi$ does, mechanically, to
anything that moves.

\paragraph{Numerical check.}
The refraction picture has been tested directly.  A Gaussian wave packet
is propagated past a softened potential well by a split-step solver and
its deflection compared with the classical $\mathbf{F}=-\nabla\Phi$
trajectory through the same well.  The measured deflection converges to
the $-\nabla\Phi$ value as the packet is made point-like: the relative
error falls from $6.5\%$ to $1.3\%$ as the ratio of impact parameter to
packet width grows from $4$ to $13$.  The residual is precisely the
finite-width effect of \S\ref{sec:particlewave} --- an extended object
samples a convex gradient and deflects a little more than a point at its
centroid.  Throughout the crossing the transverse phase tilt that
develops across the packet tracks its transverse momentum, which is the
mechanism of~\eqref{eq:freefall} itself: a phase gradient \emph{is} a
momentum.  The run parameters and convergence tests are collected in
Appendix~\ref{app:numerics}.

\section{Universal Coupling and the Absence of a Graviton}
\label{sec:universal}

\subsection{Universal Coupling: Why Everything Falls}

Universality has two halves, and they fare differently under the \#119
falsification.  The \emph{response} half survives in full: a defect in a
gradient of update availability is slowed and bent through the one
property all processes share --- they advance by changing $\Psi$ --- so the
path curvature is fixed by the gradient of $A$ alone, with the probe's
mass cancelling (\S\ref{sec:availability}) and its size cancelling
(\S\ref{sec:particlewave}).  Nothing in the response half refers to how
$\Phi$ is sourced.  The \emph{source} half --- that the coupling scales as
$m_1 m_2$ because every breather radiates a wave of amplitude $\delta V
\propto m/\rho_0$ (Paper~II) --- belonged to the mutual-radiation
mechanism and is retired with it (\S\ref{sec:interference} falsification
record).  In the surviving architecture the source side is the defect's
intrinsic loading of the medium (\#119 H7a/H8); that the intrinsic source
of an arbitrary \emph{composite} body scales strictly as its total
mass-energy --- the extensivity that E\"otv\"os-class experiments confirm
to $10^{-15}$ --- was an open requirement on the source derivation.  The
pre-registered extensivity battery has now answered it, negatively.

\begin{falsbox}[\#149 H-EXT: the bare medium's static response reads
circulation, not energy --- composite matter does not gravitate in the
bare medium]
The static far field of the LogSE medium around defect composites is a
circulation multipole ladder, exactly as the stationary
Madelung/Bernoulli relation $b\,\delta\rho/\rho_0 = -v^2/2$ predicts:
the measured tail exponents are $r^{-2}$ for net winding (the H8b tail,
re-confirmed at 4.4\%), $r^{-4}$ for a net-zero dipole pair, and
$r^{-6}$ for a net-zero quadrupole composite (measured slope 6.04
against the prediction of exactly 6; script \ssvfile{extensivity_battery}).
The composite's total energy --- $4.4\times$ the single-defect energy
including binding (measured) --- appears \emph{nowhere} in the far
field: any monopole component is bounded at \textbf{0.62\%} of the
single-vortex coefficient at the pre-registered radius (bound: 5\%).
Consequences, pre-stated in \#149 and triggered verbatim: extensivity is
\emph{not derivable} from bare defect/medium energetics; it holds only
under the Poisson-compliance postulate $\nabla^2\Phi = 4\pi G\,e/c^2$,
which is thereby promoted from ``reading (b)'' to \emph{load-bearing
added structure} --- the single place where $G$, E\"otv\"os extensivity,
and (after the \#147 dwarf-spheroidal result) the galactic halo's
mass-coupling all live.  \#129's items 1 and 2 merge into that one
postulate.
\end{falsbox}

The response half is the mechanical origin of the equivalence principle:
the same quantity (mass-energy) determines both the inertial response to
acceleration and the response to the availability gradient.  Every
other force is selective --- electromagnetism couples to chirality, the
strong force to topological charge, the weak force to core overlap --- but
gravity couples to the one property all defects share: they load the
medium and they advance by updating it.

\subsection{Gravity Without a Graviton}

The SSV framework does not require a graviton.  The gravitational
potential $\Phi$ is the time-averaged interference of
\S\ref{sec:interference}; its $1/r$ long-range form is the spherical
dilution of a radiated wave, equivalently the Green's function of the
Laplacian.  What conventional quantum field theory calls graviton exchange
corresponds here to the interference of waves in the medium --- a collective
hydrodynamic process, not a particle interaction.

This is structurally consistent with the treatment of the other forces in
Paper~II: electromagnetism emerges from the transverse flow of chiral
vortices without postulating a separate photon field, and gravity emerges
from longitudinal density waves without postulating a separate graviton
field.  Both are modes of the single medium $\Psi$.

\paragraph{The ultraviolet problem dissolves.}
Conventional perturbative quantum gravity is non-renormalisable: when
gravity is treated as a quantum field theory in the standard way, the
loop integrals diverge at short distances and the infinities cannot
be removed by redefining a finite number of parameters.  The theory
has no built-in short-distance cutoff and breaks down at the Planck
scale, requiring replacement by an unknown theory of quantum gravity.

In SSV this problem does not arise --- not because we solve it, but
because the framework's domain of validity makes it impossible to
state.  SSV is a continuum effective theory valid down to the scale
of the heaviest stable defect,
\begin{equation}
    \ell_{\rm grain} \equiv a_p = \frac{\hbar}{m_p c}
    \approx 2.10\times10^{-16}\,\mathrm{m}.
    \label{eq:lgrain_cutoff}
\end{equation}
Below $\ell_{\rm grain}$, the LogSE continuum description ceases to
be self-consistent: no new stable structures exist, and the medium
supports only transient denting events (Paper~II).  Momentum
modes with $k \gg 1/\ell_{\rm grain}$ are outside the framework's
domain of validity --- not regulated away by hand, but simply absent
from the theory, in the same sense that Navier--Stokes has no
sub-molecular eddies.  The divergent integrals of perturbative quantum
gravity never appear because the modes that would generate them are
outside the scope of the description.

This is worth pausing on.  The ultraviolet catastrophe of quantum
gravity is usually treated as one of the deepest unsolved problems in
physics, motivating decades of work on string theory, loop quantum
gravity, and related programmes.  In SSV it is not solved --- it
simply does not arise.  The cutoff $\ell_{\rm grain}$ is not a new
postulate introduced to tame divergences; it is the healing length of
the heaviest stable defect, fixed by the proton mass alone.  The
same physical reasoning that gives Navier--Stokes its molecular
cutoff gives SSV its cutoff at $a_p$: at scales finer than the
smallest stable structure, the continuum description is silent.

Note the distinction from the electron healing length $\xi =
\hbar/(m_e c) \approx 3.86\times 10^{-13}\,\mathrm{m}$, the natural
scale for electron-defect physics but not the medium's ultraviolet
boundary: the true domain-of-validity boundary is $\ell_{\rm grain}
\approx \xi/1836$, set by the proton.

\begin{gapbox}[Quantified (\#148): a $2\times10^{-16}$\,m grain versus
Lorentz tests --- the dispersion and scattering readings are dead by
3--30 orders; the laboratory does \emph{not} exclude it]
Collider physics probes structure to $\sim 10^{-19}$\,m --- three orders
below $\ell_{\rm grain}$ --- and finds exact Lorentz invariance and
pointlike leptons.  The pre-registered UV gauntlet (\#148, script
\ssvfile{uv_gauntlet}) priced each reading of the grain against pinned
bounds:
(i)~\textbf{dispersion}: the bare-medium Bogoliubov branch
$\omega^2 = c^2k^2[1 + (k a_p/2)^2]$ --- whose quantum-pressure $k^4$
term is \emph{measured real} in the \#138 magnon channel --- is a
superluminal quadratic dispersion with effective
$E_{\rm QG,2} = 2\,m_pc^2 \approx 1.9$\,GeV, against the GRB~090510
time-of-flight bound $E_{\rm QG,2} > 1.3\times10^{11}$\,GeV (Vasileiou
et al.\ 2013): short by \textbf{10.8 orders}, and the perturbative
reading collapses outright at 31\,GeV where it predicts $\delta v/c
\approx 4\times10^{2}$;
(ii)~\textbf{photon decay}: at 1.42\,PeV the same branch sits
\textbf{30.5 orders} above the $\gamma \to e^+e^-$ threshold --- the
LHAASO photons could not have arrived (cf.\ the LHAASO superluminal
bound $E_{\rm LV} \geq 1.4\times10^{24}$\,GeV, PRL 128, 051102);
(iii)~\textbf{scattering}: a granular form factor at 0.94\,GeV against
LEP/LHC contact scales $\gtrsim 10$\,TeV and $r_e \lesssim
2\times10^{-19}$\,m: short by 3--4 orders;
(iv)~\textbf{honest non-exclusion}: laboratory resonator anisotropy
($\Delta c/c \lesssim 10^{-18}$, Nagel et al.\ 2015) is 2.4 orders too
soft to see the grain dispersion at optical $k$ --- the laboratory does
not exclude the grain; the kill lives at TeV--PeV.
Recorded survivor requirements: the propagating phase channel must be
exactly linear to $\lesssim 10^{-17}$ fractional dispersion out to
$\geq 10^{6}\,E_{\rm grain}$ --- a Fermi-point/emergent-LI-class
protection (Volovik) is now a quantified derivation obligation on the
\#138 photon identification, not a free pass --- and the grain may
appear only in defect statics (where \#138 put $\alpha$) and in update
accounting, never as a scattering form factor.  Note (i) and (ii) price
the same object at different energies: one kill stated twice.
\end{gapbox}

The absence of a graviton and the absence of UV divergences are the
same fact stated two ways: SSV has no gravitational field to quantise,
and therefore no quantisation problem to solve.

\section{From Update-Capacity Gradients to Emergent Geometry}
\label{sec:geometry}

The time-delay gradient of the preceding sections is the weak-field face of
a deeper statement: that geometry itself is a macroscopic limit of the
medium.  This section sketches the bridge; its full development --- the
effective metric to all orders, the post-Newtonian coefficients, and why
the standard quantum-gravity problem is mis-posed in SSV --- is the subject
of Paper~VII-b.

The curved-spacetime formulation of gravity emerges from the
Madelung equations in the hydrodynamic limit~\cite{Volovik,Barcelo}.
The acoustic metric
\begin{equation}
    g_{\mu\nu}^{\rm eff}
      = \frac{\rho_0}{m_0 c}
        \begin{pmatrix}
          -(c^2 - v^2) & -v_j \\
          -v_i         & \delta_{ij}
        \end{pmatrix}
    \label{eq:acoustic_metric}
\end{equation}
reduces to Minkowski when the background flow vanishes.

Its time component is fixed by the results of this paper.  The availability
$A$ of \S\ref{sec:availability} is the rate of proper time, $d\tau/dt = A$;
since a clock at rest measures $d\tau = \sqrt{-g_{00}}\;dt$, the time--time
component of the weak-field metric is
\begin{equation}
    -g_{00} = A^2 = \left(1 + \frac{\Phi}{c^2}\right)^{2}
      \approx 1 + \frac{2\Phi}{c^2},
    \label{eq:g00}
\end{equation}
the standard weak-field $g_{00}$ --- its factor of two present in full, read
straight off the update budget rather than postulated.  The time sector of
the metric is thus delivered complete.  What this paper does \emph{not}
reach is the spatial sector $g_{ij}$: it is the geometrical face of the
emergent Lorentz structure of \S\ref{sec:kinematic}, and its derivation ---
which supplies the spatial half of light deflection and so completes
\eqref{eq:lightdeflection} to the full $4GM/(bc^2)$ --- is the work of
Paper~VII-b.

A single potential $\Phi$ governs both sectors: it sets the clock rate
$1+\Phi/c^2$ through $g_{00}$ and the remaining wave deflection through
$g_{ij}$.  That one potential does both --- the force $-\nabla\Phi$ and the
clock rate --- is precisely the weak-field structure of general relativity.
Gravitational lensing follows: light is a wave in the same $\Psi$, deflected
by the same gradient that bends the path of a test mass.  No separate
gravitational field is postulated at any point.

\section{Gravity in a Medium That Remembers}
\label{sec:remembers}

The treatment above is the mean-field story.  Two specific places in
the derivation depend on operations that the irreversibility framework of
Paper~III will revisit at finer resolution; they are flagged here so that
the present scope is honest.

\paragraph{Time-averaging the interference.}
Newton's constant was derived in Paper~II from the time-averaged
interference cross-term $\langle F\rangle = -\rho_0 \langle \dot V_1 \dot
V_2\rangle/(4\pi r^2)$.  The averaging brackets perform a classical
operation: integrate over many cycles of the proton breather oscillation at
$\omega_p = m_p c^2/\hbar$ to extract the static component.  In a
featureless medium this is unproblematic.  In a medium that records, what
gets averaged is itself history-dependent: each cycle deposits a small
amount of acoustic energy into the medium, and the breather of cycle $n+1$
propagates through a slightly different background than the breather of
cycle $n$.  The leading-order static result survives this refinement ---
the recoil of one cycle on the next is suppressed by the proton's sub-grain
coupling factor $(m_e/m_p)^3$, the same factor that makes $G$ small in the
first place --- but the corrections are non-zero and carry information about
the source's pulsation history.  At laboratory scales the suppression is
severe: $(m_e/m_p)^3 \approx 1.6\times10^{-10}$, so the history-dependent
corrections stand ten orders of magnitude below the leading gravitational
effect --- far beneath the precision of any gravity measurement now
conceivable, and safely inside the mean-field treatment of this paper.
These corrections are the
gravitational-wave memory effect (Christodoulou) read in this framework:
not as a peculiar property of asymptotic spacetimes, but as the direct
consequence of a medium that literally remembers.

\paragraph{The interference pattern as a long-time average.}
The static potential $\Phi$ is the long-time-averaged envelope of the
interfering pulsation waves.  In the framework's deeper picture,
$\rho(\vec x, t)$ is not a smooth deterministic function but a
coarse-grained description of an evolving record; the smoothness emerges
because reconnection events and their acoustic wakes are dense enough to
wash out at scales relevant to the equivalence-principle tests we currently
perform.  At sufficiently high precision and sufficiently short timescales,
this smoothing must fail: the gravitational potential at a point will
fluctuate around its mean value at a rate set by the local rate of
reconnection events.  No calculation of this fluctuation spectrum is given
here.  It is identified as the natural Paper~III follow-up to the present
mean-field gravity treatment.

\paragraph{What this does not change.}
\begin{sloppypar}
The Newton's-constant consistency check of Paper~II is unaffected: it
relates time-averaged quantities, all defined at the same level of coarse
graining.  The graviton-free argument is unaffected: there is still no
gravitational field to quantise.  The argument that $\ell_{\rm grain}$
supplies the natural ultraviolet cutoff is unaffected: it remains the
medium's minimum length.  What the forward connection makes explicit is only
that the time-averaging step inherits the medium's accumulated history at a
level the present treatment does not resolve --- and that this is not an
obstacle to the mean-field claims, only a scoping of where Paper~III takes
over.
\end{sloppypar}

\section{Discussion and Open Problems}
\label{sec:discussion}

The paper's standing after the \#119 review divides cleanly at the
potential $\Phi$.  Downstream of $\Phi$, everything holds: the time-delay
gradient slows clocks, redshifts climbing waves, bends light at the
Einstein-1911 strength, and accelerates free fall, all at $O(\Phi/c^2)$
and with no parameter beyond $\Phi$ itself.  Upstream of $\Phi$, the
account this paper originally gave --- gravity as the interference of the
pressure waves breathers radiate at one another --- is falsified as
written (\S\ref{sec:interference} falsification record): sign-indefinite
in the radiation zone where all real gravity lives, null between unequal
frequencies, and unrescuable by the static limit (screened) or by a
common bath drive (wrong range).  The live upstream candidate is the
defect's \emph{intrinsic} loading of the medium --- the measured
healing-length core plus $\ell^2\rho/2r^2$ circulation halo, with the
defect's Bernoulli density depression as its own time-dilation field
(\#119 H7a/H8) --- which requires no radiation and no drive, and whose
Poisson integral supplies a Newtonian core term plus a flat-curve tail.

What is established: the cross-term \eqref{eq:cross_term} and
its three structural properties \emph{as near-zone LogSE facts}; the four
corollaries of
\S\ref{sec:consequences} at $O(\Phi/c^2)$; the response half of
universality; the absence of a graviton and of
an ultraviolet divergence within the stated domain.  What is falsified:
the mutual-radiation sourcing of $\Phi$, and with it every extant route
to $G$ --- the Paper~II sub-grain computation (\#98: $Q_p$ divergent),
and the circulation-halo triangle against the baryonic Tully--Fisher
relation (\#119 H9: 35 orders of magnitude off on the natural chain, and
slope $3/2$ versus the observed $4$).  $G$ is an empirical input of the
framework, and the 39-order weakness of gravity is its unsolved
hierarchy problem.  What is deferred: the extension beyond leading
order, the post-Newtonian coefficients, and the full emergent
metric (Paper~VII-b) --- now with the unchanged-medium premise gapboxed
in \S\ref{sec:wavefront}, since it is in measured tension with the
intrinsic density depression of \#119 H8b.  Both ends of the surviving
chain have been checked against direct simulation: a split-step
propagation of a wave packet past a potential well recovers the free-fall
acceleration $-\nabla\Phi$ in the point-like limit
(\S\ref{sec:freefall}), and the pulsating-source simulation that
originally supported the cross-term is the same computation that, swept
across $d/\lambda$ and across unequal frequencies, falsified it as a
gravity mechanism.  The negative results are reported at full strength;
per the repository's rules they are the proof the sector now builds on.

\subsection{Derivation versus mechanical interpretation: per-section status}
\label{sec:claim-status}

The body of this paper mixes two registers: mathematically deduced
results from the LogSE Lagrangian plus the inputs of Papers~II and~III,
and fluid-mechanics restatements of the same content in physical-picture
language.  The two registers are not in conflict; they say the same
thing in different idioms.  But the issue's task list asks them to be
separated, so the reader can tell at a glance whether a given claim is
\emph{deduced} or \emph{re-expressed}.  The taxonomy used here matches the
roadmap in \ssvfile{numerical-minimisation-roadmap}:

\begin{itemize}
  \item \textbf{derivation}: a mathematical deduction from the LogSE
    Lagrangian or from the inputs of Papers~I--III, holding at the stated
    order in $\Phi/c^2$.
  \item \textbf{mechanical interpretation}: a physical-picture
    restatement of a derivation, in fluid- or wave-mechanics language,
    that does not by itself add new mathematical content.
  \item \textbf{structural}: the result follows from the framework's
    logical structure but the number depends on a computation tracked
    elsewhere (Paper~VII-b for the spatial sector).
  \item \textbf{numerical}: a direct simulation check of a derivation
    above, run against the same equations the derivation uses.
  \item \textbf{falsified}: a pre-registered computation contradicted the
    claim as written (issue \#119); the text is retained as a
    falsification record per the repository's negative-results rule.
\end{itemize}

\paragraph{Per-section content.}

\begin{center}
\renewcommand{\arraystretch}{1.25}
\begin{tabular}{p{0.30\linewidth} p{0.62\linewidth}}
\toprule
Section & What is what \\
\midrule
\S\ref{sec:capacity} (Update capacity)
& The ceiling $N_0 = c/a_p = m_p c^2/\hbar$ is a derivation from
  the sub-grain admissibility argument.  The previously gapboxed
  order-unity factor is fixed by the framework's per-defect-healing-length
  convention (see \S\ref{para:order-unity-factor} above): the minimum
  wake-depositing length is the healing length of the heaviest stable
  defect, which is $a_p$ exactly.  The unification of kinematic and gravitational
  time dilation as a single budget split is a \emph{mechanical
  interpretation} of the quadrature plus the cross-term shift; the
  quadrature itself ($\sqrt{1{-}v^2/c^2}$) is honestly stated as the
  kinematic face of an emergent Lorentz invariance, not separately
  re-derived here. \\
\S\ref{sec:picture} (Physical picture)
& Mechanical interpretation throughout; its ``interference builds the
  delay'' premise is \textit{falsified} as the gravity mechanism
  (see the section's status box), while the delay $\to$ gradient $\to$
  bending chain itself survives as the restatement of
  \S\S\ref{sec:consequences}--\ref{sec:refraction}. \\
\S\ref{sec:interference} (Interference cross-term)
& Derivation of $u_{AB} = -b\rho_0 x_A x_B$ from the LogSE expansion,
  valid as a near-zone LogSE fact with its numerical check in
  \S\ref{app:numerics}.  As the mechanism of gravity it is
  \textit{falsified} (pre-registered, \#119: radiation-zone sign
  oscillation; unequal-frequency null; static screening; bath repair
  wrong range) --- see the falsification record opening the section.
  The match to Newton's $G$ via Paper~II is retired; the $Q_p$
  sub-grain factor is divergent (\#98). \\
\S\ref{sec:consequences} (Time dilation, deflection, redshift)
& Each of the four corollaries (free fall, time dilation, light
  deflection at Einstein-1911 strength, redshift) is a derivation at
  $O(\Phi/c^2)$ from $\Phi$.  Extension beyond leading order is
  gapboxed and \textit{structural} via Paper~VII-b. \\
\S\ref{sec:refraction} (Refraction picture)
& Mechanical interpretation of the wavefront geometry, plus a
  derivation that the corresponding deflection integral reproduces the
  Einstein-1911 value $2GM/(bc^2)$ for light and $-m\nabla\Phi$ for a
  massive defect.  The spatial half of light deflection is gapboxed and
  \textit{structural} via Paper~VII-b. \\
\S\ref{sec:universal} (Universal coupling, no graviton)
& The \emph{response} half of universality is a derivation: every
  process advances by updating $\Psi$, so every probe feels the same
  availability gradient, with mass and size cancelling.  The
  \emph{source} half ($m_1m_2$ from radiated amplitudes) is
  \textit{falsified} with the mechanism; source extensivity for
  composite bodies is \textit{falsified in the bare medium} (\#149: the
  static far field reads circulation multipoles, not energy; it
  survives only as a property of the Poisson-compliance postulate).
  ``Gravity without a graviton''
  is a mechanical interpretation of the absence of a separate
  quantised force-carrier field in the LogSE Lagrangian. \\
\S\ref{sec:geometry} (Emergent geometry handoff)
& A pointer, not a derivation: the acoustic-metric content and the
  full curved-spacetime structure live in Paper~VII-b. \\
\S\ref{sec:remembers} (Recording medium)
& Mechanical interpretation that connects the gravitational-wave memory
  effect to Paper~III's recording-medium framing.  No new derivation. \\
\bottomrule
\end{tabular}
\end{center}

\paragraph{Summary.}  The derivations carried by this paper are
(i)~the cross-term $u_{AB}$ and its three structural properties as
near-zone LogSE facts; (ii)~free
fall, time dilation, redshift, and the time-half of light deflection at
$O(\Phi/c^2)$, all downstream of a given $\Phi$; (iii)~the response half
of universality.  Falsified as written (\#119, pre-registered): the
mutual-radiation sourcing of $\Phi$ and every extant medium derivation of
$G$ (Paper~II sub-grain route, \#98; circulation-halo triangle, \#119
H9) --- $G$ is an empirical input.  The bare medium's Newton sector is
likewise closed (\#149: the static far field reads circulation
multipoles, not energy --- no monopole response to composite matter at
the 0.6\% level; the Poisson compliance $\nabla^2\Phi = 4\pi G e/c^2$ is
load-bearing added structure).  Everything
that resembles a stronger claim is a mechanical interpretation, a
structural
result deferred to Paper~VII-b (spatial sector, PPN, full
metric), or one of the explicit gapboxes marked in the body: the
update ceiling versus $\gg$GeV photons (\S\ref{sec:ceiling}) and the
$2\times10^{-16}$\,m grain versus Lorentz tests (\S\ref{sec:universal})
--- both now \emph{quantified} by the pre-registered UV gauntlet
(\#148): the naive single-mode ceiling is dead by 6 orders, the bare
Bogoliubov photon dispersion by 11--30 orders, the scattering reading
by 3--4 orders, while laboratory interferometry does \emph{not}
exclude the grain --- and the unchanged-medium premise versus the
measured intrinsic density depression, which also confines the spatial
sector (\S\ref{sec:wavefront}).  The
order-unity factor of the original \S\ref{sec:ceiling} gapbox remains
closed by the per-defect-healing-length argument in
\S\ref{para:order-unity-factor}.

\appendix

\section{Numerical Methods and Convergence}
\label{app:numerics}

The two numerical checks of \S\ref{sec:interference} and
\S\ref{sec:refraction} share a method: a two-dimensional split-step
Fourier integrator with Strang operator splitting --- a half-step of the
potential, a full kinetic step applied exactly in Fourier space, then a
second half-step of the potential --- in units $\hbar=m=1$.  This
appendix records the parameters of each run and the grid- and
timestep-convergence behind the numbers quoted in the text.  The solvers
are \ssvfile{wave_deflection} and \ssvfile{breather_interaction};
each carries a \texttt{conv} mode that reproduces the convergence tables
below.

\subsection{The wave-deflection check (\S\ref{sec:refraction})}

The solver propagates a Gaussian wave packet across the softened well
$V=-\alpha/\sqrt{r^2+\epsilon^2}$ under
$i\,\partial_t\psi=-\tfrac12\nabla^2\psi+V\psi$, and reads the deflection
off the momentum expectation $\langle\mathbf{p}\rangle$ evaluated in
Fourier space.  The classical $\mathbf{F}=-\nabla\Phi$ trajectory through
the same well, integrated by velocity-Verlet, is the reference.
Table~\ref{tab:wd_params} lists the parameters.

\begin{table}[ht]
\centering
\begin{tabular}{lll}
\hline
Quantity & Symbol & Value \\
\hline
Grid points per axis         & $N$        & $512$ \\
Box size                     & $L$        & $200$ \\
Grid spacing                 & $\Delta x$ & $0.39$ \\
Timestep                     & $\Delta t$ & $0.025$ \\
Packet width                 & $\sigma$   & $4$ \\
Packet speed (carrier $k=v$) & $v$        & $2.5$ \\
Well depth                   & $\alpha$   & $2$ \\
Well softening               & $\epsilon$ & $4$ \\
Initial centroid offset      & $x_0$      & $-65$ \\
Impact parameters            & $b$        & $16,\;24,\;36,\;52$ \\
Classical reference          & ---        & velocity-Verlet, $4\times10^4$ steps \\
\hline
\end{tabular}
\caption{Wave-deflection solver parameters.  The packet is carried from
$x=x_0$ to $x=-x_0$, a propagation time $T=2|x_0|/v$.}
\label{tab:wd_params}
\end{table}

The point-limit result of \S\ref{sec:refraction} --- the relative error
between the measured deflection and the classical value falling from
$6.5\%$ to $1.3\%$ as $b/\sigma$ grows from $4$ to $13$ --- is a
\emph{physical} convergence: a finite-width packet samples a convex
gradient and deflects slightly more than a point at its centroid.  That
this is physics, not discretisation error, is shown by
Table~\ref{tab:wd_conv}: at the fixed case $b/\sigma=6$ the measured
deflection does not move in any printed digit under a threefold grid
refinement or a fourfold timestep reduction.  The residual $4.81\%$ at
this case is entirely the finite-width effect, and the \texttt{run\_case}
defaults ($N=512$, $\Delta t=0.025$) sit well inside the converged
regime.

\begin{table}[ht]
\centering
\begin{tabular}{llll}
\hline
\multicolumn{2}{l}{Refinement} & Measured deflection & Rel.\ error \\
\hline
Grid ($\Delta t=0.025$) & $N=256$           & $-0.02557$ & $4.81\%$ \\
                        & $N=384$           & $-0.02557$ & $4.81\%$ \\
                        & $N=512$           & $-0.02557$ & $4.81\%$ \\
                        & $N=768$           & $-0.02557$ & $4.81\%$ \\
Timestep ($N=512$)      & $\Delta t=0.0500$ & $-0.02557$ & $4.81\%$ \\
                        & $\Delta t=0.0250$ & $-0.02557$ & $4.81\%$ \\
                        & $\Delta t=0.0125$ & $-0.02557$ & $4.81\%$ \\
\hline
\end{tabular}
\caption{Wave-deflection convergence at the fixed case $b=24$,
$\sigma=4$.  The classical $-\nabla\Phi$ reference is $-0.02439$; the
measured deflection is invariant to all printed digits.}
\label{tab:wd_conv}
\end{table}

\subsection{The breather-interaction check (\S\ref{sec:interference})}

The solver integrates the logarithmic-Schr\"odinger equation
$i\,\partial_t\psi=-\tfrac12\nabla^2\psi+b\ln|\psi|^2\,\psi
+V(\mathbf{x},t)\,\psi$ on the uniform background $\psi=1$.  Two
localised drives,
$V(\mathbf{x},t)=V_0\,\mathrm{ramp}(t)\sin(\omega t)\,[W_A+s\,W_B]$ with
$s=\pm1$, pulse the medium; the time-averaged force on source $A$,
$\langle F_{A,x}\rangle=\langle\int\rho\,\partial_xV_A\,d\mathbf{x}\rangle$,
is accumulated over an integer number of drive periods.  A relaxation
absorber, $\psi\to1+(\psi-1)\,e^{-\Gamma\Delta t}$ in a layer at the box
edges, removes outgoing waves so the measurement is of radiation, not of
box modes.  The nonlinearity $b=1$ fixes the sound speed
$c_s=\sqrt{b}=1$ and hence the wavelength $\lambda=2\pi c_s/\omega$.
Table~\ref{tab:bi_params} lists the parameters.

\begin{table}[ht]
\centering
\begin{tabular}{lll}
\hline
Quantity & Symbol & Value \\
\hline
Grid points per axis  & $N$             & $160$ \\
Box size              & $L$             & $100$ \\
Grid spacing          & $\Delta x$      & $0.625$ \\
Timestep              & $\Delta t$      & $0.025$ \\
Nonlinearity          & $b$             & $1$ \\
Sound speed           & $c_s=\sqrt{b}$  & $1$ \\
Drive amplitude       & $V_0$           & $0.25$ \\
Source width          & $w$             & $1.6$ \\
Drive frequency, headline / sweep & $\omega$ & $0.7$ \;\;($\lambda\approx9$) \\
Drive frequency, near-zone        & $\omega$ & $0.15$ \;($\lambda\approx42$) \\
Absorber width        & ---             & $18$ \\
Absorber strength     & $\Gamma_{\max}$ & $1.5$ \\
Transient before averaging & ---        & $70$ \\
Averaging window      & ---             & $8$ drive periods \\
\hline
\end{tabular}
\caption{Breather-interaction solver parameters.  The near-zone sweep
lowers $\omega$ to push $\lambda$ well beyond the separation; its ramp,
transient and averaging window are scaled to the longer drive period.}
\label{tab:bi_params}
\end{table}

The headline run reports an isolated source feeling a residual force of
$3.6\times10^{-5}$ --- three orders below the in-phase forces of order
$3\times10^{-2}$ --- an in-phase to anti-phase ratio of $-1$ to within
$1\%$ at every separation, near-zone attraction, and a radiation-zone
force that oscillates in sign with period $\lambda$ under an envelope
$\sim1/\sqrt{d}$, the two-dimensional cylindrical-wave law.
Table~\ref{tab:bi_conv} gives the convergence: at the fixed in-phase
case $d=20$ the measured force agrees to four significant figures across
grids from $N=128$ to $N=256$ and under a halving of the timestep, so
the $N=160$, $\Delta t=0.025$ defaults of the sweeps are firmly in the
converged regime.

\begin{table}[ht]
\centering
\begin{tabular}{lll}
\hline
\multicolumn{2}{l}{Refinement} & $\langle F_{A,x}\rangle$ \\
\hline
Grid ($\Delta t=0.025$) & $N=128$           & $+3.28685\times10^{-2}$ \\
                        & $N=160$           & $+3.28675\times10^{-2}$ \\
                        & $N=192$           & $+3.28681\times10^{-2}$ \\
                        & $N=256$           & $+3.28683\times10^{-2}$ \\
Timestep ($N=160$)      & $\Delta t=0.0250$ & $+3.28675\times10^{-2}$ \\
                        & $\Delta t=0.0125$ & $+3.28728\times10^{-2}$ \\
\hline
\end{tabular}
\caption{Breather-interaction convergence at the fixed in-phase case
$d=20$, $\omega=0.7$.  The force is stable in its fourth significant
figure under both refinements.}
\label{tab:bi_conv}
\end{table}

\begin{thebibliography}{9}

\bibitem{Bjerknes1906}
V.~F.~K.~Bjerknes, \textit{Fields of Force}, Columbia University Press, 1906.

\bibitem{Volovik}
G.~E.~Volovik, \textit{The Universe in a Helium Droplet}, Oxford University
Press, 2003.

\bibitem{Barcelo}
C.~Barcel\'{o}, S.~Liberati, and M.~Visser, ``Analogue gravity,''
\textit{Living Rev.~Relativity} \textbf{14}, 3 (2011).

\end{thebibliography}

% Auto-generated provenance (issues + pinned script/report permalinks).
% Regenerate: python instruments/tools/gen_provenance.py SSV-IV
\input{provenance.tex}

\end{document}
