\documentclass[11pt,a4paper]{article}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{physics}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{enumitem}

\newenvironment{resultbox}[1][Result]{%
  \begin{quote}\noindent\textbf{#1.}\ }{\end{quote}}
\newenvironment{gapbox}[1][Open problem]{%
  \begin{quote}\noindent\textbf{#1.}\ }{\end{quote}}

\newcommand{\half}{\tfrac{1}{2}}
\newcommand{\mzero}{m_0}
\newcommand{\seriesname}{Saturated Superfluid Vacuum}
\newcommand{\papernumber}{Goldstone}

\title{\textbf{\seriesname~\papernumber}\\[0.4em]
The Photon as a Goldstone Mode and the Chiral-Shear Near Field}
\author{Stig Norland \\ Independent Researcher \\ Bergen, Norway}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
Paper~II of the Saturated Superfluid Vacuum (SSV) series identified a
tension in the electromagnetic sector.  The static Coulomb and magnetic
near-field mechanisms arise naturally from chiral-shear flow gradients
around toroidal charged vortices, with coupling strength set by the
dimensionless parameter $\alpha$.  But the same preliminary derivation
identified the propagating photon with the chiral-shear mode at
$c_\perp$, producing a speed mismatch: observed light in vacuum propagates
at $c$, not at $c_\perp$.

This paper isolates the correction required by that audit.  The observed
photon should be identified with the Goldstone/Bogoliubov phase mode of
the saturated order parameter $\Psi$, propagating at the medium sound speed
$c$.  The chiral-shear sector remains physically real, but it is reassigned
to the static and near-field electromagnetic structure: Coulomb pressure,
magnetic circulation, Aharonov--Bohm phase, and vortex-ring self-energy.

The central task is therefore a two-channel derivation.  One must derive
the radiative Maxwell sector from the Goldstone phase channel at speed $c$,
while preserving the static Coulomb and magnetic results of Paper~II as the
near-field limit of chiral-shear topology.  We do not claim this
derivation is complete here.  We state the consistency conditions,
identify the matching problem between the two channels, and define the
success criterion for a future full calculation.
\end{abstract}

\newpage
\tableofcontents
\newpage

% ---------------------------------------------------------------------
\section{Introduction}
\label{sec:introduction}
% ---------------------------------------------------------------------

The electromagnetic sector of SSV has split into two logically distinct
pieces:
\begin{enumerate}[label=(\roman*),leftmargin=2em]
    \item a \emph{static near-field sector}, where charge, Coulomb force,
          magnetic circulation, and AB phase arise from the
          topology and chiral-shear flow of toroidal vortices;
    \item a \emph{radiative sector}, where the observed photon must
          propagate at the measured vacuum speed $c$.
\end{enumerate}
Paper~II developed the first piece but used the wrong carrier for the
second.  It treated the chiral-shear mode as the photon and assigned it the
speed $c_\perp$.  This cannot be the final physical interpretation because
the observed photon propagates at $c$.

The correction is suggested by analogue-gravity and logarithmic-superfluid
vacuum theory: the physical photon is the Goldstone phase mode of
$\Psi$~\cite{Volovik2003,Zloshchastiev2020}.  The phase mode propagates at
the medium sound speed,
\begin{equation}
    c_s(\rho_0)=c,
    \label{eq:cs_c}
\end{equation}
and is therefore compatible with observed light propagation and with
gravitational-wave constraints such as GW170817.

The question is not simply whether one can rename the photon.  The hard
question is whether the successful static electromagnetic mechanisms of
Paper~II survive the replacement:
\[
    \hbox{chiral-shear photon at }c_\perp
    \quad\longrightarrow\quad
    \hbox{Goldstone photon at }c .
\]
The present paper formulates that problem.

\begin{resultbox}[Claim of this paper]
The observed photon should be the Goldstone/Bogoliubov phase mode of the
SSV order parameter, propagating at $c$.  The chiral-shear field should be
retained as the static near-field sector that fixes Coulomb strength and
magnetic topology.  The open calculation is the matching between these two
sectors.
\end{resultbox}

% ---------------------------------------------------------------------
\section{The Problem in Paper II}
\label{sec:problem}
% ---------------------------------------------------------------------

Paper~II recovered a set of electromagnetic structures from chiral vortex
topology:
\begin{align}
    F_C(r) &= \frac{\alpha\hbar c}{r^2}, \label{eq:coulomb}\\
    \mathbf{B} &\sim \nabla\times\mathbf{v}_\perp, \label{eq:Bcurl}\\
    \gamma_{\rm AB} &= \frac{q}{\hbar}\oint\mathbf{A}\cdot d\mathbf{l}.
    \label{eq:AB}
\end{align}
Here $\mathbf{v}_\perp$ is the chiral-shear flow associated with the
toroidal vortex, and $\alpha$ is the empirical dimensionless strength of
that sector.

The same section then tentatively identified the propagating photon with a
transverse chiral-shear phonon satisfying
\begin{equation}
    \partial_t^2\delta\theta
    =
    c_\perp^2\nabla^2\delta\theta,
    \qquad
    c_\perp \neq c.
    \label{eq:old_photon}
\end{equation}
This produces the speed problem.  If the physical photon propagated at
$c_\perp$, vacuum light speed would be $c_\perp$, not $c$.  The earlier
patch that $\alpha\to1$ in free vacuum is not available inside SSV because
vacuum \emph{is} the saturated medium.

The diagnosis is therefore:
\begin{equation}
    \hbox{static chiral-shear mechanism: keep,}
    \qquad
    \hbox{radiative photon identification: replace.}
    \label{eq:diagnosis}
\end{equation}

% ---------------------------------------------------------------------
\section{Goldstone Channel of the Saturated Medium}
\label{sec:goldstone}
% ---------------------------------------------------------------------

Write the order parameter as
\begin{equation}
    \Psi(\mathbf{x},t)
    =
    \left(\sqrt{\rho_0}
    +\frac{\delta\rho}{2\sqrt{\rho_0}}\right)
    e^{i\theta_0+i\delta\theta}.
    \label{eq:psi_expand}
\end{equation}
Linearisation gives two familiar modes:
\begin{itemize}[leftmargin=2em]
    \item the amplitude mode $\delta\rho$, gapped by the stiffness of the
          saturated medium;
    \item the phase mode $\delta\theta$, the Goldstone/Bogoliubov mode.
\end{itemize}
At long wavelengths, the phase mode obeys
\begin{equation}
    \partial_t^2\delta\theta
    =
    c^2\nabla^2\delta\theta,
    \label{eq:goldstone_wave}
\end{equation}
with dispersion
\begin{equation}
    \omega = ck .
    \label{eq:goldstone_disp}
\end{equation}

This is the natural carrier for physical radiation in SSV.  It is massless,
universal, and propagates at the observed vacuum light speed.  It is also
the same channel used in the acoustic-metric reading of gravity, where
local density variations modify the local propagation rate.

\begin{resultbox}[Channel assignment]
The Goldstone/Bogoliubov channel carries propagating radiation at $c$.
The chiral-shear channel fixes the handed near-field structure of charged
defects and the coupling strength $\alpha$.
\end{resultbox}

% ---------------------------------------------------------------------
\section{Two-Channel Electromagnetism}
\label{sec:two_channel}
% ---------------------------------------------------------------------

The corrected EM sector should be written as a two-channel theory:
\begin{align}
    \hbox{near field:}\qquad
    &\mathbf{v}_\perp,\ \Phi_\perp,\ \mathbf{A}_\perp,
    &&\hbox{chiral-shear/topological sector},\\
    \hbox{radiation:}\qquad
    &\delta\theta_{\rm G},
    &&\hbox{Goldstone phase channel}.
    \label{eq:two_channel}
\end{align}
The near field knows about charge chirality and the dimensionless coupling
$\alpha$.  The radiative field knows about propagation at $c$.

This is not unusual from the point of view of ordinary electrodynamics.  In
QED, the Coulomb field is not a freely propagating transverse photon; it is
the constrained or longitudinal part of the gauge field.  Radiation is
carried by transverse on-shell photon modes.  SSV should reproduce this
division mechanically:
\begin{align}
    \hbox{Coulomb interaction}
    &\leftrightarrow
    \hbox{static chiral-shear pressure constraint},\\
    \hbox{on-shell light}
    &\leftrightarrow
    \hbox{Goldstone phase wave at }c.
    \label{eq:qed_analogy}
\end{align}

The task is to derive the effective Maxwell system in which the source
sector is chiral-shear/topological, while the homogeneous radiative sector
propagates at $c$:
\begin{align}
    \nabla\cdot\mathbf{E} &= \rho_{\rm ch}/\epsilon_0,\\
    \nabla\cdot\mathbf{B} &= 0,\\
    \nabla\times\mathbf{E} &= -\partial_t\mathbf{B},\\
    \nabla\times\mathbf{B} &= \mu_0\mathbf{J}
       +\frac{1}{c^2}\partial_t\mathbf{E}.
    \label{eq:maxwell_target}
\end{align}
Relative to the provisional Paper~II derivation, the displacement-current
coefficient must be $1/c^2$ rather than $1/c_\perp^2$.

% ---------------------------------------------------------------------
\section{What Must Survive}
\label{sec:survive}
% ---------------------------------------------------------------------

The Goldstone correction is acceptable only if it preserves the successful
static and topological content of Paper~II.  The following structures must
survive.

\paragraph{Coulomb strength.}
The static force must remain
\begin{equation}
    F_C(r)=\frac{\alpha\hbar c}{r^2}.
    \label{eq:coulomb_survive}
\end{equation}
In the corrected reading, $\alpha$ is not a photon-speed ratio.  It is the
strength of the chiral-shear near-field coupling.

\paragraph{Magnetic topology.}
The magnetic field must remain the curl of a vector flow,
\begin{equation}
    \mathbf{B}\propto\nabla\times\mathbf{v}_\perp,
    \label{eq:B_survive}
\end{equation}
so that $\nabla\cdot\mathbf{B}=0$ remains a vector identity and magnetic
monopoles remain topologically forbidden.

\paragraph{Aharonov--Bohm phase.}
The phase shift must remain a Berry phase of the charged ring in a
background circulation:
\begin{equation}
    \gamma_{\rm AB}
    =
    \frac{q}{\hbar}\oint\mathbf{A}\cdot d\mathbf{l}.
    \label{eq:AB_survive}
\end{equation}
This is naturally near-field/topological rather than radiative.

\paragraph{Anomalous magnetic moment.}
The one-loop self-energy picture must be reinterpreted carefully.  The
virtual exchange responsible for $g-2$ may continue to involve the
chiral-shear near field, but the on-shell photon propagator must belong to
the Goldstone channel.  The calculation therefore has to distinguish
virtual constrained exchange from physical radiative quanta.

\begin{gapbox}[Open matching problem]
The central technical challenge is to show that the chiral-shear near field
and the Goldstone radiation field are not two unrelated electromagnetic
fields, but two limits of one effective EM response of the saturated medium.
\end{gapbox}

% ---------------------------------------------------------------------
\section{The Matching Calculation}
\label{sec:matching}
% ---------------------------------------------------------------------

A complete derivation should start from the full SSV functional,
\begin{equation}
    S[\Psi]
    =
    \int dt\,d^3x\,
    \left[
        \mathcal{L}_{\rm LogSE}(\Psi)
        +
        \mathcal{L}_{\perp}(\Psi,\nabla\Psi)
        +
        \mathcal{L}_{\rm defect}
    \right],
    \label{eq:full_action}
\end{equation}
expand around a charged toroidal background $\Psi_{\rm e}$, and separate
fluctuations into constrained near-field and propagating Goldstone parts:
\begin{equation}
    \Psi
    =
    \Psi_{\rm e}
    + \delta\Psi_{\rm near}
    + \delta\Psi_{\rm rad}.
    \label{eq:split}
\end{equation}
The desired effective action has the schematic form
\begin{equation}
    S_{\rm eff}
    =
    \int dt\,d^3x\,
    \left[
      \frac{\epsilon_0}{2}E^2
      -\frac{1}{2\mu_0}B^2
      + J^\mu A_\mu
    \right],
    \label{eq:Seff}
\end{equation}
with two requirements:
\begin{align}
    \hbox{static limit:}\quad
    &\nabla^2\Phi = -\rho_{\rm ch}/\epsilon_0,
    &&F_C=\alpha\hbar c/r^2,\\
    \hbox{radiative limit:}\quad
    &\partial_t^2 A_i - c^2\nabla^2 A_i = 0,
    &&\omega=ck.
    \label{eq:limits}
\end{align}

The calculation is therefore a matched asymptotic problem.  Near charged
defects, chiral-shear topology fixes the source and static response.  Far
from sources, the propagating disturbance must project onto the Goldstone
phase mode.

% ---------------------------------------------------------------------
\section{Avoiding New Circularity}
\label{sec:circularity}
% ---------------------------------------------------------------------

The Goldstone photon correction must not simply insert Maxwell theory by
hand.  The following shortcuts would not count as derivations:
\begin{enumerate}[label=(\roman*)]
    \item postulating a separate electromagnetic gauge field in addition to
          $\Psi$;
    \item replacing $c_\perp$ by $c$ in Paper~II's equations without
          deriving the channel projection;
    \item keeping Coulomb from chiral shear and radiation from Goldstone
          without deriving their matching at sources;
    \item using observed light speed to tune a coefficient in the
          chiral-shear sector.
\end{enumerate}

A valid derivation should show that the same saturated order parameter
supports both the constrained static response and the propagating radiation
response, with no additional gauge field inserted by hand.

\begin{resultbox}[Success criterion]
The Goldstone photon programme succeeds if the full SSV linearisation around
charged toroidal defects yields Maxwell electrodynamics with static Coulomb
strength $\alpha\hbar c$, magnetic topology from chiral-shear circulation,
and on-shell radiation satisfying $\omega=ck$.
\end{resultbox}

% ---------------------------------------------------------------------
\section{Consequences}
\label{sec:consequences}
% ---------------------------------------------------------------------

If the matching calculation succeeds, several tensions in Paper~II are
resolved at once.

\paragraph{Light and gravity share the correct speed.}
Both electromagnetic radiation and gravitational disturbances can be read
as Goldstone/Bogoliubov-channel excitations at $c$, consistent with
GW170817.

\paragraph{\texorpdfstring{$\alpha$}{alpha} is not a speed ratio.}
The fine-structure constant becomes a chiral-shear/topological coupling and
defect-geometry parameter, not a ratio involving the observed photon speed.
This matches the separate alpha-paper proposal
$\alpha=\xi/R^*$.

\paragraph{The old EM derivation becomes a near-field derivation.}
The Paper~II chiral-shear equations are not discarded; they are reclassified.
They describe constrained near-field structure and virtual exchange rather
than the on-shell photon.

\paragraph{Magnetic monopoles remain forbidden.}
Because the magnetic field still descends from a curl of the chiral-shear
flow, $\nabla\cdot\mathbf{B}=0$ remains topological rather than postulated.

% ---------------------------------------------------------------------
\section{Discussion}
\label{sec:discussion}
% ---------------------------------------------------------------------

The Goldstone photon problem is a cleanup calculation, but it is not a
minor one.  It decides whether the SSV electromagnetic sector can be made
internally coherent after the propagation-speed audit.  The desired result
is conceptually simple: sources live in chiral topology; radiation lives in
the Goldstone phase channel.  The hard part is proving that these are two
faces of one effective electromagnetic field.

This also clarifies the status of Paper~II.  Its static Coulomb and
magnetic results remain valuable because they identify how charge and
magnetic topology arise mechanically.  Its original radiative photon
identification should be treated as provisional scaffolding.  The present
paper states the corrected target.

% ---------------------------------------------------------------------
\section{Conclusion}
\label{sec:conclusion}
% ---------------------------------------------------------------------

The observed photon in SSV should be the Goldstone/Bogoliubov phase mode of
the saturated order parameter, propagating at $c$.  The chiral-shear sector
should be retained as the near-field and topological electromagnetic sector,
where it fixes Coulomb strength, magnetic circulation, and the role of
$\alpha$.

The open derivation is the matching calculation:
\[
    \boxed{
    \hbox{chiral-shear near field}
    \quad\longleftrightarrow\quad
    \hbox{Goldstone radiation at }c
    }
\]
If this can be derived from the full SSV functional without adding a
separate gauge field, the electromagnetic sector becomes substantially more
coherent.  If it cannot, the SSV force-sector programme must be revised at
one of its central joints.

\begin{thebibliography}{99}

\bibitem{SSV-I}
S.~Norland,
``Saturated Superfluid Vacuum I: Topological Defects in a Saturated
Superfluid Vacuum --- The Geometric Origin of Matter,''
manuscript in preparation.

\bibitem{SSV-II}
S.~Norland,
``Saturated Superfluid Vacuum II: The Force Sector,''
manuscript in preparation.

\bibitem{Volovik2003}
G.~E.~Volovik,
\textit{The Universe in a Helium Droplet},
Oxford University Press, Oxford (2003).

\bibitem{Zloshchastiev2020}
K.~G.~Zloshchastiev,
``Superfluid vacuum theory and deformed dispersion relations,''
\textit{Int.~J.~Mod.~Phys.~A} \textbf{35}, 2040032 (2020);
arXiv:2011.11897.

\bibitem{GW170817}
B.~P.~Abbott et al.,
``Multi-messenger Observations of a Binary Neutron Star Merger,''
\textit{Astrophys.~J.~Lett.} \textbf{848}, L12 (2017).

\end{thebibliography}

\end{document}
