\documentclass[11pt]{article}

\usepackage[a4paper,margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{amsmath,amssymb}
\usepackage{physics}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{setspace}

\hypersetup{colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue}
\setstretch{1.1}

\title{Saturated Superfluid Vacuum VI-b\\
Galactic Morphology as Overtone Structure}
\author{Stig Norland}
\date{Draft}

\begin{document}
\maketitle

\begin{abstract}
This paper studies galactic morphology in the Saturated Superfluid Vacuum (SSV) framework as
the azimuthal overtone structure of the disc resonance introduced in the companion galactic
paper. The central claim is that the full two-dimensional resonant structure of the disc
generates azimuthal modes that account for morphology, wiggles, and the contrast between
smooth one-dimensional rotation curves and richer two-dimensional structure.
\end{abstract}

\section{Introduction}

This paper treats galactic morphology as a downstream consequence of the disc resonance
picture, not as an independent mechanism.

\section{Two-Dimensional Disc Resonance}

Set up the 2D structure needed for morphology.

\section{Azimuthal Modes}

Describe the overtone picture and the role of azimuthal structure.

\section{Morphology Sequence}

Relate mode structure to the intended morphological classes.

\section{Why One-Dimensional Curves Look Smooth}

Explain the role of azimuthal averaging and observational projection.

\section{Predictions and Limits}

Keep the paper focused on morphology-level predictions.

\section{Conclusion}

The intended claim is that morphology is the overtone structure of the same galactic
resonance, not a separate dark-matter-driven phenomenon.

\end{document}
