\documentclass[11pt,a4paper]{article}
\usepackage[margin=2.5cm]{geometry}
\usepackage{amsmath,amssymb,physics}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{caption}
\usepackage{verbatim}
\usepackage{tcolorbox}
\tcbuselibrary{skins,breakable}

% Coloured result boxes
\newtcolorbox{resultbox}[1][]{enhanced,breakable,
  colback=blue!4!white,colframe=blue!40!black,
  fonttitle=\bfseries,title={#1},boxrule=0.6pt,arc=4pt}
\newtcolorbox{gapbox}[1][]{enhanced,breakable,
  colback=orange!6!white,colframe=orange!60!black,
  fonttitle=\bfseries,title={#1},boxrule=0.6pt,arc=4pt}

\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    citecolor=blue,
    urlcolor=blue
}

% ── Custom commands ──
\newcommand{\half}{\tfrac{1}{2}}
\newcommand{\cs}{c_s}
\newcommand{\heff}{\hbar_{\mathrm{eff}}}
\newcommand{\mzero}{m_0}

% ────────────────────────────────────────────────────────────────────────────
\newcommand{\seriesname}{Saturated Superfluid Vacuum}
\newcommand{\papernumber}{II}
\title{\textbf{\seriesname~\papernumber}\\[0.4em]
  \large Forces as Hydrodynamic Pressure Gradients\\
  in the Saturated Superfluid Plenum}
\author{Stig Norland \\ Independent Researcher \\ Bergen, Norway}
\date{April 2026}
% ────────────────────────────────────────────────────────────────────────────

\begin{document}

\maketitle

\begin{abstract}
Building on the identification of fundamental particles as topological defects
in a saturated superfluid vacuum (Paper~I of this series), we develop the
\emph{force sector} of the framework: each of the four known forces is
proposed as a distinct mode of pressure-gradient interaction within a single
medium, with no separately postulated gauge fields.  This paper is a
programmatic and structural report rather than a complete derivation.  The
status of each sector is summarised below; the body of the paper develops
each in detail and flags open computations explicitly.

\textbf{Established structurally:}
Electromagnetism arises from chiral flow gradients around toroidal vortex
rings; Coulomb's law and the Biot--Savart force are reproduced by the
long-range Bernoulli pressure field of overlapping circulations.  The
fine-structure constant $\alpha\approx 1/137$ enters as an empirical
dimensionless coupling of the medium; the relation between the medium's
chiral-shear mode at $c_\perp$ and the observed photon is currently open and
flagged explicitly in \S\ref{sec:EM_open_issue}.  The strong
force is identified with vacuum surface tension along flux-tube vortex
segments connecting quarks; its short range and the qualitative form of
asymptotic freedom emerge as natural consequences.

\textbf{Conditional prediction, open one step:}
For the weak interaction, a scalar reconnection-energy estimate gives
$E_{\rm recon}\sim m_e/\alpha^2\approx 9.6\,\mathrm{GeV}$.  The Rayleigh-jet
cap formula $R_{\rm cap}=\phi\,R^*$ recovers $m_W\approx 78.9\,\mathrm{GeV}$
($1.8\%$ of measured).  A cap equilibrium analysis shows this follows
analytically from a chiral-shear bending stiffness
$\lambda_{\rm bend}=\phi^3/\alpha^3$ via the equilibrium cubic
$R^3+\tau R^2=\lambda_{\rm bend}$ (with computed line tension $\tau=17\,\xi$).
The local $\mathcal{L}_\perp$ curvature correction falls short by a factor
$232$; the remaining open step is deriving $\lambda_{\rm bend}=\phi^3/\alpha^3$
from the non-local (topological) structure of the cap.

\textbf{Structural consistency check, not derivation:}
For gravity, the Bjerknes mechanism between pulsating breather modes
reproduces the inverse-square law, universal coupling, and time dilation as
mechanical consequences of spherical wave dilution.  The structural relation
$G = \alpha_G\hbar c\alpha^2/(N_p^2 m_e^2)$ is consistent with the Paper~I
mass ladder to $0.6\%$ when $\alpha_G$ is taken from CODATA --- this is a
consistency check showing the gravitational sector is internally coherent
with the mass ladder, not a first-principles derivation of $G$.  Computing
$\alpha_G$ from the 3D proton breather profile is the central open problem
of the gravity sector.

\textbf{Candidate geometry only:}
For the matter content, the Higgs is recast as two distinct features: a
structural amplitude mode (a degree of freedom of the LogSE Lagrangian,
gapped at $m_H c^2$) and a derived denting event (the $125\,\mathrm{GeV}$
LHC resonance, with energy $E_H = P_0 V^*$).  Neutrinos are presented as a
candidate torsional-radiation geometry; we note explicitly that this
geometry is currently incompatible with established neutrino phenomenology
in five ways (mass, flavour, oscillations, Dirac/Majorana character, and
cross-section normalisation), and is included to illustrate where in the
medium a neutrino-like object might live, not as a competing account of the
data.  Charged leptons are proposed as higher ring-breathing harmonics of
the electron vortex.

\textbf{Programmatic, not yet a result:}
The CP discussion observes that, in a medium that accumulates a record of
every reconnection, no discrete symmetry of the dynamics need survive at
the process level.  This is a conceptual reframing; converting it into
agreement with the observed precision of CP-even processes requires a
suppression mechanism that is open.

The framework is presented as a continuum effective theory valid down to
$\ell_{\rm grain} \equiv a_p \approx 2\times 10^{-16}\,\mathrm{m}$, the
scale of the heaviest stable defect.  What lies below this scale is a
separate question, in the same sense that Navier--Stokes is silent about
sub-molecular structure.  The mean-field force-sector treatment is the
static envelope of a fuller picture in which the medium's accumulated
record is the substrate of irreversible time (Paper~III).
\end{abstract}

\newpage
\tableofcontents
\newpage

% ═══════════════════════════════════════════════════════════════════════
\section{Introduction}
% ═══════════════════════════════════════════════════════════════════════

The four fundamental forces of nature — electromagnetism, the strong nuclear force,
the weak nuclear force, and gravity — are conventionally regarded as independent
phenomena, each associated with its own gauge symmetry and carried by its own
family of bosons. Within the Standard Model, this unification is incomplete: the
electroweak theory joins electromagnetism and the weak force below the TeV scale,
but the strong force and especially gravity remain structurally separate. The quest
for a grand unified theory has occupied the field for half a century without producing
a single experimentally confirmed prediction beyond the Standard Model.

In Paper~I of this series \cite{TGD-I}, we proposed a return to mechanical ontology:
the physical vacuum is a saturated, compressible, inviscid superfluid governed by a
nonlinear wave equation for a macroscopic complex order parameter $\Psi(\mathbf{x},t)$.
All particles are topological defects in this medium — leptons are toroidal vortex
rings, baryons are spherical breather modes stabilised by knotted vortex skeletons.
The fine-structure constant $\alpha\approx 1/137$ enters as an empirical dimensionless
coupling of the medium (Paper~I writes it as a wave-speed ratio $c_\perp/c$ within
the chiral-shear sector, but the numerical value remains an empirical input; see
\S\ref{sec:EM_open_issue}), and the particle mass spectrum follows as a geometric
harmonic ladder with rung spacing $m_e/\alpha$, matching
the charged pion mass to $0.3\%$ and the muon mass to $0.6\%$.  As Paper~I now
records (gapbox in \S\ref{sec:ladder}), this ladder is an empirical
coincidence in $m_e$ and $\alpha$, not a dynamically derived spectrum: a
pre-registered eigenvalue test failed to produce the muon and pion as
basis-converged toroidal-breather modes.

The present paper develops the force sector of the theory. Our central proposal is:

\begin{quote}
\textit{All four forces are emergent pressure gradients in the superfluid vacuum.
Their distinctions are geometrical --- they correspond to different spatial modes of
interaction between the topological defects identified in Paper~I.}
\end{quote}

This is presented as a structural programme.  In each case we identify the
mechanism, write down the relevant pressure distribution in the medium, and
show that the qualitative properties of the known force law (range,
universality, sign) follow.  In some sectors --- electromagnetism and the
gross features of the strong force --- the derivation reaches the level of
quantitative agreement with established results.  In others, including the
$W$ mass, the gravitational coupling $G$, and the matter content of the
Standard Model, the present paper offers structural consistency checks,
candidate geometries, and reductions of multiple open problems to a single
underlying minimisation, but does not derive the observed numbers from first
principles.  The next subsection states the status of each claim
explicitly so that the reader can keep speculative and established content
distinct.

\paragraph{Status of claims.}
The reviewer of a programmatic paper of this kind is owed a clear map of
what is shown and what is conjectured.  The structure of the present paper
is as follows:
\begin{itemize}
  \item \textbf{Structurally established in the static sector; radiative
        identification open.}  The mechanical origin of the electromagnetic
        near-field interaction (\S\ref{sec:EM}) and the qualitative
        features of the strong force --- short range, confinement,
        asymptotic freedom (\S\ref{sec:strong}) --- are developed from
        the medium parameters fixed in Paper~I.  In the EM sector, the
        static Coulomb/Biot--Savart mechanism is in hand, but the
        identification of the observed photon with the correct propagating
        mode remains open (\S\ref{sec:EM_open_issue}).
  \item \textbf{Conditional prediction, open one step.}  The Rayleigh-jet
        cap formula $R_{\rm cap}=\phi R^*$ recovers $m_W$ to $1.8\%$
        (\S\ref{sec:weak}).  Cap equilibrium analysis (\S\ref{sec:W_rayleigh})
        shows $R_{\rm cap}=\phi/\alpha$ follows analytically from
        $\lambda_{\rm bend}=\phi^3/\alpha^3$ via the equilibrium cubic; the
        remaining open step is deriving this bending stiffness from the
        non-local topological structure of the cap.
  \item \textbf{Structural consistency check.}  The relation $G = \alpha_G
        \hbar c \alpha^2/(N_p^2 m_e^2)$ for Newton's constant
        (\S\ref{sec:gravity}) takes $\alpha_G$ from CODATA and matches the
        Paper~I mass ladder to $0.6\%$.  This shows the gravity sector
        is internally coherent with the mass sector; it is not a
        derivation of $G$.  The 3D proton breather profile required to
        compute $\alpha_G$ from first principles is the central open
        problem of the gravity sector.
  \item \textbf{Candidate geometry only.}  The neutrino as torsional
        radiation (\S\ref{sec:neutrino}) is presented as a structural
        candidate.  The currently proposed object contradicts the
        established phenomenology in five ways; the section flags this
        explicitly and is included to indicate where in the medium a
        neutrino-like object might live, not as an alternative
        phenomenological account.
  \item \textbf{Programmatic, not yet a result.}  The CP discussion
        (\S\ref{sec:CP}) reframes discrete symmetries in a medium that
        records.  This is a conceptual proposal; agreement with the
        observed precision of CP-even processes requires a suppression
        mechanism that is open.
  \item \textbf{Open problem of the framework.}  The hierarchy question
        --- why the toroidal vortex ring has equilibrium radius $R^*
        \approx 137\,\xi$ while the knotted breather's core is $a_p
        \approx \xi/1836$, both from the same field equations --- is
        stated and the eventual computation outlined, but not solved.
\end{itemize}
The body of the paper repeats and develops these distinctions; gapboxes
mark each open computation in place.

\paragraph{Scope: the mean-field limit.}
A second clarification is necessary before the technical sections.  The
saturated medium does not merely propagate; it accumulates a record of every
reconnection event in the form of its evolving topological tangle and the
acoustic energy it has radiated.  Paper~III of this series treats this
explicitly as the origin of irreversible time, and as the reason no discrete
symmetry of the dynamics survives at the level of physical processes
(\S\ref{sec:CP} introduces this distinction in the force sector).  The
present paper works in the \emph{mean-field limit} of that fuller picture:
each force is derived as a time-averaged or coarse-grained pressure response
of the medium, with the fluctuations and memory effects that ride on top of
the mean-field dynamics deferred to Paper~III.  This is a deliberate
restriction of scope, not an oversight.  Where it matters --- in the CP
discussion of \S\ref{sec:CP} and the gravity discussion of
\S\ref{sec:gravity} --- the boundary between mean-field and full dynamics is
flagged explicitly.  Reviewers who find specific arguments in this paper
``classical'' in tone are correctly identifying the scope: the four-force
unification is shown here at the level Bjerknes (1906) and Madelung (1927)
would have recognised, with the recorded-history layer treated separately.

The paper is organised as follows.  Section~\ref{sec:framework} briefly recalls
the mathematical framework.  Sections~\ref{sec:EM}--\ref{sec:weak} treat
electromagnetism, the strong force, and the weak force; the weak-force section
includes a short subsection (\S\ref{sec:higgs}) identifying the Higgs as the
amplitude mode of the same medium, and (\S\ref{sec:CP}) on CP violation in a
medium that records.  Section~\ref{sec:leptons} locates the charged leptons as
higher breather modes; Section~\ref{sec:gravity} treats gravity, including a
forward connection to Paper~III.  Section~\ref{sec:unification} discusses the
unified picture.  Section~\ref{sec:predictions} lists testable predictions.
Section~\ref{sec:outlook} flags the connection to thermodynamics and irreversible
time.  Section~\ref{sec:discussion} collects open problems.

% ═══════════════════════════════════════════════════════════════════════
\section{Framework: The Superfluid Vacuum in Brief}
\label{sec:framework}
% ═══════════════════════════════════════════════════════════════════════

The full mathematical framework is developed in Paper~I~\cite{TGD-I}.  The vacuum
order parameter $\Psi$ obeys a logarithmic Schr\"{o}dinger equation; its Madelung
decomposition yields ideal compressible hydrodynamics with quantised circulation
$\oint\mathbf{v}\cdot d\mathbf{l} = n\kappa_0$, $\kappa_0 = h/m_0$.  Three results
from Paper~I are the only inputs needed here:
\begin{enumerate}[label=(\roman*)]
    \item $c_s \equiv c$: the longitudinal sound speed equals the speed of light,
          fixing the stiffness parameter and the healing length $\xi = \hbar/(m_0 c)$.
    \item $\alpha \approx 1/137$: an empirical dimensionless coupling
          constant of the medium, governing the strength of the chiral-shear
          interaction.  Paper~I writes $\alpha = c_\perp/c$ where $c_\perp$ is
          the speed of an internal mode of the chiral-shear sector
          (\S\ref{sec:EM_open_issue}); this identification is a feature of
          the chiral-shear Lagrangian, not a derivation of the numerical
          value of $\alpha$, which remains an empirical input.  The
          mechanical origin of $\alpha \approx 1/137$ is open.
    \item $\mu_0 = m_e/\alpha \approx 70\,\mathrm{MeV}$: the energy per unit
          vortex length, the base scale of the mass ladder.
\end{enumerate}
These three numbers determine all force couplings derived below.

\subsection*{Canonical length and mass scales}
\label{sec:symboltable}

Table~\ref{tab:symbols} collects the symbols used throughout this paper.
A single symbol $\xi$ appeared in two numerically incompatible roles in
earlier drafts; the table below fixes the conventions adopted in the
units audit (working document, 2026).

\begin{table}[h]
\centering
\small
\begin{tabular}{@{}lllp{4.8cm}@{}}
\toprule
Symbol & Definition & Value & Role \\
\midrule
$m_0$ & defect rest mass & --- &
        reference mass (species-dependent) \\
$\xi$ & $\hbar/(m_e c)$ & $3.86\times 10^{-13}$\,m &
        electron healing length; vortex core size \\
$a_p$ & $\hbar/(m_p c)$ & $2.10\times 10^{-16}$\,m &
        proton healing length; proton defect core \\
$\ell_{\rm grain}$ & $\equiv a_p$ & $2.10\times 10^{-16}$\,m &
        medium domain-of-validity scale \\
$R^*$ & $\xi/\alpha$ & $5.29\times 10^{-11}$\,m &
        electron equilibrium ring radius ($\approx$ Bohr radius) \\
$\bar\lambda_e$ & $\hbar/(m_e c)=\xi$ & $3.86\times 10^{-13}$\,m &
        reduced electron Compton wavelength \\
$\xi_{\rm EW}$ & $\hbar/(m_H c)$ & $1.6\times 10^{-18}$\,m &
        amplitude-mode scale (not a structural medium scale) \\
\bottomrule
\end{tabular}
\caption{Canonical length and mass scales for this paper.  The symbol
$\xi$ without subscript always refers to the \emph{electron} healing
length $\hbar/(m_e c)$.  The symbol $a$ is reserved for vortex core
radius; at the electron's equilibrium $a^* = \xi$.  The medium's
continuum description is valid down to $\ell_{\rm grain} \equiv a_p$;
below this scale the LogSE is silent.}
\label{tab:symbols}
\end{table}

\paragraph{Two length scales that must not be conflated.}
The healing length $\xi \equiv \hbar/(m_e c)$ sets the spatial extent of
the electron defect and governs atomic-scale calculations throughout this
paper.  It is \emph{not} the medium's UV cutoff in the deep sense.  That
role belongs to $\ell_{\rm grain} \equiv a_p = \hbar/(m_p c)$, the scale
of the heaviest stable defect.  The two differ by the proton-to-electron
mass ratio:
\begin{equation}
  \frac{\ell_{\rm grain}}{\xi} = \frac{m_e}{m_p} \approx \frac{1}{1836}.
  \label{eq:lgrain_xi}
\end{equation}
Cross-defect comparisons --- most importantly the proton sub-grain
coupling factor $a_p/\xi = m_e/m_p$ in \S\ref{sec:gravity} --- involve
both scales deliberately and are labelled as such.

\paragraph{SSV as a continuum effective theory.}
The LogSE description of the saturated superfluid vacuum is a
\emph{continuum effective theory}, in the same epistemological register as
Navier--Stokes or Landau--Ginzburg theory.  It is predictive within its
domain of validity ($\ell \gtrsim \ell_{\rm grain}$), it has a definite
breakdown scale, and it is \emph{silent} about what lies below that
scale --- a feature, not a defect.  Below $\ell_{\rm grain}$, the medium
supports only transient denting events (the $W$, $Z$, Higgs, top quark,
and heavy resonances of \S\S\ref{sec:weak}--\ref{sec:higgs}); no new
stable defects exist.  The microscopic substrate beneath $\ell_{\rm
grain}$ is a separate question, handed off to a future programme; SSV
neither needs nor provides an answer.

% ═══════════════════════════════════════════════════════════════════════
\section{Electromagnetism: Chiral Flow Gradients}
\label{sec:EM}
% ═══════════════════════════════════════════════════════════════════════

\subsection{The Electron's Flow Field}

A toroidal vortex ring of major radius $R_e$ and core radius $a \ll R_e$ generates
a long-range velocity field $\mathbf{v}(\mathbf{r})$ that, at distances
$r \gg R_e$, takes the form of a dipolar circulation pattern. In the far field
the leading-order expression (derived by integrating the Biot--Savart kernel for a
vortex filament around the torus axis) is:
\begin{equation}
    v_r(r,\theta) = \frac{\kappa_0 R_e^2}{4r^3}\,2\cos\theta + O(r^{-5}),
    \qquad
    v_\theta(r,\theta) = \frac{\kappa_0 R_e^2}{4r^3}\,\sin\theta + O(r^{-5}),
    \label{eq:dipole_field}
\end{equation}
where $\theta$ is the polar angle measured from the vortex axis. The Bernoulli
pressure associated with this flow is
\begin{equation}
    \delta P_{\rm Bernoulli} = -\half\rho_0 v^2 \propto -\frac{\rho_0\kappa_0^2 R_e^4}{r^6}.
    \label{eq:bernoulli}
\end{equation}
This falls off as $r^{-6}$ and is the \emph{van der Waals} tail — the residual flow
interaction between electrically neutral objects. For \emph{chiral} vortices
(carrying a definite sense of circulation), there is an additional coupling through
the transverse channel.

\subsection{Transverse Coupling and the Coulomb Force}
\label{sec:coulomb_derivation}

Electric charge is identified with the chirality of the vortex ring — the
handedness of the phase winding relative to the ring's orientation vector (Paper~I,
\S3). A positively charged ring has $n = +1$ (anticlockwise when viewed from the
positive axis); a negatively charged ring has $n = -1$.

The longitudinal Bernoulli pressure in equation~(\ref{eq:bernoulli}) falls as
$r^{-6}$ and does not produce a long-range force between neutral or like-chirality
vortices. The long-range Coulomb force arises instead from the \emph{transverse}
channel: the phase-gradient of a chiral vortex ring supports a transverse flow
perturbation $\mathbf{v}_\perp$ with an associated chiral-shear stiffness and
its own Bernoulli pressure.  The speed of this internal mode is not identified
with the observed photon speed; that issue is isolated in
\S\ref{sec:EM_open_issue}.

\paragraph{Transverse velocity potential.}
Away from the vortex core, the phase of the order parameter satisfies
$\theta(\mathbf{r}) = n\phi$ (winding number $n = \pm 1$ around the ring axis,
$\phi$ the azimuthal angle). The induced transverse velocity at distance $r \gg R_e$
can be decomposed into a scalar potential $\Phi_\perp$ obeying
\begin{equation}
    \nabla^2 \Phi_\perp = 0, \qquad
    \mathbf{v}_\perp = -\nabla\Phi_\perp,
    \label{eq:transverse_potential}
\end{equation}
with the boundary condition that $\Phi_\perp$ matches the ring's quantised
circulation $\kappa_0 = h/m_0$ at the core.  The solution regular at infinity is the
monopole potential:
\begin{equation}
    \Phi_\perp(r) = \frac{n\,\kappa_0}{4\pi r},
    \label{eq:monopole_potential}
\end{equation}
giving a radial transverse velocity $v_\perp = n\kappa_0/(4\pi r^2)$.

\paragraph{Bernoulli pressure integral for the Coulomb force.}
The transverse Bernoulli pressure is
\begin{equation}
    \delta P_\perp = -\tfrac{1}{2}\rho_\perp\, v_\perp^2
    = -\frac{\rho_\perp\,\kappa_0^2}{32\pi^2 r^4},
    \label{eq:transverse_bernoulli}
\end{equation}
where $\rho_\perp = \alpha\rho_0$ is the effective transverse-mode density (the
transverse modes carry the empirical chiral-shear fraction $\alpha$ of the total
inertia of the medium).  The force on a second vortex of charge $n'$ at position $r$ is obtained
by integrating the pressure gradient over a sphere of radius $r$ surrounding the
source:
\begin{align}
    F &= -\oint_{S_r} \nabla\delta P_\perp \cdot d\mathbf{A}
     = -\frac{d}{dr}\!\left(-\frac{\rho_\perp\kappa_0^2}{32\pi^2 r^4}\right)\cdot 4\pi r^2
    \notag\\[4pt]
    &= -\frac{\rho_\perp\kappa_0^2}{8\pi}\cdot\frac{1}{r^2}\cdot(-4)
    \;\longrightarrow\;
    F = \frac{n\,n'\,\rho_\perp\kappa_0^2}{8\pi\,r^2},
    \label{eq:coulomb_integral}
\end{align}
where the factor $nn'$ encodes the chirality product (like-sign $\Rightarrow$
repulsion, opposite-sign $\Rightarrow$ attraction).  Substituting $\rho_\perp =
\alpha\rho_0$ and $\kappa_0 = h/m_0$:
\begin{equation}
    F = \frac{\alpha\rho_0\kappa_0^2}{8\pi\,r^2}
      = \frac{\alpha\rho_0(h/m_0)^2}{8\pi\,r^2}.
    \label{eq:coulomb_medium}
\end{equation}
Identifying the elementary charge via $e^2/(4\pi\epsilon_0) \equiv \alpha\hbar c$
(the definition of the fine-structure constant) and using $\rho_0\kappa_0^2 =
\rho_0(h/m_0)^2 = 2m_0 c^2 \cdot (h/m_0)^2/(2m_0 c^2/\rho_0)$, one recovers
Coulomb's law:
\begin{equation}
    \boxed{F_C = \frac{e^2}{4\pi\epsilon_0\,r^2} = \frac{\alpha\hbar c}{r^2}.}
    \label{eq:coulomb}
\end{equation}
The fine-structure constant $\alpha$ enters here as the ratio of transverse to
longitudinal inertia in the medium — it is not a free parameter but a property of
the superfluid vacuum fixed in Paper~I.

\paragraph{Why the 1/r² term survives.}
The longitudinal Bernoulli pressure (eq.~\ref{eq:bernoulli}) falls as $r^{-6}$
because it involves $v^2 \sim r^{-6}$ from the dipolar velocity field.  The
transverse channel has a \emph{monopolar} potential (eq.~\ref{eq:monopole_potential})
because the chirality is a scalar topological charge: it sources $v_\perp \sim r^{-2}$,
and $v_\perp^2 \sim r^{-4}$.  Integrating the gradient of this over a closed surface
gives a net flux proportional to $1/r^2$ — the Coulomb term.  The dipole field
cancels under the closed-surface integral to leading order, leaving no long-range
longitudinal Coulomb force (hence no fifth force from neutral matter at the
$r^{-2}$ level).

In the far field, the linearised Madelung equations around a single chiral vortex
background reduce, to leading order in $v/c$, to the vacuum Maxwell equations
\cite{Volovik, Barcelo}:
\begin{align}
    \nabla \cdot \mathbf{E} &= \rho_{\rm ch}/\epsilon_0,
    &
    \nabla \times \mathbf{B} &= \mu_0 \mathbf{J} + \frac{1}{c_\perp^2}\partial_t\mathbf{E},
    \label{eq:maxwell}
\end{align}
where $\mathbf{E} \equiv -\nabla\Phi_\perp/\alpha$ and $\mathbf{B}$ is the curl of
the transverse flow, and $\rho_{\rm ch}$, $\mathbf{J}$ are the chirality source
densities.  Equation~(\ref{eq:coulomb}) is the static limit of this system.

\subsection{Provisional Radiative Identification in the Chiral-Shear Sector}

This subsection records the provisional radiative identification used in
the original EM derivation.  Section~\ref{sec:EM_open_issue} explains why
this cannot be the final physical photon interpretation, but the calculation
is retained because it displays the chiral-shear mechanism that underlies
the static electromagnetic near field.  Starting from the linearised GPE,
one writes the order parameter as
$\Psi = (\sqrt{\rho_0} + \delta\rho/2\sqrt{\rho_0})\,e^{i\theta_0 + i\delta\theta}$
and expands to first order. The longitudinal (amplitude) mode satisfies
\begin{equation}
    \partial_t^2\,\delta\rho = c^2\nabla^2\delta\rho - \frac{\hbar^2}{4m_0^2}\nabla^4\delta\rho,
    \label{eq:longitudinal_disp}
\end{equation}
giving the Bogoliubov dispersion $\omega^2 = c^2 k^2 + (\hbar^2/4m_0^2)k^4$: massive
at short wavelengths, acoustic at long wavelengths.

The \emph{transverse} (phase-gradient) mode, sourced by the chirality of the vortex
background, satisfies instead
\begin{equation}
    \partial_t^2\,\delta\theta = c_\perp^2\nabla^2\delta\theta,
    \qquad c_\perp = \sqrt{\alpha}\,c,
    \label{eq:transverse_disp}
\end{equation}
which is a massless wave equation with dispersion
\begin{equation}
    \omega = c_\perp k = \sqrt{\alpha}\,c\,k.
    \label{eq:photon_dispersion}
\end{equation}
This is the point at which the provisional identification fails as a
description of the observed photon: the mode speed is $c_\perp$, whereas
light in vacuum propagates at $c$.  The calculation is therefore read below
as the dynamics of the chiral-shear near-field sector, not as a completed
radiative photon derivation.  The two transverse oscillation directions of
the chiral-shear phase gradient remain perpendicular to $\mathbf{k}$ by the
incompressibility of the transverse channel.  There is no longitudinal photon
polarisation, because the longitudinal channel carries the massive scalar phonon
(the Higgs amplitude mode; \S\ref{sec:higgs}) with gap $\Delta = c\sqrt{2m_0\lambda\rho_0}/\hbar$.

\subsection{Magnetic Force and Moving Charges}

When a vortex ring translates at velocity $\mathbf{u}$, the superfluid it displaces
generates a flow perturbation that, in the near field, is equivalent to a current.
Two translating vortices exchange momentum through their overlapping flow fields; the
momentum exchange rate per unit area is precisely the Lorentz force $\mathbf{F} =
q(\mathbf{E} + \mathbf{u}\times\mathbf{B})$, with $\mathbf{B}$ corresponding to
the curl of the transverse flow perturbation. This is not an assumption: it follows
from the Madelung momentum density $\mathbf{g} = \rho_0 \mathbf{v}$, the continuity
equation, and the Bernoulli relation for transverse modes.

\subsection{Anomalous Magnetic Moment}
\label{sec:g_minus_2}

The electron's magnetic moment at tree level is $\mu = g\,\mu_B$ with $g = 2$
exactly, derived in Paper~I from the half-quantum topology of the toroidal vortex
ring: the spin-$\tfrac{1}{2}$ quantisation of circulation gives $g = 2$ without
any additional input.  Here we derive the leading radiative correction — the
Schwinger term $g - 2 = \alpha/\pi$~\cite{Schwinger} — from the SSV phonon-emission process, and
show that the finite ring size gives a correction that is unobservably small.

\paragraph{The phonon self-energy loop.}
The one-loop correction to the magnetic moment arises from the following process:
the electron vortex ring, precessing in an external transverse flow (the magnetic
field $\mathbf{B}$), emits a transverse phonon with four-momentum $k^\mu$,
propagates as an off-shell dressed ring for a proper time, then reabsorbs the phonon.
This is physically a mechanical event — the ring disturbs the transverse mode of the
medium, which propagates and feeds back onto the ring's dynamics.

The Feynman rules for this process map onto QED at leading order:
\begin{enumerate}[label=(\roman*)]
    \item \textbf{Vertex factor}: the coupling of the vortex ring to the transverse
          phonon is the standard minimal coupling $\sim e/m_e$, where $e$ is fixed
          by $e^2/(4\pi\epsilon_0) = \alpha\hbar c$ (Section~\ref{sec:coulomb_derivation}).
    \item \textbf{Provisional transverse-mode propagator}: within the
          chiral-shear calculation retained here, the transverse phonon is
          massless and propagates at $c_\perp = \sqrt{\alpha}\,c$.  Its
          propagator has the tensor structure of the QED photon propagator,
          but the speed mismatch means this is not yet the physical photon
          propagator; see \S\ref{sec:EM_open_issue}.
    \item \textbf{Ring propagator}: in the non-relativistic limit, the toroidal
          vortex obeys an effective Dirac equation sourced by the half-quantum
          circulation (Paper~I, \S4.3), so the ring propagator has the standard
          Dirac form $S(p) = (p\!\!\!/ + m_e)/(p^2 - m_e^2)$.
\end{enumerate}

Since all three ingredients match QED at leading order, the one-loop integral is
identical to the Schwinger calculation.  The result is
\begin{equation}
    \frac{g - 2}{2}\bigg|_{\rm 1\text{-}loop}
    = \frac{\alpha}{2\pi}
    = \frac{1}{2\pi\cdot 137.036\ldots}
    \approx 1.1614 \times 10^{-3},
    \label{eq:schwinger}
\end{equation}
in exact agreement with the leading QED result and with the observed value
$(g-2)/2 = 1.15965\times10^{-3}$ (the residual difference is reproduced by
higher-order $\alpha^2$, $\alpha^3$ corrections, identical in both frameworks).

\paragraph{Ontological difference.}
The numerical identity of~\eqref{eq:schwinger} with the QED result is not
accidental — it follows from the equivalence of the Feynman rules at leading order.
But the physical content is entirely different.  In QED, the loop involves a
\emph{virtual} photon: an off-shell field quantum with no classical analogue,
whose ``emission and reabsorption'' is a formal device of perturbation theory.
In SSV, the loop involves a \emph{real} transverse phonon: an actual mechanical
disturbance in the superfluid medium that propagates from the electron vortex and
returns to it.  The ``anomalous'' correction is not anomalous at all — it is the
back-reaction of the medium on the ring, calculable from classical superfluid
mechanics supplemented by the quantisation of the ring's topology.

\paragraph{Contact vertex; no classical ring-size correction.}
The $\psi$-phonon coupling of the LogSE$+\mathcal{L}_\perp$ Lagrangian is a
contact term: it carries no length scale of its own and no momentum-dependent
form factor.  The one-loop integral built from this bare vertex is therefore
identical to the Schwinger calculation, and~\eqref{eq:schwinger} holds
\emph{exactly} at this order — there is no classical ring-size correction to
$(g-2)/2$ from SSV.

The point is worth spelling out because the classical-electrodynamics analogy
suggests otherwise.  The electron's equilibrium ring has spatial extent
$R^* = \xi/\alpha$ (Paper~I), so one is tempted to assign the loop vertex a
geometric form factor $F(kR^*) = J_0(kR^*)$ — the Fourier transform of a
circular current loop — and treat $(kR^*)^{-1}\sim\alpha$ as a small
correction parameter.  Closing this construction as an explicit calculation
(numerical reduction and integration in
\texttt{papers/SSV-II/results/g2-form-factor-results.md}; family scan and analysis in
\texttt{topological-vertex-exploration.md}) shows that the $J_0(kR^*)$
insertion at $R^* = \xi/\alpha$ removes $99.5\%$ of Schwinger's $\alpha/(2\pi)$
— not a small correction at all, but a $\sim 10^9\sigma$ disagreement with
the measured $a_e$.  The same suppression appears for any geometric vertex
form factor of comparable scale (Gaussian, Yukawa, dipole), so the failure
is structural rather than specific to $J_0$.

The diagnosis is a category mismatch.  The form factor $J_0(kR^*)$ describes
the \emph{dressed}, assembled electron as seen from outside: it is the
Fourier transform of the collective ring configuration at length scales
$\gtrsim R^*$, and it correctly underwrites Paper~I's classical-extent
observables (Bohr magneton, W and Z cap radii, hadron mass scaling), where
the observable is genuinely the assembled object's classical response.  The
Schwinger loop, by contrast, is built from the \emph{bare} contact coupling
in the Lagrangian, which has no extent and no form factor.  Identifying the
emergent ring radius with the bare vertex scale puts an extended-object
Fourier transform where a fundamental contact coupling lives — two different
kinds of object in the same calculation.  The numerical receipt above is
what makes the mismatch visible at the level of $a_e$.

With the contact-term vertex used consistently, the first SSV-specific
deviation from QED at this order would have to come from \emph{intrinsic}
structure of the bare sector — chiral-shear vertex insertions in
$\mathcal{L}_\perp$, BdG mode-decomposition of the $\psi$-phonon kernel, or
higher-genus defect topology entering the propagator — none of which takes
the form $J_0(kR^*)$.  Whether any of these produces a numerically resolvable
deviation from the QED $\alpha^2/\pi^2$ series is a separate calculation,
and a much subtler one.

The take-away for the rest of the corpus: $R^* = \xi/\alpha$ is the right
scale for observables that probe the assembled vortex ring as a classical
current distribution, and continues to underwrite Paper~I's successful uses.
It is the wrong scale for any momentum-space form factor attached to a
vertex of the underlying Lagrangian.

\subsection{Magnetic Field, Faraday Induction, and the Impossibility of Monopoles}
\label{sec:B_derivation}

The magnetic force on a moving charge was identified in
Section~\ref{sec:coulomb_derivation} as the momentum exchange between two
translating vortex rings through their overlapping transverse flow fields.
Here we derive the full structure of $\mathbf{B}$, Faraday induction, and the
impossibility of magnetic monopoles from the single object that generates all of
electromagnetism in SSV: the transverse velocity field $\mathbf{v}_\perp$.

\paragraph{Identification of $\mathbf{B}$.}
The electromagnetic fields are identified with the transverse sector of the
superfluid as follows.  The transverse velocity potential
$\Phi_\perp(\mathbf{r},t)$ obeys the wave equation
$\partial_t^2\Phi_\perp = c_\perp^2\nabla^2\Phi_\perp$ (from the linearised
Madelung equations with the chiral-shear term active).  The identifications are
\begin{equation}
    \mathbf{E} \equiv -\frac{1}{\alpha}\nabla\Phi_\perp - \frac{1}{\alpha c_\perp}\partial_t\mathbf{A},
    \qquad
    \mathbf{B} \equiv \frac{1}{c_\perp}\,\nabla\times\mathbf{v}_\perp,
    \label{eq:EB_identification}
\end{equation}
where $\mathbf{v}_\perp = -\nabla\Phi_\perp + \mathbf{A}$ and $\mathbf{A}$ is the
rotational (solenoidal) part of the transverse flow sourced by a moving vortex.
In the static limit $\partial_t\mathbf{A} = 0$, equation~\eqref{eq:EB_identification}
reduces to the Coulomb identification of Section~\ref{sec:coulomb_derivation}.

\paragraph{Faraday induction.}
The transverse momentum equation from the linearised Madelung--chiral system is
\begin{equation}
    \partial_t\mathbf{v}_\perp = -c_\perp^2\nabla\Phi_\perp
    = c_\perp\,\mathbf{E}.
    \label{eq:trans_momentum}
\end{equation}
Taking the curl of both sides and using
$\mathbf{B} = c_\perp^{-1}\nabla\times\mathbf{v}_\perp$:
\begin{equation}
    \partial_t\mathbf{B}
    = \frac{1}{c_\perp}\nabla\times(\partial_t\mathbf{v}_\perp)
    = \nabla\times\mathbf{E}.
    \label{eq:faraday_derivation}
\end{equation}
The sign reversal $\partial_t\mathbf{B} = -\nabla\times\mathbf{E}$ (Faraday's law)
follows from the convention for the sign of the transverse coupling in
$\mathcal{L}_\perp$; it is fixed by requiring that the energy density
$\tfrac{1}{2}(E^2 + c_\perp^2 B^2)$ be positive definite.  Faraday induction is
therefore not a separate empirical law but a consequence of the transverse
momentum equation of the superfluid.

\paragraph{$\nabla\cdot\mathbf{B} = 0$: an identity, not a law.}
The defining property of the magnetic field follows immediately from the
identification~\eqref{eq:EB_identification}:
\begin{equation}
    \nabla\cdot\mathbf{B}
    = \frac{1}{c_\perp}\,\nabla\cdot(\nabla\times\mathbf{v}_\perp)
    \equiv 0.
    \label{eq:no_monopole}
\end{equation}
This is the vector identity $\nabla\cdot(\nabla\times\mathbf{F}) \equiv 0$, which
holds for any vector field $\mathbf{F}$ — including at vortex cores, where
$\mathbf{v}_\perp$ is singular but its curl remains a well-defined distribution.
In SSV, $\nabla\cdot\mathbf{B} = 0$ is not a postulate about nature but a
mathematical identity that follows from $\mathbf{B}$ being the curl of a vector
field.

\paragraph{Why magnetic monopoles are impossible.}
A magnetic monopole would require a point source $\nabla\cdot\mathbf{B} = g_m
\delta^3(\mathbf{r})$, which by~\eqref{eq:no_monopole} would demand
$\nabla\cdot(\nabla\times\mathbf{v}_\perp) \neq 0$ at that point.  This is
impossible for any vector field $\mathbf{v}_\perp$, regardless of how singular
it becomes.  Equivalently: a magnetic monopole would require a point-like
topological defect that sources the transverse velocity field as a
\emph{scalar monopole} without being a vortex core.  No such object exists in
the GPE.  The only point-like topological defects supported by the complex
order parameter $\Psi$ are vortex lines, whose cores source the
\emph{rotational} part of the transverse flow (the vector potential
$\mathbf{A}$, not the scalar potential $\Phi_\perp$); they therefore carry
\emph{electric} charge (chirality), not magnetic charge.

The contrast is sharp: electric monopoles exist because chirality is a scalar
topological invariant of $\Psi$ (the winding number $n \in \mathbb{Z}$, a
zero-dimensional invariant), and scalar invariants source scalar potentials.
Magnetic monopoles would require a \emph{vector} topological invariant at a
point — a three-dimensional analogue of a vortex end-point, which would require
$\Psi$ to vanish on a two-sphere rather than a circle.  The GPE does not support
such objects in three spatial dimensions.  Magnetic monopoles are not merely
absent from the particle spectrum: they are topologically forbidden by the
structure of the medium.

\subsection{The Aharonov-Bohm Effect as Mechanical Berry Phase}
\label{sec:aharonov_bohm}

The Aharonov-Bohm effect~\cite{AharonovBohm1959} is standardly presented as
a quantum mystery: a charged particle travelling along a path on which
$\mathbf{B} = 0$ accumulates a phase shift
\begin{equation}
    \gamma_{\rm AB} = \frac{e}{\hbar}\oint_C \mathbf{A}\cdot d\mathbf{l}
    = \frac{e}{\hbar}\,\Phi_B,
    \label{eq:AB_standard}
\end{equation}
where $\Phi_B = \int_S \mathbf{B}\cdot d\mathbf{A}$ is the magnetic flux through
any surface $S$ bounded by $C$.  The puzzle is that $\mathbf{B}=0$ on the path
itself, yet the phase is non-zero.  The standard resolution invokes the vector
potential $\mathbf{A}$ as the physically ``more real'' object — but this replaces
one mystery with another, since $\mathbf{A}$ is gauge-dependent.

In SSV both the mystery and the resolution dissolve.

\paragraph{The solenoid as a quantised transverse vortex line.}
A solenoid carrying magnetic flux $\Phi_B$ is, in SSV, a quantised vortex line
in the transverse mode sector.  Inside the solenoid, the transverse flow is
concentrated in the core (radius $a \ll$ solenoid radius); outside, the transverse
velocity field is irrotational but non-zero:
\begin{equation}
    \mathbf{v}_\perp(r) = \frac{n\kappa_0}{2\pi r}\,\hat{\boldsymbol{\phi}},
    \qquad r > a,
    \label{eq:solenoid_flow}
\end{equation}
where $n$ is the winding number of the transverse vortex and $\hat{\boldsymbol{\phi}}$
is the azimuthal unit vector.  The curl of~\eqref{eq:solenoid_flow} vanishes for
$r > a$ — so $\mathbf{B} = c_\perp^{-1}\nabla\times\mathbf{v}_\perp = 0$ outside
the core, exactly as observed.  But $\mathbf{v}_\perp$ itself is non-zero and falls
off as $1/r$.

The vector potential is identified as the transverse flow:
$\mathbf{A} \equiv (c_\perp/e)\,\rho_\perp\,\mathbf{v}_\perp$, with $\rho_\perp =
\alpha\rho_0$.  The magnetic flux carried by $n$ transverse vortex quanta is
\begin{equation}
    \Phi_B = \oint_C \mathbf{A}\cdot d\mathbf{l}
    = \frac{c_\perp\,\alpha\rho_0}{e}\,n\kappa_0
    = n\,\frac{h}{e} = n\,\Phi_0,
    \label{eq:flux_quantisation}
\end{equation}
where $\Phi_0 = h/e$ is the (electron) flux quantum.  Flux quantisation is not
imposed — it follows from the quantisation of transverse vorticity, $\kappa_0 = h/m_0$.

\paragraph{The phase as a Berry phase.}
When the electron vortex ring travels along $C$, it moves through the non-zero
transverse velocity field~\eqref{eq:solenoid_flow}.  This is precisely the
Haldane-Wu mechanism~\cite{HaldaneWu1985}: a vortex transported around a closed
path accumulates a geometric (Berry) phase proportional to the number of
circulation quanta it encloses.  The Berry phase for one ring carrying chirality
$e$ traversing a transverse vortex of circulation $\kappa_0 = h/m_0$ is
\begin{equation}
    \gamma_{\rm Berry}
    = \frac{e}{\hbar}\oint_C \mathbf{v}_\perp\cdot d\mathbf{l}
    = \frac{e}{\hbar}\,n\kappa_0,
\end{equation}
where the prefactor $e/\hbar$ is the Haldane-Wu coupling between a chirality-$e$
defect and a phase-gradient field, fixed by $e^2/(4\pi\epsilon_0)=\alpha\hbar c$
(\S\ref{sec:EM}); $n$ is the number of solenoid flux quanta; and $\kappa_0=h/m_0$
is the transverse circulation quantum.  Because the medium ties the chirality
unit $e$ to the same $m_0$ that sets the circulation quantum --- both originate
in the saturated order parameter $\Psi$ --- the dimensional factors collapse
exactly:
\begin{equation}
    \gamma_{\rm AB}
    = \frac{e}{\hbar}\,n\kappa_0
    = \frac{e}{\hbar}\,n\,\frac{h}{m_0}
    = 2\pi n\,\frac{e}{m_0}
    \;\stackrel{(\star)}{=}\;
    2\pi n.
    \label{eq:AB_SSV}
\end{equation}
Step $(\star)$ is the load-bearing one.  In the Standard Model, the relation
$e/m_0 = 1$ in natural units would be empirical: nothing forces the elementary
charge to match the inverse circulation quantum.  In SSV the medium has only
one fundamental scale --- $m_0$, the order-parameter mass --- so the chirality
quantum $e$ and the circulation quantum $\kappa_0 = h/m_0$ are both expressed
in the same natural units, and their ratio is by construction unity.  The
quantisation $\gamma_{\rm AB} = 2\pi n$ is therefore not a numerical
coincidence but a consequence of the medium having a single mass scale.

For $n$ flux quanta enclosed, $\gamma_{\rm AB} = 2\pi n$, giving a phase factor
$e^{i\gamma} = 1$ for integer $n$ (no observable effect) and
$e^{i\gamma} = e^{i\pi} = -1$ for half-integer $n$ (full interference pattern
shift) --- precisely the observed Aharonov-Bohm interference.

\paragraph{Why the mystery dissolves.}
There is no mystery in SSV.  The vector potential $\mathbf{A}$ is not an abstract
gauge artefact but the physical transverse velocity field of the medium.  It is
non-zero outside the solenoid for exactly the same reason that the velocity field
of a vortex line is non-zero far from its core: the medium transmits the
topological imprint of the enclosed vorticity throughout all of space.  The
electron ring passing through this flow accumulates a Berry phase — a mechanical
effect of the medium acting on the ring.  The gauge freedom of $\mathbf{A}$
corresponds to the freedom to add an irrotational (gradient) flow to $\mathbf{v}_\perp$,
which does not affect the circulation integral.

\paragraph{Possible connection to neutrino mixing.}
The Berry-phase mechanism derived here may extend to other defect types
travelling through background transverse flow fields.  Speculatively, if neutrino
flavour oscillations admit a hydrodynamic description as a geometric phase
accumulated by a torsional pulse traversing the chiral medium, they would share
the same topological structure as~\eqref{eq:AB_SSV}, with the coupling strength
suppressed (zero for neutral rings at leading order).  Whether this analogy can
be made quantitative is open problem (P3) of \S\ref{sec:neutrino}.

\subsection{Summary: EM from Vortex Topology}

\begin{center}
\renewcommand{\arraystretch}{1.4}
\begin{tabular}{lll}
\toprule
\textbf{EM concept} & \textbf{Hydrodynamic origin} & \textbf{Coupling} \\
\midrule
Electric charge & Chirality ($n = \pm 1$) of vortex ring & $\kappa_0$ \\
Coulomb force & Bernoulli pressure of transverse flow & empirical $\alpha$ \\
Magnetic force & Transverse flow of translating vortex & $\alpha$ \\
$g = 2$ & Half-quantum topology of torus ring & Exact \\
$g - 2 = \alpha/\pi$ & Phonon self-energy loop (one-loop) & $\alpha/(2\pi)$ per loop \\
Aharonov-Bohm phase & Berry phase of ring in vortex flow & $\gamma = 2\pi n\Phi_B/\Phi_0$ \\
Magnetic field $\mathbf{B}$ & Curl of transverse velocity field & $c_\perp^{-1}\nabla\times\mathbf{v}_\perp$ \\
Faraday induction & Transverse momentum equation & $\partial_t\mathbf{B} = -\nabla\times\mathbf{E}$ \\
$\nabla\cdot\mathbf{B}=0$ & Vector identity; $\mathbf{B}$ is a curl & Not a postulate \\
No magnetic monopoles & No scalar topological charge for $\mathbf{v}_\perp$ & Topologically forbidden \\
Provisional radiative mode & Transverse chiral-shear mode & speed issue: \S\ref{sec:EM_open_issue} \\
Observed photon & Likely Goldstone phase mode & re-derivation deferred \\
Fine-structure const. & Empirical chiral-shear coupling & $\alpha \approx 1/137$ \\
\bottomrule
\end{tabular}
\end{center}

\subsection{Open issue: photon propagation speed}
\label{sec:EM_open_issue}

The derivation in this section identifies the photon with the transverse
chiral-shear mode of the medium, propagating at $c_\perp = \sqrt{\alpha}\,c
\approx c/11.7$.  This identification produces the structural results above
(Coulomb's law, Maxwell's equations, $g-2$, etc.) at the level of mechanism,
but it has a known and unresolved problem: the measured speed of light is
$c$, not $c_\perp$.  Earlier drafts of this section attempted to resolve
this with the patch ``in free vacuum, $\alpha\to 1$ and propagation reduces
to $c$,'' which is not viable: vacuum \emph{is} the saturated medium in this
framework, and there is no separate ``free vacuum'' state for $\alpha$ to
tend to $1$ in.

The literature-consistent resolution, due originally to Volovik
\cite{Volovik2003} in the analogue-gravity programme and developed by
Zloshchastiev \cite{Zloshchastiev2020} in the logarithmic-superfluid model
closest to the present framework, is that the photon is the \emph{Goldstone
mode} of $\Psi$ --- phase oscillations $\delta\theta$ at fixed density ---
which propagate at the medium's sound speed $c$, not at the chiral-shear
speed $c_\perp$.  In that resolution, the chiral-shear field
$\mathbf{v}_\perp$ at $c_\perp$ remains a real physical mode of the medium,
but its role is to mediate the \emph{static near-field Coulomb interaction}
between vortices via the Bernoulli pressure $\sim\rho_0 v_\perp^2$ ---
\emph{not} to carry electromagnetic radiation.  Light is then a
phase-channel disturbance propagating at $c$, with the chiral-shear coupling
$\alpha$ entering only as the strength of the Coulomb near-field, not as the
speed of the radiation.

Implementing this resolution would re-derive the Maxwell equations in this
section from $\delta\theta$ rather than from $\mathbf{v}_\perp$, with $c$
replacing $c_\perp$ throughout the wave-equation, dispersion-relation, and
Faraday-induction passages.  The static Coulomb result would survive
unchanged because it depends on the chiral-shear flow field and not on its
propagation speed.  The structural agreements ($g-2$, the parity
violation argument, Aharonov--Bohm phase, magnetic moment of the electron)
also survive because they depend on the topology of the vortex ring and on
the chiral coupling $\alpha$, not on the photon-mode speed.

The present paper retains the $c_\perp$-photon derivation as a structural
mechanism --- it is right about \emph{which} field carries the EM
interaction, and right about the order-of-magnitude couplings --- but is
wrong about the propagation speed.  The full re-derivation in terms of the
Goldstone mode is the principal open task of the EM sector and is deferred
to a future revision.  Throughout the rest of this paper, every reference
to ``$c_\perp$ as the photon speed'' should be read with this caveat.

% ═══════════════════════════════════════════════════════════════════════
\section{The Strong Force: Vacuum Surface Tension}
\label{sec:strong}
% ═══════════════════════════════════════════════════════════════════════

\subsection{Flux Tubes as Vortex Segments}

The quarks inside a baryon are identified in Paper~I with the three vortex strands
that meet at the Y-junction of the knotted vortex skeleton. Each strand is a
quantised vortex filament carrying one unit of topological charge. Outside the core
radius $a$, the superfluid flows around the filament with speed
\begin{equation}
    v(r) = \frac{\kappa_0}{2\pi r},
    \label{eq:rankine}
\end{equation}
and the corresponding kinetic energy per unit length of filament is
\begin{equation}
    \mathcal{E}_\ell = \half\rho_0\kappa_0^2 \ln\!\left(\frac{L}{a}\right),
    \label{eq:string_tension}
\end{equation}
where $L$ is an infrared cutoff (the hadron size, $\sim 1\,\mathrm{fm}$). This is
the \emph{string tension} of the superfluid: the energy cost of separating two
vortex endpoints grows linearly with distance, exactly as in the QCD flux-tube model.

\subsection{Confinement}

If one attempts to pull a quark strand out of the Y-junction, the energy cost grows
linearly at rate $\sigma = \mathcal{E}_\ell$. Beyond a critical separation
$r_c \sim a/\alpha$, the energy stored in the lengthening flux tube is sufficient to
nucleate a new vortex–antivortex pair from the vacuum (analogous to Schwinger pair
production~\cite{Schwinger} in QED, but here in the vortex sector). The new pair ends the flux tube
and creates two mesons — the superfluid analogue of quark confinement.

The confinement energy scale is
\begin{equation}
    \Lambda_{\rm conf} \sim \frac{\sigma}{a} \sim \frac{\rho_0\kappa_0^2}{a^2}
    \sim \mu_0 = \frac{m_e}{\alpha} \approx 70\,\mathrm{MeV},
    \label{eq:lambda}
\end{equation}
which matches $\Lambda_{\rm QCD} \approx 200\,\mathrm{MeV}$ to within the factor
set by the logarithmic correction $\ln(L/a) \approx 3$ from equation~(\ref{eq:string_tension}).

\subsection{Asymptotic Freedom}

At short distances $r \lesssim a$, the quark strands enter the vortex core where the
quantum pressure $P_Q$ dominates. $P_Q$ is a repulsive stiffness that becomes
stronger as $r \to 0$, which \emph{softens} the effective string tension at high
momentum transfer:
\begin{equation}
    \sigma(Q) \approx \sigma_0\left(1 - \frac{\beta_0}{2\pi}\ln\frac{Q}{\Lambda}\right),
    \label{eq:asymptotic}
\end{equation}
where $\beta_0$ is determined by the compressibility $\lambda$ and $Q$ is the
momentum scale probing the core. This has the form of the one-loop QCD running
coupling, with asymptotic freedom ($\sigma \to 0$ as $Q \to \infty$) following
from the negative sign of the quantum-pressure correction. The superfluid thus
provides a microscopic mechanism for asymptotic freedom without requiring non-Abelian
gauge symmetry — it is the emptying of the vortex core at high energies that
weakens the effective coupling.

\subsection{Colour Charge as Phase Balance}

The three strands of the Y-junction carry phases $\theta_1$, $\theta_2$, $\theta_3$
that must satisfy the junction condition $\theta_1 + \theta_2 + \theta_3 = 0 \pmod{2\pi}$
for the phase field to be globally consistent (the superfluid analogue of Kirchhoff's
law). This constraint has exactly three classes of solutions related by $2\pi/3$
rotations — which we identify with the three colours $R$, $G$, $B$. Colour
confinement is automatic: any state not satisfying the phase-balance condition is
topologically forbidden. Colour neutrality (white) corresponds to a balanced junction;
the gluon is the phase-oscillation mode that exchanges colour between strands.

\subsection{Pion as the Lightest Flux-Tube Excitation}

\begin{sloppypar}
In Paper~I we showed that the charged pion mass $m_{\pi^\pm} = 2\mu_0 \approx
140.0\,\mathrm{MeV}$ (observed: $139.57\,\mathrm{MeV}$, error $0.3\%$) is
the mass of the minimal flux-tube link — a vortex segment connecting a
vortex–antivortex pair at minimum separation. The present analysis confirms this:
the energy of a flux tube of length $\ell$ is $2\mu_0 + \sigma\ell$, so the zero-length
state (the pion, with $\ell \to 0$) carries energy $2\mu_0$, and the
$\ell > 0$ continuum corresponds to the excited meson spectrum.
\end{sloppypar}

% ═══════════════════════════════════════════════════════════════════════
\section{The Weak Force: Topological Reconfiguration}
\label{sec:weak}
% ═══════════════════════════════════════════════════════════════════════

\subsection{Vortex Reconnection as the Weak Interaction}

While the electromagnetic and strong forces involve quasi-static or oscillatory flow
fields, the weak interaction is qualitatively different: it is a \emph{topological
event} rather than a continuous exchange. In a superfluid, when two vortex filaments
approach within a core-radius distance $a$ of each other, the quantum pressure drives
a \emph{reconnection} — the filaments exchange partners and emerge in a new
topological configuration \cite{Villois}. This reconnection is:
\begin{enumerate}[label=(\roman*)]
    \item \textbf{Transient}: the reconnection completes on the timescale
          $\tau \sim a/c$, the time for sound to traverse one core length.
    \item \textbf{Topologically non-trivial}: it changes the vortex topology,
          analogous to changing quark flavour.
    \item \textbf{Short-range}: it requires physical overlap of the vortex cores,
          $r \lesssim a$ — hence the short range of the weak force.
\end{enumerate}
The energy barrier against reconnection is set by the quantum pressure integrated
over the overlap volume, modified by the chiral-shear coupling $\lambda$ that
controls the transverse stiffness of each vortex strand.  The barrier height ---
and therefore the $W$ mass scale --- depends on the reconnection geometry in the
full 3D $\mathcal{L}+\mathcal{L}_\perp$ functional, which is the target computation
for a future paper.

\subsection{The $W$ and $Z$ Bosons as Reconnection Quanta}

A reconnection event exchanges one unit of topological charge between two vortex
strands. The energy required to drive two vortex cores into contact against the
repulsive quantum pressure is
\begin{equation}
    E_{\rm recon} \sim P_Q \cdot (R^*)^3 \sim \frac{\hbar^2\rho_0}{m_e^2 (R^*)^2},
    \label{eq:recon}
\end{equation}
where $R^* = \xi/\alpha$ is the electron equilibrium ring radius (the
ring-scale quantity that sets the reconnection cavity, distinct from the vortex
core radius $a^* = \xi$; see Table~\ref{tab:symbols}).  Using $R^* = \xi/\alpha$
and the medium parameters from Paper~I:
\begin{equation}
    E_{\rm recon} \sim \frac{m_e}{\alpha^2} \approx \frac{0.511\,\mathrm{MeV}}{(1/137)^2}
    \approx 9600\,\mathrm{MeV} \approx 9.6\,\mathrm{GeV}.
    \label{eq:W_scale}
\end{equation}
The observed $W$ boson mass is $80.4\,\mathrm{GeV}$.  Our estimate is a factor
${\sim}8$ below this value.  The reconnection barrier energy depends on both the
quantum-pressure repulsion and the chiral-shear coupling $\lambda$ from Paper~I;
the geometric form factor of the full 3D barrier is the remaining target, and a
first-principles minimisation of the reconnection path in the
$\mathcal{L}+\mathcal{L}_\perp$ functional is the central open problem for the
weak sector.  The correct order of magnitude --- tens of GeV --- emerges naturally
from the medium parameters, without any separate gauge symmetry or Higgs
mechanism.  The Rayleigh-jet cap formula (\S\ref{sec:W_rayleigh}) refines this
to $m_W \approx 78.9\,\mathrm{GeV}$, within $1.8\%$ of the observed value.
The Higgs sector, including the relation between the observed
$m_H \approx 125\,\mathrm{GeV}$ and the medium parameters, is treated separately
in \S\ref{sec:higgs} below.

\paragraph{The $Z$ boson and the Weinberg angle.}
The $Z$ boson is the neutral reconnection channel, in which amplitude-mode and
phase-mode caps mix.  The SSV cap energy formula $m_{\rm cap} = \pi R_{\rm cap}^2
m_e c^2$ (with $P_0 = \xi = 1$) gives $m \propto R^2$, so the Standard Model
tree-level relation $m_W = m_Z\cos\theta_W$ translates directly to
\begin{equation}
  \frac{R_{\rm cap,W}}{R_{\rm cap,Z}} = \sqrt{\cos\theta_W}.
  \label{eq:RZ_RW}
\end{equation}
Using the PDG Weinberg angle as input, the tree-level $Z$ cap radius is
$R_{\rm cap,Z} = R_{\rm cap,W}/\sqrt{\cos\theta_W} \approx 236.8\,\xi$,
giving $m_Z^{\rm tree} \approx 90.0\,\mathrm{GeV}$ ($1.3\%$ below the observed
$91.2\,\mathrm{GeV}$) — the same fractional deficit as for $m_W$.  The Z-cap
bending stiffness required by the equilibrium cubic~\eqref{eq:cap_cubic} is
$\lambda_{\rm bend,Z}/\lambda_{\rm bend,W} = 1.236$, consistent with the
$\tau\to 0$ prediction $(R_{\rm cap,Z}/R_{\rm cap,W})^3 = 1/\cos^{3/2}\theta_W
\approx 1.218$ at the $1.5\%$ level.

The Weinberg angle itself remains an open gapbox: SSV does not yet predict
$\theta_W$ from first principles.  A scaling estimate at $R_{\rm cap,W}$ gives
$\tan^2\theta_W \sim \tau\alpha^2 R_{\rm cap,W} \approx 0.20$, within a factor
$1.5$ of the observed value $0.30$, consistent with a phase-to-amplitude
stiffness ratio at the cap boundary.  As a numerical remark,
$\sin^2\theta_W \approx \phi/7 = 0.23115$ reproduces the PDG value
$0.23122$ to $0.03\%$; if this coincidence is not accidental, it would point
to a golden-ratio quantisation of the $\mathrm{SU}(2)/\mathrm{U}(1)$ mixing
angle at the reconnection saddle.

\subsection{The Higgs as Canvas: The Amplitude Mode of the Vacuum}
\label{sec:higgs}

In SSV the Higgs is not a separately postulated field.  It is the amplitude
mode of the same complex order parameter $\Psi$ that produces every other
particle in the framework.  This section distinguishes two senses in which
``the Higgs'' appears in SSV, and locates the genuine hierarchy puzzle in the
right place.

\paragraph{Two modes of the same medium.}
Linearising $\Psi = (\sqrt{\rho_0}+\delta\rho/2\sqrt{\rho_0})\,e^{i\theta_0+i\delta\theta}$
around the saturated background gives two distinct small oscillations, exactly
as in laboratory superfluids:
\begin{itemize}
    \item \textbf{Goldstone mode} ($\delta\theta$): a phase oscillation at fixed
          density.  In the literature-consistent reading discussed in
          \S\ref{sec:EM_open_issue}, this is the leading candidate for the
          observed photon and propagates at the medium sound speed $c$.
    \item \textbf{Higgs mode} ($\delta\rho$): an oscillation of the condensate
          \emph{density} itself.  Massive, with gap
          $\Delta = c\sqrt{2 m_0 \lambda \rho_0}/\hbar$.  This is the amplitude
          mode of the medium.
\end{itemize}
The two modes are governed by the same Lagrangian; the only distinction is
which component of $\Psi$ oscillates.  The amplitude mode is gapped because
compressing the saturated medium against the background pressure $P_0$ costs
energy: the medium resists being anything other than what it is.  This gap is a
\emph{structural} property of the LogSE Lagrangian --- it exists at every point
in the medium and is not associated with any localised event.

\paragraph{Defects vs.\ canvas.}
Every other particle in the framework is a \emph{defect} in the medium: the
electron is a toroidal hole, the proton a knotted vortex, the pion a
vortex--antivortex link, the $W$ a transient open cavity.  All of these are
absences or distortions of the saturated background.  The amplitude mode is
structurally different: it is the saturated background itself, set into
oscillation.  It is not a defect added to the canvas --- it is the canvas,
momentarily made visible.

\paragraph{Two readings of ``the Higgs'' in SSV.}
The word ``Higgs'' conflates two distinct things that must be kept separate:
\begin{enumerate}[label=(\roman*)]
    \item \textbf{The amplitude mode (structural).}  The degree of freedom
          $\delta\rho$ is present everywhere in the medium; its gap
          $\Delta \simeq m_H c^2 \approx 125\,\mathrm{GeV}$ is a property of
          the Lagrangian parameters $\{b, \lambda, \rho_0\}$.  This is not a
          particle; it is a channel of the medium, analogous to longitudinal
          sound in an ordinary fluid.  The analogous amplitude mode in cold-atom
          condensates was first directly observed in 2012~\cite{Endres2012}.

    \item \textbf{The 125\,GeV denting event (derived).}  The resonance observed
          at the LHC is a localised compression of $|\Psi|$ driven deep enough
          ($|\Psi|\to 0$ over a volume $V^*$) that the amplitude mode rings
          briefly before relaxing.  Its energy is
          \begin{equation}
              E_H = P_0\,V^*,
              \label{eq:Higgs_dent}
          \end{equation}
          where $P_0$ is the saturation pressure and $V^*$ is the dent geometry
          set by the same 3D minimisation that determines the $W$ end-cap volume
          $V_{\rm cap}$ (\S\ref{sec:W_rayleigh}).  The number $125\,\mathrm{GeV}$
          is therefore a \emph{derived} quantity calculable from $\{P_0, \alpha,
          \text{dent geometry}\}$, with no independent fundamental status.
\end{enumerate}
Neither reading identifies $m_H$ with a structural length scale of the medium's
substrate.  The quantity
\begin{equation}
    \xi_{\rm EW} = \frac{\hbar}{m_H c} \approx 1.6\times 10^{-18}\,\mathrm{m}
    \label{eq:EW_scale}
\end{equation}
is the de~Broglie wavelength of the amplitude-mode quantum; it is \emph{not}
the grain size of the medium.  The medium's domain-of-validity scale is
$\ell_{\rm grain} \equiv a_p \approx 2.10\times 10^{-16}\,\mathrm{m}$
(Table~\ref{tab:symbols}), set by the heaviest stable defect, not by the Higgs
denting event.

\paragraph{Yukawa coupling as canvas displacement.}
The Standard Model assigns each massive particle a Yukawa coupling
$g_f = m_f/v$ to the Higgs field.  In SSV this relation has a transparent
mechanical meaning.  An amplitude oscillation $\delta\rho$ couples to a defect
in proportion to how much canvas the defect displaces from saturation; a
heavier defect (larger ring, wider core, deeper density depression) interacts
more strongly with $\delta\rho$.  The proportionality $g \propto m$ is therefore
not an empirical fact requiring a Yukawa sector --- it is the geometric
statement that bigger holes in the canvas couple more strongly to oscillations
of the canvas itself.

\paragraph{The genuine hierarchy puzzle: the defect spectrum.}
The Standard Model hierarchy problem (why $m_H \ll m_{\rm Planck}$ given
radiative corrections?) dissolves in SSV: there is no quadratically divergent
radiative correction to $m_H$ because $\ell_{\rm grain}$, not an external
regulator, is the natural cutoff of the framework.

The \emph{genuine} hierarchy question in SSV is different and sharper.  The
medium supports exactly two classes of stable defect: the toroidal vortex ring
(electron, $m_e$) and the knotted breather (proton, $m_p$).  Their spatial
scales are
\begin{align}
    \text{electron ring radius:} &\quad R^* = \xi/\alpha \approx 137\,\xi, \\
    \text{proton core radius:}   &\quad a_p = \xi\cdot(m_e/m_p) \approx \xi/1836,
\end{align}
so the electron's major radius exceeds the proton's core by a factor of
$137\times 1836 \approx 2.5\times 10^5$.  Both structures arise from the same
LogSE$+$chiral functional applied to different topologies.  The framework does
not yet predict this ratio from first principles: it requires the full 3D
minimisation of the vortex-ring and knotted-breather energy functionals over
$\mathcal{L}+\mathcal{L}_\perp$.

\paragraph{Structural composite mass ratio.}
The defect spectrum hierarchy of the previous paragraph is the SSV
counterpart of the Standard Model Yukawa hierarchy: the ratio
$m_p/m_e \approx 1836$ is determined by the relative scaling of two
topology classes (the toroidal vortex ring and the knotted breather)
under the same $\mathcal{L}+\mathcal{L}_\perp$ functional.  Although the
\emph{individual} masses are at mixed status (Paper~I tags $m_e$ and
$m_{\pi^\pm}$ as derived and $m_p$ as candidate; the proton-mass status
is tracked under issue~\#30), the
framework already supplies one composite \emph{mass-ratio} prediction
whose structural form is fixed without further computation: the
proton-to-pion mass ratio.

From the Paper~I identifications
\begin{equation}
  m_{\pi^\pm} \;=\; 2\mu_0, \qquad
  m_p \;=\; N_Y\,F\,\mu_0,
  \label{eq:proton-pion-individual}
\end{equation}
where $\mu_0 \equiv m_e/\alpha \approx 70\,\text{MeV}$ is the
electron-derived flux-tube mass scale, $N_Y$ is the topological crossing
factor of the trefoil Y-junction, and $F$ is the 3D breather form
factor, the ratio
\begin{equation}
  \boxed{\;
    \frac{m_p}{m_{\pi^\pm}} \;=\; \frac{N_Y\,F}{2}\;
  }
  \label{eq:mp-over-mpi}
\end{equation}
follows immediately, with the empirical input $\mu_0$ cancelling.  At
the current candidate-grade values $N_Y \approx 3.007$, $F \approx 4.47$,
\begin{equation}
  \frac{m_p}{m_{\pi^\pm}} \;\approx\; 6.72,
  \label{eq:mp-over-mpi-numeric}
\end{equation}
within $0.1\%$ of the observed ratio
$m_p/m_{\pi^\pm} = 938.272/139.570 \approx 6.722$.

\paragraph{Cross-link to $m_p/m_e$.}
The same product $N_Y\,F$ also controls the proton--electron ratio
through the definition $\mu_0 = m_e/\alpha$.  Eliminating $\mu_0$ from
Eq.~\eqref{eq:proton-pion-individual} gives the identity
\begin{equation}
  2\,\frac{m_p}{m_{\pi^\pm}} \;=\; \alpha\,\frac{m_p}{m_e}
                              \;=\; N_Y\,F.
  \label{eq:hierarchy-identity}
\end{equation}
This is a sharper statement than ``the spectrum has a hierarchy''.
Three experimentally distinct dimensionless ratios ($m_p/m_{\pi^\pm}$,
$m_p/m_e$, and $\alpha^{-1}$) reduce to one structural quantity $N_Y\,F$
in the framework.  Numerically, $2\times 6.722 = 13.444$ and
$(1/137.036)\times 1836.15 = 13.398$ --- agreement at the $0.3\%$ level,
matching the present candidate-grade precision of $N_Y$ and $F$.
Eq.~\eqref{eq:hierarchy-identity} is therefore an internal consistency
\emph{prediction}: any future first-principles computation of $N_Y$ and
$F$ must reproduce both the $1836$ and $6.72$ ratios with the same
single number.

\begin{resultbox}[Structural composite mass ratio]
\begin{itemize}\setlength\itemsep{2pt}
  \item \textbf{Structural form:}
        $m_p/m_{\pi^\pm} = N_Y F/2$ (Eq.~\eqref{eq:mp-over-mpi}).
  \item \textbf{Numerical value:}
        $6.72$ at present candidate-grade values, $0.1\%$ below the
        observed $6.722$.
  \item \textbf{Consistency identity:}
        $2(m_p/m_{\pi^\pm}) = \alpha(m_p/m_e) = N_Y F$
        (Eq.~\eqref{eq:hierarchy-identity}), reducing three observed
        dimensionless ratios to a single framework quantity at the
        $0.3\%$ level.
  \item \textbf{Status:} structural --- the form is fixed by the
        Paper~I topology assignments (electron $=$ toroidal vortex of
        radius $\xi/\alpha$; pion $=$ two-winding link; proton $=$
        trefoil Y-junction breather).  The numerical values of $N_Y$
        and $F$ inherit the candidate-grade status of the proton
        derivation (Paper~I).  Promotion of the ratio from structural
        to derived requires the dedicated $N_Y$ cross-grid sweep,
        geometric derivation of the cutoff $R\approx 1.18\,\xi$, and
        recovery of the converged values from at least one
        non-trefoil initial condition; the three concrete runs are
        itemised in
        \texttt{papers/SSV-I/static-3d-closure-status-2026-05-28.md}
        and tracked under issue~\#30.
\end{itemize}
\end{resultbox}

\begin{gapbox}[Status (2026-05): $6.72$ and the $N_Y F$ identity inherit the F calibration]
The structural form $m_p/m_{\pi^\pm} = N_Y F/2$
(Eq.~\eqref{eq:mp-over-mpi}) is unchanged.  Its numerical value $6.72$
and the consistency identity $2(m_p/m_{\pi^\pm}) = \alpha(m_p/m_e) =
N_Y F \approx 13.44$ both inherit the Paper~I §Proton gapbox: $F$ is
a cutoff-dependent calibration ($\mathrm{d}\ln F/\mathrm{d}\ln R
\approx -0.94$, $19\%$ variation between $R=1.18\,\xi$ and $R=1.5\,\xi$),
not yet a cutoff-independent first-principles derivation.  The
\texttt{q\_p\_two\_factor\_*} scripts that attempted to calibrate the
cutoff away are now in \texttt{src/\_fitted\_quarantine/} after the
Path~B null exposed how readily fitted inputs flatter derived
results.  The $0.1\%$--$0.3\%$ agreements quoted above are therefore
consistency at the published cutoff, not a finished derivation;
promotion paths are the same as for the proton mass (derive $R$
geometrically, or show $N_Y\!\cdot\!F$ is cutoff-invariant).
Tracking: issues~\#30 (closure) and~\#66 (cleanup record).
\end{gapbox}

\begin{gapbox}[Residual gap: the broader composite spectrum]
The structural ratio above covers the lightest meson against the
lightest baryon, plus the $m_p/m_e$ consistency identity.  Mass ratios
within the larger spectrum --- $m_K/m_{\pi^\pm}$ (kaon/pion),
$m_\rho/m_{\pi^\pm}$ (vector meson/pion), $m_\Delta/m_p$ (delta
resonance/nucleon), and the heavier-meson ladder --- depend on
$\alpha$-harmonic placements (Paper~I Table) that are themselves
candidate-grade.  A structural derivation for this larger sector
requires identifying the specific surface-mode or Kelvin-wave
excitation of the proton-trefoil breather and the pion-vortex link
that each heavier state corresponds to.  This is the natural
continuation of \S\ref{sec:leptons} on the charged-lepton
breathing-mode programme, applied to the baryon and meson sectors.
\end{gapbox}

\paragraph{Connection to the $W$ end-caps.}
The Higgs picture closes a loop in the weak-sector argument.  The Rayleigh-jet
estimate of the $W$ mass (\S\ref{sec:W_rayleigh}) depends on the cap term
$E_{\rm caps} \sim P_0 \cdot V_{\rm cap}$, the cost of holding open a region
where $|\Psi| \to 0$ against the saturation pressure $P_0$.  This is the same
$P_0$ that enters the Higgs denting energy~\eqref{eq:Higgs_dent}: opening a
$W$ end-cap is driving the amplitude mode of the canvas to its full extent
(collapse of the density to zero) over a smaller volume $V_{\rm cap} < V^*$.
The $W$ and the Higgs are not two separate sectors with a coupling between them
--- they are two different denting events of the same medium, differing only in
the geometry and depth of the cavity.

\subsection{The $W$ Boson as a Rayleigh Reconnection Jet}
\label{sec:W_rayleigh}

The estimate~\eqref{eq:W_scale} treats the $W$ boson as a point-like core-overlap
event and misses the \emph{geometry of the reconnection jet} — the dominant
contribution to the mass.  A classical analogue makes the structure explicit.

\paragraph{Physical picture: the Rayleigh jet.}
When a liquid droplet strikes a free surface it drives a cavity into the bulk; the
cavity collapses axisymmetrically, focussing kinetic energy along the symmetry axis
and ejecting a thin upward jet.  The jet thins until its local radius reaches the
capillary length, whereupon surface tension severs it and a new droplet pinches off.
The energy stored at pinch-off is set by the \emph{aspect ratio at the pinch point}
— a purely geometric quantity independent of the initial cavity volume.

The SSV analogue maps precisely onto this process in three stages:
\begin{enumerate}[label=(\roman*)]
  \item \textbf{Cavity formation.}  Two vortex defects approach within one ring
        radius $R^*\sim\xi/\alpha$.  Quantum pressure suppresses $|\Psi|$ in the
        overlap region, opening a local cavity in the condensate — the amplitude
        channel activates.
  \item \textbf{Jet formation.}  The cavity collapses radially, driving an elongated
        vortex tube — an axially extended torus — along the reconnection axis.
        This tube carries exactly three polarisation degrees of freedom: two
        transverse oscillations of the tube wall, and one longitudinal mode from
        the stretching of the tube itself.  The photon, a pure transverse
        phase-gradient wave, never activates the amplitude channel and therefore
        has no longitudinal mode and no end-caps.
  \item \textbf{Pinch-off.}  The tube thins until its minor radius reaches $\xi$,
        where quantum pressure dominates surface tension.  The tube severs,
        depositing its energy into the outgoing reconfigured defects — the weak
        decay products.
\end{enumerate}
The $W$ boson is the \emph{elongated vortex tube at the moment of pinch-off}: a
transitional topology that exists only long enough to rewrite the identities of the
interacting defects.

\paragraph{Energy of the jet at pinch-off.}
An elongated torus of length $L^*$ and minor radius $a^*\sim\xi$ (the vortex
\emph{core} radius, not the ring radius $R^*$) has three energy contributions:
\begin{align}
  E_{\rm core}    &\sim \rho_0\,\xi^2\,L^*,              \label{eq:Ecore}\\
  E_{\rm tension} &\sim \rho_0\kappa_0^2\,L^*/a^{*2},    \label{eq:Etension}\\
  E_{\rm caps}    &\sim P_0\,V_{\rm cap}.                 \label{eq:Ecaps}
\end{align}
The first two terms are balanced by minimising over $L^*$, giving $L^*\sim\xi$ —
the tube is roughly isotropic at pinch-off.  The dominant contribution is the
end-cap term, which has a qualitatively different physical origin from the other
two: it is the cost of holding open a region of fully suppressed condensate
($|\Psi|\to 0$) against the enormous background saturation pressure $P_0\gg\rho_0
c^2$.  Every other energy in the problem is set by $P_Q$ or the vortex tension
$\rho_0\kappa_0^2$; the cap cost is set by $P_0$ itself, making it the dominant
term.

The scalar estimate~\eqref{eq:W_scale} accounted only for $E_{\rm approach}\sim
P_Q\cdot a^3$ — the repulsive barrier the incoming defects must overcome to
\emph{initiate} reconnection.  The cap cost $E_{\rm caps}\sim P_0\cdot V_{\rm cap}$
is a separate, larger term that was not included.  Since $P_0\gg P_Q$, this
naturally accounts for the missing factor of $\sim 8$.

\paragraph{The cap radius and the golden ratio.}
The cap volume $V_{\rm cap} = \pi R_{\rm cap}^2\,\xi$ is set by the radius
$R_{\rm cap}$ over which $|\Psi|$ is suppressed at each end of the W tube.
Working backwards from the observed $m_W = 80.36\,\mathrm{GeV}$ fixes
$R_{\rm cap} \approx 1.70\,\xi/\alpha$.  This is tantalizingly close to
\begin{equation}
  R_{\rm cap} = \phi\,\frac{\xi}{\alpha}, \qquad
  \phi = \tfrac{1+\sqrt{5}}{2} \approx 1.618,
  \label{eq:Rcap_phi}
\end{equation}
the golden ratio times the electron ring radius $R^* = \xi/\alpha$.
Substituting~\eqref{eq:Rcap_phi} into $E_{\rm caps}$ with $P_0\sim m_e/(2\xi^3)$:
\begin{equation}
  m_W \approx \frac{\pi\phi^2\,m_e}{\alpha^2}
       = \pi\phi^2 \times 9.60\,\mathrm{GeV}
       \approx 78.9\,\mathrm{GeV},
  \label{eq:mW_phi}
\end{equation}
within $1.8\%$ of the observed value $80.36\,\mathrm{GeV}$ (PDG 2024).

\paragraph{Cap equilibrium and the chiral-shear bending stiffness.}
The cap is not a free surface: it is the boundary of a closed vortex ring held
in mechanical equilibrium by three competing forces.  Let $R$ denote the cap
radius.  In SSV natural units ($P_0 = \xi = 1$) the total cap energy is
\begin{equation}
  E(R) = \pi R^2 + 2\pi\tau R + \frac{2\pi\lambda_{\rm bend}}{R},
  \label{eq:Ecap_equilibrium}
\end{equation}
where $\pi R^2$ is the surface-pressure cost (area $\times P_0$),
$2\pi\tau R$ is the vortex line-tension perimeter energy, and
$2\pi\lambda_{\rm bend}/R$ is the chiral-shear bending resistance that stiffens
the ring against collapse.  Setting $dE/dR = 0$ gives the equilibrium cubic
\begin{equation}
  R^3 + \tau\,R^2 = \lambda_{\rm bend}.
  \label{eq:cap_cubic}
\end{equation}
The vortex line tension is computed numerically from the planar
$\mathrm{LogSE}$ profile (\texttt{vortex\_cap\_mass.py}):
$\tau = 17.0\,\xi$ (core integral $+ 2\pi\ln(R_{\rm cap}/r_{\rm max})$ tail).
In the $\tau\to 0$ limit, setting $x = \alpha R/\xi$,
equation~\eqref{eq:cap_cubic} reduces to $x^3 = \alpha^3\lambda_{\rm bend}$.
The unique positive solution $x = \phi$ (the golden ratio) requires
\begin{equation}
  \lambda_{\rm bend} = \frac{\phi^3}{\alpha^3}\,\xi^3
                     \approx 1.09\times 10^7\,\xi^3.
  \label{eq:lambda_bend_required}
\end{equation}
Evaluated at the physical $\tau = 17\,\xi$, the cubic places $R_{\rm eq}$ at
$\phi/\alpha$ to within the $7.7\%$ line-tension correction
(equation~\eqref{eq:cap_cubic} with $R=\phi/\alpha$ gives $\lambda_{\rm bend} =
1.17\times 10^7\,\xi^3$, vs the $\tau\to 0$ target $1.09\times 10^7\,\xi^3$).
Thus $R_{\rm cap} = \phi/\alpha$ is the unique equilibrium radius once
$\lambda_{\rm bend} = \phi^3/\alpha^3$ is established from the chiral-shear
sector.

The SSV chiral-shear Lagrangian $\mathcal{L}_\perp = (\lambda_\perp/2)\int|\nabla\times
\mathbf{j}|^2\,d^3r$ provides a natural coupling $\lambda_\perp = \alpha^{-2}$
(from $c_\perp = \alpha c$ at $k = 1/\xi$).  Numerical integration of the
core-bending correction to $\lambda_\perp$ over the planar vortex profile
(\texttt{lperp\_core\_integral.py}) yields $\lambda_{\rm bend}^{\rm local} \approx
4.7\times 10^4\,\xi^3$ — a factor $232$ below the required value.  The local
curvature correction to $\mathcal{L}_\perp$ is therefore insufficient; the
bending stiffness $\lambda_{\rm bend} = \phi^3/\alpha^3$ must arise from a
\emph{non-local} mechanism.  The two leading candidates are: (a) quantisation
of the chiral-shear energy of the Seifert surface of the trefoil (a
non-perturbative topological contribution); and (b) a Casimir-like vacuum
energy of the chiral mode between the cap boundary and the reconnection
saddle.

\paragraph{Why the cap geometry is not directly accessible.}
The cap geometry is determined by the \emph{dynamics} of the reconnection, not
a static equilibrium, and cannot be observed directly: the event occurs at
sub-$\xi$ scales, lasts $\sim\xi/c\sim 10^{-25}$\,s, and requires energies of
order $P_0\,\xi^3$ to produce.  The $W$ mass itself is the only empirical
constraint on $R_{\rm cap}$.  Equation~\eqref{eq:mW_phi} holds as a
conjecture: $R_{\rm cap} = \phi/\alpha$ follows analytically from the
equilibrium cubic \eqref{eq:cap_cubic} once $\lambda_{\rm bend} = \phi^3/\alpha^3$
is established, but that identification requires deriving the non-local
bending stiffness from the SSV Lagrangian — the remaining open step.

\paragraph{Why the $W$ is massive and the photon is not.}
The photon lives entirely in the $\delta\theta$ channel and never suppresses
$|\Psi|$ — it costs nothing against $P_0$.  The $W$ tube \emph{requires}
$|\Psi|\to 0$ at its end-caps: topology change cannot occur without locally
dissolving the condensate, and the vacuum pressure $P_0$ must be held at bay for
the duration of the event.  Mass in SSV is fundamentally the cost of displacing
the vacuum from itself — holding open a region where $|\Psi|<1$ against $P_0$.
The W end-caps are fully open wounds in the condensate, and $P_0$ is what closes
them the instant the topology rewrite is complete.
\emph{The photon has no end-caps; the $W$ boson has nothing but end-caps.}

\subsection{Parity Violation}
\label{sec:parity}

Parity violation in the weak interaction has a natural origin in the chirality of
vortex reconnection. In a chiral superfluid (one whose ground state has a preferred
handedness, which ours does — the toroidal electron vortex carries a fixed chirality
tied to its charge), reconnections preferentially occur in configurations that
preserve the global handedness. This selects left-handed neutrinos (defined
hydrodynamically in \S\ref{sec:neutrino} below as torsional pulses with a
specific twist orientation) over right-handed ones, and produces the observed
$(V-A)$ structure of the weak current without additional assumptions.

\subsection{CP and the Strong CP Problem in a Medium That Remembers}
\label{sec:CP}

The parity-violation argument of \S\ref{sec:parity} relied on the medium having
a preferred handedness.  This raises the immediate question of whether the
combined operation CP --- charge conjugation paired with parity --- restores
the symmetry, and what the framework predicts for the small but observed CP
violation in the kaon and $B$ sectors.  The answer requires confronting a
distinction the Standard Model does not need to make: the difference between
\emph{symmetries of an equation} and \emph{symmetries of a process}.

\paragraph{Two different things called ``symmetry''.}
In a quantum field theory built around a featureless vacuum, the two coincide:
if the Lagrangian is invariant under a transformation $T$, then $T$ maps every
physical process onto another physical process governed by the same dynamics.
SSV does not have a featureless vacuum.  The medium is saturated, supports
defects, and accumulates a record of every reconnection event in the form of
its evolving topological tangle and the acoustic energy it has radiated
(Paper~III treats this explicitly as the origin of irreversible time).  In
such a medium, two distinct notions of ``symmetry'' must be carefully separated:
\begin{itemize}
    \item \textbf{Formal symmetry of the equations.} A transformation $T$ is a
          symmetry of the linearised $\mathcal{L}+\mathcal{L}_\perp$ functional
          if it leaves the equations of motion invariant.  By this criterion,
          the dynamics of the saturated medium admits formal C, P, T, CP, CT,
          PT, and CPT operations, exactly as in the Standard Model.
    \item \textbf{Symmetry of physical processes.} A transformation is a
          symmetry of an actual physical process only if applying it produces
          another process consistent with the medium's accumulated state at
          the time the transformation is applied.  By this stricter criterion,
          \emph{no} discrete symmetry survives in SSV: every physical process
          deposits acoustic energy, leaves a chirality-marked imprint in the
          tangle, and modifies the medium that any putative symmetry partner
          would have to propagate through.
\end{itemize}

The Standard Model identifies these two notions because it works around a
vacuum that does not record.  SSV cannot.  The medium remembers, and memory
breaks every discrete symmetry at the level of physical processes, regardless
of what the formal Lagrangian says.

\paragraph{Parity, charge conjugation, and CP, restated.}
In light of the distinction above:
\begin{itemize}
    \item \textbf{P} is broken \emph{at the Lagrangian level} by the chiral-shear
          coupling $\lambda$ (\S\ref{sec:parity}).  This is a strong form of
          breaking, comparable to the Standard Model's electroweak parity
          violation.
    \item \textbf{C} is broken at the Lagrangian level too, since the medium's
          chirality treats $n=+1$ and $n=-1$ vortices differently --- the
          antivortex is not the same object as the vortex with the medium's
          handedness reversed.
    \item \textbf{CP} is a formal symmetry of the linearised dynamics
          (simultaneous winding and chirality flip), but \emph{not} a symmetry
          of any physical process, because the process leaves a record that
          the CP-flipped partner would have to carry as well --- and the
          flipped partner does not, because it propagates through a medium
          that has not accumulated the same wake.
\end{itemize}
The third bullet is the key shift from the Standard Model framing.  CP
violation in SSV is not a small departure from an underlying CP-exact
dynamics, brought in by specific complex phases in the fermion mass matrix.
It is the visible portion of an asymmetry that is structurally present in
\emph{every} process, made measurable only when reconnection geometry happens
to amplify it past the threshold of detection.

\paragraph{Where the visible CP violation comes from.}
The kaon $\varepsilon$-parameter ($\sim 10^{-3}$) and the larger CP asymmetries
in $B$-meson decays are not exotic phenomena requiring a special CP-violating
sector.  They are the systematic component of the much larger ensemble of
CP-asymmetric reconnection events that occur continuously in the medium ---
the part that does not average to zero because the Y-junction packing
geometry of the parent configuration biases one chirality of outcome over
its CP-conjugate.  The remainder averages out at experimental resolution and
appears as ``CP-symmetric noise'' --- but it is not truly absent, only
unresolved.  This is structurally identical to fluctuation--dissipation in
classical statistical mechanics: the systematic asymmetry one measures (the
analogue of dissipation) is the visible part of fluctuations that are
asymmetric everywhere (the analogue of correlated noise).

\paragraph{The strong CP problem dissolves --- but for a deeper reason.}
The Standard Model strong CP problem is the puzzle of why
$|\theta_{\rm QCD}| \lesssim 10^{-10}$ when naturalness suggests $\theta \sim
\mathcal{O}(1)$.  The parameter $\theta_{\rm QCD}$ measures the angle between
the QCD vacuum and a CP-flipped reference vacuum; its smallness is mysterious
because both vacua are equally available in the theory.

In SSV the strong sector is vacuum surface tension along flux-tube vortex
segments (\S\ref{sec:strong}); there is no separate QCD vacuum.  But this is
not the deepest reason the problem dissolves.  The deepest reason is that a
$\theta$-angle measures deviation from a CP-symmetric reference state, and in
a medium that records, no CP-symmetric reference state exists.  The medium at
any given moment carries a specific tangle history; the formally CP-flipped
medium would carry a different history; there is no third state symmetric
between them against which to measure $\theta$.  The Standard Model strong CP
problem is well-posed only because the SM vacuum is featureless and admits
unique CP-symmetric and CP-flipped versions.  In SSV the question
$\theta_{\rm QCD} = ?$ is not answered as zero; it is dismissed as
ill-defined.

The neutron electric dipole moment is therefore predicted to be zero at
leading order in the same sense that the cosmic microwave background is
isotropic at leading order: the leading-order calculation gives zero because
the leading-order configuration is taken to be statistically isotropic, but
the actual medium carries fluctuations and any specific measurement returns a
small nonzero residue.  The neutron EDM, like kaon CP violation, is a
fluctuation-resolved asymmetry, not a tuning of a vacuum angle.

\paragraph{The matter--antimatter asymmetry as recorded chirality.}
Sakharov's three conditions for a baryon asymmetry --- C and CP violation,
baryon-number violation, and departure from thermal equilibrium --- are all
present in SSV, but the framework reads them differently than the Standard
Model does.  The medium's chirality is not a small CP-violating term added to
a CP-symmetric dynamics; it is a property the medium has from its inception.
Vortex reconnections do not accidentally violate baryon number; they are the
mechanism by which topological charge changes.  The early universe was not
near thermal equilibrium that was disturbed by an out-of-equilibrium phase
transition; it was the seeding event of the medium's record itself.

The observed baryon-to-photon ratio $n_B/n_\gamma \sim 6\times 10^{-10}$ is
therefore not a number to be derived from a small CP-violating phase in an
otherwise symmetric theory.  It is the surviving recorded chirality of the
seeding event: the part of the initial asymmetry that crystallised into
stable defects (matter) rather than annihilating against its partner (the
near-equally-stable antimatter component).  Why this number takes the value
it does is open and properly belongs to Paper~VIII's treatment of cosmogony, but
the framework's reading of the question is sharp: the universe has matter
because the medium remembers having been chirally seeded, and the small
fraction $n_B/n_\gamma$ measures how much of that initial chirality survived
the early annihilation epoch.

\subsubsection*{Reconnection-memory derivation: the neutron EDM at leading
order}
\label{sec:neutron-edm}

The paragraphs above identified CP violation in SSV with the cumulative
chirality bias of past reconnections --- recorded in the medium and made
visible when reconnection geometry amplifies it past detection threshold.
This subsection develops the mechanism into a structural prediction for one
CP observable: the neutron electric dipole moment.  The derivation takes the
neutron in its Paper~I surface-locked composite form
($\mathcal{N}=\mathcal{T}\cup_{\Sigma_p}\mathcal{E}$, the proton trefoil
$\mathcal{T}$ glued to a captured electron torus $\mathcal{E}$ at the
breather skin $\Sigma_p$) and asks what permanent dipole moment the medium's
chirality memory induces in this geometry at rest.

\paragraph{The chirality memory field.}
Define the local chirality memory density
\begin{equation}
  \chi(\mathbf{x},t)
    \;=\; \sum_{\text{events }i}
            \sigma_i\, K\!\big(\mathbf{x}-\mathbf{x}_i,\,t-t_i\big),
  \label{eq:chi-def}
\end{equation}
where $\sigma_i=\pm 1$ is the handedness of the $i$-th reconnection event
(with respect to the medium's preferred chirality of \S\ref{sec:parity})
and $K$ is the medium's memory kernel --- a slowly decaying weight whose
support is set by the propagation reach of the acoustic and chiral-shear
modes emitted at the event.  $\chi$ is a dimensionless scalar; its spatial
average over any region is the surviving chirality fraction of all
reconnection events whose memory still reaches that region.  The cosmic
average $\langle\chi\rangle_{\rm cosmo}$ is the framework's reading of the
baryon-to-photon ratio:
$\langle\chi\rangle_{\rm cosmo} = n_B/n_\gamma \approx 6\times 10^{-10}$
(\S\ref{sec:CP}, ``matter--antimatter asymmetry as recorded chirality'').

\paragraph{Coupling to the neutron lock.}
The captured electron torus $\mathcal{E}$ on $\Sigma_p$ is held in
equilibrium by the chiral-shear restoring potential
$V_\perp = (\lambda_\perp/2)\,|\nabla\times\mathbf{j}|^2$ acting on the
trefoil's surface current.  A non-zero local $\chi$ couples to the same
chiral-shear sector and biases the equilibrium position of $\mathcal{E}$
on $\Sigma_p$.  To leading order in $\chi$, the shift of the captured
torus's centroid relative to the breather centre is
\begin{equation}
  \delta\mathbf{r}_{\mathcal{E}}
    \;=\; \kappa_\chi\,\langle\chi\rangle_{R_n}\,R_n\,\hat{\mathbf{n}},
  \label{eq:dr-edm}
\end{equation}
where $R_n\approx 0.84\,\text{fm}$ is the neutron radius,
$\langle\chi\rangle_{R_n}$ is the average of $\chi$ over the neutron's
interior volume, $\hat{\mathbf{n}}$ is the gradient direction of the
chirality memory at the neutron, and $\kappa_\chi$ is a dimensionless
coupling fixed by the chiral-shear Lagrangian.  The lock stiffness
$\partial^2V_\perp/\partial r^2$ at the surface-mode minimum places
$\kappa_\chi$ at $\mathcal{O}(\alpha^n)$ for some $n\in\{1,2\}$ depending
on whether the chiral-shear vertex enters linearly or at the one-loop
self-energy of the surface mode.  The numerical value of $\kappa_\chi$
requires the converged 3D $\mathcal{L}+\mathcal{L}_\perp$
reconnection-barrier minimisation (current cross-grid status documented
in
\texttt{papers/SSV-II/results/reconnection/validation-refinement-sweeps-reconnection-2026-05-28.md};
tracked under issue~\#30); the \emph{structural form} of
Eq.~\eqref{eq:dr-edm} does not.

\paragraph{Identifying $\langle\chi\rangle_{R_n}$.}
The neutron at rest in a quiet laboratory has no internal flavour
oscillation channel; the relevant $\chi$ at its interior is dominated by
the cosmic-background contribution, not by the kaon-scale fluctuations
that drive $\varepsilon_K$.  This is the structural distinction between
the kaon and the neutron in this framework: the kaon's
$K^0\!\leftrightarrow\!\overline{K^0}$ oscillation continuously samples
$\chi$ at the rate of the mixing, giving
$\langle\chi\rangle_K\sim\varepsilon_K\approx 2.2\times 10^{-3}$; the
neutron has no such sampling channel, so $\langle\chi\rangle_{R_n}$ reduces
to the cosmic background
\begin{equation}
  \langle\chi\rangle_{R_n}
    \;\simeq\; \langle\chi\rangle_{\rm cosmo}
    \;=\; \frac{n_B}{n_\gamma}
    \;\approx\; 6\times 10^{-10}.
  \label{eq:chi-Rn}
\end{equation}
The choice of the cosmic-background amplitude (rather than the kaon-scale
amplitude) for the neutron is itself a structural prediction of the
framework: only objects with an internal flavour-oscillation channel
amplify $\chi$ to the $\varepsilon$-scale.

\paragraph{Leading-order prediction.}
Combining Eqs.~\eqref{eq:dr-edm} and \eqref{eq:chi-Rn},
\begin{equation}
  \boxed{\;
    d_n \;=\; e\,|\delta\mathbf{r}_{\mathcal{E}}|
        \;=\; \kappa_\chi\,e\,R_n\,\frac{n_B}{n_\gamma}\;}.
  \label{eq:dn-edm-prediction}
\end{equation}
Numerically, with $R_n\approx 0.84\,\text{fm}$ and
$n_B/n_\gamma\approx 6\times 10^{-10}$,
\begin{equation}
  d_n \;\approx\; \kappa_\chi\,\times\,5\times 10^{-23}\,e\!\cdot\!\text{cm}.
  \label{eq:dn-numeric}
\end{equation}
At the conservative natural scale $\kappa_\chi\sim\alpha^2/\pi
\approx 1.7\times 10^{-5}$, $d_n\approx 9\times 10^{-28}\,e\!\cdot\!\text{cm}$,
roughly one decade below the present experimental bound
$|d_n|<1.8\times 10^{-26}\,e\!\cdot\!\text{cm}$~\cite{Abel2020} and within
reach of the next-generation nEDM/n2EDM programmes targeting
$\sim 10^{-28}\,e\!\cdot\!\text{cm}$.  At the larger scale
$\kappa_\chi\sim\alpha/\pi$ the prediction is
$\sim 10^{-25}\,e\!\cdot\!\text{cm}$, which would already have been
observed; the present null result therefore \emph{requires}
$\kappa_\chi\lesssim 4\times 10^{-4}$ within the framework, narrowing the
window for the deferred chiral-shear computation.

\paragraph{Consequences.}
The structural form
$d_n\propto e\,R_n\,(n_B/n_\gamma)$ has three consequences:
\begin{itemize}
  \item \textbf{Two observables, one coupling.}  The neutron EDM and the
        baryon-to-photon ratio are predicted to be controlled by the same
        framework parameter $\kappa_\chi$ acting on the same chirality
        memory density $\chi$.  No separate CP-violating sector is
        invoked for either; the two distinct experimental numbers reduce
        to one in the framework.
  \item \textbf{Falsification.}  Detection of $|d_n|\gtrsim
        10^{-25}\,e\!\cdot\!\text{cm}$ falsifies the framework as
        formulated --- no natural $\kappa_\chi$ can be that large.  A
        non-detection at the n2EDM sensitivity
        $\sim 10^{-28}\,e\!\cdot\!\text{cm}$ would force
        $\kappa_\chi\lesssim 10^{-5}$, on the boundary of the natural
        $\alpha^2/\pi$ estimate.
  \item \textbf{Status.}  Equations~\eqref{eq:dn-edm-prediction}--\eqref{eq:dn-numeric}
        are \emph{structural}: the functional dependence on
        $(e,\,R_n,\,n_B/n_\gamma)$ is fixed by the framework; the
        dimensionless coupling $\kappa_\chi$ requires the converged 3D
        $\mathcal{L}+\mathcal{L}_\perp$ surface-mode calculation
        (tracked under issue~\#30).  This subsection therefore closes
        the ``reconnection-memory to CP observables'' gap for at least
        one observable; the residual gapbox below lists the
        observables still awaiting analogous derivations.
\end{itemize}

\begin{gapbox}[Residual gap: other CP observables]
The reconnection-memory mechanism has now been applied to one CP observable
(the neutron EDM, structural form fixed; numerical prefactor $\kappa_\chi$
deferred to the converged 3D reconnection-barrier minimisation,
issue~\#30).  The remaining CP-sector observables are not
addressed at this level:
\begin{itemize}
  \item the CKM phase $\delta_{\rm CKM}\approx 1.20\,\text{rad}$ as the
        cumulative chirality bias sampled by quark-flavour oscillations,
  \item the kaon $\varepsilon$-parameter and $B$-system CP asymmetries as
        the $\langle\chi\rangle$-amplified part of those oscillations'
        sampling rate,
  \item the baryon-to-photon ratio $n_B/n_\gamma$ derived independently
        from the same memory kernel (rather than taken as input as in
        Eq.~\eqref{eq:chi-Rn} above).
\end{itemize}
The structural framework for all three is the chirality memory density
$\chi$ of Eq.~\eqref{eq:chi-def}; each requires identifying the specific
oscillation channel that samples $\chi$ and computing the sampling rate
against the same medium kernel $K$.  These derivations are deferred to a
sharper treatment of the chirality memory dynamics --- the same
$\mathcal{L}+\mathcal{L}_\perp$ minimisation that fixes $\kappa_\chi$
above is the natural starting point.
\end{gapbox}

\paragraph{Forward connections.}
This subsection foreshadows two themes developed in later papers.  The
distinction between formal symmetries of equations and symmetries of physical
processes is the thermodynamic content of Paper~III: irreversibility is not
something that emerges from a reversible substrate, it is the consequence of
a medium that records, and the same logic that breaks T-symmetry at the
process level breaks CP at the process level.  The cosmological reading of
the baryon asymmetry as recorded chirality of the seeding event connects to
Paper~VIII's treatment of cosmogony, where the universe's history is the
medium's accumulated tangle and the question ``what existed before?'' becomes
a question about the medium prior to its first records.  The same whole-state
coherence of the medium that breaks process-symmetries also underlies the
holographic correlations of entangled particles: these correlations are
properties of the medium's global saturated state, not signals propagating
between measurement sites, and the experimental lower bounds on any
``speed of spooky action'' \cite{Salart2008,Yin2013} constrain hypothetical
signal velocities and are vacuous in the present framework --- they fall
outside the propagation-speed analysis of \S\ref{sec:EM} and
\S\ref{sec:gravity} entirely.

\subsection{Neutrinos as Torsional Radiation}
\label{sec:neutrino}

\begin{gapbox}[Subsection status: speculative spectrum, with falsifiers]
\textbf{Claim-status taxonomy for this subsection} (following Paper~I
\S Claim~Status).  The treatment below mixes four levels of claim, each
flagged explicitly:
\begin{itemize}\setlength\itemsep{2pt}
  \item \textit{Structural} (framework-determined, deferred prefactor only):
        the neutrino has charge zero, is produced only in topology-changing
        events, propagates as a torsional pulse in the longitudinal channel,
        and carries the chirality of its parent reconnection.  These four
        identifications follow from the same chiral-superfluid Lagrangian
        that fixes the photon and the charged-lepton sectors and are not
        adjustable.
  \item \textit{Structural prediction, sharply falsifiable}: neutrinos are
        \emph{Dirac}, not Majorana --- the medium's preferred chirality
        breaks C, so a self-conjugate state cannot be a stable mode of the
        chiral channel.  Falsifier: positive observation of neutrinoless
        double-beta decay ($0\nu\beta\beta$) at any experiment
        (CUORE~\cite{CUORE2022}, KamLAND-Zen~\cite{KamLANDZen2023},
        LEGEND~\cite{LEGEND2021}) directly contradicts this prediction.
  \item \textit{Speculative} (no derivation in the present framework): the
        non-zero mass spectrum, the threefold flavour count, the
        oscillation pattern, the PMNS angles and CP phase, the
        $G_F$ normalisation.  Each of these is listed below as a numbered
        open problem (P1)--(P5) with an explicit falsifier attached.
  \item \textit{Cross-link to Paper~I}: the antineutrino is identified
        with the torsional recoil pulse released when the surface-locked
        electron of the neutron composite
        ($\mathcal{N}=\mathcal{T}\cup_{\Sigma_p}\mathcal{E}$,
        Paper~I~\S Neutron) unwraps into a bulk defect during $\beta$-decay.
        The decay endpoint is the surface-unbinding energy of that lock
        (\S\ref{sec:CP}, Eq.~\eqref{eq:chi-def} ff., and Paper~I).
\end{itemize}
The body below identifies the geometric object, develops the structural
identifications, and lists the (P1)--(P5) problems with attached
falsification conditions sharp enough to retire the subsection if any
single experiment returns the wrong answer.  Closure of (P1)--(P5)
\emph{as derivations} (rather than as falsifiers) is deferred to a
subsequent paper in the series.
\end{gapbox}

\subsubsection{Geometric identification}

A vortex reconnection is not silent.  When two filaments exchange partners on the
timescale $\tau \sim a/c$, the local twist density along each strand is
discontinuously rearranged, and the excess torsion must be radiated away to
satisfy conservation of angular momentum about the reconnection point.  In a
chiral superfluid the radiated mode is not a longitudinal phonon (that channel
carries the photon and the gravitational signal) and not a transverse shear
oscillation of a stable core (that channel carries the charged leptons): it is a
\emph{torsional pulse} --- a propagating twist of the vortex-line phase field,
unattached to any standing structure.

We tentatively identify the neutrino with this torsional radiation:
\begin{itemize}
    \item \textbf{Charge}: zero, because the pulse carries no closed circulation
          loop --- only a twist along an open propagation direction.
    \item \textbf{Helicity}: fixed by the chirality of the parent reconnection
          (see \S\ref{sec:parity}), giving the observed
          left-handed $\nu$ / right-handed $\bar\nu$ asymmetry without further
          assumption.
    \item \textbf{Range}: long, because a torsional pulse in the unsaturated
          longitudinal channel does not couple to the chiral-shear sector that
          gives charged leptons their rest mass --- so it propagates at $c$ to
          excellent approximation, with negligible interaction cross-section.
    \item \textbf{Production}: only in events that change vortex topology
          (reconnections) --- i.e.\ in weak processes, never in EM or strong
          processes alone.
\end{itemize}

This identification is consistent with the qualitative picture sketched in
Paper~I (neutron decay as ``straitjacket release''): the antineutrino is the
torsional recoil pulse emitted when the compressed electron vortex unwraps from
the proton breather.

\subsubsection{What the framework does \emph{not} yet predict}

The torsional-pulse picture predicts a strictly massless, structureless
radiation mode.  This contradicts the established phenomenology in five distinct
ways, each of which constitutes an open problem for the SSV programme:

\begin{enumerate}[label=(P\arabic*)]
    \item \textbf{Non-zero mass.} Atmospheric and solar oscillation experiments
          require $\Delta m^2_{21} \approx 7.5\times 10^{-5}\,\mathrm{eV}^2$ and
          $|\Delta m^2_{32}| \approx 2.5\times 10^{-3}\,\mathrm{eV}^2$, so at
          least two mass eigenstates have $m_\nu \gtrsim 10^{-2}\,\mathrm{eV}$.
          A purely massless torsional pulse cannot accommodate this.  A possible
          resolution is that the pulse is weakly self-localising in the chiral
          channel, acquiring a small effective mass from the same chiral-shear
          coupling $\lambda$ that sets $\alpha$ --- but no calculation
          currently exists.\\
          \textbf{Falsifier.}  The self-localisation route gives a mass scale
          tied to $\alpha^n$ times a structural energy; a measured sum
          $\sum m_\nu \gtrsim 1\,\mathrm{eV}$ at KATRIN~\cite{KATRIN2022} or
          from cosmology~\cite{Planck2020}, taken with the framework's
          structural ratios, cannot be accommodated by any reasonable
          chiral-shear suppression and would falsify the torsional-pulse
          identification.  The present bounds $\sum m_\nu < 0.45\,\mathrm{eV}$
          (KATRIN) and $\sum m_\nu < 0.12\,\mathrm{eV}$ (cosmology) leave the
          identification viable but already constrain the self-localisation
          mechanism more tightly than its leading-order natural scale.
    \item \textbf{Three flavours.} The Standard Model has $\nu_e, \nu_\mu,
          \nu_\tau$, paired with the three charged-lepton generations.  In the
          present geometry a torsional pulse is characterised by its twist rate
          and its parent reconnection, neither of which has an obvious threefold
          structure.  This is plausibly the same threefold-generation problem
          that affects the charged leptons (the muon and tau remain
          unidentified as specific breathing modes; see \S\ref{sec:leptons}); a
          unified resolution of both would be desirable.\\
          \textbf{Falsifier.}  Discovery of a fourth active neutrino flavour
          --- whether as a fourth oscillation channel beyond
          $(\nu_e,\nu_\mu,\nu_\tau)$ in long-baseline or atmospheric data
          \cite{MicroBooNE2022}, or as a fourth eV-scale state at reactor
          experiments --- would falsify the unified eigenmode identification:
          the SSV framework ties the neutrino count to the charged-lepton
          breathing-mode count, so a fourth $\nu$ requires a fourth
          breathing-mode charged lepton, which is independently constrained
          by precision electroweak fits.  Sterile candidates with
          $\Delta m^2 \sim 1\,\mathrm{eV}^2$ are not direct falsifiers: the
          framework allows torsional sub-modes of the existing flavours to
          appear as sterile-like states, and the present data are
          inconclusive.
    \item \textbf{Oscillations.} Flavour oscillation requires a coherent
          superposition of distinct mass eigenstates with different propagation
          phases.  A classical torsional pulse has no obvious analogue of this;
          the natural objects in the medium are eigenmodes of the linearised
          GP+chiral operator, and whether these admit a three-state mixing
          structure with the observed PMNS angles is unknown.\\
          \textbf{Falsifier.}  The PMNS matrix and the CKM matrix should both
          arise from the same chirality-memory mechanism that produces the
          neutron EDM (\S\ref{sec:neutron-edm}, Eq.~\eqref{eq:chi-def}).  A
          measured statistical \emph{independence} of the PMNS Dirac CP phase
          $\delta_{\rm PMNS}$ from the CKM phase $\delta_{\rm CKM}$ ---
          specifically, a precision-determined $\delta_{\rm PMNS}$ from
          DUNE~\cite{DUNE2020} or Hyper-Kamiokande~\cite{HK2018} that lies
          outside the SSV-predicted relation to $\delta_{\rm CKM}$ once the
          converged 3D chirality-memory calculation is in hand
          (issue~\#30) --- would falsify the
          unified-chirality reading.  Until that calculation lands, the
          falsifier is statable but not numerically testable.
    \item \textbf{Dirac vs.\ Majorana.} A torsional pulse and its
          parity-conjugate differ only in the sign of the twist.  In a chiral
          medium the two are not related by~$C$: the medium's preferred
          chirality (\S\ref{sec:parity}) breaks $C$ explicitly, so a
          self-conjugate state cannot be a stable eigenmode of the chiral
          channel.\\
          \textbf{Structural prediction.}  Neutrinos are \emph{Dirac}, not
          Majorana.  This is one of the few sharply-falsifiable structural
          predictions of the neutrino subsection.\\
          \textbf{Falsifier.}  A positive observation of neutrinoless
          double-beta decay ($0\nu\beta\beta$) at CUORE~\cite{CUORE2022},
          KamLAND-Zen~\cite{KamLANDZen2023}, LEGEND~\cite{LEGEND2021}, or any
          successor experiment falsifies the Dirac prediction.  Present
          bounds on the effective Majorana mass
          $\langle m_{\beta\beta}\rangle \lesssim 0.04$--$0.2\,\mathrm{eV}$
          (KamLAND-Zen) are consistent with the SSV prediction $\langle
          m_{\beta\beta}\rangle = 0$; the half-life lower bounds will tighten
          in the next decade.
    \item \textbf{Cross-section normalisation.} The Fermi constant
          $G_F = 1.166\times 10^{-5}\,\mathrm{GeV}^{-2}$ should follow from the
          same reconnection-barrier calculation that fixes the $W$ mass.  Until
          that calculation is in hand (see \S\ref{sec:weak}), the predicted
          neutrino interaction cross-section is unconstrained.\\
          \textbf{Falsifier.}  The structural relation
          $G_F/\sqrt{2} = g_W^2/(8 m_W^2)$ is inherited from electroweak
          theory; in SSV $g_W$ comes from the chiral-shear vertex strength
          and $m_W$ from the cap geometry of \S\ref{sec:W_rayleigh}.  Both
          are fixed by the same 3D reconnection-barrier minimisation tracked
          under issue~\#30.  A measured precision $G_F$ at low energy
          incompatible with the framework's predicted relation between the
          cap-geometry coupling and the chiral-shear coupling, once that
          calculation lands, would falsify the reconnection identification
          of the weak sector.  Until then, the falsifier is structural rather
          than numerical.
\end{enumerate}

The neutrino subsection is therefore presented as a geometric \emph{hypothesis}
in the same spirit as the neutron section of Paper~I: a motivated identification
that defers most quantitative claims to the full 3D reconnection computation.
The four structural identifications (charge, helicity, range, production),
the structural Dirac prediction (P4), and the falsifiers attached to
(P1)--(P5) constitute the testable content of the present treatment.
Problems (P1)--(P5) as positive derivations are deliverables for a subsequent
paper in this series.

% ═══════════════════════════════════════════════════════════════════════
\section{Charged Leptons: Higher Breather Modes}
\label{sec:leptons}
% ═══════════════════════════════════════════════════════════════════════

\begin{gapbox}[Status: muon dynamically placed, tau topologically identified]
The electron is derived from first principles in Paper~I.  The muon is
identified there with the lowest core-breathing $\oplus$ Kelvin hybrid mode
of the toroidal vortex (Paper~I~\cite{TGD-I}, \S Muon and
Appendix~H), giving $m_\mu = (3/2)\mu_0$ to $0.59\%$.  The tau is identified
in \S\ref{sec:tau} below with the lowest Hopf-linked pair of trefoil-skeleton
baryon breathers bound by a shared muon-class ring-breathing quantum, giving
the leading-order prediction $m_\tau = 2m_p - m_\mu = 1770.9~\mathrm{MeV}$
to $0.34\%$ of the observed value.  Both identifications are structural
hypotheses to be confirmed by the same $\mathcal{L}+\mathcal{L}_\perp$
eigenmode minimisation that closes the muon eigenfrequency (Paper~I open
problem~2); the tau identification additionally requires the two-component
link of trefoils to be admissible in the SSV functional, which is a Borromean-class
extension of the (2,3)-torus knot calculation in Paper~I.
\end{gapbox}

\subsection{The Muon: Ring-Breathing Mode of the Electron}

The muon is identified in Paper~I~\cite{TGD-I} (\S Muon) with the lowest
ring-breathing mode of the toroidal electron vortex --- a periodic oscillation of the major
radius $R_e$ about its equilibrium value $R^*=\xi/\alpha$.  The muon is therefore
not a new defect type but the same toroidal vortex as the electron, set into a
specific oscillatory mode by the chiral-shear coupling.  Its production in weak
processes is the same reconnection event that produces an electron, with the
outgoing ring excited rather than ground-state.

The mass assignment $m_\mu=(3/2)\mu_0$ reproduces the observed muon mass to
$0.61\%$.  The open problem (Paper~I, gapbox~2) is the first-principles
computation of the ring-breathing eigenfrequency $\omega_{\rm ring}$ in the
3D $\mathcal{L}+\mathcal{L}_\perp$ functional; the predicted frequency ratio
$\omega_{\rm ring}/\omega_c = m_\mu/m_e \approx 206.77$ is the target.  Paper~II
contributes no new computation here; the muon is mentioned for completeness,
since otherwise the force-sector treatment would discuss neutrinos as the
torsional companion of a charged lepton without ever introducing the charged
lepton itself.

\subsection{The Tau: Hopf-Linked Trefoil Pair Bound by a Muon Quantum}
\label{sec:tau}

\begin{gapbox}[Status (2026-05): identification inherits both the muon ladder and the proton F calibration]
The mass relation $m_\tau = 2 m_p - m_\mu$ derived below uses the
\emph{measured} muon mass (PDG) as a numerical input, so the predicted
$1770.89$~MeV is arithmetically correct.  The \emph{structural}
identification however imports two demoted claims: (i) the muon as the
$3/2\,\mu_0$ ladder rung carried by a core-breathing $\oplus$ Kelvin
hybrid (Paper~I §The Muon gapbox: this hybrid mode is not basis-converged
under the Path~B null; the muon is a candidate carried by empirical
agreement on $m_e/\alpha$, not by a derived eigenmode), and
(ii) the proton mass $m_p = N_Y F \mu_0$ (Paper~I §The Proton gapbox:
$F$ is a cutoff-dependent calibration).  The Hopf-link binding-energy
argument therefore stands or falls with the muon eigenmode and proton
form-factor derivations; it is not an independent identification.
Tracking: issue~\#66.
\end{gapbox}

The tau lands on rung $25\tfrac12$ of the $\alpha$-harmonic ladder
($m_\tau = 25.375\,\mu_0$, $0.49\%$).  We identify it structurally with the
\emph{lowest Hopf-linked pair of trefoil-skeleton breathers bound by a shared
core-breathing $\oplus$ Kelvin hybrid quantum}, equivalently a $(2,3)\oplus(2,3)$
two-component link of vortex trefoils joined by one muon-class ring-mode
in the linkage region.  The mass follows directly from the topological
building-block accounting:
\begin{equation}
  \boxed{m_\tau \;=\; 2\,m_p \;-\; m_\mu
    \;=\; 2(13.5)\mu_0 - (3/2)\mu_0
    \;=\; \tfrac{51}{2}\mu_0,}
  \label{eq:tau-mass}
\end{equation}
which lands on rung $25\tfrac12$ by construction and evaluates to
\begin{equation}
  m_\tau^\mathrm{pred} \;=\; 2(938.272) - 105.658 \;=\; 1770.89~\mathrm{MeV},
  \label{eq:tau-numeric}
\end{equation}
to compare with the observed $1776.86~\mathrm{MeV}$ ($-0.336\%$).  Using the
isospin-averaged baryon mass $(m_p + m_n)/2$ instead of $m_p$ alone gives
$m_\tau^\mathrm{pred} = 1772.18~\mathrm{MeV}$ ($-0.263\%$), tightening the
agreement since the two trefoils in a Hopf link are not isospin-distinguished
in the symmetry channel of the linkage.  Either choice is more accurate
than the bare ladder coincidence $25\tfrac12\,\mu_0 = 1785.6$~MeV
($+0.494\%$).

\paragraph{Topological reading.}
The (2,3)-trefoil is the simplest non-trivial knot and the topological skeleton
of the proton in Paper~I~\cite{TGD-I}.  Two trefoils linked through each other
once (Hopf link) form the simplest non-trivial two-component link and are
the natural next building block above the proton in the SSV defect spectrum.
The linkage region carries a shared Kelvin-helicity standing wave; the lowest
such mode has the same eigenfrequency as the muon's core-breathing $\oplus$
Kelvin hybrid (Paper~I, \S Muon), so the binding energy is exactly one muon
mass.  The minus sign in~\eqref{eq:tau-mass} is the binding-energy convention:
two free trefoils ($2 m_p$) bound by a Kelvin quantum lower their total energy
by $m_\mu$.

\paragraph{Structural verification path.}
The full eigenmode closure of~\eqref{eq:tau-mass} requires the same
$\mathcal{L} + \mathcal{L}_\perp$ minimisation that fixes the muon
eigenfrequency (Paper~I open problem~2) but on the $(2,3)\oplus(2,3)$
two-component link instead of on the single toroidal vortex.  The crossing
count is six (each trefoil contributes three, the link contributes a global
linking number of one without crossing additions), so the relevant N-junction
count is $N_Y^\mathrm{link} = 6$ and the form-factor extraction is the
direct extension of the trefoil computation in Paper~I~\S A.  The
linkage-region Kelvin quantum carrying $-m_\mu$ binding energy is the same
$m_\varphi = \pm 1$ helicity-sector hybrid that drives the muon rung; no
new mode type is required.  This identification therefore relies on no
additional theoretical input beyond what already closes the proton mass
formula and the muon eigenmode hypothesis.

\paragraph{Consequence for the generation problem.}
The three charged leptons now have a tower of building-block decompositions:
$e = 1$ toroidal vortex; $\mu = e + $ core-breathing $\oplus$ Kelvin hybrid;
$\tau = 2$ proton-trefoils $-$ muon quantum.  The threefold structure is no
longer a coincidence of ladder rungs but reflects the three lowest topological
classes admissible in the SSV functional: pure torus (electron),
torus-with-internal-mode (muon, pion), and linked trefoils (tau).  This
sharpens the threefold-generation problem (\S\ref{sec:leptons}) into a
specific topological question: is the SSV functional admissible on the
$(2,3)\oplus(2,3)$ link, and at what link spacing does the muon-class
binding mode go off-shell?

\subsection{The Generation Problem}

The Standard Model has three charged-lepton generations
($e, \mu, \tau$) paired with three neutrino flavours
($\nu_e, \nu_\mu, \nu_\tau$).  The framework presents a single unified version
of this puzzle, since charged leptons and neutrinos in SSV are not independent
objects but two facets of the same vortex configuration --- the standing
toroidal mode and the torsional pulse it emits during reconnection.  A first-principles
account of the threefold structure should therefore appear simultaneously in:

\begin{itemize}
    \item \textbf{Charged leptons.}  The eigenmode spectrum of the
          $\mathcal{L}+\mathcal{L}_\perp$ functional on the toroidal vortex
          background should yield the muon eigenfrequency (Paper~I open
          problem 2) and, at the next harmonic, an identification of the tau.
          Whether the spectrum has exactly three accessible modes below the
          confinement scale --- and why three --- is unknown.
    \item \textbf{Neutrinos.}  The same eigenmode structure should determine
          the flavour content of the radiated torsional pulse during
          reconnection (\S\ref{sec:neutrino}, P2 and P3), giving the three
          neutrino flavours and their oscillation phenomenology as a
          downstream consequence rather than as a separately tunable sector.
\end{itemize}

The threefold-generation problem is therefore one problem, not two.  Its
resolution is deferred to the same 3D minimisation that closes the W-mass gap
(\S\ref{sec:weak}) and the $\alpha_G$ derivation (\S\ref{sec:gravity}); all
three open problems share the same $\mathcal{L}+\mathcal{L}_\perp$
computation as their target.

% ═══════════════════════════════════════════════════════════════════════
\section{Gravity: Time-Delay Gradients from Interfering Pressure Waves}
\label{sec:gravity}
% ═══════════════════════════════════════════════════════════════════════

\subsection{The Physical Picture}

Gravity is the gradient of a time delay.  Every massive defect is a breather
(Paper~I) --- a localised region of $\Psi$ oscillating at its Compton
frequency --- and it radiates a longitudinal pressure wave into the medium,
with amplitude falling as $1/r$ from spherical dilution.  A second breather,
or a light wave, propagating through that radiated wave interferes with it;
the time-averaged interference, through the medium's logarithmic
nonlinearity, slows the local rate of change of $\Psi$.  Since physical time
is that rate of change (Paper~III), the interference imposes a time delay,
strongest near the mass and fading with distance.  Its gradient bends every
trajectory that crosses it toward the region of greatest delay --- and that
bending is gravity.

A single potential $\Phi(\mathbf{x})$ summarises the delay: the acceleration
of free fall is $-\nabla\Phi$, and the fractional slowing of clocks is
$\Phi/c^2$.  The present section establishes gravity as the fourth
force-sector mode and fixes Newton's constant from the medium; the full
development of the mechanism --- the chain from interference to path
bending, and the gravitational consequences that follow --- is the subject
of Paper~IV.

\subsection{The Proton as a Pulsating Breather}

The source of the radiated wave is the breather itself.  In
Paper~I the proton is a spherical breather mode --- a periodically
oscillating, localised density perturbation of the superfluid vacuum,
sustained by its own knotted vortex skeleton and not by any external
driving.  Its eigenfrequency is the Compton frequency:
\begin{equation}
    f_p = \frac{m_p c^2}{h} \approx 2.27\times 10^{23}\,\mathrm{Hz},
    \label{eq:proton_freq}
\end{equation}
with corresponding wavelength $\lambda_p = c/f_p \approx
1.3\times10^{-15}\,\mathrm{m}$.  The breather radiates a longitudinal
density wave into the medium at this frequency; the fractional density
modulation $x(\mathbf{x},t) \equiv \delta\rho/\rho_0$ falls as $1/r$ from
spherical dilution.  Every massive defect does the same, scaled by its
mass-energy.  Gravity is the interference of these radiated waves, and the
next subsection computes it.

\subsection{The Interference Cross-Term}
\label{sec:interference}

The mechanism is contained in the LogSE.  Where the radiated waves of
several breathers overlap, the local density is $\rho/\rho_0 = 1 + \sum_i
x_i$ with $x_i \equiv \delta\rho_i/\rho_0$, and the medium's logarithmic
energy term expands as
\begin{equation}
    \ln\!\Big(1 + \sum_i x_i\Big)
      = \sum_i x_i - \frac{1}{2}\sum_i x_i^2
      - \sum_{i<j} x_i x_j + \cdots
    \label{eq:log_expansion}
\end{equation}
The single-wave terms are each breather's own self-energy; the interaction
between two breathers $A$ and $B$ is carried by the cross term
\begin{equation}
    u_{AB} \;=\; -\,b\,\rho_0\,x_A\,x_B ,
    \label{eq:cross_term}
\end{equation}
an \emph{interference} term --- the product of two waves.  Three properties
fix the character of gravity.  Its time average $\langle u_{AB}\rangle =
-b\rho_0\langle x_A x_B\rangle$ survives only for phase-correlated waves, so
coherent breathers gravitate and incoherent disturbances do not.  Its sign
is fixed negative by the concavity of the logarithm: \emph{gravity is
attractive because the medium's nonlinearity is logarithmic}.  And it is
bilinear --- with $x_A \propto m_A/r$ diluted to $B$'s location and $x_B
\propto m_B$ --- so $\langle u_{AB}\rangle \propto -m_A m_B/r$, the
Newtonian potential, with the inverse-square force its gradient.  The
coefficient is fixed by the breather's acoustic output in
\S\ref{sec:Newton}; the full development is given in Paper~IV.

\subsection{The Bjerknes Mechanism}

This time-averaged interference of pulsating waves is, classically, the
Bjerknes interaction~\cite{Bjerknes1906}.  For two bodies pulsating in a
compressible medium with volume rates $\dot V_1$ and $\dot V_2$, Bjerknes
obtained a time-averaged force
\begin{equation}
    \langle F \rangle
      = -\,\rho_0\,\langle \dot{V}_1 \dot{V}_2 \rangle\,\frac{1}{4\pi r^2},
    \label{eq:bjerknes_classical}
\end{equation}
with $r$ the separation.  The cross-correlation $\langle \dot V_1 \dot
V_2\rangle$ is exactly the interference term $\langle x_A x_B\rangle$ of
\S\ref{sec:interference} written in volume-velocity variables, and the
geometric $1/(4\pi r^2)$ is the spherical dilution of a wave in three
dimensions --- not an independently postulated force law.  We use Bjerknes'
result and his volume-velocity form below to compute $G$; ``force'' is
retained only as shorthand for the gradient of the interference.  Its
reinterpretation --- why this is not a force across an inert gap, and why
the breathers, as self-sustaining eigenmodes of $\Psi$, need no external
driving --- is developed in Paper~IV.

\subsection{Newton's Constant from the Medium}
\label{sec:Newton}

The quantitative interference calculation uses the breather's acoustic
output to fix the coefficient of $\Phi$ left open in \S\ref{sec:interference}.
It proceeds in four steps: the wave a breather radiates, the
cross-term a second breather feels in it, the sub-grain suppression that
makes that coupling weak, and the value of $G$ that results.

\paragraph{Step 1: Acoustic wave of a pulsating breather.}
The proton pulsates at angular frequency $\omega_p = m_p c^2/\hbar$
with volume oscillation $V(t) = V_0 + \delta V_p\sin(\omega_p t)$,
so $\dot{V}(t) = \omega_p\delta V_p\cos(\omega_p t)$.  Both the static
Bjerknes force and its radiative part are governed by the medium's
sound speed $c$ (\S\ref{sec:EM_open_issue}).  The standard
Landau--Lifshitz result for a pulsating sphere
\cite{Landau_Fluid} gives:
\begin{equation}
    \delta P(r,t)
      = \frac{\rho_0}{4\pi r}\,\ddot{V}\!\!\left(t - \frac{r}{c}\right)
      = -\frac{\rho_0\,\omega_p^2\,\delta V_p}{4\pi r}
         \sin\!\left(\omega_p t - \frac{\omega_p r}{c}\right).
    \label{eq:acoustic_pressure}
\end{equation}
Since the proton Compton wavelength $\lambda_p \approx
1.3\times10^{-15}\,\mathrm{m}$ is smaller than all macroscopic
separations, gravitational interactions always occur in the radiation
zone ($r \gg \lambda_p$).  The time-averaged Bjerknes force
nevertheless retains its $1/r^2$ form at all distances for in-phase
oscillators, because the static component of the time-averaged force
is independent of zone.

\paragraph{Step 2: The interference cross-term.}
A second identical proton at separation $r$ is a resonant oscillator
at $\omega_p$, driven in phase by the same medium.  The time-averaged
cross-term of \S\ref{sec:interference}, in volume-velocity variables, is
the average over one cycle:
\begin{equation}
    \langle \mathbf{F}_{12} \rangle
      = -\frac{\rho_0}{4\pi r^2}
        \langle \dot{V}_1(t)\,\dot{V}_2(t) \rangle\,\hat{r},
    \label{eq:bjerknes_formula}
\end{equation}
with
\begin{equation}
    \langle \dot{V}_1 \dot{V}_2 \rangle
      = \tfrac{1}{2}\omega_p^2 (\delta V_p)^2 > 0,
    \label{eq:in_phase}
\end{equation}
giving an attractive force $\propto 1/r^2$.  For two objects of
different masses the pulsation amplitude scales as $\delta V \propto
m/\rho_0$ (Step~3), so the force is proportional to $m_1 m_2/r^2$
--- Newton's law.

\paragraph{Step 3: Pulsation amplitude and sub-grain suppression.}
The fractional volume excursion of a breather is set by the ratio of
its rest-mass energy to the medium's compressive energy over the
breather volume:
\begin{equation}
    \varepsilon_p
      \equiv \frac{\delta V_p}{V_0}
      = \frac{m_p c^2}{\rho_0 c^2 \cdot V_0}
      = \frac{m_p}{\rho_0 V_0},
    \label{eq:epsilon_def}
\end{equation}
where $P_0 \approx \rho_0 c^2$ and $V_0 = \tfrac{4}{3}\pi a_p^3$
with $a_p = \hbar/(m_p c)$ the proton Compton radius.  Since $\delta
V \propto m$, the Bjerknes force scales as $m_1 m_2$ as required.

The proton is a \emph{sub-grain} oscillator: its Compton radius $a_p$
is smaller than the medium healing length $\xi = \hbar/(m_e c)$ by
the mass ratio
\begin{equation}
    \frac{a_p}{\xi} = \frac{m_e}{m_p} \approx \frac{1}{1836}.
    \label{eq:subgrain}
\end{equation}
The medium cannot resolve structure below $\xi$; consequently the
proton couples to the long-range density wave with a suppressed
\emph{acoustic monopole moment} $Q_p$.  For a sub-grain spherical
oscillator, coupling to wavelengths $\lambda \gg \xi$ is suppressed
by the ratio of source volume to grain volume:
\begin{equation}
    Q_p \sim \delta V_p \cdot \left(\frac{a_p}{\xi}\right)^3
           = \delta V_p \cdot \left(\frac{m_e}{m_p}\right)^3.
    \label{eq:Qp}
\end{equation}
The coupling hierarchy across the three stable objects is:
\begin{equation}
    \eta_\gamma = 1
    \;\gg\; \eta_e \sim \alpha
    \;\gg\; \eta_p \sim \left(\frac{m_e}{m_p}\right)^3,
    \label{eq:eta_hierarchy}
\end{equation}
where the photon couples with unit efficiency (it \emph{is} the
medium wave), the electron couples through its toroidal topology
(suppressed by $\alpha$), and the proton couples as a sub-grain
monopole (suppressed by $(m_e/m_p)^3$).  Gravity is weak not because
of a separate fine-tuning, but because the proton is small relative
to the medium grain.

\paragraph{Step 4: Newton's constant from the medium.}
Substituting $Q_p$ into~\eqref{eq:bjerknes_formula} and matching to
$\langle F_{12}\rangle = Gm_p^2/r^2$:
\begin{equation}
    G = \frac{\rho_0\,\omega_p^2\,Q_p^2}{8\pi\,m_p^2}.
    \label{eq:G_from_medium}
\end{equation}
A full evaluation of $Q_p$ requires the 3D proton breather profile
and is deferred to a companion paper.  A structural consistency check
uses the gravitational coupling constant $\alpha_G \equiv
Gm_p^2/(\hbar c)$ and the proton mass from the Paper~I ladder,
$m_p = N_p\,m_e/\alpha$ with $N_p \approx 13.44$:
\begin{equation}
    G = \frac{\alpha_G\,\hbar c\,\alpha^2}{N_p^2\,m_e^2}.
    \label{eq:G_ladder}
\end{equation}
This is a structural consistency check: $\alpha_G$ is taken from CODATA rather
than derived from first principles, confirming that the Paper~I mass ladder is
internally coherent with the gravitational sector.  Inserting
$\alpha_G = 5.906\times10^{-39}$ (CODATA) and $N_p = 13.44$:
\begin{equation}
    G_{\rm pred}
      = \frac{5.906\times10^{-39}\;\hbar c\,\alpha^2}{(13.44)^2\,m_e^2}
      = 6.634\times10^{-11}\,\mathrm{m^3\,kg^{-1}\,s^{-2}},
    \label{eq:G_numerical}
\end{equation}
within $0.6\%$ of $G_{\rm obs} = 6.674\times10^{-11}
\,\mathrm{m^3\,kg^{-1}\,s^{-2}}$.  The residual is consistent with
the imprecision in $N_p$ (form factor $F \approx 4.47$ is
numerically determined; see Paper~I).

\begin{gapbox}[Open problem: first-principles derivation of $\alpha_G$]
$G$ is not an independent constant of the framework.
Equation~\eqref{eq:G_ladder} expresses it in terms of $\{\hbar, c,
\alpha, m_e, \alpha_G\}$, where the first four are fixed in Paper~I.
The remaining open problem is computing the sub-grain suppression
factor $(a_p/\xi)^n$ from the 3D proton breather profile, which
would yield $\alpha_G$ without empirical input.  The $0.6\%$
structural agreement is evidence of framework consistency, not a
completed derivation.  Two further dependencies that the $0.6\%$
figure inherits, not visible from this expression alone:
(i) the proton mass $m_p = N_p\,m_e/\alpha$ with $N_p = N_Y F \approx
13.44$ inherits the SSV-I §Proton F-cutoff calibration (issue~\#66);
(ii) any tightening of $\alpha_G$ via the sub-grain suppression
will need the same basis-converged 3D toroidal solve that the
Path~B null (\S Muon, Paper~I) showed has not yet been achieved.
\end{gapbox}

\paragraph{Gravitational and electromagnetic radiation share the Goldstone channel.}
In the literature-consistent reading of the medium (\S\ref{sec:EM_open_issue};
Volovik \cite{Volovik2003}, Zloshchastiev \cite{Zloshchastiev2020}),
electromagnetic and gravitational radiation are both phase-channel
($\delta\theta$) disturbances of $\Psi$, propagating at the medium's sound
speed $c$.  They differ only in their multipole-source structure:
electromagnetic radiation arises from the chirality dipole of an accelerating
toroidal vortex, while gravitational radiation arises from the mass-energy
quadrupole of an accelerating pulsating breather.  Because they share the
carrier channel, both propagate at exactly $c$ and arrive simultaneously from
any astrophysical source.

This is consistent with the GW170817 observation.  The neutron-star merger
GW170817 produced gravitational and electromagnetic signals separated by
$\sim 1.7\,\mathrm{s}$.  In SSV this delay is attributed entirely to
source physics --- the relativistic-jet launch, cocoon formation, and jet
breakout that delay the gamma-ray burst relative to the moment of merger
--- and carries no implication for any difference in propagation speed.
This is in full agreement with the mainstream astrophysical interpretation
\cite{GW170817_MMA}.  The chiral-shear mode at $c_\perp$ is an internal
mode of the medium that mediates the static near-field Coulomb interaction
between vortices (\S\ref{sec:EM_open_issue}); it is not the carrier of any
astrophysical radiation.

\subsection{Scope: the Full Treatment in Paper~IV}

The present section has established gravity as the fourth force-sector mode
--- the interference of breather pressure waves, building a time-delay
gradient that bends trajectories --- and has fixed Newton's constant from
the medium.  The conceptual development of the mechanism is carried out in
Paper~IV: the formal account of update capacity, the four gravitational
consequences (free fall, time dilation, light deflection, redshift), the
universality of the coupling, the absence of a graviton, the bridge to
emergent geometry (Paper~VII-b), and the connection to the recording medium
of Paper~III.

% ═══════════════════════════════════════════════════════════════════════
\section{Unified Picture: One Medium, Four Modes}
\label{sec:unification}
% ═══════════════════════════════════════════════════════════════════════

\begin{table}[h]
\centering
\small
\renewcommand{\arraystretch}{1.4}
\caption{The four forces and the Higgs mode as facets of the saturated
superfluid vacuum.  The four forces describe how defects in the medium
interact; the Higgs row is structurally different --- it is the medium
itself oscillating, not an interaction between defects.}
\label{tab:forces}
\begin{tabular}{@{}lllll@{}}
\toprule
\textbf{Sector} & \textbf{Mode type} & \textbf{Source} & \textbf{Carrier} & \textbf{Range} \\
\midrule
Electromagnetism & Transverse flow    & Chirality ($\pm$)      & Transverse phonon         & $\infty$        \\
Strong           & Surface tension    & Topological charge     & Phase oscillation         & $\sim a/\alpha$ \\
Weak             & Reconnection       & Core overlap           & Amplitude mode ($W,Z$)    & $\sim \xi$      \\
Gravity          & Pressure-wave interference & Mass-energy (breather) & Longitudinal density wave & $\infty$        \\
\midrule
Higgs            & Amplitude mode     & The medium itself      & $\delta\rho$ oscillation  & $\sim \xi_{\rm EW}$ \\
\bottomrule
\end{tabular}
\end{table}

We can now survey the four forces from a single vantage point. Table~\ref{tab:forces}
summarises the correspondence and adds the Higgs as the one remaining mode of the
medium that is not an interaction between defects.

Several structural features of this table deserve comment.

\paragraph{Range.} The infinite-range forces (EM and gravity) are both mediated by
massless phonon modes --- transverse and longitudinal, respectively. The short-range
forces involve massive excitations: the gluon acquires an effective mass from the
string tension at distances exceeding $a/\alpha$, and the $W/Z$ are the high-energy
amplitude modes of the core. The hierarchy of ranges $r_W \ll r_{\rm strong} \ll
r_{\rm EM}, r_{\rm gravity}$ follows directly from the hierarchy $\xi \ll a/\alpha \ll
\infty$.

\paragraph{Coupling strengths.} All couplings are fixed by the same two numbers: $\alpha$
and $\mu_0 = m_e/\alpha$. The strong coupling is $\sim 1$ (order-unity surface
tension, no small factor), the electromagnetic coupling is $\alpha \approx 1/137$
(transverse-to-longitudinal wave speed ratio), the weak coupling is $\alpha^2$ (two
powers of the core penetration suppression), and the gravitational coupling is
$Gm_p^2/\hbar c \approx 6\times 10^{-39}$ (set by the tiny ratio of the
proton's acoustic cross-section to $c^4/G$, which encodes the vast hierarchy between
$P_0$ and $m_p c^2$).

\paragraph{Universality of gravity.} Every other force is selective: EM acts on
chirality, strong on topological charge, weak on core overlap. Gravity (Bjerknes)
acts on \emph{acoustic cross-section}, which is determined by mass-energy alone.
This is the mechanical root of the equivalence principle.

\paragraph{Defects vs.\ canvas.} The four force rows of Table~\ref{tab:forces}
share a common pattern: each describes how some property of a defect (chirality,
topological charge, core overlap, mass-energy) sources a pressure response of
the medium that another defect feels.  The Higgs row breaks this pattern.  Its
source is not a defect but the medium itself; its carrier is the same density
mode that, when sourced by a moving breather, transmits gravity --- but
intrinsically rather than as an interaction.  A reader who finds the Higgs
``out of place'' in a force-unification table is correctly identifying the
asymmetry: the Higgs is not a fifth force, it is the medium's own amplitude
mode made briefly visible.  Its inclusion makes explicit that the same
$\Psi$-field substrate produces both the four-force interaction structure
and the matter / canvas distinction that underlies it.

\paragraph{Particle content on the same substrate.} The same medium produces
the full matter content of the Standard Model: electrons as toroidal vortex
rings, baryons as knotted vortex breathers, charged leptons as higher
breathing harmonics (\S\ref{sec:leptons}), neutrinos as torsional radiation
from reconnection (\S\ref{sec:neutrino}), and the discrete symmetry structure
(P-violation, the absence of exact CP, the absence of a strong CP problem)
as direct consequences of the medium's chirality and recorded history
(\S\ref{sec:CP}).  Particles, forces, and symmetries are not three separate
sectors with mutual couplings; they are three different ways of asking
questions about the same hydrodynamic substrate.

\paragraph{No new constants.} The framework introduces no coupling constants
beyond those already fixed by the mass spectrum of Paper~I.  The four forces,
the Higgs sector, the particle content, and the discrete-symmetry structure
share the same parameters $\{c, \alpha, \mu_0\}$; they are not independent
sectors with mutual couplings, but four different windows into the same
hydrodynamic pressure field.

% ═══════════════════════════════════════════════════════════════════════
\section{Testable Predictions}
\label{sec:predictions}
% ═══════════════════════════════════════════════════════════════════════

The framework makes a number of predictions that go beyond what is currently
explained within the Standard Model.  We organise them by sector to make
clear which experiments would test which structural claims.

\subsection*{Force-sector and gravitational predictions}

\begin{enumerate}

\item \textbf{Gravitational-wave polarisation.} The longitudinal acoustic phonon
  of the superfluid is spin-0, and there is no graviton in SSV.  Nevertheless,
  we predict that gravitational radiation carries an \emph{effective} spin-2
  polarisation pattern emerging from the transverse-traceless projector in the
  linearised acoustic metric --- a testable statement about the polarisation
  modes of gravitational waves.  The ratio of breathing-mode to quadrupolar-mode
  amplitude in binary inspirals should carry a small superfluid correction at
  order $v/c$ in the post-Newtonian expansion.

\item \textbf{Gravitational Casimir correction.} The Bjerknes force between a proton
  and any other mass includes a next-to-leading-order correction from the
  \emph{quantum pressure} term, analogous to the Casimir effect. At sub-micron
  separations, this produces a deviation from the $1/r^2$ law of order
  $(\delta r/r)^2 \sim (a/r)^2$, where $a \sim 10^{-15}\,\mathrm{m}$ is the vortex
  core size. This is below current precision for laboratory experiments but could
  in principle be detected at hadron-scale separations in neutron scattering
  experiments.

\item \textbf{Gravitational-wave memory.}  The gravitational-wave memory effect
  (a permanent shift in detector test-mass separation following the passage of
  a strong GW signal) is predicted to be present at the Christodoulou level
  --- not as a peculiar property of asymptotic spacetimes but as a direct
  consequence of a medium that records (Paper~IV).  The
  framework sets the same magnitude as conventional GR; the structural claim
  is that memory is the natural state of affairs in SSV, not an exotic
  signal.  Distinguishing SSV from GR on this point requires precision
  detection of the fluctuation spectrum around the memory mean, which is
  beyond current sensitivity.

\item \textbf{$W$-mass geometric fit awaiting derivation.} Our scalar estimate $E_{\rm recon} \sim m_e/\alpha^2
  \approx 9.6\,\mathrm{GeV}$ is a factor of $\sim 8$ below the observed $W$ mass.
  The Rayleigh-jet cap formula (\S\ref{sec:W_rayleigh}, equation~\eqref{eq:Rcap_phi})
  gives $m_W \approx 78.9\,\mathrm{GeV}$ --- within $1.8\%$ of the observed value.
  The cap equilibrium analysis (\S\ref{sec:W_rayleigh}) shows that
  $R_{\rm cap} = \phi/\alpha$ follows analytically from the chiral-shear bending
  stiffness $\lambda_{\rm bend} = \phi^3/\alpha^3$ via the equilibrium cubic
  $R^3 + \tau R^2 = \lambda_{\rm bend}$.  The local $\mathcal{L}_\perp$
  curvature correction falls short by a factor $232$; the remaining open step is
  deriving the non-local bending stiffness $\phi^3/\alpha^3$ from the
  topological structure of the cap (Seifert surface of the trefoil, or
  chiral-mode Casimir energy at the reconnection saddle).

\item \textbf{Running of $\alpha$ at high energy.} The chiral-shear coupling
  $\alpha$ is a medium property, not a bare constant. At momentum scales
  approaching $\mu_0 = m_e/\alpha$, the chiral-shear sector begins to couple
  to the longitudinal modes through the nonlinear term in the
  Gross--Pitaevskii equation. This produces a logarithmic running of $\alpha$
  that should match the QED renormalisation group equation --- a
  self-consistency check that has not yet been completed in detail.

\end{enumerate}

\subsection*{Particle-content predictions}

\begin{enumerate}\setcounter{enumi}{5}

\item \textbf{Higgs as condensed-matter analogue.}  Since the Higgs is identified
  as the amplitude mode of the same order parameter that carries every other
  particle (\S\ref{sec:higgs}), its phenomenology should match that of the
  amplitude mode in cold-atom condensates near saturation~\cite{Endres2012},
  scaled up by the medium parameters.  The decay channels of the Higgs reflect
  the eigenmode structure of the medium's amplitude oscillations; small
  deviations from Standard Model branching ratios at high precision (above
  current LHC sensitivity) would be expected if the amplitude mode is
  genuinely a hydrodynamic excitation rather than a fundamental scalar.  The
  framework does not predict a specific deviation magnitude; the structural
  claim is that the Higgs is not a separately tunable sector.

\item \textbf{No fourth-generation lepton.}  The framework predicts that the
  toroidal vortex background supports a finite eigenmode spectrum; the muon and
  tau are placeholders for the lowest two ring-breathing harmonics
  (\S\ref{sec:leptons}).  Whether the spectrum has exactly three accessible
  modes below the confinement scale is open, but the threefold Standard Model
  generation structure --- if it follows from this eigenmode structure --- would
  imply that searches for a fourth charged lepton or a fourth neutrino flavour
  should return null at all accessible energies.  Any positive result would
  falsify the eigenmode interpretation of the lepton generations.

\item \textbf{Neutrino emission only in topology-changing events.}
  In the SSV picture, neutrinos are torsional pulses emitted only when vortex
  topology is rewritten --- i.e.\ in weak-sector reconnections.  No neutrino
  emission accompanies any purely electromagnetic or strong-sector process,
  including high-energy QED vertices, hadronic resonance decays that proceed
  through the strong channel, or coherent photon scattering, regardless of
  available phase space.  This is a sharper statement than the Standard Model
  selection rule (which is a flavour-current statement); here it is a
  structural impossibility, since no torsional pulse can be sourced without a
  reconnection.  Any observation of neutrino emission unaccompanied by
  topology change in the source would falsify the framework.

\item \textbf{No magnetic monopoles.}
  The magnetic field is defined as $\mathbf{B} \equiv c_\perp^{-1}\nabla\times
  \mathbf{v}_\perp$ (Section~\ref{sec:B_derivation}), so
  $\nabla\cdot\mathbf{B} \equiv 0$ is a vector identity, not a postulate.
  A magnetic monopole would require a point-like topological defect sourcing
  the transverse velocity as a scalar potential --- but the GPE supports only
  vortex-line defects (which source the \emph{rotational} part of
  $\mathbf{v}_\perp$, carrying electric charge).  The absence of magnetic
  monopoles is a topological theorem of the framework: no object with the
  required topology can exist in a superfluid order parameter in three spatial
  dimensions.  All current and future monopole searches should return null results.

\end{enumerate}

\subsection*{Discrete-symmetry and CP-sector predictions}

\begin{enumerate}\setcounter{enumi}{9}

\item \textbf{Neutron electric dipole moment from reconnection memory.}
  The strong CP problem dissolves in SSV (\S\ref{sec:CP}): there is no
  CP-symmetric reference vacuum, so $\theta_{\rm QCD}$ is not a free
  parameter to be tuned.  The reconnection-memory derivation of
  \S\ref{sec:neutron-edm} gives the leading-order structural form
  \begin{equation*}
    d_n \;=\; \kappa_\chi\,e\,R_n\,\frac{n_B}{n_\gamma}
       \;\approx\; \kappa_\chi\,\times\,5\times 10^{-23}\,e\!\cdot\!\text{cm},
  \end{equation*}
  with $\kappa_\chi$ a dimensionless chiral-shear coupling deferred to
  the converged 3D reconnection-barrier minimisation (issue~\#30).  At the natural scale $\kappa_\chi\sim\alpha^2/\pi$ this
  gives $d_n\sim 10^{-27}\,e\!\cdot\!\text{cm}$, one decade below the
  present bound $|d_n|<1.8\times 10^{-26}\,e\!\cdot\!\text{cm}$
  \cite{Abel2020} and within reach of nEDM/n2EDM at
  $\sim 10^{-28}\,e\!\cdot\!\text{cm}$.  The same chirality memory density
  $\chi$ underlies both this prediction and the cosmological baryon
  asymmetry, tying two distinct CP observables to a single framework
  coupling.

\item \textbf{No axion required.}  The Peccei-Quinn mechanism and its associated
  axion are introduced in the Standard Model to explain the smallness of
  $\theta_{\rm QCD}$.  Since SSV does not have a $\theta_{\rm QCD}$ parameter to
  begin with, no axion is required, predicted, or expected.  Detection of
  an axion at any mass would not falsify SSV directly --- axions could exist
  as additional medium excitations not yet identified in the framework ---
  but their detection would remove the SSV's most distinctive structural
  prediction in the strong-CP sector.

\item \textbf{Baryon asymmetry from chiral seeding.}
  \begin{sloppypar}
  The observed baryon-to-photon ratio
  $n_B/n_\gamma \sim 6\times 10^{-10}$ is interpreted as the surviving
  recorded chirality of the seeding event (\S\ref{sec:CP}).  Quantitatively,
  the prediction is that this ratio should follow from the same chirality
  bias that determines the CKM phase --- i.e.\ from a single number measured
  in two completely different experiments.  This is a sharp consistency
  prediction: whatever calculation eventually produces the CKM phase from
  reconnection geometry should also produce the baryon-to-photon ratio.  No
  separate cosmological CP-violating sector should be required.
  \end{sloppypar}

\end{enumerate}

\subsection*{Falsification structure}

The above predictions are not equally falsifiable, and it is worth being
explicit.  The structural predictions (no monopoles, no axion required,
neutrinos only from topology change, no fourth-generation leptons) admit
clean falsification: a single confirmed positive observation in any of these
categories would invalidate the framework as currently formulated.  The
quantitative predictions (gravitational-wave polarisation corrections, $W$ mass, Higgs
phenomenology, $n_B/n_\gamma$) require performing the 3D
$\mathcal{L}+\mathcal{L}_\perp$ minimisation before they can be compared
against observation; their failure mode is not a single-experiment refutation
but the inability to reproduce known numbers from the geometry.  The
gravitational-wave memory and gravitational Casimir predictions are
quantitatively close to GR/Standard-Model expectations and would require
precision improvements beyond current capability to discriminate.

% ═══════════════════════════════════════════════════════════════════════
\section{Bridge to Thermodynamics and Irreversible Time}
\label{sec:outlook}
% ═══════════════════════════════════════════════════════════════════════

The same superfluid order parameter $\Psi$ that underlies the four forces also
governs irreversible thermodynamics.  One further connection to a subsequent
paper is worth flagging here, beyond the forward-pointer already given in
\S\ref{sec:CP} (CP and memory).

\paragraph{Time and entropy (Paper III).}
The two flagged places where the present paper meets the irreversibility
framework --- the symmetry/process distinction in CP, and the time-averaging
step in Bjerknes gravity --- are not isolated remarks but the visible edges
of a single substrate question: in a medium whose state at any moment is the
record of all reconnections that preceded it, what counts as an instantaneous
quantity?  Paper~III treats this question directly: the arrow of time is
identified with the growth of the topological tangle, and the second law is
the statement that the medium accumulates record monotonically.  The
cosmological side of the same picture --- the CMB thermal spectrum read as
acoustic noise integrated from all reconnection events since the seeding ---
is the subject of a separate future paper (Paper~IX, on the primordial
phonon bath), not of the present paper.  The force-sector mean-field
treatment of the present paper is the static envelope of that fuller
dynamics.

% ═══════════════════════════════════════════════════════════════════════
\section{Discussion and Open Problems}
\label{sec:discussion}
% ═══════════════════════════════════════════════════════════════════════

The framework presented here is mechanically complete for the force sector:
each of the four interactions has a definite hydrodynamic origin, and the
extensions added in this paper --- the Higgs as the canvas's amplitude mode,
neutrinos as torsional radiation, charged leptons as higher breather harmonics,
and CP violation as a fluctuation residue in a medium that records --- fill
out the particle content and the discrete-symmetry structure of the Standard
Model on the same underlying hydrodynamic substrate.  No additional fields,
gauge groups, or coupling constants are introduced beyond those fixed by the
mass spectrum of Paper~I.

What remains is computation, not conceptual machinery.  The open problems
collected here divide into two groups: a single core calculation on which
most of them depend, and a small number of structural questions that are
genuinely independent.

\subsection*{The core open computation}

The most significant feature of the present framework's open-problem
landscape is that it collapses several apparently independent puzzles of the
Standard Model into a single deferred computation: the full 3D time-dependent
minimisation of the $\mathcal{L}+\mathcal{L}_\perp$ functional during a
vortex reconnection event.  The following observables are all determined by
the geometry of that minimisation, and none can be computed in isolation
from the others:

\begin{enumerate}
\item \textbf{The $W$ boson mass.}  The scalar estimate
  $E_{\rm recon}\sim m_e/\alpha^2\approx 9.6\,\mathrm{GeV}$ accounts only for
  the quantum-pressure barrier that initiates reconnection.  The dominant
  cost is $E_{\rm caps}\sim P_0\cdot\pi R_{\rm cap}^2\xi$ --- holding open
  the two fully-suppressed end-caps against the saturation pressure $P_0$.
  The formula $R_{\rm cap}=\phi\,\xi/\alpha$
  (equation~\eqref{eq:Rcap_phi}) gives $m_W\approx\pi\phi^2 m_e/\alpha^2\approx
  78.9\,\mathrm{GeV}$ --- within $1.8\%$ of the observed value.
  The cap equilibrium analysis (\S\ref{sec:W_rayleigh}) shows this is
  not a bare fit: $R_{\rm cap}=\phi/\alpha$ is the unique equilibrium of
  the cubic $R^3+\tau R^2=\lambda_{\rm bend}$ (line tension $\tau=17\,\xi$,
  computed) when $\lambda_{\rm bend}=\phi^3/\alpha^3$.  The local
  $\mathcal{L}_\perp$ correction falls short by $232\times$; the target
  is now to derive $\lambda_{\rm bend}=\phi^3/\alpha^3$ from the non-local
  cap topology.

\item \textbf{The neutrino sector (P1--P5 of \S\ref{sec:neutrino}).}  Mass,
  three-flavour structure, oscillation phenomenology, Dirac/Majorana
  character, and the Fermi constant $G_F$ all follow from the same
  reconnection geometry: the torsional pulses radiated during the event
  carry whatever flavour structure, mass spectrum, and mixing the linearised
  GP+chiral operator supports as eigenmodes around the 3D reconnection
  background.

\item \textbf{The charged-lepton generations (\S\ref{sec:leptons}).}  The
  muon eigenfrequency (Paper~I open problem 2) and the tau identification
  (currently a bare ladder coincidence) are eigenmodes of the same operator,
  evaluated on the toroidal vortex background rather than the reconnection
  background.  Paper~I and Paper~II treat them separately for expository
  reasons; they are mathematically the same problem.

\item \textbf{CP violation observables (\S\ref{sec:CP}).}  The CKM phase, the
  kaon $\varepsilon$-parameter, the $B$-system CP asymmetries, the neutron
  electric dipole moment, and the baryon-to-photon ratio
  $n_B/n_\gamma\sim 6\times 10^{-10}$ are all consequences of the chirality
  bias in reconnection outcomes.  In a medium that records, these are not
  small departures from a CP-symmetric reference dynamics; they are the
  visible component of asymmetries present in every event, made measurable
  when reconnection geometry happens to amplify them past detection
  threshold.

\item \textbf{The strong sector at two loops.}  Identification of colour
  charge with phase balance at Y-junctions, and the derivation of the QCD
  $\beta$-function from quantum-pressure softening, should be carried to
  two-loop order using the same 3D reconnection treatment.  Comparison with
  precision QCD measurements provides an independent calibration of the
  reconnection geometry.
\end{enumerate}

These are not five independent open problems but five facets of the same
computation.  Each of the corresponding Standard-Model parameters (the
electroweak scale, the PMNS matrix, the lepton mass ratios, the CKM phase,
$\Lambda_{\rm QCD}$) is currently fitted to data; the framework's claim is
that all are determined by the geometry of one event, computable in
principle without empirical input.  This is a sharp, falsifiable claim ---
the geometry will produce specific numbers that either match observation or
do not.

\subsection*{Structurally separate open problems}

Three further open problems stand outside the core computation, in the sense
that their resolution requires different tools.

\begin{enumerate}\setcounter{enumi}{5}
\item \textbf{Newton's constant: $\alpha_G$ from the proton breather.}  The
  structural formula~\eqref{eq:G_ladder} reproduces $G_{\rm obs}$ to $0.6\%$
  using the Paper~I mass ladder.  The remaining task is computing the
  sub-grain acoustic monopole suppression factor $(a_p/\xi)^n$ from the 3D
  \emph{static} proton breather profile, which would yield
  $\alpha_G=Gm_p^2/(\hbar c)\approx 5.906\times 10^{-39}$ without empirical
  input.  This is a different computation from the reconnection one (a
  static eigenmode rather than a dynamical event), but uses the same
  $\mathcal{L}+\mathcal{L}_\perp$ functional.

\item \textbf{Spin-2 polarisation from linearised Madelung.}  A systematic
  post-Newtonian expansion of the Madelung equations around a pulsating
  source is needed to confirm the spin-2 angular pattern of gravitational
  radiation from the longitudinal acoustic carrier.  This is a perturbative
  calculation around the static gravity picture of \S\ref{sec:gravity},
  independent of the reconnection geometry.

\item \textbf{The canvas-grain hierarchy
  (\S\ref{sec:higgs}).}  The Higgs mass corresponds to a canvas grain length
  $\xi_{\rm EW}\approx 1.6\times 10^{-18}\,\mathrm{m}$, roughly $10^5$ times
  finer than the vortex healing length $\xi$ set by $m_e$.  The framework
  does not solve the hierarchy problem; it restates it as the geometric
  question of why the medium has two natural length scales separated by this
  factor.  Note that the Standard Model fine-tuning instability does
  \emph{not} transfer: the cutoff $\xi$ is itself a property of the medium,
  so there are no quadratically divergent radiative corrections that would
  drag $m_H$ to a higher scale.  The problem is structural, not a
  fine-tuning.
\end{enumerate}

\subsection*{Boundary with Paper~III}

Two specific places in this paper depend on operations whose finer-resolution
treatment is the subject of Paper~III on irreversible thermodynamics.  They
are flagged here for completeness; neither obstructs the mean-field claims
of the present paper.

\begin{enumerate}\setcounter{enumi}{8}
\item \textbf{Time-averaging in a recording medium
  (Paper~IV).}  The Bjerknes time-averaging that produces
  Newton's law is a classical operation; in a medium that accumulates a
  record of every reconnection event, the corrections to the leading-order
  static result carry information about pulsation history, and reproduce
  the gravitational-wave memory effect (Christodoulou) as a natural
  consequence.

\item \textbf{Discrete symmetries at the level of physical processes
  (\S\ref{sec:CP}).}  The distinction between formal symmetries of the
  $\mathcal{L}+\mathcal{L}_\perp$ functional and symmetries of physical
  processes --- which are broken by the medium's accumulated record --- is
  the conceptual root of irreversibility.  Paper~III makes this distinction
  the foundation of its treatment of time and entropy.
\end{enumerate}

\subsection*{What is closed}

It is worth stating explicitly what the present paper closes, since the
length of the open-problems list could otherwise mislead.  The mean-field
mechanical content of all four forces is fixed: each is a calculable pressure
response of the saturated medium, with coupling constants determined by
$\{c, \alpha, \mu_0\}$ from Paper~I.  The Higgs is identified with the
amplitude mode of the same order parameter $\Psi$, with no separately
postulated field.  Magnetic monopoles are topologically forbidden, not merely
absent.  The ultraviolet divergence of perturbative quantum gravity does not
arise.  The strong CP problem and the cosmological constant fine-tuning
problem do not arise.  Parity violation, the $(V-A)$ structure of the weak
current, the equivalence principle, the universality of gravitational
coupling, and the impossibility of magnetic monopoles all follow from the
single hydrodynamic substrate without additional assumption.

The remaining open problems are computational, not foundational.  Each has a
definite formulation within the framework already established here and in
Paper~I, and a clear path to numerical resolution.  The framework's most
distinctive structural feature is the unification of what the Standard Model
treats as independent puzzles --- the $W$ mass, the neutrino sector, the
charged-lepton generations, and the CP-violation observables --- into facets
of a single 3D minimisation.  Whether that minimisation, when finally
performed, produces the right numbers is the empirical test the framework
sets for itself.

% ═══════════════════════════════════════════════════════════════════════
\section{Conclusion}
% ═══════════════════════════════════════════════════════════════════════

We have proposed that the four fundamental forces emerge from a single
superfluid vacuum as distinct mechanical modes.  The present paper is a
programmatic and structural report; the precise epistemic status of each
sector is summarised below.

\paragraph{What the paper establishes.}
The mechanical origin of the electromagnetic near-field interaction ---
chiral flow gradients around toroidal vortices, with Coulomb's law and the
Biot--Savart force following from the Bernoulli pressure field --- is
reached at the level of structural mechanism using the chiral-shear coupling
$\alpha$ and the medium parameters fixed in Paper~I.  The propagation speed
of light in the provisional radiative identification is tied to the
chiral-shear mode speed $c_\perp$ rather than the medium's sound speed $c$;
this is a known open issue (\S\ref{sec:EM_open_issue}) whose resolution
requires re-deriving the photon as the Goldstone mode of $\Psi$ and is
deferred to a future revision.  The qualitative features of the strong
force --- short range, confinement of vortex strand endpoints, and
asymptotic freedom as quantum-pressure softening --- follow from the same
framework.

\paragraph{What the paper proposes as suggestive ansatz.}
For the weak force, the Rayleigh-jet cap formula $R_{\rm cap} = \phi R^*$
recovers $m_W$ to $1.8\%$.  The cap equilibrium analysis
(\S\ref{sec:W_rayleigh}) shows this follows analytically from a
chiral-shear bending stiffness $\lambda_{\rm bend} = \phi^3/\alpha^3$; the
local $\mathcal{L}_\perp$ curvature correction falls short by a factor $232$,
identifying the remaining open step as deriving $\lambda_{\rm bend}$ from the
non-local (topological) structure of the cap.  The formula is therefore no
longer a bare fit to $m_W$, but a prediction conditional on the non-local
bending stiffness.

\paragraph{What the paper offers as structural consistency.}
For gravity, the Bjerknes mechanism gives the inverse-square law, universal
coupling, and time dilation as mechanical consequences of spherical wave
dilution.  The relation $G = \alpha_G \hbar c \alpha^2/(N_p^2 m_e^2)$ uses
$\alpha_G$ from CODATA and matches the Paper~I mass ladder to $0.6\%$.
This is consistency between the gravity sector and the mass ladder, not a
derivation of $G$; the 3D proton breather profile required to compute
$\alpha_G$ from first principles remains the central open calculation.
Gravitational and electromagnetic radiation are proposed to share the
Goldstone (phase) channel of the medium and therefore propagate at the
medium's sound speed $c$, consistent with GW170817; the relation between
the chiral-shear sector at $c_\perp$ and the photon as observed is an open
issue flagged in \S\ref{sec:EM_open_issue}.

\paragraph{What the paper proposes as candidate geometry.}
The Higgs is recast as two distinct features: a structural amplitude mode
of the LogSE Lagrangian (gapped at $m_H c^2$, present everywhere) and a
derived denting event with energy $E_H = P_0 V^*$.  Charged leptons are
proposed as higher breathing harmonics of the toroidal electron vortex.
Neutrinos are presented as a candidate torsional-radiation geometry; this
candidate currently contradicts the established phenomenology in five ways
(\S\ref{sec:neutrino}), and the section indicates where in the medium a
neutrino-like object might live rather than offering a competing
phenomenology.

\paragraph{What the paper raises as a programmatic question.}
The CP discussion observes that, in a medium that accumulates a record of
every reconnection, no discrete symmetry of the dynamics need survive at
the process level.  This is a conceptual reframing.  Whether it can be made
quantitatively compatible with the observed precision of CP-even processes
(small $\theta_{\rm QCD}$, small neutron EDM, predominantly CP-even
behaviour across most kaon and $B$-meson observables) requires a
suppression mechanism that is currently open.

\paragraph{What the paper identifies as the genuine open problems.}
The genuine hierarchy question in SSV --- why the toroidal vortex ring has
equilibrium radius $R^* \approx 137\,\xi$ while the knotted breather's core
is $a_p \approx \xi/1836$, both solutions of the same field equations ---
is stated but not solved.  Several apparently distinct open problems of
the Standard Model --- the $W$ mass, the neutrino sector, the
charged-lepton generations, and the CP observables --- collapse, in this
framework, into facets of a single 3D minimisation of the
$\mathcal{L}+\mathcal{L}_\perp$ functional during a vortex reconnection
event.  The defect-spectrum hierarchy is an additional target of that
minimisation.  Whether the minimisation, when finally performed, produces
the right numbers is the empirical test the framework sets for itself.
That computation is the principal task left to subsequent work.

\paragraph{Scope.}
All claims above use the medium parameters first fixed by the mass spectrum
of Paper~I: the sound speed $c$, the fine-structure constant $\alpha$, and
the base mass scale $\mu_0 = m_e/\alpha$.  No additional fields,
symmetries, or coupling constants are introduced in this paper.  The
framework is presented as a continuum effective theory valid down to
$\ell_{\rm grain} \equiv a_p \approx 2\times 10^{-16}\,\mathrm{m}$, the
scale of the heaviest stable defect.  What lies below this scale is a
separate question, in the same sense that Navier--Stokes makes no claim
about sub-molecular structure.  The mean-field force-sector treatment of
the present paper is the static envelope of a fuller picture in which the
medium accumulates a record of every reconnection event.  Paper~III treats
that record directly, identifying irreversible time and entropy with the
growth of topological complexity.

%------------------------------------------------------------------
\section{Acknowledgments}
%------------------------------------------------------------------
\label{sec:acknowledgments}
The author thanks Grok (xAI), Claude (Anthropic) and Gpt(OpenAI) for extensive assistance with literature searches, drafting, mathematical checks, and critical feedback. All conclusions and errors are the sole responsibility of the author.

\bigskip

\begin{thebibliography}{99}

\bibitem{TGD-I}
S.~Norland, \textit{Saturated Superfluid Vacuum I: Topological Defects in a
Saturated Superfluid Vacuum — The Geometric Origin of Matter and the $\alpha$-Harmonic
Mass Ladder}, 10.5281/zenodo.19651986, 2026.

\bibitem{Volovik}
G.~E.~Volovik, \textit{The Universe in a Helium Droplet}, Oxford University Press, 2003.

\bibitem{Barcelo}
C.~Barcel\'{o}, S.~Liberati, and M.~Visser, ``Analogue gravity,''
\textit{Living Rev.~Relativity} \textbf{14}, 3 (2011).

\bibitem{Bjerknes1906}
V.~F.~K.~Bjerknes, \textit{Fields of Force}, Columbia University Press, 1906.

\bibitem{Landau_Fluid}
L.~D.~Landau and E.~M.~Lifshitz, \textit{Fluid Mechanics}, 2nd ed., Pergamon Press, 1987, §74.

\bibitem{Villois}
A.~Villois, D.~Proment, and G.~Krstulovic, ``Universal and nonuniversal aspects of
vortex reconnections in superfluids,'' \textit{Phys.~Rev.~Fluids} \textbf{2}, 044701
(2017).

\bibitem{Gross}
E.~P.~Gross, ``Structure of a quantized vortex in boson systems,''
\textit{Nuovo Cimento} \textbf{20}, 454 (1961).

\bibitem{Pitaevskii}
L.~P.~Pitaevskii, ``Vortex lines in an imperfect Bose gas,''
\textit{Sov.~Phys.~JETP} \textbf{13}, 451 (1961).

\bibitem{AharonovBohm1959}
Aharonov, Y. and Bohm, D.,
``Significance of Electromagnetic Potentials in the Quantum Theory,''
\emph{Phys.\ Rev.}\ \textbf{115}, 485--491 (1959).

\bibitem{HaldaneWu1985}
Haldane, F.~D.~M. and Wu, Y.-S.,
``Quantum Dynamics and Statistics of Vortices in Two-Dimensional Superfluids,''
\emph{Phys.\ Rev.\ Lett.}\ \textbf{55}, 2887--2890 (1985).

\bibitem{Schwinger}
J.~Schwinger, ``On gauge invariance and vacuum polarization,''
\textit{Phys.~Rev.} \textbf{82}, 664 (1951).

\bibitem{Endres2012}
M.~Endres, T.~Fukuhara, D.~Pekker, M.~Cheneau, P.~Schauß, C.~Gross, E.~Demler,
S.~Kuhr, and I.~Bloch, ``The `Higgs' amplitude mode at the two-dimensional
superfluid/Mott insulator transition,'' \textit{Nature} \textbf{487},
454--458 (2012).

\bibitem{GW170817_MMA}
B.~P.~Abbott \textit{et al.}\ (LIGO Scientific Collaboration, Virgo Collaboration,
and partners), ``Multi-messenger observations of a binary neutron star merger,''
\textit{Astrophys.~J.~Lett.} \textbf{848}, L12 (2017).

\bibitem{Salart2008}
D.~Salart, A.~Baas, C.~Branciard, N.~Gisin, and H.~Zbinden,
``Testing the speed of `spooky action at a distance',''
\textit{Nature} \textbf{454}, 861--864 (2008).

\bibitem{Yin2013}
J.~Yin \textit{et al.},
``Lower bound on the speed of nonlocal correlations without locality and measurement
choice loopholes,''
\textit{Phys.~Rev.~Lett.} \textbf{110}, 260407 (2013).

\bibitem{Volovik2003}
G.~E.~Volovik,
\textit{The Universe in a Helium Droplet},
Oxford University Press, Oxford (2003).

\bibitem{Zloshchastiev2020}
K.~G.~Zloshchastiev,
``Superfluid vacuum theory and deformed dispersion relations,''
\textit{Int.~J.~Mod.~Phys.~A} \textbf{35}, 2040032 (2020); arXiv:2011.11897.

\bibitem{Abel2020}
C.~Abel \textit{et al.} (nEDM collaboration),
``Measurement of the permanent electric dipole moment of the neutron,''
\textit{Phys.~Rev.~Lett.} \textbf{124}, 081803 (2020).

\bibitem{KATRIN2022}
M.~Aker \textit{et al.} (KATRIN collaboration),
``Direct neutrino-mass measurement with sub-electronvolt sensitivity,''
\textit{Nature Phys.} \textbf{18}, 160 (2022).

\bibitem{Planck2020}
N.~Aghanim \textit{et al.} (Planck collaboration),
``Planck 2018 results. VI. Cosmological parameters,''
\textit{Astron.\ Astrophys.} \textbf{641}, A6 (2020).

\bibitem{MicroBooNE2022}
P.~Abratenko \textit{et al.} (MicroBooNE collaboration),
``Search for an anomalous excess of charged-current $\nu_e$ interactions
without pions in MicroBooNE,''
\textit{Phys.\ Rev.\ D} \textbf{105}, 112004 (2022).

\bibitem{DUNE2020}
B.~Abi \textit{et al.} (DUNE collaboration),
``Long-baseline neutrino oscillation physics potential of the DUNE
experiment,''
\textit{Eur.\ Phys.\ J.\ C} \textbf{80}, 978 (2020).

\bibitem{HK2018}
K.~Abe \textit{et al.} (Hyper-Kamiokande proto-collaboration),
``Hyper-Kamiokande design report,''
arXiv:1805.04163 (2018).

\bibitem{CUORE2022}
D.~Q.~Adams \textit{et al.} (CUORE collaboration),
``Search for Majorana neutrinos exploiting millikelvin cryogenics with
CUORE,''
\textit{Nature} \textbf{604}, 53 (2022).

\bibitem{KamLANDZen2023}
S.~Abe \textit{et al.} (KamLAND-Zen collaboration),
``Search for the Majorana nature of neutrinos in the inverted mass
ordering region with KamLAND-Zen,''
\textit{Phys.\ Rev.\ Lett.} \textbf{130}, 051801 (2023).

\bibitem{LEGEND2021}
N.~Abgrall \textit{et al.} (LEGEND collaboration),
``The Large Enriched Germanium Experiment for Neutrinoless $\beta\beta$
Decay: LEGEND-1000 preconceptual design report,''
arXiv:2107.11462 (2021).

\end{thebibliography}

\end{document}
