\documentclass[11pt]{article}

\usepackage[a4paper,margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{amsmath,amssymb}
\usepackage{physics}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{setspace}
\usepackage{mdframed}

\hypersetup{colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue}
\setstretch{1.1}

\newmdenv[linewidth=1.1pt,innertopmargin=5pt,innerbottommargin=5pt,%
  innerleftmargin=8pt,innerrightmargin=8pt]{resultbox}

\title{SSV Conclusions~B\\[2pt]
\large A concrete superfluid holographic screen undershoots Newton's\\
constant by 38 orders: a mode-counting negative}
\author{Stig Norland}
\date{Closing note, 2026-06-16}

\begin{document}
\maketitle

\begin{abstract}
This note records the sharpest quantitative result the Saturated Superfluid
Vacuum (SSV) programme produced about emergent gravity, and it is a negative.
Entropic and holographic gravity (Jacobson, Verlinde, Padmanabhan) keep the
holographic screen \emph{abstract}. SSV's distinctive move was to give the screen
a concrete microscopic substrate --- a logarithmic-Schr\"odinger condensate whose
self-defined acoustic horizon carries an area-law entanglement entropy that can be
\emph{computed} rather than assumed. The form survives: the entanglement entropy
is area-law, as expected for a local quadratic Hamiltonian. The magnitude does
not. Counting the actual degrees of freedom the substrate offers per grain gives
$\mathcal{O}(10)$ Bogoliubov modes --- a few bits --- against the
$N \simeq 1.7\times10^{38}$ required to reproduce $G$. The screen, made literal
and counted, misses Newton's constant by roughly $10^{37}$. This is the
Sakharov / Susskind--Uglum species problem, here \emph{demonstrated by direct
mode-counting} rather than argued, and it is a clean constraint on any reading
that takes a holographic screen to be a real reservoir of local degrees of
freedom. Per the repository's rule, the negative is stated at full strength.
\end{abstract}

\section{The setup: making the screen literal}

In the thermodynamic route to gravity, the field equation follows from
$\delta Q = T\,dS$ applied to local causal horizons (Jacobson~1995), with the
entropy taken to be area-law, $S = A/4\ell_P^2$. Verlinde's entropic-force reading
and Padmanabhan's surface-thermodynamics share this skeleton. In all of them the
holographic screen is a bookkeeping surface: one assumes an area-law entropy with
the Planck-scale coefficient and does not ask what physical degrees of freedom
supply it.

SSV's one genuinely distinctive contribution is to refuse that abstraction. The
vacuum is a real medium with order parameter $\Psi=\sqrt{\rho}\,e^{i\theta}$
governed by a logarithmic Schr\"odinger equation; the screen is a self-defined
acoustic horizon, the surface where the background flow turns sonic (Visser).
Because the medium is concrete, two things that are normally assumed become
\emph{computable}: the area law, and the coefficient.

\section{The form survives (and is the cheap half)}

Quantising the LogSE Bogoliubov fluctuations and computing the vacuum
entanglement entropy by the correlator method (Peschel; Casini--Huerta) via the
Srednicki radial decomposition gives an area law. Across a static acoustic
inhomogeneity (a ``dumb hole'') the scaling exponent tracks a provably area-law
gapped-scalar calibrator and sits a full $\sim\!1.2$ in log--log slope below a
volume-law thermal control; the entropy-per-area is universal at the few-percent
level; and the chiral-shear term is \emph{exactly} silent
(\texttt{instruments/paper\_v/horizon\_entanglement\_entropy.py}).

This must be stated honestly (repository rule~1): the area law is the
\emph{expected} side. It follows from the LogSE quadratic Hamiltonian being
\emph{local} --- generic for any local free field (Srednicki~1993;
Bombelli--Koul--Lee--Sorkin~1986; Jacobson's entanglement-equilibrium, 2016). So
the computation \emph{earns} (turns assumed into derived-from-locality) the area
law that the SSV gravity papers had merely asserted by cell-counting, but it is
not a surprise and not a discovery. The discriminator was shown to be
negative-capable --- it correctly rejects volume-law data --- which is what makes
the next step trustworthy.

\section{The magnitude fails (and is the whole point)}

The coefficient is the test. Writing $\delta Q = T_H\,dS$ with the computed
area-law $dS = \eta\,dA$ closes the Jacobson construction to
$\nabla^2\Phi = 4\pi G_{\mathrm{eff}}\,e/c^2$ with
$G_{\mathrm{eff}} = 1/(4\hbar c\,\eta)$
(\texttt{instruments/paper\_v/dumb\_hole\_surface\_gravity.py}). To recover the
measured $G$, the entropy density $\eta$ must correspond to
\begin{equation}
  N \;=\; \frac{1}{\alpha_G}
        \;=\; \frac{\hbar c}{G\,m_p^2}
        \;=\; \left(\frac{m_P}{m_p}\right)^{2}
        \;\simeq\; 1.7\times 10^{38}
\end{equation}
microstates per grain --- this is just the Planck-to-proton hierarchy.

Now count what the substrate actually provides. The localised excitations of a
LogSE vortex core --- the bound Bogoliubov (BdG) modes --- number
$\mathcal{O}(10)$, i.e.\ a few bits, per healing-length cell
(\texttt{instruments/paper\_i/vortex\_core\_mode\_spectrum.py}). The continuum
modes are delocalised and do not localise entropy on the grain. So the concrete
screen supplies of order $10^{1}$ states where $10^{38}$ are needed: a shortfall
of about $10^{37}$, and correspondingly $G_{\mathrm{eff}}$ overshoots the
Newtonian value by $(a_p/\ell_P)^2 \approx 1.7\times10^{38}$.

\section{What the negative establishes}

\begin{resultbox}
\noindent When a holographic screen is given a real microscopic substrate and
its localised degrees of freedom are \emph{counted}, the area law emerges for
free but the Planck-scale coefficient does not: the count is short by $\sim$38
orders of magnitude. A holographic screen that is a literal reservoir of local
degrees of freedom cannot, in this construction, reproduce $G$. Newton's constant
must enter below the effective theory's cutoff --- as a trans-Planckian input,
not an emergent output.
\end{resultbox}

This is the Sakharov induced-gravity problem and the Susskind--Uglum species
problem, which Volovik identifies as the structural weak point of superfluid
gravity. The contribution here is not the existence of the gap but its
\emph{concrete demonstration}: rather than estimating an induced Newton constant
from a cutoff, one builds the medium, computes its horizon entanglement entropy,
counts its grain degrees of freedom, and reads off the shortfall directly. That
makes the failure model-specific and checkable rather than a heuristic worry.

\section{Caveats, stated plainly}

\begin{itemize}[leftmargin=1.4em]
\item The identity $(a_p/\ell_P)^2 = 1/\alpha_G$ is a \emph{tautology}: it
  restates the hierarchy and is not a derivation. It must not be read backwards as
  ``deriving $N$.'' The negative is the \emph{measured} count
  ($\mathcal{O}(10)$), confronted with this target.
\item Bound modes are finite by construction; continuum modes are delocalised.
  A genuinely different substrate (e.g.\ a rank-2, $^3$He-A-like order parameter
  with sub-grain structure) is not excluded by this argument --- it simply is not
  the LogSE medium, and would have to supply its own $10^{38}$ states.
\item The area-law computation assumes irrotational flow for the exact chiral
  silence; a rotational/vortex-sink horizon carries a bounded residual risk that
  was not closed.
\end{itemize}

\noindent None of these weaken the headline: in the medium actually proposed, the
screen is $\sim$37 orders too sparse to be the source of $G$.

\begin{thebibliography}{9}
\bibitem{jacobson} T.\ Jacobson, \emph{Thermodynamics of spacetime: the Einstein
  equation of state}, Phys.\ Rev.\ Lett.\ \textbf{75}, 1260 (1995).
\bibitem{jacobson16} T.\ Jacobson, \emph{Entanglement equilibrium and the
  Einstein equation}, Phys.\ Rev.\ Lett.\ \textbf{116}, 201101 (2016).
\bibitem{verlinde} E.\ Verlinde, \emph{On the origin of gravity and the laws of
  Newton}, JHEP \textbf{04}, 029 (2011).
\bibitem{padmanabhan} T.\ Padmanabhan, \emph{Thermodynamical aspects of gravity:
  new insights}, Rep.\ Prog.\ Phys.\ \textbf{73}, 046901 (2010).
\bibitem{visser} M.\ Visser, \emph{Acoustic black holes: horizons, ergospheres
  and Hawking radiation}, Class.\ Quantum Grav.\ \textbf{15}, 1767 (1998);
  gr-qc/9712010.
\bibitem{srednicki} M.\ Srednicki, \emph{Entropy and area}, Phys.\ Rev.\ Lett.\
  \textbf{71}, 666 (1993).
\bibitem{bkls} L.\ Bombelli, R.\,K.\ Koul, J.\ Lee, R.\,D.\ Sorkin,
  \emph{Quantum source of entropy for black holes}, Phys.\ Rev.\ D \textbf{34},
  373 (1986).
\bibitem{peschel} I.\ Peschel, \emph{Calculation of reduced density matrices from
  correlation functions}, J.\ Phys.\ A \textbf{36}, L205 (2003).
\bibitem{casinihuerta} H.\ Casini and M.\ Huerta, \emph{Entanglement entropy in
  free quantum field theory}, J.\ Phys.\ A \textbf{42}, 504007 (2009).
\bibitem{sakharov} A.\,D.\ Sakharov, \emph{Vacuum quantum fluctuations in curved
  space and the theory of gravitation}, Sov.\ Phys.\ Dokl.\ \textbf{12}, 1040
  (1968).
\bibitem{susskinduglum} L.\ Susskind and J.\ Uglum, \emph{Black hole entropy in
  canonical quantum gravity and superstring theory}, Phys.\ Rev.\ D \textbf{50},
  2700 (1994).
\bibitem{volovik} G.\,E.\ Volovik, \emph{The Universe in a Helium Droplet}
  (Oxford, 2003).
\end{thebibliography}

\end{document}
