\documentclass[11pt,a4paper]{article}
\usepackage[margin=2.5cm]{geometry}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amsmath,amssymb,physics}
\usepackage{booktabs}
\usepackage{float}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{caption}
\usepackage{tcolorbox}
\tcbuselibrary{skins,breakable}

\newtcolorbox{resultbox}[1][]{enhanced,breakable,
  colback=blue!4!white,colframe=blue!40!black,
  fonttitle=\bfseries,title={#1},boxrule=0.6pt,arc=4pt}
\newtcolorbox{gapbox}[1][]{enhanced,breakable,
  colback=orange!6!white,colframe=orange!60!black,
  fonttitle=\bfseries,title={#1},boxrule=0.6pt,arc=4pt}

\hypersetup{colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue}

\title{\textbf{Saturated Superfluid Vacuum~VII-a}\\[0.4em]
\large Quantum Mechanics from Hydrodynamics}
\author{Stig Norland\\
\small Independent Researcher, Bergen, Norway}
\date{Draft}

\begin{document}
\maketitle

\begin{abstract}
This paper derives the quantum-mechanical limit of the Saturated Superfluid
Vacuum (SSV) framework.  The scope is deliberately narrow: the Madelung
transformation, the linear Schr\"odinger limit, uncertainty as a hydraulic
localisation bound, and measurement as a physical reconnection channel.

The formal result is the recovery of the Schr\"odinger equation from the
low-amplitude hydrodynamic regime of the SSV medium.  Quantisation enters as
the single-valuedness condition on the order-parameter phase, tying discrete
winding numbers to circulation topology rather than to a separate quantisation
rule.  Uncertainty is interpreted as the hydraulic cost of localising the
order parameter.  Measurement is modelled as an irreversible reconnection or
wake-writing event in the medium.

Claims about emergent geometry, Einstein equations, cosmogony, and full-series
synthesis are left to companion papers.
\end{abstract}

\newpage
\tableofcontents
\newpage

%%──────────────────────────────────────────────────────────────────────────────
\section{Introduction}
\label{sec:intro}
%%──────────────────────────────────────────────────────────────────────────────

The SSV programme treats the vacuum as a saturated superfluid described by a
complex order parameter $\Psi$.  Papers~I--IV establish the medium, its
topological defects, thermodynamic arrow, and gravitational limit.  Paper~V
extends the same picture to strong gravity and condensate black holes.  The
present paper asks a narrower question: under what conditions does ordinary
quantum mechanics appear as the low-amplitude hydrodynamic limit of this medium?

The working thesis is that quantum behaviour is not an additional foundational
axiom.  It is the wave-mechanical regime of the order parameter when
perturbations are small, gradients are gentle, and defects can be treated
through an effective potential.  In that regime the Madelung variables map the
complex wave equation to continuity plus Euler-type flow equations, and the
inverse map recovers the Schr\"odinger equation for a slowly varying envelope.

This paper separates three claim levels explicitly.

\begin{enumerate}[label=\textbf{C\arabic*.}]
  \item \textbf{Formal derivation.}  The Madelung transformation and its inverse
    show how a Schr\"odinger-type equation is equivalent to hydrodynamic flow
    with a quantum-pressure term.  This is an exact algebraic equivalence, not
    an approximation.

  \item \textbf{Controlled limit.}  In the low-amplitude, long-wavelength regime,
    the nonlinear medium terms can be absorbed into background phase and
    effective potentials, leaving the ordinary linear Schr\"odinger equation.
    This is the limit that makes SSV contact with standard quantum mechanics.

  \item \textbf{Physical interpretation.}  Quantisation, uncertainty, and
    measurement are read as circulation topology, localisation stiffness, and
    irreversible reconnection respectively.  These are model commitments that
    arise naturally from the medium picture; they are not independent mathematical
    theorems.
\end{enumerate}

\subsection*{Roadmap}

Section~\ref{sec:madelung} carries out the Madelung transformation and derives
the hydrodynamic equations.  Section~\ref{sec:schrodinger} takes the
Schr\"odinger limit and shows how quantisation follows from phase topology.
Section~\ref{sec:uncertainty} interprets the uncertainty relation as a hydraulic
localisation bound.  Section~\ref{sec:measurement} models measurement as a
reconnection event and discusses Born weights.  Section~\ref{sec:discussion}
places the results in the context of the series and lists open calculations.

%%──────────────────────────────────────────────────────────────────────────────
\section{Madelung Transformation}
\label{sec:madelung}
%%──────────────────────────────────────────────────────────────────────────────

\subsection{Polar decomposition}

Let the local state of the saturated medium be the complex order parameter
\begin{equation}
  \Psi(\mathbf{x},t) = \sqrt{\rho(\mathbf{x},t)}\,e^{iS(\mathbf{x},t)/\hbar},
  \label{eq:polar}
\end{equation}
where $\rho$ is the mass density and $S$ is the phase action.  The associated
irrotational velocity field is
\begin{equation}
  \mathbf{v} = \frac{1}{m}\nabla S.
  \label{eq:velocity}
\end{equation}

For the logarithmic-Schr\"odinger (LogSE) medium of Papers~I--IV~\cite{SSV-I,SSV-IV},
the order parameter obeys
\begin{equation}
  i\hbar\,\partial_t\Psi =
  \left[-\frac{\hbar^2}{2m}\nabla^2 + V_\mathrm{ext} + U(\rho)\right]\Psi,
  \label{eq:medium_wave}
\end{equation}
where $V_\mathrm{ext}$ is a slowly varying effective potential and $U(\rho)$
is the medium self-interaction.  The specific equation of state is fixed in
earlier papers; the present paper uses only the hydrodynamic structure common
to this class of order-parameter equations.

\subsection{Hydrodynamic form}

Substituting \eqref{eq:polar} into \eqref{eq:medium_wave} and separating real
and imaginary parts gives the Madelung pair
\begin{align}
  \partial_t\rho + \nabla\cdot(\rho\mathbf{v}) &= 0,
  \label{eq:continuity}\\
  \partial_t S + \frac{(\nabla S)^2}{2m} + V_\mathrm{ext} + U(\rho) + Q &= 0,
  \label{eq:hamilton_jacobi}
\end{align}
with the quantum pressure
\begin{equation}
  Q(\mathbf{x},t) = -\frac{\hbar^2}{2m}\frac{\nabla^2\sqrt{\rho}}{\sqrt{\rho}}.
  \label{eq:Q}
\end{equation}
Taking a gradient of \eqref{eq:hamilton_jacobi} gives the Euler equation
\begin{equation}
  m\bigl(\partial_t + \mathbf{v}\cdot\nabla\bigr)\mathbf{v}
  = -\nabla\bigl(V_\mathrm{ext} + U(\rho) + Q\bigr).
  \label{eq:euler}
\end{equation}

\begin{resultbox}[Madelung equivalence]
Equations~\eqref{eq:continuity}--\eqref{eq:euler} are the Madelung equations:
compressible irrotational hydrodynamics with an additional quantum-pressure
contribution $Q$.  The transformation is exact and bidirectional: starting from
the wave equation \eqref{eq:medium_wave} yields the hydrodynamic pair; starting
from the hydrodynamic pair with the irrotationality condition~\eqref{eq:velocity}
reconstructs the wave equation.
\end{resultbox}

%%──────────────────────────────────────────────────────────────────────────────
\section{Schr\"odinger Limit}
\label{sec:schrodinger}
%%──────────────────────────────────────────────────────────────────────────────

\subsection{Low-amplitude long-wavelength limit}

Write the density as $\rho = \rho_0 + \delta\rho$, with $|\delta\rho|\ll\rho_0$,
and assume gradients are long compared with the healing length of the medium.
Expand the self-interaction around the saturated background:
\begin{equation}
  U(\rho) = U(\rho_0) + U'(\rho_0)\,\delta\rho + \cdots
  \label{eq:U_expand}
\end{equation}
The constant term shifts the global phase; the linear term is either negligible
for a sufficiently weak perturbation or is absorbed into an effective potential.
The slowly varying envelope $\psi$ then obeys
\begin{equation}
  i\hbar\,\partial_t\psi =
  \left[-\frac{\hbar^2}{2m}\nabla^2 + V_\mathrm{eff}\right]\psi,
  \label{eq:schrodinger}
\end{equation}
the ordinary Schr\"odinger equation for the effective degree of freedom.

\begin{resultbox}[Schr\"odinger limit]
The SSV medium in its low-amplitude, long-wavelength regime reduces exactly to
linear Schr\"odinger dynamics.  The claim is not that all of quantum theory is
derived in one step; it is that whenever the conditions above hold, the same
hydrodynamic variables reduce to the linear dynamics used in standard quantum
mechanics.
\end{resultbox}

\begin{gapbox}[Deferred: bound-state spectra from LogSE]
Equation~\eqref{eq:schrodinger} recovers the Schr\"odinger form, but the
effective potential $V_\mathrm{eff}$ must be computed from the full LogSE
background for each physical situation.  Deriving the hydrogen-like bound-state
spectrum directly from the LogSE---without invoking $V_\mathrm{eff}$ as an
external input---is part of the Paper~I numerical programme~\cite{SSV-I}.
\end{gapbox}

\subsection{Quantisation as circulation topology}

The phase in \eqref{eq:polar} must be single-valued around any closed loop
that does not cross a zero of $\Psi$.  Therefore
\begin{equation}
  \oint_C \nabla S\cdot d\mathbf{l} = 2\pi n\hbar, \qquad n\in\mathbb{Z}.
  \label{eq:phase_quantisation}
\end{equation}
Using \eqref{eq:velocity}, this is the circulation condition
\begin{equation}
  \oint_C \mathbf{v}\cdot d\mathbf{l} = \frac{2\pi n\hbar}{m}.
  \label{eq:circulation}
\end{equation}
Quantisation is therefore tied to the topology of the phase field.  It is not
inserted as a separate rule for allowed orbits; it is the consequence of a
single-valued order parameter with vortex cores acting as excluded phase
singularities.  The detailed spectrum of any bound system depends on its
effective potential and boundary conditions, but the origin of discrete winding
numbers is hydrodynamic.  This is consistent with the topological spin derivation
of Paper~V~\cite{SSV-V} and the vortex-defect picture of Paper~I~\cite{SSV-I}.

%%──────────────────────────────────────────────────────────────────────────────
\section{Uncertainty as Hydraulic Bound}
\label{sec:uncertainty}
%%──────────────────────────────────────────────────────────────────────────────

The standard uncertainty relation follows from Fourier analysis of wave packets:
\begin{equation}
  \Delta x\,\Delta p \geq \frac{\hbar}{2}.
  \label{eq:uncertainty}
\end{equation}
SSV adds a physical interpretation of the same relation.  A perturbation
localised to a short length scale must include large gradients of $\sqrt{\rho}$
and phase.  Through \eqref{eq:Q}, those gradients carry a quantum-pressure cost
$\Delta E_Q \sim \hbar^2/(2m\,\Delta x^2)$.  Pushing $\Delta x$ downward
therefore forces a larger spread in momentum modes and a larger hydraulic
stiffness penalty, arriving at \eqref{eq:uncertainty}.

This section does not replace the standard proof.  The proof remains the
wave-packet Fourier theorem.  The SSV claim is that the theorem has a
medium-level interpretation: localisation is limited because the order parameter
cannot be squeezed arbitrarily without exciting the pressure and phase-gradient
channels of the fluid.  The uncertainty principle is the hydraulic stiffness of
the vacuum medium made quantitative.

\begin{gapbox}[Deferred: precise $\hbar/2$ prefactor from LogSE]
The hydraulic-stiffness argument recovers the form of \eqref{eq:uncertainty} and
the $\hbar$ scaling, but deriving the precise prefactor $1/2$ from the LogSE
equation of state---without importing it from the standard wave-packet
calculation---requires computing the minimum uncertainty product for a LogSE
coherent state.  This is a tractable calculation deferred to the Paper~I
programme~\cite{SSV-I}.
\end{gapbox}

%%──────────────────────────────────────────────────────────────────────────────
\section{Measurement as Reconnection}
\label{sec:measurement}
%%──────────────────────────────────────────────────────────────────────────────

\subsection{Irreversibility from topology change}

The measurement argument is kept narrow.  A measurement is a physical interaction
that correlates a microscopic degree of freedom with a macroscopic apparatus by
driving some part of the combined medium into an irreversible reconnection or
wake-writing event.

Before measurement, a superposition is represented by a coherent field
configuration,
\begin{equation}
  \Psi = \sum_k c_k\psi_k.
  \label{eq:superposition}
\end{equation}
As long as the relevant branches remain phase-coherent, interference terms are
available.  A measurement interaction is the regime in which the probe, apparatus,
and surrounding medium cross a threshold where topology changes, phonons are
emitted, or persistent wake degrees of freedom are written---the same
irreversibility mechanism described for time in Paper~III~\cite{SSV-III}.
Reversing the event would require restoring the original phase relation across
all of those channels, not merely evolving a small isolated wave packet backward.

\subsection{Born weights as a working model}

The natural branch weights in \eqref{eq:superposition} are $|c_k|^2$ because
the local energy and intensity available to drive an apparatus response are
quadratic in the wave amplitude.  This gives the usual Born form
\begin{equation}
  P_k = |c_k|^2.
  \label{eq:born}
\end{equation}
At the present stage this is a consistency argument, not a full derivation.

\begin{gapbox}[Deferred: Born-rule basin calculation]
The missing calculation is a reconnection-threshold model showing that the
basin volume for outcome $k$ is proportional to the incoming branch intensity
$|c_k|^2$ under realistic apparatus coupling.  This would promote the Born-rule
argument from physical interpretation to formal derivation.  It requires a
reduced model of a two-channel apparatus coupled to the SSV medium and is
deferred to a companion analysis.
\end{gapbox}

\subsection{What this does and does not claim}

This picture does not require consciousness, an external classical realm, or a
new postulate of collapse.  It also does not yet derive a complete theory of
detector statistics.  Its useful claim is more modest: SSV supplies a candidate
physical channel for irreversibility during measurement, namely topology change
and wake formation in the medium.  This is consistent with, but not identical
to, decoherence programmes that source irreversibility from environmental
entanglement; the SSV version sources it from physical medium events.

%%──────────────────────────────────────────────────────────────────────────────
\section{Discussion}
\label{sec:discussion}
%%──────────────────────────────────────────────────────────────────────────────

\subsection{Position in the series}

Paper~VII-a is the QM-limit paper in the SSV series.  Its inputs are the
Madelung structure of the LogSE (Papers~I--IV) and the topological defect picture
(Paper~I, Paper~V).  Its output---the Schr\"odinger limit, circulation
quantisation, hydraulic uncertainty, and reconnection measurement---is passed
forward as a standing result.  Claims about emergent geometry and the
Einstein-equation limit are in Paper~VII-b; cosmogony is in Paper~VIII.

\subsection{Relation to the Madelung and pilot-wave literature}

The Madelung~\cite{Madelung1927} representation is well-established; this is not
a new mathematical result.  The SSV contribution is the physical identification
of the medium: the Madelung fluid is not an abstract probability fluid but the
actual vacuum condensate, with the same medium that supports topological defects,
irreversible wakes, and gravitational gradients.  The difference from the de
Broglie--Bohm pilot-wave interpretation is that SSV does not introduce a guiding
wave on top of classical trajectories; the order parameter IS the physical medium
in which all dynamics occurs.

\subsection{Testable consequences}

Two near-term checks follow from the reconnection model.

\begin{enumerate}[label=\textbf{Q\arabic*.}]
  \item \textbf{Reconnection signatures in controlled measurement systems.}
    If detector clicks are accompanied by irreversible medium events, then
    sufficiently sensitive optomechanical or superconducting measurement
    platforms should show excess correlated acoustic, thermal, or phase-noise
    signatures at the transition from reversible coupling to recorded outcome.

  \item \textbf{Born-rule basin scaling.}
    A reduced reconnection-threshold model should test whether branch basin
    volumes scale as $|c_k|^2$ for simple two-channel apparatus couplings.
    This is the calculation needed before the measurement section can be
    promoted from physical interpretation to formal derivation.
\end{enumerate}

\subsection{Open problems}

\begin{enumerate}
  \item \emph{Bound-state spectra.}  Deriving the hydrogen-like spectrum directly
    from the LogSE, without treating $V_\mathrm{eff}$ as an external input, is
    part of the Paper~I programme.

  \item \emph{Precise uncertainty prefactor.}  The $1/2$ in \eqref{eq:uncertainty}
    should emerge from the LogSE coherent-state minimum; the calculation is
    deferred to the Paper~I programme.

  \item \emph{Multi-particle states.}  The present treatment addresses
    single-order-parameter modes.  Multi-particle states and exchange statistics
    require a field-theoretic extension of the Madelung picture, not developed
    here.

  \item \emph{Spin-1/2 and Fermi statistics.}  The order parameter $\Psi$ is a
    scalar; half-integer spin and fermion anti-symmetry require an extension to
    spinor order parameters or a separate topological argument.  This is noted as
    a gap, not filled here.
\end{enumerate}

%%──────────────────────────────────────────────────────────────────────────────
\section{Conclusion}
\label{sec:conclusion}
%%──────────────────────────────────────────────────────────────────────────────

SSV~VII-a isolates the quantum-mechanics limit from the geometry and cosmogony
claims elsewhere in the series.  The paper's formal core is the Madelung bridge
and the Schr\"odinger limit.  Its interpretive core has three parts:

\begin{enumerate}
  \item Quantisation reflects phase topology: discrete winding numbers follow
    from the single-valuedness of the order parameter, not from a separate
    quantisation axiom.

  \item Uncertainty reflects hydraulic localisation stiffness: the quantum
    pressure~\eqref{eq:Q} penalises tight localisation, recovering the
    Heisenberg bound from the equation of state.

  \item Measurement reflects irreversible reconnection or wake-writing in the
    medium, sourcing irreversibility from topology change rather than from an
    external classical realm.
\end{enumerate}

That is sufficient to make the paper standalone without requiring it to carry
the Einstein-equation or cosmogony arguments.

%%──────────────────────────────────────────────────────────────────────────────
\begin{thebibliography}{99}

\bibitem{SSV-I}
S.\ Norland, \textit{Saturated Superfluid Vacuum~I: Topological Defects in a
Saturated Superfluid Vacuum---The Geometric Origin of Matter and the
$\alpha$-Harmonic Mass Ladder}, 10.5281/zenodo.19651986, 2026.

\bibitem{SSV-II}
S.\ Norland, \textit{Saturated Superfluid Vacuum~II: Forces as Hydrodynamic
Pressure Gradients in the Saturated Superfluid Plenum}, SSV programme, 2026.

\bibitem{SSV-III}
S.\ Norland, \textit{Saturated Superfluid Vacuum~III: Irreversible Time and
Wake Entropy}, SSV programme, 2026.

\bibitem{SSV-IV}
S.\ Norland, \textit{Saturated Superfluid Vacuum~IV: Gravity as Update-Capacity
Gradient}, SSV programme, 2026.

\bibitem{SSV-V}
S.\ Norland, \textit{Saturated Superfluid Vacuum~V: Condensate Black Holes},
SSV programme, 2026.

\bibitem{Gross1961}
E.~P.\ Gross, ``Structure of a Quantised Vortex in Boson Systems,''
\textit{Nuovo Cim.}\ \textbf{20}, 454--457 (1961).

\bibitem{Madelung1927}
E.\ Madelung, ``Quantentheorie in hydrodynamischer Form,''
\textit{Z.\ Phys.}\ \textbf{40}, 322--326 (1927).

\bibitem{Onsager1949}
L.\ Onsager, ``Statistical Hydrodynamics,''
\textit{Nuovo Cim.\ Suppl.}\ \textbf{6}, 279--287 (1949).

\bibitem{Pitaevskii1961}
L.~P.\ Pitaevskii, ``Vortex Lines in an Imperfect Bose Gas,''
\textit{Sov.\ Phys.\ JETP}\ \textbf{13}, 451--454 (1961).

\bibitem{Zloshchastiev2011}
K.~G.\ Zloshchastiev, ``Logarithmic nonlinearity in theories of quantum
gravity,'' \textit{Grav.\ Cosmol.}\ \textbf{16}, 288 (2010).

\end{thebibliography}

\end{document}
